### 2015-09-28 doc/dispRity.demo.Rnw
vignette INIT

\documentclass{article}
\usepackage[sc]{mathpazo}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm}
\setcounter{secnumdepth}{2}
\setcounter{tocdepth}{2}
\usepackage{url}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=true,bookmarksnumbered=true,bookmarksopen=true,bookmarksopenlevel=2,
 breaklinks=false,pdfborder={0 0 1},backref=false,colorlinks=false]
 {hyperref}
\hypersetup{
 pdfstartview={XYZ null null 1}}

\newcommand{\dispRity}{\texttt{dispRity} }
\newcommand{\R}{\texttt{R} }

\begin{document}
<<setup, include=FALSE, cache=FALSE>>=
library(knitr)
# set global chunk options
opts_chunk$set(fig.path='figure/minimal-', fig.align='center', fig.show='hold')
options(formatR.arrow=TRUE,width=90)
@

\title{dispRity demo}

\author{Thomas Guillerme}

\maketitle

This a quick demo to go through the beta version of the \dispRity package (v.\Sexpr{packageVersion("dispRity")}).
Many parts are still missing (see section \ref{whatsleft}) but this version should be running all right and give an idea of what's implemented in the package.

\section{Before starting}

\dispRity is a package for calculating disparity in \R.
What is disparity? Well that's a more complex question...
To keep it short, this package allows to summarise ordinated matrices (e.g. MDS, PCA, PCO, PCoA) into single values.
% More explanations to put here?

\subsection{Installation}
You can install this package pretty easily if you use the latest version of \R and \texttt{devtools}.
Just copy past the following
<<installation, eval=FALSE>>=
install.packages("devtools")
library(devtools)
install_github("TGuillerme/dispRity", ref="release")
library(dispRity)
@

Note that we use here the \texttt{release} branch which is version \Sexpr{packageVersion("dispRity")}.
If you want the piping-hot (and full of bugs) version, change the option \texttt{ref="release"} to \texttt{ref="master"}.

This package depends heavily on the \texttt{ape} package as well as the nice function \texttt{timeSliceTree::paleotree}.
However, I recommend to also install the packages \texttt{geomorph} or \texttt{Claddis} if you want to try this example on your own morphometric or cladistic data.

\subsection{Getting the data}
You'll need to do this and that

\section{A super quick go through}
This package allows to:
\begin{enumerate}
    \item Splitting the data
    \item Bootstrapping the data
    \item Calculating disparity
    \item Summarising the results
\end{enumerate}

<<Quick go, fig.width=4, fig.height=4, out.width='.4\\linewidth'>>=
## Separate your ordinated data into different series for example
1+1
## Bootstrap this data
1+1
## Calculate a disparity metric from it
1+1
# Summarise and plot the disparity
summary(1)
plot(1)
@

\section{In more details}

\subsection{Splitting the data}

\subsection{Bootstrapping the data}

\subsection{Calculating disparity}

\subsection{Summarising the results}

\section{What's left?}
\label{whatsleft}

\end{document}