### 2015-10-06 doc/dispRity-ecology-demo.Rnw
Added ecological data example

\documentclass{article}
\usepackage[sc]{mathpazo}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm}
\setcounter{secnumdepth}{2}
\setcounter{tocdepth}{2}
\usepackage{url}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=true,bookmarksnumbered=true,bookmarksopen=true,bookmarksopenlevel=2,
 breaklinks=false,pdfborder={0 0 1},backref=false,colorlinks=false]
 {hyperref}
\hypersetup{
 pdfstartview={XYZ null null 1}}

\newcommand{\dispRity}{\texttt{dispRity} }
\newcommand{\R}{\texttt{R} }

\begin{document}
<<setup, include=FALSE, cache=FALSE>>=
library(knitr)
# set global chunk options
opts_chunk$set(fig.path='figure/minimal-', fig.align='center', fig.show='hold')
options(formatR.arrow=TRUE,width=90)
@

\title{dispRity demo for ecological data}

\author{Thomas Guillerme}

\maketitle

This is a quick demo for using the \dispRity package (v.\Sexpr{packageVersion("dispRity")}) in ecological analysis.
See the \href{https://github.com/TGuillerme/dispRity/blob/master/dispRity-tutorial.0.1.0.pdf}{dispRity-tutorial} vignette for a general demo of the \dispRity package.

\section{Before starting}

\dispRity is a package for calculating disparity in \texttt{R}.
To keep it short, this package allows to summarise ordinated matrices (e.g. MDS, PCA, PCO, PCoA) into single values.
These ordinated matrices can be called the morphospace (when using morphometric data) or the cladisto-space (for cladistic data) and will represent the total spread of the forms available in the dataset.
However, keep in mind that if you are using other type of data, say ecological data for example, then the ordinated matrix will represent the eco-space which represents the total spread of ecological traits available in the dataset.
In the rest of this demo, I will sometimes refer to the ordinated matrix as the any-o-space where you can replace \textit{any} by any prefix you want your space to be based on!

\subsection{Installation}
You can install this package easily if you use the latest version of \R and \texttt{devtools}.

<<installation, eval=FALSE>>=
install.packages("devtools")
library(devtools)
install_github("TGuillerme/dispRity", ref = "master")
library(dispRity)
@

WARNING: this is going to load the latest version, it might contain bugs.
A further version of this vignette will be based on the release version.

Note that we use the \texttt{release} branch here which is version \Sexpr{packageVersion("dispRity")}.
If you want the piping-hot (and full of bugs) version, change the option \texttt{ref = "release"} to \texttt{ref = "master"}.
This package depends mainly on the \texttt{ape} package and the \texttt{timeSliceTree::paleotree} function.

\subsection{Data}

<<data, fig.width=8, fig.height=8, out.width='.8\\linewidth'>>=
## Loading demo and the package data
library(dispRity)
## let's use the matrix from McClean (unpubl.)
data(McClean_data)
## This dataset contains an ordinated matrix
ord_matrix <- McClean_data$ordination
## As well as two list of factors for separating the rows of the matrix
treatments <- McClean_data$treatment
depth <- McClean_data$depth
@

\section{A classic PCA}
A classical way to represent this ordinated data will be to use PCA plots

<<PCAplot, fig.width=8, fig.height=8, out.width='.8\\linewidth'>>=
## The x and y axis represent the two first dimensions of the PCA
x <- ord_matrix[, 1]
y <- ord_matrix[, 2]
## The colors will represent the treatments
cols <- sub("treat1", "red", treatments)
cols <- sub("treat2", "blue", cols)
## The symbols will represent the depth
pchs <- sub(1, 16, depth)
pchs <- as.numeric(sub(2, 17, pchs))
## Graphical option
par(bty = "n")
## A classic PCA plot
plot(x, y, col = cols, pch = pchs, xlab = "PCA1", ylab = "PCA2",
    xlim = range(x) + c(0, 100))
@

However, the problem, is it ignores the 18 other axis of the ordination and the PCA axis 1 and 2 do not represent a biological reality \textit{per se} but more some ordinations of correlations between the data and some factors.
Therefore, one might want to approach this problem without getting stuck in only two dimensions and consider the whole dataset.

\section{Splitting the data}

First let's make a factor table.

<<factors>>=
## Making the factor table
factors <- as.data.frame(matrix(data = c(treatments, depth), nrow = nrow(ord_matrix),
    ncol = 2, byrow = FALSE, dimnames = list(rownames(ord_matrix))))
names(factors)<-c("Treat", "Depth")
## And here is what it looks like
head(factors)
@

Second, let's split the data according to these factors to create the subsamples of the ordinated space by using the \texttt{cust.series} function:

<<cust.series>>=
## Splitting the ordinated space into four subsamples
customised_series <- cust.series(ord_matrix, factors)
## Note that the output of dispRity functions are dispRity objects
class(customised_series)
## These objects are automatically printed in a summary method (calling S3 print.dispRity)
## giving information about the object
customised_series
@

For more details on the \dispRity objects, see the \href{https://github.com/TGuillerme/dispRity/blob/master/dispRity-tutorial.0.1.0.pdf}{dispRity-tutorial} vignette.
Basically the idea is to avoid jamming the \R console such as when using:

<<dispRity class, eval=FALSE>>=
## Summarise the object
str(customised_series)
@

\subsection{Calculating disparity}

Now we're going to see the functionalities of the core function of this package: the \dispRity function.
This function is a modulable function that allow to simply (and quickly!) calculate disparity from a matrix.
Disparity can be calculated in many ways, this function is a tool to measure disparity \textit{as defined by the user} (and here's where the modulable part comes in).
For more details on disparity, see the \href{https://github.com/TGuillerme/dispRity/blob/master/dispRity-tutorial.0.1.0.pdf}{dispRity-tutorial} vignette.

One can usually decompose the disparity metrics into two elements:
\begin{enumerate}
\item the \textbf{class metric} that is a descriptor of the matrix. For example describing the ranges of each column in the matrix or the euclidean distances between each row and the centroid of the matrix.
\item the \textbf{summary metric} that is a summary of the class metric values. For example, the sum of the ranges or the median of the euclidean distances. 
\end{enumerate}
Basically the combination can be infinite between the class and summary metrics.
For example, people might want to measure the median variances of the axis or the product of the distances from the centroid.
However, it is probable that some metrics are better to reflect some biological aspects of the any-o-space than others (see section \ref{whatsleft})...

In practice, the \dispRity function intakes a pair of class and summary metrics as a definition of disparity.
Several of these metrics are implemented in other packages (like \texttt{stats::median}, \texttt{base::sum}, etc.) and this package proposes several metrics listed in \texttt{dispRity.metric} (see \texttt{?dispRity.metric}).
It is even possible to use your very own class and summary metrics!
This will be actually heavily encouraged and facilitate with the \texttt{make.metric} function in a future version (see \ref{whatsleft}).

To use these metrics pairs in the \dispRity function, it's pretty easy:

<<dispRity1>>=
## For example, let's calculate the median distance between each plot and the
## centroid of the ordinated space
disparity <- dispRity(customised_series, metric = c(median, centroids))
## Note that disparity is a dispRity object and printing it just gives details
## on the object, not the results. We need to use summary.dispRity (S3) to get
## the results.
summary(disparity)
@

Note that we calculated the median distance between plots and the centroid but the output displays mean values.
This is because summary will, by default, summarise the data using the mean value.
Here the mean represents the mean of the median distance for each series which is a bit useless (i.e. the mean of one value is that same value).
We can display the median as well using:

<<dispRity2>>=
summary(disparity, cent.tend=median)
## Or even the product! It won't affect the results
summary(disparity, cent.tend=prod)
@

This is because we did calculate the noise within our data.
We can classically do so by bootstrapping the data!

\subsection{Bootstrapping the data}

The \dispRity pacakge also provides easy way to bootstrap the data via the \texttt{boot.matrix} function.
We can even rarefy the data to see the effect of the number of plots per series:
For more details on the bootstrapping options, see the \href{https://github.com/TGuillerme/dispRity/blob/master/dispRity-tutorial.0.1.0.pdf}{dispRity-tutorial} vignette.

<<boot.matrix>>=
bootstrapped_data <- boot.matrix(customised_series, bootstraps=100)
rarefied_data <- boot.matrix(customised_series, bootstraps=100, rarefaction=TRUE)
## Note that the output is a dispRity object giving some details on the series and the bootstraps
bootstrapped_data
@

We can now rerun a more robust disparity analysis:

<<dispRity3>>=
disparity_BS <- dispRity(bootstrapped_data, metric = c(median, centroids))
disparity_rare <- dispRity(rarefied_data, metric = c(median, centroids))
## Note that calculation time is increased!
@

We can now summarise the data using various options such as the confidence intervals levels and the central tendency.

<<summary>>=
## The default:
summary(disparity_BS)
## The CIs are calculated as 50 and 95 and the central tendency is the mean by default.
## But we can specify different options
summary(disparity_BS, CI=90, cent.tend=median)

## Finally we can see the results of the rarefaction analysis:
head(summary(disparity_rare))
## This outputs a longer table with all the variations of plots down to 3 plots per series.
@

<<disparity plots, fig.width=8, fig.height=8, out.width='.8\\linewidth'>>=
## Graphical option
par(bty = "n")
## Plotting the score for each groups
plot(disparity_BS)
@

<<rare plots, fig.width=10, fig.height=5, out.width='1\\linewidth'>>=
## Graphical options
quartz(width = 10, height = 5) ; par(mfrow = (c(1,2)), bty = "n")
## The same but looking at the number of plots
plot(disparity_rare, diversity = TRUE)
## With the same number of plots per group
plot(disparity_rare, diversity = TRUE, rarefaction = 17)

@

<<rare plots2, fig.width=8, fig.height=8, out.width='.8\\linewidth'>>=
## Graphical option
par(bty = "n")
## Plotting the rarefaction curves
plot(disparity_rare, rarefaction = "plot")
@

\end{document}