### 2016-05-25 doc/dispRity-ecology-demo.Rmd
Red manuals INIT

---
title: "dispRity ecology demo"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: References.bib
bst: sysbio.bst
vignette: >
  %\VignetteIndexEntry{dispRity ecology demo}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

This demo aims to give quick overview of the <span>`dispRity`
</span>package (v.0.2) for ecological analysis. Please refer to [GitHub
page:
github.com/TGuillerme/dispRity](https://github.com/TGuillerme/dispRity)
for other vignettes, namely the `dispRity` tutorial that
explains the functions in more detail.

To keep it short, this package allows the use of all of the dimensions
of ordinated matrices (i.e. PCA, MDS, PCO) for statistical analysis
rather than just a sub-set of dimensions. For example, one might want to
know whether a specific kind of water treatment at certain water depth
alters invertebrate communities and composition in natural habitats.

Before starting
===============

Installing `dispRity` 
-----------------------------------

You can install this package easily if you are using the latest version
of <span>`R` </span>and `devtools`.

` `\
`   `\

Data
----

For this example with ecological data we are going to use data from
McClean (unpubl.). This data is the ordination of a distance matrix
based on nutrient enrichment and depth and freshwater benthic
invertebrates.

\

\
\
\
\
\
\
\
\
`  `\
\
\
\
`  `\
`  `

A classical two dimensional approach
====================================

A classical way to represent this ordinated data will be to use two
dimensional plots.

\
\
`   `\
`   `\
\
`   `\
`   `\
\
`   `\
`   `\
\
\
`  `\
\
`          `\
`         `

![image](figure/minimal-PCAplot-1){width=".8\linewidth"}

This shows the distribution of the experimental plots along the two
first axis of variance of the ordinated distance matrix (i.e. the first
two dimensions). At a first glance, it seems difficult to see a clear
effect from the treatments (blue or red) or the depth (rounds or
triangles) since the different experimental plots with the same
parameters don’t seem to cluster. However, the problem, is that this
plot ignores the 18 other dimensions of the ordination! Additionally,
these two represented dimensions do not represent a biological reality
*per se*; i.e. the values on the first dimension do not represent some
continuous traits (e.g. depth) but just the ordinations of correlations
between the data and some factors.

Therefore, one might want to approach this problem without getting stuck
in only two dimensions and consider the whole dataset as a
*n*-dimensional object.

A multidimensional approach with `dispRity` 
=========================================================

Splitting the data
------------------

The first step in such analysis will be to create different series that
are two sub-samples of the ordinated space (i.e. the *n*-dimensional
object). Each of these series contain a certain number of elements (i.e.
the 40 experimental field plots) that have some attributes in common. In
our example, we are going to group the elements according to their depth
and treatment.

\
`       `\
`            `\
`   `\

    ##     Treat Depth
    ## 1       a     1
    ## 1.1     a     2
    ## 2       b     1
    ## 2.1     b     2
    ## 3       a     1
    ## 3.1     a     1

Second, let’s split the data according to these factors to create the
series of the ordinated space by using the `cust.series` function:

\
`  `\
\
\

    ## [1] "dispRity"

\
\

    ## 4 custom series for 40 elements 
    ## Series:
    ## Treat.a, Treat.b, Depth.1, Depth.2.

For more details on the `dispRity` objects, see the
[<span>`dispRity`
</span>manual](https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-manual.pdf).
Basically the idea is to avoid jamming the <span>`R` </span>console such
as when using:

\
\
\
\
`   `

Calculating disparity
---------------------

Disparity can be calculated in many ways, and therefore, the
`dispRity` function allows to measure disparity *as defined
by the user*. For more details on disparity, see the [other
vignettes](https://github.com/TGuillerme/dispRity).

In this example, we are going to define disparity as the median distance
between the different experimental plots and the centroid of the
ordinated space. High values of disparity will indicate a general high
dispersal from this centroid (i.e. on average, the experimental plots
are far apart in the ordinated space). We can define the metrics easily
in the `dispRity` function by feeding them to the `metric`
argument. Here we are going to feed the functions `stats::median` and
`dispRity::centroids` which calculates distances between elements and
centroid.

\
`     `\
\
\
\
\

    ## Disparity measurements across 4 series for 40 elements 
    ## Series:
    ## Treat.a, Treat.b, Depth.1, Depth.2.
    ## Disparity calculated as: c(median, centroids) for 20 dimensions.
    ## Data was split using custom method.

    ##    series  n observed
    ## 1 Treat.a 21    82.38
    ## 2 Treat.b 19    94.68
    ## 3 Depth.1 23    83.98
    ## 4 Depth.2 17    89.72

Bootstrapping the data
----------------------

One might also want to bootstrap the data to test the robustness of the
measured disparity to outliers. Also, as we can see, each series has
different numbers of elements. It might be interesting to rarefy the
data as well to have only series with the same number of elements. Both
steps are easily doable through the `boot.matrix` function.

\
`     `\
\
\
`     `\
`      `\
\
\
\

    ## Bootstrapped ordinated matrix with 40 elements 
    ## Series:
    ## Treat.a, Treat.b, Depth.1, Depth.2.
    ## Data was split using custom method.
    ## Data was bootstrapped 100 times, using the full bootstrap method.

    ## Bootstrapped ordinated matrix with 40 elements 
    ## Series:
    ## Treat.a, Treat.b, Depth.1, Depth.2.
    ## Data was split using custom method.
    ## Data was bootstrapped 100 times, using the full bootstrap method.
    ## Data was rarefied with a maximum of 17 elements

We can now rerun a more robust disparity analysis using the bootstrapped
data:

\
`     `\
\
\
`     `

Summarising the data
--------------------

We can now summarise the data using various options such as the
confidence intervals levels and the central tendency.

\

    ##    series  n observed  mean  2.5%   25%   75% 97.5%
    ## 1 Treat.a 21    82.38  79.3 64.83 74.45  84.1  95.4
    ## 2 Treat.b 19    94.68 100.2 79.63 90.99 106.8 133.6
    ## 3 Depth.1 23    83.98  82.4 69.04 78.32  85.9  92.7
    ## 4 Depth.2 17    89.72 100.9 69.58 85.03 110.6 163.4

\
\
`     `

    ##    series  n observed median    5%   95%
    ## 1 Treat.a 21    82.38  78.50 68.34  91.2
    ## 2 Treat.b 19    94.68  97.02 83.17 128.3
    ## 3 Depth.1 23    83.98  83.21 72.79  91.3
    ## 4 Depth.2 17    89.72  93.04 75.77 149.4

\

    ##    series  n observed  mean  2.5%   25%   75% 97.5%
    ## 1 Treat.a 17       NA  79.4 67.47 74.52  83.1  96.7
    ## 2 Treat.b 17       NA  99.8 79.24 88.04 106.9 147.4
    ## 3 Depth.1 17       NA  82.4 69.41 77.08  86.3 106.1
    ## 4 Depth.2 17    89.72 102.7 77.36 86.44 113.7 159.0

\

We can also plot the results and have a look at the effect of the number
of experimental plots:

\
`          `\
\
\
`   `\
\
\
`   `

![image](figure/minimal-rare_plots-1){width="1\linewidth"}

As we can see, there seems to be no strong effect of the number of
experimental plots in each series (i.e. the rarefied plot is really
similar to the bootstrapped plot) which is a good thing!

Testing the hypothesis
----------------------

Finally, we can test our hypothesis (whether the sort of water treatment
at certain depth alters invertebrate communities and composition in
natural habitats) by using the `test.dispRity` function.

\
`     `

    ##              Df Sum Sq Mean Sq F value Pr(>F)    
    ## series        3  39184   13061   61.17 <2e-16 ***
    ## Residuals   396  84553     214                   
    ## ---
    ## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

\
`     `

    ##              Df Sum Sq Mean Sq F value Pr(>F)    
    ## series        3  42229   14076   63.85 <2e-16 ***
    ## Residuals   396  87301     220                   
    ## ---
    ## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

\
\
`  `\
`       `

    ##                      W      p.value
    ## Treat.a - Treat.b  695 7.170082e-26
    ## Depth.1 - Depth.2 2064 7.360379e-13

\
`  `\
`       `

    ##                      W      p.value
    ## Treat.a - Treat.b  880 7.845518e-24
    ## Depth.1 - Depth.2 1599 9.672040e-17

As we can see, there is strong support for an effect of the treatment
and the depth on the median distance between each experimental plots and
the centroid of the ordinated space. In other words, it seems that with
the second treatment and the second level of depth, the invertebrate
communities and composition were further apart (i.e. more dispersed in
the ordinated space) than with the first treatment and the first level
of depth.