### 2016-05-29 doc/dispRity-palaeo-demo.Rmd
Added code to the Rmd docs (removed redundant Rnw)

---
title: "dispRity palaeo demo"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: References.bib
bst: sysbio.bst
vignette: >
  %\VignetteIndexEntry{dispRity palaeo demo}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

This demo aims to give quick overview of the <span>`dispRity`
</span>package (v.0.2) for palaeobiology analysis. Please refer to
[GitHub page:
github.com/TGuillerme/dispRity](https://github.com/TGuillerme/dispRity)
for other vignettes, namely the [<span>`dispRity`
</span>manual](https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-manual.pdf)
that explains the functions in more detail.

To keep it short, this package allows to easily perform
**disparity-through-time** analysis. This type of analysis often starts
by ordinating **morphometric** or **cladistic** data into a
multidimensional object hereafter called the **morphospace**. One might
be interested in studying how the occupancy of the morphospace changed
through time by measuring a summary metric of this morphospace through
time, called **disparity**. Disparity can then be seen as a value
summarising the diversity of morphologies.

This demo showcases a typical disparity-through-time analyse: we are
going to test whether the disparity changed through time in a subset of
Eutheria (Mammalia) since the last 100 million years.

Before starting
===============

Installing `dispRity` 
-----------------------------------

You can install this package easily if you are using the latest version
of <span>`R` </span>and `devtools`.

```{r, eval = FALSE}
if(!require(devtools)) install.packages("devtools")
install_github("TGuillerme/dispRity", ref = "release")
library(dispRity)
```

The morphospace
---------------

In this example, we are going to use a subset of the data from
@beckancient2014 that contains an ordinated cladistic matrix and a
phylogenetic tree containing 50 taxa along side with a table containing
some first and last occurence data for several taxa. The ordinated
matrix will represent our full morphospace, i.e. all the morphologies
that ever existed through time (for this dataset).

```{r}
## Loading demo and the package data
library(dispRity)

## Setting the random seed for repeatability
set.seed(123)

## The morphospace:
data(BeckLee_mat50)
head(BeckLee_mat50[,1:5])

dim(BeckLee_mat50)
## The morphospace contains 50 taxa and has 48 dimensions (or axis)

## A list of first and last occurrences data for some fossils
data(BeckLee_ages)
head(BeckLee_ages)

## And a phylogeny:
data(BeckLee_tree)
plot(BeckLee_tree, cex=0.7); axisPhylo(root=140)
```

![image](figure/minimal-data-1){width=".8\linewidth"}

Note that you can have an even nicer tree if you use the `strap`
package:

```{r, eval = FALSE}
if(!require(strap)) install.packages("strap")
library(strap)
geoscalePhylo(BeckLee_tree, cex.tip=0.7, cex.ts=0.6)
```

A disparity-through-time analysis
=================================

Splitting the morphospace through time
--------------------------------------

One of the crucial steps in disparity-through-time analysis is to split
the full morphospace into smaller time seriess that contain the total
amount of morphologies at certain points in time (time-slicing) or
during certain periods in time (time-binning). Basically, the full
morphospace in the previous step represents the total amount of
morphologies in all time and is likely to be greater than any time
series of the morphospace.

The `dispRity` package provides a `time.series` function
that allows to split the morphospace time slices (using
`method = continuous`) or into time bins (using `method = discrete`). In
this example, we are going to split the morphospace into 5 equal time
bins of 20 million years long from 100 million years ago (Mya) to the
present. Note that we will also provide to the function the table
containing the first and last occurrences data for some fossils to take
into account the fact that some fossils might span between our different
time bins.

```{r}
## Creating the vector of time bins ages
time_bins <- rev(seq(from = 0, to = 100, by = 20))

## Splitting the morphospace using the time.series function
binned_morphospace <- time.series(data = BeckLee_mat50, tree = BeckLee_tree,
    method = "discrete", time = time_bins, inc.nodes = FALSE,
    FADLAD = BeckLee_ages)
```

The output object is a `dispRity` object. For details about
this object class, please refer to the [<span>`dispRity`
</span>manual](https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-manual.pdf).
In brief, `dispRity` objects are lists of different
elements (i.e. disparity results, morphospace time seriess, etc...) that
display only a summary of the object when calling the object for
avoiding jamming the <span>`R` </span>console.

```{r}
## The object class
class(binned_morphospace)

## What's in the object (messy version for the console!)
str(binned_morphospace)
names(binned_morphospace)

## How it's printed by default
binned_morphospace
```

Note that these objects will gradually contain more information when
reaching the following steps in the disparity-through-time analysis.

Bootstrapping the data
----------------------

Once we obtain our different time series, we can bootstrap and rarefy
them (i.e. pseudo-replicating the data). The bootstrapping will allow to
make each series more robust to outliers and the rarefaction will allow
us to compare slices with the same number of taxa to get rid of eventual
sampling problems (i.e. more taxa in one bin than the other). The
`boot.matrix` function that allows to bootstrap the <span>`dispRity`
</span>object and the `rarefaction` option within allows to rarefy it.

```{r}
## Bootstrapping each time series 100 times
boot_bin_morphospace <- boot.matrix(binned_morphospace, bootstraps = 100)

## Getting the minimum number of rows (=taxa) in the time seriess
min(unlist(lapply(binned_morphospace$data, nrow)))

## Bootstrapping each time series 100 times and rarefying them 
rare_bin_morphospace <- boot.matrix(binned_morphospace, bootstraps = 100,
    rarefaction = 6)

## Note that the output objects are both from the "dispRity" class ...
class(boot_bin_morphospace); class(rare_bin_morphospace)

## ... and display some more information on the previous step when printed:
boot_bin_morphospace
rare_bin_morphospace
```
Calculating disparity
---------------------

We can now calculate the disparity within each time series along side
with some confidence intervals generated by the pseudo-replication step
above (bootstraps/rarefaction). Disparity can be calculated in many ways
and this package allows user to come up with their own disparity
metrics. For more info, please refer to the [<span>`dispRity`
</span>manual and the metric
vignette](https://github.com/TGuillerme/dispRity).

In this example, we are going to calculate the spread of the data in
each time series by calculating the disparity as the sum of the variance
of each dimension of the morphospace in each time series using the
`dispRity` function. Thus, in this example, the disparity
is defined by the multi-dimensional variance of each time series (i.e.
the spread of the taxa within the morphospace). Note that this metric
comes with a caveat not solved here since it ignores the covariance
among the dimensions of the morphospace.

```{r}
## Disparity for the bootstrapped data
boot_disparity <- dispRity(boot_bin_morphospace, metric = c(sum, variances))

## Disparity for the rarefied data
rare_disparity <- dispRity(rare_bin_morphospace, metric = c(sum, variances))

## Again, these objects are both from the "dispRity" class and just display
## some information when called, note the results!
boot_disparity
rare_disparity

## To display the actual results we can use the summary function
summary(boot_disparity)
summary(rare_disparity)
```

As for any <span>`R` </span>package, you can refer yourself to each
individual function for more details.

```{r}
## See more options on the summary function:
?summary.dispRity
```

Plotting the results
--------------------

It is sometimes easier to visualise the results in a plot than in a
table. For that we can use the `plot` function to plot the
`dispRity` objects.

```{r}
## Graphical options
quartz(width = 10, height = 5) ; par(mfrow = (c(1,2)), bty = "n")

## Plotting the bootstrapped and rarefied results
plot(boot_disparity, type = "continuous", main = "bootstrapped results")
plot(rare_disparity, type = "continuous", main = "rarefied results")
```

![image](figure/minimal-plot-1){width="1\linewidth"}

As we can see in this case, the rarefaction tends to increase the
confidence intervals around the mean.

Testing differences
-------------------

Finally, to draw some conclusions from these results, we need to apply
some statistical testing. As stated before, we wanted to test whether
the disparity changed through time in this dataset since the last 100
million years. To do so, we can apply some comparisons of the means
between each time-bins in a sequential manner to see whether the
disparity in bin X is equal to the disparity in bin Y which is in turn
equal to the disparity in bin Z, etc. Because our data is temporally
auto-correlated (i.e. what happens in bin Y depends on what happened in
bin X) and was pseudo-replicated (i.e. each bootstrap draw creates
non-independent time seriess because they are all based on the same time
series), we are going to apply a non-parametric mean comparison: the
`wilcox.test`. Also, we are going to apply a p-value correction to avoid
type I error inflation due to performing multiple test for the same
hypothesis (see `?p.adjust` for more details).

```{r}
## testing the differences between bins in the bootstrapped dataset.
test.dispRity(boot_disparity, test = wilcox.test, comparison = "sequential",
    correction = "bonferroni")

## testing the differences between bins in the rarefied dataset.
test.dispRity(rare_disparity, test = wilcox.test, comparison = "sequential",
    correction = "bonferroni")
```

Here our results show significant changes in disparity through time
between all time bins during the last 100 million years. However, when
looking at the rarefied results, there is no significant difference
between the time bins in the Cenozoic (60-40, 40-20 and 20-0 Mya),
suggesting that the differences detected in the first test might just be
due to the differences in number of taxa sample (respectively 13, 6 and
10 in each time bin).