### 2016-09-06 doc/dispRity-simulate_data.Rmd
simulate data vignette

---
title: "dispRity simulate data"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: References.bib
bst: sysbio.bst
vignette: >
  %\VignetteIndexEntry{dispRity metric}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

This vignette details how to simulate morphological data in `dispRity` (v.0.2.1).

Simulating morphological data in `dispRity`
=================

Before starting
===============

Installing `dispRity` 
-----------------------------------

You can install this package easily if you are using the latest version of `R` and `devtools`.

```{r, eval = FALSE}
if(!require(devtools)) install.packages("devtools")
install_github("TGuillerme/dispRity", ref = "release")
library(dispRity)
```

Quick go through
=================================
This is a quick example on how it works

```{r}
library(dispRity)
set.seed(123)
## Setting the starting tree with 15 taxa as a random coalescent tree
my_tree <- rcoal(15)

## Generating a matrix with 100 characters (85% binary and 15% three states) and
## an equal rates model with a gamma rates distribution (0.5, 1) with no 
## invariant characters.
my_matrix <- sim.morpho(tree = my_tree, characters = 100, states = c(0.85,
    0.15), rates = c(rgamma, 0.5, 1), invariant = FALSE)

## The head of the matrix
my_matrix[1:5, 1:10]

## Checking the matrix properties with a quick Maximum Parsimony tree search
check.morpho(my_matrix, my_tree)
```

In details
=================================
Basically, `sim.morpho` is really flexible and intakes a lot of different arguments to allow to simulate realistic matrix.
It has two implemented models: `"ER"` for Equal Rates (M*k* model) or the `"HKY"` one which is the molecular HKY model but transforms pyrines in 0's and purimidines in 1's.
It can also intake some specific distribution for the rates or the substitutions, allowing to modify these basic models.

The `check.morpho` runs a quick MP tree using the `phangorn` parsimony algorithm.
It quickly calculates the parsimony score, the consistency and retention indices and, if a tree is provided (e.g. the tree used to generate the matrix) it calculates the Robinson-Foulds distance between the MP tree and the provided tree.

Good parameters for a good matrix
-----------------------------------

I was testing the effect of different parameters on the ability to generate a ``realistic'' matrix (meaning not to different from the input tree with a consistency and retention index close to what's in the literature).
The matrix seems the most ``realistic'' with a starting coalescent tree, equal rates model with 0.85 binary characters and 0.15 three states characters, a gamma distribution with a shape parameter ($\alpha$) of 0.5 and no scaling ($\beta$ = 1).

Something like that should give a decent matrix:

```{r}
set.seed(1)
## tree
my_tree <- rcoal(15)
## matrix
morpho_mat <- sim.morpho(my_tree, characters = 100, model = "ER",
    rates = c(rgamma, 0.5, 1), invariant = TRUE)
check.morpho(morpho_mat, my_tree)
```