### 2017-01-13 doc/dispRity-simulate_data.Rmd
Updated data simulation manual

---
title: "dispRity simulate data"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: References.bib
bst: sysbio.bst
vignette: >
  %\VignetteIndexEntry{dispRity metric}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

Simulating morphological data in `dispRity`
=================

Before starting
===============

Installing `dispRity` 
-----------------------------------

You can install this package easily if you are using the latest version of `R` and `devtools`.

```{r, eval = FALSE}
if(!require(devtools)) install.packages("devtools")
install_github("TGuillerme/dispRity", ref = "release")
library(dispRity)
```

Quick go through
=================================
The function basically intakes a phylogenetic tree, the number of required characters and the evolutionary model and a function from which to draw the rates.
The package then proposes a function for quickly checking the matrix' phylogenetic signal using parsimony.

```{r}
library(dispRity)
set.seed(123)
## Setting the starting tree with 15 taxa as a random coalescent tree
my_tree <- rcoal(15)

## Generating a matrix with 100 characters (85% binary and 15% three states) and
## an equal rates model with a gamma rates distribution (0.5, 1) with no 
## invariant characters.
my_matrix <- sim.morpho(tree = my_tree, characters = 100, states = c(0.85,
    0.15), rates = c(rgamma, 0.5, 1), invariant = FALSE)

## The head of the matrix
my_matrix[1:5, 1:10]

## Checking the matrix properties with a quick Maximum Parsimony tree search
check.morpho(my_matrix, my_tree)
```

In details
=================================
Basically, `sim.morpho` is really flexible and intakes a lot of different arguments to allow to simulate realistic matrix.
It has three implemented models: `"ER"` for Equal Rates (the M*k* model); the `"HKY"` one which is the molecular HKY model but transforms pyrines in 0's and purimidines in 1's; or the `"mixed"` model that randomly uses an `"ER"` or and `"HKY"` for the binary characters and `"ER"` for the multistates (>2) characters.
Both models intakes specific distributions for their rates or substitution models.
These distributions should be passed to these arguments in the format of `c(sampler_function, distribution_parameters)` where the the sampler function is a the random generation function of that distribution (e.g. `rnorm`, `runif`, etc...) and the parameters are any parameters to be passed to this function.

The `check.morpho` runs a quick maximum parsimony tree using the `phangorn` parsimony algorithm.
It quickly calculates the parsimony score, the consistency and retention indices and, if a tree is provided (e.g. the tree used to generate the matrix) it calculates the Robinson-Foulds distance between the most parsimonious tree and the provided tree.

Parameters for a good(ish) matrix
-----------------------------------
There are many parameters that can realise a ``realistic'' matrix (meaning not to different from the input tree with a consistency and retention index close to what's in the literature) but because of the randomness of the generation (the code generates random matrices after all!) not all end up in creating good matrices.
The following parameters however, seems to generate a fairly ``realistic'' with a starting coalescent tree, equal rates model with 0.85 binary characters and 0.15 three states characters, a gamma distribution with a shape parameter ($\alpha$) of 5 and no scaling ($\beta$ = 1) with a rate of 100.

```{r}
set.seed(1)
## tree
my_tree <- rcoal(15)
## matrix
morpho_mat <- sim.morpho(my_tree, characters = 100, model = "ER",
    rates = c(rgamma, rate = 100, shape = 5), invariant = FALSE)
check.morpho(morpho_mat, my_tree)
```