### 2017-01-17 doc/dispRity-metrics.Rmd
Updated metrics vignette

---
title: "dispRity metrics"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: References.bib
bst: sysbio.bst
vignette: >
  %\VignetteIndexEntry{dispRity metric}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

Disparity metrics
=================

There are many ways for measuring disparity!
In brief, disparity is a summary metric that will represent an aspect of an ordinated space (e.g. a MDS, PCA, PCO, PCoA).
For example, one can look at ellipsoid hyper-volume of the ordinated space [@DonohueDim], the sum and the product of the ranges and variances [@Wills1994] or the median position of the elements relative to their centroid [@GuillermeSTD].
Of course, there are many more examples of metrics one can use for describing some aspect of the ordinated space, with some performing better than other ones at particular descriptive tasks or some being more generalist.

Because of this great diversity of metric, the package `dispRity` has not a unique way to measure disparity but rather proposes to facilitate users to define their own disparity metric that will best suit to their particular analysis.
In fact, the core function of the package, `dispRity`, allows the user to define any metric with the `metric` argument.
However, because of the infrastructure of the package (especially the one of the `dispRity` objects – see the [`dispRity` manual](https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-manual.Rmd)), the `metric` argument has to follow certain rules:

1.  it must be composed from one to three `function` objects;

2.  these functions can only intake a `matrix` or a `vector`;

3.  each of these functions must be of one of the three levels described below;

4.  at least one of the functions must be of level 1 or 2 (see below).

The function levels
===================

The metric functions levels determine the dimensional level of decomposition of the input `matrix`.
In other words, each level designate the dimensions of the output: either three (a `matrix`); two (a `vector`); or one (a single `numeric` value) dimensions.

![Illustration of the different levels of output of a function with an input `matrix`](dispRity_fun.png)

A level 1 function
------------------

A level 1 function will decompose a `matrix` or a `vector` into a single value:

```{r}
## First, let's create a dummy matrix
dummy_matrix <- matrix(rnorm(12), 4, 3)

## Any summary metric such as mean or median are good examples of level 1
## functions as they reduce the matrix to a single dimension (i.e. one value):
mean(dummy_matrix)
median(dummy_matrix)
```

A level 2 function
------------------

A level 2 function will decompose a `matrix` into a `vector`.

```{r}
## We can define a function that outputs the products of each row of the matrix:
prod.rows <- function(matrix) apply(matrix, 1, prod)

## This function is a level 2 function as it reduces the matrix into a two
## dimensional element (a vector of values):
prod.rows(dummy_matrix)
```

Several level 2 functions are implemented in `dispRity` (see `?dispRity.metric`) such as the `variances` function that calculates the variance of each dimension of the ordinated matrix or the `ranges` one that calculates the range of each dimension.

A level 3 function
------------------

Finally a level 3 function will transform the `matrix` into another `matrix`.
Note that the dimension of the output matrix don't need to match the the input one:

```{r}
## The var functions computes the variance/covariance matrix which will be a 3x3 matrix
var(dummy_matrix)
```

`make.metric`
=============

Of course, functions can be more complex and involve multiple operations such as the `centroids` function (see `?dispRity.metric`) that calculates the euclidean distance between each elements and the ordinated space centroid.
The `make.metric` function implemented in `dispRity` is designed to help testing and finding the level of the functions.
This function tests:

1.  if your function can deal with a `matrix` or a `vector` as an input;

2.  which is your function's level according to it's output (level 1, 2 or 3, see above);

3.  whether the function can properly be implemented in the `dispRity` function (the function is fed into a `lapply` loop).

For example, let's see if the functions described above are the right levels:

```{r}
## First we need to load the package
library(dispRity)

## Which level is the mean function? And is able to be used in the dispRity?
make.metric(mean)

## Same questions from the prod.rows function:
make.metric(prod.rows)

## Same questions for the var function:
make.metric(var)
```

metric argument in `dispRity`
===========================================

Using this metric structure, we can easily use any disparity metric in the `dispRity`function as follows:

```{r}
## First we need to load the data
data(BeckLee_mat50)

## Measuring disparity as the standard deviation of all the value of the
## ordinated matrix (level 1 function).
summary(dispRity(BeckLee_mat50, metric = sd))

## Measuring disparity as the standard deviation of the variance of each axis of
## the ordinated matrix (level 1 and level 2 functions).
summary(dispRity(BeckLee_mat50, metric = c(sd, variances)))

## Measuring disparity as the standard deviation of the variance of each axis of
## the variance covariance matrix (level 1, level 2 and level 3 functions).
summary(dispRity(BeckLee_mat50, metric = c(sd, variances, var)), round = 10)
```

Note that the order of each function in the metric argument does not matter, the `dispRity` function will automatically detect the function levels (using `make.metric`) and apply them to the data in decreasing order (level 3 > level 2 > level 1).

```{r}
## Disparity as the standard deviation of the variance of each axis of the
## variance covariance matrix:
disparity1 <- summary(dispRity(BeckLee_mat50, metric = c(sd, variances, var)), round = 10)

## Same but using a different functions order for the metric argument
disparity2 <- summary(dispRity(BeckLee_mat50, metric = c(variances, sd, var)), round = 10)

## Both ways output the same disparity values:
disparity1 == disparity2
```

What level is the disparity?
------------------

In these examples, we considered disparity to be a single value.
For example, in the previous one, we defined disparity as the standard deviation of the variances of each column of the variance/covariance matrix (`metric = c(variances, sd, var)`).
This results in single value to be calculated.

In some cases however, it might also be interesting to calculate disparity as a distribution.
For example, one of the metrics commonly used in palaeobiology is the sum of variances of each dimensions (`metric = c(sum, variances)`) [@Wills1994].
However, it might be interesting to look directly at the distribution of the variances itself.

```{r}
## A one dimensional disparity value
(sum_variances1 <- summary(dispRity(BeckLee_mat50, metric = c(median, variances))))

## A two dimensional disparity
(summary(all_variances <- (dispRity(BeckLee_mat50, metric = variances))))
```

Looking at all the variances rather than simply their median can tell us that (by the look of it) the variances of each dimensions seems to be normally distributed.

> It is also possible to reuse an already calculated disparity distribution and summarise it using a level 1 metric:

```{r}
## Transforming the two dimensional one in a one dimensional one
sum_variances2 <- summary(dispRity(all_variances, metric = median))

## Both ways output the same disparity values:
sum_variances1 == sum_variances2
```

Metrics implemented in `dispRity`
===========================================

Several disparity metrics are directly implemented in the `dispRity` package.
The detailed list can be found in `?dispRity.metric` along with some explanations on each metric.

Level | Name | Description | Source | 
------|------|---------------------------------------------------|--------|
3 | `ellipse.volume`<sup>1</sup> | The volume of the ellipsoid of the space | [@DonohueDim] |
3 | `convhull.surface` | The surface of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln` |
3 | `convhull.volume` | The volume of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln` |
3 | `hypervolume` | The volume of the ordinated space | [`hypervolume`](https://cran.r-project.org/web/packages/hypervolume/index.html)`::hypervolume` |
3 | `diagonal` | The longest distance in the ordinated space (like the diagonal in two dimensions) | `dispRity` |
2 | `ranges` | The range of each dimension | `dispRity` |
2 | `variances` | The variance of each dimension | `dispRity` |
2 | `centroids`<sup>2</sup> | The distance between each element and the centroid of the ordinated space | `dispRity` |
1 | `mode.val` | The modal value | `dispRity` |

1: The functions estimation of the eigen value only works for MDS or PCoA ordinations (not PCA).

2: Note that by default, the centroid is the centroid of the elements.
It can however be fixed to an arbitrary value by using the `centroid` argument `centroids(space, centroid = rep(0, ncol(space)))`, for example the origin of the ordinated space.

## Equations and implementations:
Some of the functions described below are directly implemented in the `dispRity` package and do not require any dependencies  ([see implementation here](https://github.com/TGuillerme/dispRity/blob/master/R/dispRity.metric.R)).

\begin{equation}
    ellipse.volume = \frac{\pi^{k/2}}{\Gamma(\frac{k}{2}+1)}\displaystyle\prod_{i=1}^{k} (\lambda_{i}^{0.5})
\end{equation}
With *k* being the number of dimensions, $\lambda_i$ is the eigen value of each dimension.

\begin{equation}
    diagonal = \sqrt{\sum_{i=1}^{k}|max(k_i) - min(k_i)|}
\end{equation}
With *k* being the dimensions.

\begin{equation}
    ranges = |max(k_i) - min(k_i)|
\end{equation}
With *k* being the dimensions.

\begin{equation}
    variances = \sigma^{2}{k_i}
\end{equation}
With *k* being the dimensions and $\sigma^{2}$ their variance.

\begin{equation}
    centroids = \sqrt{\sum_{i=1}^{n}{({k}_{n}-Centroid_{k})^2}}
\end{equation}
With *n* being each element in the ordinated space, *k* being its dimensions and $Centroid_{k}$ their mean (or can be set to an arbitrary value) and.

## Metrics run through
Here is a run through of the main metrics implemented in `dispRity`.
First, we will create a dummy ordinated space using the `space.maker` utility function:

```{r}
## Creating a 10*5 normal space
set.seed(1)
dummy_space <- space.maker(10, 5, rnorm)
```

### Volumes and surface metrics
The functions `ellipse.volume`, `convhull.surface`, `convhull.volume` and `hyper.volume` all measure the surface or the volume of the ordinated space occupancy:

```{r}
## Calculating the ellipsoid volume
summary(dispRity(dummy_space, metric = ellipse.volume))
```
> Because there is only one series (i.e. one matrix) in the dispRity object, this operation is the equivalent of doing simply `ellipse.volume(dummy_space)` (with rounding).

```{r}
## Calculating the convex hull surface
summary(dispRity(dummy_space, metric = convhull.surface))

## Calculating the convex hull volume
summary(dispRity(dummy_space, metric = convhull.volume))
```

The convex hull functions make a (good) estimation of the multidimensional properties of the ordinated space.
For the full and correct calculation of the volume of the ordinated space, it is possible to us the `hyper.volume` function that intakes more options but takes longer to calculate.

```{r}
## Calculating the true multidimensional volume
summary(dispRity(dummy_space, metric = hyper.volume))

## Calculating the volume with different options to be passed to hypervolume::hypervolume
summary(dispRity(dummy_space, metric = hyper.volume, repsperpoint = 100, quantile = 0.5))
```
> Cautionary note: measuring volumes in a high number of dimensions can be strongly affected by the [Curse of dimensionality](https://en.wikipedia.org/wiki/Curse_of_dimensionality) which often results in near 0 disparity values.

### Ranges, variances and diagonal

The functions `ranges`, `variances` and `diagonal` all measure properties of the ordinated space base on it's dimensions properties (they are also less affected by the Curse of dimensionality):

`ranges` and `variances` both work on the same principle and measure the range/variance of each dimensions:

```{r}
## Calculating the ranges of each dimensions in the ordinated space
ranges(dummy_space)

## Calculating disparity as the distribution of these ranges
summary(dispRity(dummy_space, metric = ranges))

## Calculating disparity as the sum and the product of these ranges
summary(dispRity(dummy_space, metric = c(sum, ranges)))
summary(dispRity(dummy_space, metric = c(prod, ranges)))

## Calculating the variances of each dimensions in the ordinated space
variances(dummy_space)

## Calculating disparity as the distribution of these variances
summary(dispRity(dummy_space, metric = variances))

## Calculating disparity as the sum and the product of these variances
summary(dispRity(dummy_space, metric = c(sum, variances)))
summary(dispRity(dummy_space, metric = c(prod, variances)))
```

The `diagonal` function measures the multidimensional diagonal of the whole space (i.e. in our case the longest euclidean distance in our 5 dimensional space):

```{r}
## Calculating the ordinated space's diagonal
summary(dispRity(dummy_space, metric = diagonal))
```

> This metric is only mathematically valid if the dimensions within the space are all orthogonal!

### Centroids metrics

The `centroids` metric allows to measure the position of the different elements compared to a fixed point in the ordinated space.
By default, this function measures the distance between each elements and the central point between them:

```{r}
## Calculating the distribution of the distances between each elements and the centroid
summary(dispRity(dummy_space, metric = centroids))

## Calculating disparity as the median value of these distances
summary(dispRity(dummy_space, metric = c(median, centroids)))
```

It is however possible to fix the coordinates of a specific point in the ordinated space as long as it has the correct number of dimensions:

```{r}
## Calculating the distance between the elements and the origin of the ordinated space
summary(dispRity(dummy_space, metric = centroids, centroid = c(0,0,0,0,0)))

## Calculating disparity as the distance between the elements and a specific point
summary(dispRity(dummy_space, metric = centroids, centroid = c(0,1,2,3,4)))
```

# References