### 2017-08-15 inst/gitbook/_book/manual.tex

\documentclass[]{book}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={dispRity manual},
            pdfauthor={Thomas Guillerme (guillert@tcd.ie) and Natalie Cooper (natalie.cooper@nhm.ac.uk)},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{dispRity manual}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Thomas Guillerme
(\href{mailto:guillert@tcd.ie}{\nolinkurl{guillert@tcd.ie}}) and Natalie
Cooper
(\href{mailto:natalie.cooper@nhm.ac.uk}{\nolinkurl{natalie.cooper@nhm.ac.uk}})}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{2017-08-15}

\usepackage{booktabs}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{lemma}{Lemma}[chapter]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[chapter]
\newtheorem{corollary}{Corollary}[chapter]
\newtheorem{proposition}{Proposition}[chapter]
\theoremstyle{definition}
\newtheorem{example}{Example}[chapter]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\begin{document}
\maketitle

{
\setcounter{tocdepth}{1}
\tableofcontents
}
\chapter{\texorpdfstring{\texttt{dispRity}}{dispRity}}\label{disprity}

This is a package for measuring disparity in \texttt{R}. It allows users
to summarise ordinated matrices (e.g.~MDS, PCA, PCO, PCoA) into single
values so they can easily be compared. This manual is based on the
version \texttt{1.0}.

\section{\texorpdfstring{What is
\texttt{dispRity}?}{What is dispRity?}}\label{what-is-disprity}

This is a package for measuring disparity in \texttt{R}. It allows users
to summarise ordinated matrices (e.g.~MDS, PCA, PCO, PCoA) to perform
some multidimensional analysis. Typically, these analysis are used in
palaeobiology and evolutionary biology to study the changes in
morphology through time. However, there are many more applications in
ecology, evolution and beyond.

\section{Installing and running the
package}\label{installing-and-running-the-package}

You can install this package easily if you use the latest version of
\texttt{R} (\textgreater{} 4.0) and \texttt{devtools}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Checking if devtools is already installed}
\ControlFlowTok{if}\NormalTok{(}\OperatorTok{!}\KeywordTok{require}\NormalTok{(devtools)) }\KeywordTok{install.packages}\NormalTok{(}\StringTok{"devtools"}\NormalTok{)}

\NormalTok{## Installing the latest released version directly from GitHub}
\KeywordTok{install_github}\NormalTok{(}\StringTok{"TGuillerme/dispRity"}\NormalTok{, }\DataTypeTok{ref =} \StringTok{"release"}\NormalTok{)}

\NormalTok{## Loading the package}
\KeywordTok{library}\NormalTok{(dispRity)}
\end{Highlighting}
\end{Shaded}

Note this uses the \texttt{release} branch (version\texttt{0.4}). For
the piping-hot (but potentially unstable) version, you can change the
argument \texttt{ref\ =\ release} to \texttt{ref\ =\ master}.
\texttt{dispRity} depends mainly on the \texttt{ape} package and uses
functions from several other packages (\texttt{ade4}, \texttt{geometry},
\texttt{grDevices}, \texttt{hypervolume}, \texttt{paleotree},
\texttt{snow}, \texttt{Claddis}, \texttt{geomorph} and \texttt{RCurl}).

\section{Why not CRAN?}\label{why-not-cran}

This package is not available on CRAN. This is mainly because some parts
are still in development and that the reactivity of GitHub is better for
implementing new suggestions from users. However, the package follows
the strict (and useful!) CRAN standards via Travis. \#\# Help

If you need help with the package, hopefully the following manual will
be useful. However, parts of this package are still in development and
some other parts are probably not covered. Thus if you have suggestions
or comments on on what has already been developed or will be developed,
please send me an email
(\href{mailto:guillert@tcd.ie}{\nolinkurl{guillert@tcd.ie}}) or if you
are a GitHub user, directly create an issue on the
\href{https://github.com/TGuillerme/dispRity}{GitHub page}.

\section{Citations}\label{citations}

You can cite both the package or this manual with the following
citation:

\begin{quote}
Guillerme, T. (2016). dispRity: a package for measuring disparity in R.
Zenodo. 10.5281/zenodo.55646
\end{quote}

Note that this citation is only temporary (but can still be used!). A
future proper version of the latest package release, this manual and an
associated methods paper should be submitted soon(ish!).

\chapter{Glossary}\label{glossary}

\begin{itemize}
\item
  \textbf{Ordinated space}. The mathematical multidimensional object
  that will be analysed with this package. In morphometrics, this is
  often referred to as the morphospace. However it may also be referred
  to as the cladisto-space for cladistic data or the eco-space for
  ecological data etc. In practice, this term designates an ordinated
  matrix where the columns represent the dimensions of the space (often
  -- but not necessarily - \textgreater{} 3!) and the rows represent the
  elements within this space.
\item
  \textbf{Elements}. The rows of the ordinated space. Elements can be
  taxa, field sites, countries etc.
\item
  \textbf{Dimensions}. The columns of the ordinated space. The
  dimensions are referred to as axes of variation, or principal
  components, for ordinated spaces obtained from a PCA.
\item
  \textbf{Subsamples}. Sub-samples of the ordinated space. A sub-sample
  (or subsamples) contains the same number of dimensions as the
  ordinated space but may contain a smaller subset of elements. For
  example, if our ordinated space is composed of birds and mammals (the
  elements) and 50 principal components (the dimensions), we can create
  two subsamples containing just mammals or birds, but with the same 50
  dimensions, to compare disparity in the two clades.
\end{itemize}

\chapter{Getting started with
dispRity}\label{getting-started-with-disprity}

\section{What sort of data does dispRity work
with?}\label{what-sort-of-data-does-disprity-work-with}

Disparity can be estimated from pretty much any matrix. Classically,
however, it is measured from ordinated matrices. These matrices can be
from any type of ordination (PCO, PCA, PCoA, MDS, etc.) as long as they
have your element names as rows (taxa, experiments, countries etc.) and
your ordination axes as columns (the dimensions of your dataset).

\subsection{\texorpdfstring{Ordination matrices from
\texttt{Claddis}}{Ordination matrices from Claddis}}\label{ordination-matrices-from-claddis}

\begin{verbatim}
## Warning: package 'phytools' was built under R version 3.4.1
\end{verbatim}

\texttt{dispRity} package can easily take data from \texttt{Claddis}
using the \texttt{Claddis.ordination} function. For this, simply input a
matrix in the \texttt{Claddis} format to the function and it will
automatically calculate and ordinate the distances among taxa:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{require}\NormalTok{(Claddis)}

\NormalTok{## Ordinating the example data from Claddis}
\KeywordTok{Claddis.ordination}\NormalTok{(Michaux1989) }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                      [,1]          [,2]       [,3]
## Ancilla      7.252259e-17  4.154578e-01  0.2534942
## Turrancilla -5.106645e-01 -4.566150e-16 -0.2534942
## Ancillista   5.106645e-01 -8.153839e-16 -0.2534942
## Amalda      -3.207162e-16 -4.154578e-01  0.2534942
\end{verbatim}

Note that several options are available, namely which type of distance
should be computed. See more info in the function manual
(\texttt{?Claddis.ordination}). Alternatively, it is of course also
possible to manual calculate the ordination matrix using the functions
\texttt{Claddis::MorphDistMatrix} and \texttt{stats::cmdscale}.

\subsection{\texorpdfstring{Ordination matrices from
\texttt{geomorph}}{Ordination matrices from geomorph}}\label{ordination-matrices-from-geomorph}

You can also easily use data from \texttt{geomorph} using the
\texttt{geomorph.ordination} function. This function simply takes
Procrustes aligned data and performs an ordination:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{require}\NormalTok{(geomorph)}

\NormalTok{## Loading the plethodon dataset}
\KeywordTok{data}\NormalTok{(plethodon)}

\NormalTok{## Performing a Procrustes transform on the landmarks}
\NormalTok{procrustes <-}\StringTok{ }\KeywordTok{gpagen}\NormalTok{(plethodon}\OperatorTok{$}\NormalTok{land, }\DataTypeTok{PrinAxes =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{print.progress =} \OtherTok{FALSE}\NormalTok{)}

\NormalTok{## Ordinating this data}
\KeywordTok{geomorph.ordination}\NormalTok{(procrustes)[}\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{,}\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                PC1        PC2           PC3          PC4          PC5
## [1,] -0.0369931363 0.05118247 -0.0016971082 -0.003128809 -0.010936371
## [2,] -0.0007493738 0.05942082  0.0001371715 -0.002768680 -0.008117383
## [3,]  0.0056004654 0.07419599 -0.0052612103 -0.005034566 -0.002746592
## [4,] -0.0134808572 0.06463959 -0.0458436015 -0.007887369  0.009816827
## [5,] -0.0334696244 0.06863518  0.0136292041  0.007359409  0.022347225
\end{verbatim}

Options for the ordination (from \texttt{?prcomp}) can be directly
passed to this function to perform customised ordinations. Additionally
you can give the function a \texttt{geomorph.data.frame} object. If the
latter contains sorting information (i.e.~factors), they can be directly
used to make a customised \texttt{dispRity} object
\protect\hyperlink{customised-subsamples}{customised \texttt{dispRity}
object}!

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Using a geomorph.data.frame}
\NormalTok{geomorph_df <-}\StringTok{ }\KeywordTok{geomorph.data.frame}\NormalTok{(procrustes,}
     \DataTypeTok{species =}\NormalTok{ plethodon}\OperatorTok{$}\NormalTok{species, }\DataTypeTok{site =}\NormalTok{ plethodon}\OperatorTok{$}\NormalTok{site)}

\NormalTok{## Ordinating this data and making a dispRity object}
\KeywordTok{geomorph.ordination}\NormalTok{(geomorph_df)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 4 customised subsamples for 40 elements:
##     species.Jord, species.Teyah, site.Allo, site.Symp.
\end{verbatim}

More about these \texttt{dispRity} objects below!

\subsection{Other kinds of ordination
matrices}\label{other-kinds-of-ordination-matrices}

If you are not using the packages mentioned above (\texttt{Claddis} and
\texttt{geomorph}) you can easily make your own ordination matrices by
using the following functions from the \texttt{stats} package. Here is
how to do it for the following types of matrices:

\begin{itemize}
\tightlist
\item
  Multivariate matrices (principal components analysis; PCA)
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## A multivariate matrix}
\KeywordTok{head}\NormalTok{(USArrests)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##            Murder Assault UrbanPop Rape
## Alabama      13.2     236       58 21.2
## Alaska       10.0     263       48 44.5
## Arizona       8.1     294       80 31.0
## Arkansas      8.8     190       50 19.5
## California    9.0     276       91 40.6
## Colorado      7.9     204       78 38.7
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Ordinating the matrix using `prcomp` }
\NormalTok{ordination <-}\StringTok{ }\KeywordTok{prcomp}\NormalTok{(USArrests)}

\NormalTok{## Selecting the ordinated matrix}
\NormalTok{ordinated_matrix <-}\StringTok{ }\NormalTok{ordination}\OperatorTok{$}\NormalTok{x}
\KeywordTok{head}\NormalTok{(ordinated_matrix)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                  PC1        PC2        PC3        PC4
## Alabama     64.80216 -11.448007 -2.4949328 -2.4079009
## Alaska      92.82745 -17.982943 20.1265749  4.0940470
## Arizona    124.06822   8.830403 -1.6874484  4.3536852
## Arkansas    18.34004 -16.703911  0.2101894  0.5209936
## California 107.42295  22.520070  6.7458730  2.8118259
## Colorado    34.97599  13.719584 12.2793628  1.7214637
\end{verbatim}

This results in a ordinated matrix with US states as elements and four
dimensions (PC 1 to 4). For an alternative method, see the
\texttt{?princomp} function.

\begin{itemize}
\tightlist
\item
  Distance matrices (classical multidimensional scaling; MDS)
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## A matrix of distances between cities}
\KeywordTok{str}\NormalTok{(eurodist)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Class 'dist'  atomic [1:210] 3313 2963 3175 3339 2762 ...
##   ..- attr(*, "Size")= num 21
##   ..- attr(*, "Labels")= chr [1:21] "Athens" "Barcelona" "Brussels" "Calais" ...
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Ordinating the matrix using cmdscale() with k = 5 dimensions }
\NormalTok{ordinated_matrix <-}\StringTok{ }\KeywordTok{cmdscale}\NormalTok{(eurodist, }\DataTypeTok{k =} \DecValTok{5}\NormalTok{)}
\KeywordTok{head}\NormalTok{(ordinated_matrix)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                 [,1]      [,2]       [,3]       [,4]       [,5]
## Athens    2290.27468 1798.8029   53.79314 -103.82696 -156.95511
## Barcelona -825.38279  546.8115 -113.85842   84.58583  291.44076
## Brussels    59.18334 -367.0814  177.55291   38.79751  -95.62045
## Calais     -82.84597 -429.9147  300.19274  106.35369 -180.44614
## Cherbourg -352.49943 -290.9084  457.35294  111.44915 -417.49668
## Cologne    293.68963 -405.3119  360.09323 -636.20238  159.39266
\end{verbatim}

This results in a ordinated matrix with European cities as elements and
five dimensions.

Of course any other method for creating the ordination matrix is totally
valid, you can also not use any ordination at all! The only requirements
for the \texttt{dispRity} functions is that the input is a matrix with
elements as rows and dimensions as columns.

\section{Performing a simple dispRity
analysis}\label{performing-a-simple-disprity-analysis}

Two \texttt{dispRity} functions allow users to run an analysis pipeline
simply by inputting an ordination matrix. These functions allow users to
either calculate the disparity through time
(\texttt{dispRity.through.time}) or the disparity of user-defined groups
(\texttt{dispRity.per.group}).

\textbf{IMPORTANT}

Note that \texttt{disparity.through.time} and
\texttt{disparity.per.group} are wrapper functions (i.e.~they
incorporate lots of other functions) that allow users to run a basic
disparity-through-time, or disparity among groups, analysis without too
much effort. As such they use a lot of default options. These are
described in the help files for the functions that are used to make the
wrapper functions, and not described in the help files for
\texttt{disparity.through.time} and \texttt{disparity.per.group}. These
defaults are good enough for \textbf{data exploration}, but for a proper
analysis you should consider the \textbf{best parameters for your
question and data}. For example, which metric should you use? How many
bootstraps do you require? What model of evolution is most appropriate
if you are time slicing? Should you rarefy the data? See
\protect\hyperlink{time-slicing}{\texttt{time.subsamples}},
\protect\hyperlink{customised-subsamples}{\texttt{custom.subsamples}},
\protect\hyperlink{bootstraps-and-rarefactions}{\texttt{boot.matrix}}
and \protect\hyperlink{disparity-metrics}{\texttt{dispRity.metric}} for
more details of the defaults used in each of these functions. Note that
any of these default arguments can be changed within the
\texttt{disparity.through.time} or \texttt{disparity.per.group}
functions.

\hypertarget{example-data}{\subsection{Example
data}\label{example-data}}

To illustrate these functions, we will use data from
\citet{beckancient2014}. This dataset contains an ordinated matrix of 50
discrete characters from mammals (\texttt{BeckLee\_mat50}), another
matrix of the same 50 mammals and the estimated discrete data characters
of their descendants (thus 50 + 49 rows, \texttt{BeckLee\_mat99}), a
dataframe containing the ages of each taxon in the dataset
(\texttt{BeckLee\_ages}) and finally a phylogenetic tree with the
relationships among the 50 mammals (\texttt{BeckLee\_tree}).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Loading the ordinated matrices}
\KeywordTok{data}\NormalTok{(BeckLee_mat50)}
\KeywordTok{data}\NormalTok{(BeckLee_mat99)}

\NormalTok{## The first five taxa and dimensions of the 50 taxa matrix}
\KeywordTok{head}\NormalTok{(BeckLee_mat50[, }\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{])}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                    [,1]          [,2]        [,3]       [,4]      [,5]
## Cimolestes   -0.5319679  0.1117759259  0.09865194 -0.1933148 0.2035833
## Maelestes    -0.4087147  0.0139690317  0.26268300  0.2297096 0.1310953
## Batodon      -0.6923194  0.3308625215 -0.10175223 -0.1899656 0.1003108
## Bulaklestes  -0.6802291 -0.0134872777  0.11018009 -0.4103588 0.4326298
## Daulestes    -0.7386111  0.0009001369  0.12006449 -0.4978191 0.4741342
## Uchkudukodon -0.5105254 -0.2420633915  0.44170317 -0.1172972 0.3602273
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The first five taxa and dimensions of the 99 taxa + ancestors matrix}
\NormalTok{BeckLee_mat99[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{98}\NormalTok{, }\DecValTok{99}\NormalTok{), }\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                   [,1]       [,2]       [,3]        [,4]        [,5]
## Cimolestes -0.60824375 -0.0323683 0.08458885 -0.43384481 -0.30536875
## Maelestes  -0.57302058 -0.2840361 0.01308847 -0.12588477  0.06123611
## n48        -0.05529018  0.4799330 0.04118477  0.04944912 -0.35588301
## n49        -0.13067785  0.4478168 0.11956268  0.13800340 -0.32227852
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Loading a list of first and last occurrence dates for the fossils}
\KeywordTok{data}\NormalTok{(BeckLee_ages)}
\KeywordTok{head}\NormalTok{(BeckLee_ages)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##             FAD  LAD
## Adapis     37.2 36.8
## Asioryctes 83.6 72.1
## Leptictis  33.9 33.3
## Miacis     49.0 46.7
## Mimotona   61.6 59.2
## Notharctus 50.2 47.0
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Loading and plotting the phylogeny}
\KeywordTok{data}\NormalTok{(BeckLee_tree)}
\KeywordTok{plot}\NormalTok{(BeckLee_tree, }\DataTypeTok{cex =} \FloatTok{0.8}\NormalTok{) }
\KeywordTok{axisPhylo}\NormalTok{(}\DataTypeTok{root =} \DecValTok{140}\NormalTok{)}
\KeywordTok{nodelabels}\NormalTok{(}\DataTypeTok{cex =} \FloatTok{0.5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{manual_files/figure-latex/unnamed-chunk-10-1.pdf}

Of course you can use your own data as detailed in the
\protect\hyperlink{What-sort-of-data-does-dispRity-work-with}{previous
chapter}.

\subsection{Disparity through time}\label{disparity-through-time}

The \texttt{dispRity.through.time} function calculates disparity through
time, a common analysis in palaeontology. This function (and the
following one) uses an analysis pipeline with a lot of default
parameters to make the analysis as simple as possible. Of course all the
defaults can be changed if required, more on this later.

For a disparity through time analysis, you will need:

\begin{itemize}
\tightlist
\item
  An ordinated matrix (we covered that above)
\item
  A phylogenetic tree: this must be a \texttt{phylo} object (from the
  \texttt{ape} package) and needs a \texttt{root.time} element. To give
  your tree a root time (i.e.~an age for the root), you can simply do
  \texttt{my\_tree\$root.time\ \textless{}-\ my\_age}.
\item
  The required number of time subsamples (here \texttt{time\ =\ 3})
\item
  Your favourite disparity metric (here the sum of variances)
\end{itemize}

Using the \citet{beckancient2014} data described
\protect\hyperlink{example-data}{above}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Measuring disparity through time}
\NormalTok{disparity_data <-}\StringTok{ }\KeywordTok{dispRity.through.time}\NormalTok{(BeckLee_mat50, BeckLee_tree,}
                                        \DataTypeTok{time =} \DecValTok{3}\NormalTok{, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, variances))}
\end{Highlighting}
\end{Shaded}

This generates a \texttt{dispRity} object (see
\protect\hyperlink{manipulating-dispRity-objects}{here} for technical
details). When displayed, these \texttt{dispRity} objects provide us
with information on the operations done to the matrix:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Print the disparity_data object}
\NormalTok{disparity_data}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 3 discrete time subsamples for 50 elements with 48 dimensions:
##     133.51104 - 89.00736, 89.00736 - 44.50368, 44.50368 - 0.
## Data was bootstrapped 100 times (method:"full").
## Disparity was calculated as: metric.
\end{verbatim}

We asked for three subsamples (evenly spread across the age of the
tree), the data was bootstrapped 100 times (default) and the metric used
was the sum of variances.

We can now summarise or plot the \texttt{disparity\_data} object, or
perform statistical tests on it (e.g.~a simple \texttt{lm}):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Summarising disparity through time}
\KeywordTok{summary}\NormalTok{(disparity_data)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##             subsamples  n   obs bs.median  2.5%   25%   75% 97.5%
## 1 133.51104 - 89.00736  5 1.575     1.305 0.729 1.118 1.420 1.509
## 2  89.00736 - 44.50368 29 1.922     1.867 1.775 1.830 1.889 1.922
## 3         44.50368 - 0 16 1.990     1.871 1.716 1.831 1.914 1.942
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Plotting the results}
\KeywordTok{plot}\NormalTok{(disparity_data, }\DataTypeTok{type =} \StringTok{"continuous"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{manual_files/figure-latex/unnamed-chunk-13-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Testing for an difference among the time bins}
\NormalTok{disp_lm <-}\StringTok{ }\KeywordTok{test.dispRity}\NormalTok{(disparity_data, }\DataTypeTok{test =}\NormalTok{ lm, }\DataTypeTok{comparisons =} \StringTok{"all"}\NormalTok{)}
\KeywordTok{summary}\NormalTok{(disp_lm)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## test(formula = data ~ subsamples, data = data)
## 
## Residuals:
##      Min       1Q   Median       3Q      Max 
## -0.56623 -0.04160  0.01049  0.05507  0.31886 
## 
## Coefficients:
##                               Estimate Std. Error t value Pr(>|t|)    
## (Intercept)                    1.25647    0.01270   98.97   <2e-16 ***
## subsamples44.50368 - 0         0.60863    0.01795   33.90   <2e-16 ***
## subsamples89.00736 - 44.50368  0.60169    0.01795   33.51   <2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 0.127 on 297 degrees of freedom
## Multiple R-squared:  0.8361, Adjusted R-squared:  0.835 
## F-statistic: 757.5 on 2 and 297 DF,  p-value: < 2.2e-16
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Note that ANOVA assumptions were not checked here!}
\end{Highlighting}
\end{Shaded}

Please refer to the \protect\hyperlink{specific-tutorial}{specific
tutorials} for (much!) more information on the nuts and bolts of the
package. You can also directly explore the specific function help files
within R and navigate to related functions.

\hypertarget{disparity-among-groups}{\subsection{Disparity among
groups}\label{disparity-among-groups}}

The \texttt{dispRity.per.group} function is used if you are interested
in looking at disparity among groups rather than through time. For
example, you could ask if there is a difference in disparity between two
groups?

To perform such an analysis, you will need:

\begin{itemize}
\tightlist
\item
  An matrix with rows as elements and columns as dimensions (always!)
\item
  A list of group members: this list should be a list of numeric vectors
  or names corresponding to the row names in the matrix. For example
  \texttt{list("a"\ =\ c(1,2),\ "b"\ =\ c(3,4))} will create a group
  \emph{a} containing elements 1 and 2 from the matrix and a group
  \emph{b} containing elements 3 and 4. Note that elements can be
  present in multiple groups at once.
\item
  Your favourite disparity metric (here the sum of variances)
\end{itemize}

Using the \citet{beckancient2014} data described
\protect\hyperlink{example-data}{above}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating the two groups (crown versus stem) as a list}
\NormalTok{mammal_groups <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}\StringTok{"crown"}\NormalTok{ =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{16}\NormalTok{, }\DecValTok{19}\OperatorTok{:}\DecValTok{41}\NormalTok{, }\DecValTok{45}\OperatorTok{:}\DecValTok{50}\NormalTok{),}
                      \StringTok{"stem"}\NormalTok{ =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{15}\NormalTok{, }\DecValTok{17}\OperatorTok{:}\DecValTok{18}\NormalTok{, }\DecValTok{42}\OperatorTok{:}\DecValTok{44}\NormalTok{))}

\NormalTok{## Measuring disparity for each group}
\NormalTok{disparity_data <-}\StringTok{ }\KeywordTok{dispRity.per.group}\NormalTok{(BeckLee_mat50, }\DataTypeTok{group =}\NormalTok{ mammal_groups,}
                                     \DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, variances))}
\end{Highlighting}
\end{Shaded}

We can display the disparity of both groups by simply looking at the
output variable (\texttt{disparity\_data}) and then summarising the
\texttt{disparity\_data} object and plotting it, and/or by performing a
statistical test to compare disparity across the groups (here a Wilcoxon
test).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Print the disparity_data object}
\NormalTok{disparity_data}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 2 customised subsamples for 50 elements with 48 dimensions:
##     crown, stem.
## Data was bootstrapped 100 times (method:"full").
## Disparity was calculated as: metric.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Summarising disparity in the different groups}
\KeywordTok{summary}\NormalTok{(disparity_data)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs bs.median  2.5%   25%   75% 97.5%
## 1      crown 30 1.995     1.936 1.866 1.914 1.947 1.971
## 2       stem 20 1.715     1.632 1.541 1.604 1.667 1.695
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Plotting the results}
\KeywordTok{plot}\NormalTok{(disparity_data)}
\end{Highlighting}
\end{Shaded}

\includegraphics{manual_files/figure-latex/unnamed-chunk-15-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Testing for a difference between the groups}
\KeywordTok{test.dispRity}\NormalTok{(disparity_data, }\DataTypeTok{test =}\NormalTok{ wilcox.test, }\DataTypeTok{details =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## $`crown : stem`
## $`crown : stem`[[1]]
## 
##  Wilcoxon rank sum test with continuity correction
## 
## data:  dots[[1L]][[1L]] and dots[[2L]][[1L]]
## W = 10000, p-value < 2.2e-16
## alternative hypothesis: true location shift is not equal to 0
\end{verbatim}

\chapter{Details of specific
functions}\label{details-of-specific-functions}

The following section contains information specific to some functions.
If any of your questions are not covered in these sections, please refer
to the function help files in R, send me an email
(\href{mailto:guillert@tcd.ie}{\nolinkurl{guillert@tcd.ie}}), or raise
an issue on GitHub. The several tutorials below describe specific
functionalities of certain functions; please always refer to the
function help files for the full function documentation!

\hypertarget{time-slicing}{\section{Time slicing}\label{time-slicing}}

The function \texttt{time.subsamples} allows users to divide the matrix
into different time subsamples or slices given a dated phylogeny that
contains all the elements (i.e.~taxa) from the matrix. Each subsample
generated by this function will then contain all the elements present at
a specific point in time or during a specific period in time.

Two types of time subsamples can be performed by using the
\texttt{method} option:

\begin{itemize}
\tightlist
\item
  Discrete time subsamples (or time-binning) using
  \texttt{method\ =\ discrete}
\item
  Continuous time subsamples (or time-slicing) using
  \texttt{method\ =\ continuous}
\end{itemize}

For the time-slicing method details see \citet{GuillermeSTD}. For both
methods, the function takes the \texttt{time} argument which can be a
vector of \texttt{numeric} values for:

\begin{itemize}
\tightlist
\item
  Defining the boundaries of the time bins (when
  \texttt{method\ =\ discrete})
\item
  Defining the time slices (when \texttt{method\ =\ continuous})
\end{itemize}

Otherwise, the \texttt{time} argument can be set as a single
\texttt{numeric} value for automatically generating a given number of
equidistant time-bins/slices. Additionally, it is also possible to input
a dataframe containing the first and last occurrence data (FAD/LAD) for
taxa that span over a longer time than the given tips/nodes age, so taxa
can appear in more than one time bin/slice.

Here is an example for \texttt{method\ =\ discrete}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Generating three time bins containing the taxa present every 40 Ma}
\KeywordTok{time.subsamples}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat50, }\DataTypeTok{tree =}\NormalTok{ BeckLee_tree, }\DataTypeTok{method =} \StringTok{"discrete"}\NormalTok{,}
                \DataTypeTok{time =} \KeywordTok{c}\NormalTok{(}\DecValTok{120}\NormalTok{, }\DecValTok{80}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{0}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 3 discrete time subsamples for 50 elements:
##     120 - 80, 80 - 40, 40 - 0.
\end{verbatim}

Note that we can also generate equivalent results by just telling the
function that we want three time-bins as follow:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Automatically generate three equal length bins:}
\KeywordTok{time.subsamples}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat50, }\DataTypeTok{tree =}\NormalTok{ BeckLee_tree, }\DataTypeTok{method =} \StringTok{"discrete"}\NormalTok{,}
                \DataTypeTok{time =} \DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 3 discrete time subsamples for 50 elements:
##     133.51104 - 89.00736, 89.00736 - 44.50368, 44.50368 - 0.
\end{verbatim}

In this example, the taxa were split inside each time-bin according to
their age. However, the taxa here are considered as single points in
time. It is totally possible that some taxa could have had longer
longevity and that they exist in multiple time bins. In this case, it is
possible to include them in more than one bin by providing a table of
first and last occurrence dates (FAD/LAD). This table should have the
taxa names as row names and two columns for respectively the first and
last occurrence age:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Displaying the table of first and last occurrence dates for each taxa}
\KeywordTok{head}\NormalTok{(BeckLee_ages)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##             FAD  LAD
## Adapis     37.2 36.8
## Asioryctes 83.6 72.1
## Leptictis  33.9 33.3
## Miacis     49.0 46.7
## Mimotona   61.6 59.2
## Notharctus 50.2 47.0
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Generating time bins including taxa that might span between them}
\KeywordTok{time.subsamples}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat50, }\DataTypeTok{tree =}\NormalTok{ BeckLee_tree, }\DataTypeTok{method =} \StringTok{"discrete"}\NormalTok{,}
                \DataTypeTok{time =} \KeywordTok{c}\NormalTok{(}\DecValTok{120}\NormalTok{, }\DecValTok{80}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{FADLAD =}\NormalTok{ BeckLee_ages)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 3 discrete time subsamples for 50 elements:
##     120 - 80, 80 - 40, 40 - 0.
\end{verbatim}

When using this method, the oldest boundary of the first bin (or the
first slice, see below) is automatically generated as the root age plus
1\% of the tree length, as long as at least three elements/taxa are
present at that point in time. The algorithm adds an extra 1\% tree
length until reaching the required minimum of three elements. It is also
possible to include nodes in each bin by using
\texttt{inc.nodes\ =\ TRUE} and providing a matrix that contains the
ordinated distance among tips \emph{and} nodes.

For the time-slicing method (\texttt{method\ =\ continuous}), the idea
is fairly similar. This option, however, requires a matrix that contains
the ordinated distance among taxa \emph{and} nodes and an extra argument
describing the assumed evolutionary model (via the \texttt{model}
argument). This model argument is used when the time slice occurs along
a branch of the tree rather than on a tip or a node, meaning that a
decision must be made about what the value for the branch should be. The
model can be one of the following:

\begin{itemize}
\tightlist
\item
  \texttt{acctran} where the data chosen along the branch is always the
  one of the descendant
\item
  \texttt{deltran} where the data chosen along the branch is always the
  one of the ancestor
\item
  \texttt{punctuated} where the data chosen along the branch is randomly
  chosen between the descendant or the ancestor
\item
  \texttt{gradual} where the data chosen along the branch is either the
  descendant or the ancestor depending on branch length
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Generating four time slices every 40 million years under a model of gradual evolution}
\KeywordTok{time.subsamples}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat99, }\DataTypeTok{tree =}\NormalTok{ BeckLee_tree, }
    \DataTypeTok{method =} \StringTok{"continuous"}\NormalTok{, }\DataTypeTok{model =} \StringTok{"gradual"}\NormalTok{, }\DataTypeTok{time =} \KeywordTok{c}\NormalTok{(}\DecValTok{120}\NormalTok{, }\DecValTok{80}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{0}\NormalTok{),}
    \DataTypeTok{FADLAD =}\NormalTok{ BeckLee_ages)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 4 continuous (gradual) time subsamples for 99 elements:
##     120, 80, 40, 0.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Generating four time slices automatically}
\KeywordTok{time.subsamples}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat99, }\DataTypeTok{tree =}\NormalTok{ BeckLee_tree,}
    \DataTypeTok{method =} \StringTok{"continuous"}\NormalTok{, }\DataTypeTok{model =} \StringTok{"gradual"}\NormalTok{, }\DataTypeTok{time =} \DecValTok{4}\NormalTok{, }\DataTypeTok{FADLAD =}\NormalTok{ BeckLee_ages)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 4 continuous (gradual) time subsamples for 99 elements:
##     133.51104, 89.00736, 44.50368, 0.
\end{verbatim}

\hypertarget{customised-subsamples}{\section{Customised
subsamples}\label{customised-subsamples}}

Another way of separating elements into different categories is to use
customised subsamples as briefly explained
\protect\hyperlink{disparity-among-groups}{above}. This function simply
takes the list of elements to put in each group (whether they are the
actual element names or their position in the matrix).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating the two groups as a list}
\NormalTok{mammal_groups <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}\StringTok{"crown"}\NormalTok{ =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{16}\NormalTok{, }\DecValTok{19}\OperatorTok{:}\DecValTok{41}\NormalTok{, }\DecValTok{45}\OperatorTok{:}\DecValTok{50}\NormalTok{),}
                      \StringTok{"stem"}\NormalTok{ =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{15}\NormalTok{, }\DecValTok{17}\OperatorTok{:}\DecValTok{18}\NormalTok{, }\DecValTok{42}\OperatorTok{:}\DecValTok{44}\NormalTok{))}

\NormalTok{## Separating the dataset into two different groups}
\KeywordTok{custom.subsamples}\NormalTok{(BeckLee_mat50, }\DataTypeTok{group =}\NormalTok{ mammal_groups)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 2 customised subsamples for 50 elements:
##     crown, stem.
\end{verbatim}

Elements can easily be assigned to different groups if necessary!

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating the three groups as a list}
\NormalTok{mammal_groups <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}\StringTok{"crown"}\NormalTok{ =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{16}\NormalTok{, }\DecValTok{19}\OperatorTok{:}\DecValTok{41}\NormalTok{, }\DecValTok{45}\OperatorTok{:}\DecValTok{50}\NormalTok{),}
                      \StringTok{"stem"}\NormalTok{ =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{15}\NormalTok{, }\DecValTok{17}\OperatorTok{:}\DecValTok{18}\NormalTok{, }\DecValTok{42}\OperatorTok{:}\DecValTok{44}\NormalTok{).}
                      \StringTok{"all"}\NormalTok{ =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{50}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{bootstraps-and-rarefactions}{\section{Bootstraps and
rarefactions}\label{bootstraps-and-rarefactions}}

One important step in analysing ordinated matrices is to
pseudo-replicate the data to see how robust the results are, and how
sensitive they are to outliers in the dataset. This can be achieved
using the function \texttt{boot.matrix} to bootstrap and/or rarefy the
data. The default options will bootstrap the matrix 100 times without
rarefaction using the ``full'' bootstrap method (see below):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Default bootstrapping}
\KeywordTok{boot.matrix}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat50)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 50 elements with 48 dimensions.
## Data was bootstrapped 100 times (method:"full").
\end{verbatim}

The number of bootstrap replicates can be defined using the
\texttt{bootstraps} option. The method can be modified by controlling
which bootstrap algorithm to use through the \texttt{boot.type}
argument. Currently two algorithms are implemented:

\begin{itemize}
\tightlist
\item
  \texttt{full} where the bootstrapping is entirely stochastic (\emph{n}
  elements are replaced by any \emph{m} elements drawn from the data)
\item
  \texttt{single} where only one random element is replaced by one other
  random element for each pseudo-replicate
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Bootstrapping with the single bootstrap method}
\KeywordTok{boot.matrix}\NormalTok{(BeckLee_mat50, }\DataTypeTok{boot.type =} \StringTok{"single"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 50 elements with 48 dimensions.
## Data was bootstrapped 100 times (method:"single").
\end{verbatim}

This function also allows users to rarefy the data using the
\texttt{rarefaction} argument. Rarefaction allows users to limit the
number of elements to be drawn at each bootstrap replication. This is
useful if, for example, one is interested in looking at the effect of
reducing the number of elements on the results of an analysis.

This can be achieved by using the \texttt{rarefaction} option that draws
only \emph{n-x} at each bootstrap replicate (where \emph{x} is the
number of elements not sampled). The default argument is \texttt{FALSE}
but it can be set to \texttt{TRUE} to fully rarefy the data (i.e.~remove
\emph{x} elements for the number of pseudo-replicates, where \emph{x}
varies from the maximum number of elements present in each subsample to
a minimum of three elements). It can also be set to one or more
\texttt{numeric} values to only rarefy to the corresponding number of
elements.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Bootstrapping with the full rarefaction}
\KeywordTok{boot.matrix}\NormalTok{(BeckLee_mat50, }\DataTypeTok{bootstraps =} \DecValTok{20}\NormalTok{, }\DataTypeTok{rarefaction =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 50 elements with 48 dimensions.
## Data was bootstrapped 20 times (method:"full") and fully rarefied.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Or with a set number of rarefaction levels}
\KeywordTok{boot.matrix}\NormalTok{(BeckLee_mat50, }\DataTypeTok{bootstraps =} \DecValTok{20}\NormalTok{, }\DataTypeTok{rarefaction =} \KeywordTok{c}\NormalTok{(}\DecValTok{6}\OperatorTok{:}\DecValTok{8}\NormalTok{, }\DecValTok{3}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 50 elements with 48 dimensions.
## Data was bootstrapped 20 times (method:"full") and rarefied to 6, 7, 8, 3 elements.
\end{verbatim}

One additional important argument is \texttt{dimensions} that specifies
how many dimensions from the matrix should be used for further analysis.
When missing, all dimensions from the ordinated matrix are used.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Using the first 50% of the dimensions}
\KeywordTok{boot.matrix}\NormalTok{(BeckLee_mat50, }\DataTypeTok{dimensions =} \FloatTok{0.5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 50 elements with 24 dimensions.
## Data was bootstrapped 100 times (method:"full").
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Using the first 10 dimensions}
\KeywordTok{boot.matrix}\NormalTok{(BeckLee_mat50, }\DataTypeTok{dimensions =} \DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 50 elements with 10 dimensions.
## Data was bootstrapped 100 times (method:"full").
\end{verbatim}

Of course, one could directly supply the subsamples generated above
(using \texttt{time.subsamples} or \texttt{custom.subsamples}) to this
function.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating subsamples of crown and stem mammals}
\NormalTok{crown_stem <-}\StringTok{ }\KeywordTok{custom.subsamples}\NormalTok{(BeckLee_mat50,}
                                \DataTypeTok{group =} \KeywordTok{list}\NormalTok{(}\StringTok{"crown"}\NormalTok{ =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{16}\NormalTok{, }\DecValTok{19}\OperatorTok{:}\DecValTok{41}\NormalTok{, }\DecValTok{45}\OperatorTok{:}\DecValTok{50}\NormalTok{), }
                                             \StringTok{"stem"}\NormalTok{ =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{15}\NormalTok{, }\DecValTok{17}\OperatorTok{:}\DecValTok{18}\NormalTok{, }\DecValTok{42}\OperatorTok{:}\DecValTok{44}\NormalTok{)))}
\NormalTok{## Bootstrapping and rarefying these groups}
\KeywordTok{boot.matrix}\NormalTok{(crown_stem, }\DataTypeTok{bootstraps =} \DecValTok{200}\NormalTok{, }\DataTypeTok{rarefaction =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 2 customised subsamples for 50 elements with 48 dimensions:
##     crown, stem.
## Data was bootstrapped 200 times (method:"full") and fully rarefied.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating time slice subsamples}
\NormalTok{time_slices <-}\StringTok{ }\KeywordTok{time.subsamples}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat99, }\DataTypeTok{tree =}\NormalTok{ BeckLee_tree, }
                               \DataTypeTok{method =} \StringTok{"continuous"}\NormalTok{, }\DataTypeTok{model =} \StringTok{"gradual"}\NormalTok{, }
                               \DataTypeTok{time =} \KeywordTok{c}\NormalTok{(}\DecValTok{120}\NormalTok{, }\DecValTok{80}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{0}\NormalTok{),}
                               \DataTypeTok{FADLAD =}\NormalTok{ BeckLee_ages)}

\NormalTok{## Bootstrapping the time slice subsamples}
\KeywordTok{boot.matrix}\NormalTok{(time_slices, }\DataTypeTok{bootstraps =} \DecValTok{100}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 4 continuous (gradual) time subsamples for 99 elements with 97 dimensions:
##     120, 80, 40, 0.
## Data was bootstrapped 100 times (method:"full").
\end{verbatim}

\hypertarget{disparity-metrics}{\section{Disparity
metrics}\label{disparity-metrics}}

There are many ways of measuring disparity! In brief, disparity is a
summary metric that will represent an aspect of an ordinated space
(e.g.~a MDS, PCA, PCO, PCoA). For example, one can look at ellipsoid
hyper-volume of the ordinated space \citet{DonohueDim}, the sum and the
product of the ranges and variances \citet{Wills1994} or the median
position of the elements relative to their centroid \citet{Wills1994}.
Of course, there are many more examples of metrics one can use for
describing some aspect of the ordinated space, with some performing
better than other ones at particular descriptive tasks, and some being
more generalist.

Because of this great diversity of metrics, the package
\texttt{dispRity} does not have one way to measure disparity but rather
proposes to facilitate users in defining their own disparity metric that
will best suit their particular analysis. In fact, the core function of
the package, \texttt{dispRity}, allows the user to define any metric
with the \texttt{metric} argument. However the \texttt{metric} argument
has to follow certain rules:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  It must be composed from one to three \texttt{function} objects;
\item
  The function(s) must take as a first argument a \texttt{matrix} or a
  \texttt{vector};
\item
  The function(s) must be of one of the three dimension-levels described
  below;
\item
  At least one of the functions must be of dimension-level 1 or 2 (see
  below).
\end{enumerate}

\subsection{The function
dimension-levels}\label{the-function-dimension-levels}

The metric function dimension-levels determine the ``dimensionality of
decomposition'' of the input matrix. In other words, each
dimension-level designates the dimensions of the output, i.e.~either
three (a \texttt{matrix}); two (a \texttt{vector}); or one (a single
\texttt{numeric} value) dimension.

\begin{figure}
\centering
\includegraphics{dispRity_fun.png}
\caption{Illustration of the different dimension-levels of functions
with an input \texttt{matrix}}
\end{figure}

\subsubsection{Dimension-level 1
functions}\label{dimension-level-1-functions}

A dimension-level 1 function will decompose a \texttt{matrix} or a
\texttt{vector} into a single value:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating a dummy matrix}
\NormalTok{dummy_matrix <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\KeywordTok{rnorm}\NormalTok{(}\DecValTok{12}\NormalTok{), }\DecValTok{4}\NormalTok{, }\DecValTok{3}\NormalTok{)}

\NormalTok{## Example of dimension-level 1 functions}
\KeywordTok{mean}\NormalTok{(dummy_matrix)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] -0.3227241
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{median}\NormalTok{(dummy_matrix)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] -0.2690165
\end{verbatim}

Any summary metric such as mean or median are good examples of
dimension-level 1 functions as they reduce the matrix to a single
dimension (i.e.~one value).

\subsubsection{Dimension-level 2
functions}\label{dimension-level-2-functions}

A dimension-level 2 function will decompose a \texttt{matrix} into a
\texttt{vector}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Defining the function as the product of rows}
\NormalTok{prod.rows <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(matrix) }\KeywordTok{apply}\NormalTok{(matrix, }\DecValTok{1}\NormalTok{, prod)}

\NormalTok{## A dimension-level 2 metric}
\KeywordTok{prod.rows}\NormalTok{(dummy_matrix)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] -1.2630799  0.2148864 -0.1797556 -0.2421790
\end{verbatim}

Several dimension-level 2 functions are implemented in \texttt{dispRity}
(see \texttt{?dispRity.metric}) such as the \texttt{variances} or
\texttt{ranges} functions that calculate the variance or the range of
each dimension of the ordinated matrix respectively.

\subsubsection{Dimension-level 3
functions}\label{dimension-level-3-functions}

Finally a dimension-level 3 function will transform the matrix into
another matrix. Note that the dimension of the output matrix doesn't
need to match the the input matrix:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## A dimension-level 3 metric}
\KeywordTok{var}\NormalTok{(dummy_matrix)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##          [,1]      [,2]      [,3]
## [1,] 1.916420 1.9501955 0.2907060
## [2,] 1.950196 3.2948635 0.3958234
## [3,] 0.290706 0.3958234 0.6704976
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## A dimension-level 3 metric with a forced matrix output}
\KeywordTok{as.matrix}\NormalTok{(}\KeywordTok{dist}\NormalTok{(dummy_matrix))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##          1        2        3        4
## 1 0.000000 5.140790 4.005068 3.120827
## 2 5.140790 0.000000 2.179318 2.982267
## 3 4.005068 2.179318 0.000000 2.174896
## 4 3.120827 2.982267 2.174896 0.000000
\end{verbatim}

\subsection{\texorpdfstring{\texttt{make.metric}}{make.metric}}\label{make.metric}

Of course, functions can be more complex and involve multiple operations
such as the \texttt{centroids} function (see \texttt{?dispRity.metric})
that calculates the Euclidean distance between each element and the
centroid of the ordinated space. The \texttt{make.metric} function
implemented in \texttt{dispRity} is designed to help test and find the
dimension-level of the functions. This function tests:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  If your function can deal with a \texttt{matrix} or a \texttt{vector}
  as an input;
\item
  Your function's dimension-level according to its output
  (dimension-level 1, 2 or 3, see above);
\item
  Whether the function can be implemented in the \texttt{dispRity}
  function (the function is fed into a \texttt{lapply} loop).
\end{enumerate}

For example, let's see if the functions described above are the right
dimension-levels:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Which dimension-level is the mean function? And can it be used in dispRity?}
\KeywordTok{make.metric}\NormalTok{(mean)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## mean outputs a single value.
## mean is detected as being a dimension-level 1 function.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Which dimension-level is the prod.rows function? And can it be used in dispRity?}
\KeywordTok{make.metric}\NormalTok{(prod.rows)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## prod.rows outputs a matrix object.
## prod.rows is detected as being a dimension-level 2 function.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Which dimension-level is the var function? And can it be used in dispRity?}
\KeywordTok{make.metric}\NormalTok{(var)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## var outputs a matrix object.
## var is detected as being a dimension-level 3 function.
## Additional dimension-level 2 and/or 1 function(s) will be needed.
\end{verbatim}

A non verbose version of the function is also available. This can be
done using the option \texttt{silent\ =\ TRUE} and will simply output
the dimension-level of the metric.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Testing whether mean is dimension-level 1}
\ControlFlowTok{if}\NormalTok{(}\KeywordTok{make.metric}\NormalTok{(mean, }\DataTypeTok{silent =} \OtherTok{TRUE}\NormalTok{) }\OperatorTok{!=}\StringTok{ "level1"}\NormalTok{) \{}
    \KeywordTok{message}\NormalTok{(}\StringTok{"The metric is not dimension-level 1."}\NormalTok{)}
\NormalTok{\}}
\NormalTok{## Testing whether var is dimension-level 1}
\ControlFlowTok{if}\NormalTok{(}\KeywordTok{make.metric}\NormalTok{(var, }\DataTypeTok{silent =} \OtherTok{TRUE}\NormalTok{) }\OperatorTok{!=}\StringTok{ "level1"}\NormalTok{) \{}
    \KeywordTok{message}\NormalTok{(}\StringTok{"The metric is not dimension-level 1."}\NormalTok{)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## The metric is not dimension-level 1.
\end{verbatim}

\subsection{\texorpdfstring{Metrics in the \texttt{dispRity}
function}{Metrics in the dispRity function}}\label{metrics-in-the-disprity-function}

Using this metric structure, we can easily use any disparity metric in
the \texttt{dispRity} function as follows:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Measuring disparity as the standard deviation of all the values of the}
\NormalTok{## ordinated matrix (dimension-level 1 function).}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(BeckLee_mat50, }\DataTypeTok{metric =}\NormalTok{ sd))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs
## 1          1 50 0.201
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Measuring disparity as the standard deviation of the variance of each axis of}
\NormalTok{## the ordinated matrix (dimension-level 1 and 2 functions).}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(BeckLee_mat50, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sd, variances)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs
## 1          1 50 0.028
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Measuring disparity as the standard deviation of the variance of each axis of}
\NormalTok{## the variance covariance matrix (dimension-level 1, 2 and 3 functions).}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(BeckLee_mat50, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sd, variances, var)), }\DataTypeTok{round =} \DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n          obs
## 1          1 50 0.0001025857
\end{verbatim}

Note that the order of each function in the metric argument does not
matter, the \texttt{dispRity} function will automatically detect the
function dimension-levels (using \texttt{make.metric}) and apply them to
the data in decreasing order (dimension-level 3 \textgreater{} 2
\textgreater{} 1).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Disparity as the standard deviation of the variance of each axis of the}
\NormalTok{## variance covariance matrix:}
\NormalTok{disparity1 <-}\StringTok{ }\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(BeckLee_mat50, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sd, variances, var)),}
                      \DataTypeTok{round =} \DecValTok{10}\NormalTok{)}

\NormalTok{## Same as above but using a different function order for the metric argument}
\NormalTok{disparity2 <-}\StringTok{ }\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(BeckLee_mat50, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(variances, sd, var)),}
                      \DataTypeTok{round =} \DecValTok{10}\NormalTok{)}

\NormalTok{## Both ways output the same disparity values:}
\NormalTok{disparity1 }\OperatorTok{==}\StringTok{ }\NormalTok{disparity2}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##      subsamples    n  obs
## [1,]       TRUE TRUE TRUE
\end{verbatim}

In these examples, we considered disparity to be a single value. For
example, in the previous example, we defined disparity as the standard
deviation of the variances of each column of the variance/covariance
matrix (\texttt{metric\ =\ c(variances,\ sd,\ var)}). It is, however,
possible to calculate
\protect\hyperlink{disparity-as-a-distribution}{disparity as a
distribution}.

\subsection{\texorpdfstring{Metrics implemented in
\texttt{dispRity}}{Metrics implemented in dispRity}}\label{metrics-implemented-in-disprity}

Several disparity metrics are implemented in the \texttt{dispRity}
package. The detailed list can be found in \texttt{?dispRity.metric}
along with some description of each metric.

\begin{longtable}[]{@{}llll@{}}
\toprule
\begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
Level\strut
\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
Name\strut
\end{minipage} & \begin{minipage}[b]{0.61\columnwidth}\raggedright\strut
Description\strut
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Source\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
3\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{ellipse.volume}1\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The volume of the ellipsoid of the space\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\citep{DonohueDim}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
3\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{convhull.surface}\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The surface of the convex hull formed by all the elements\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\href{https://cran.r-project.org/web/packages/geometry/index.html}{\texttt{geometry}}\texttt{::convhulln}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
3\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{convhull.volume}\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The volume of the convex hull formed by all the elements\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\href{https://cran.r-project.org/web/packages/geometry/index.html}{\texttt{geometry}}\texttt{::convhulln}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
3\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{hypervolume}\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The volume of the ordinated space\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\href{https://cran.r-project.org/web/packages/hypervolume/index.html}{\texttt{hypervolume}}\texttt{::hypervolume}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
3\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{diagonal}\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The longest distance in the ordinated space (like the diagonal in two
dimensions)\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\texttt{dispRity}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
2\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{ranges}\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The range of each dimension\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\texttt{dispRity}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
2\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{variances}\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The variance of each dimension\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\texttt{dispRity}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
2\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{centroids}2\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The distance between each element and the centroid of the ordinated
space\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\texttt{dispRity}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
1\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{mode.val}\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The modal value\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\texttt{dispRity}\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

1: This function uses an estimation of the eigenvalue that only works
for MDS or PCoA ordinations (not PCA).

2: Note that by default, the centroid is the centroid of the elements.
It can, however, be fixed to a different value by using the
\texttt{centroid} argument
\texttt{centroids(space,\ centroid\ =\ rep(0,\ ncol(space)))}, for
example the origin of the ordinated space.

\subsection{Equations and
implementations}\label{equations-and-implementations}

Some of the functions described below are implemented in the
\texttt{dispRity} package and do not require any other packages to
calculate
(\href{https://github.com/TGuillerme/dispRity/blob/master/R/dispRity.metric.R}{see
implementation here}).

\begin{equation}
    ellipse.volume = \frac{\pi^{k/2}}{\Gamma(\frac{k}{2}+1)}\displaystyle\prod_{i=1}^{k} (\lambda_{i}^{0.5})
\end{equation}

Where \emph{k} is the number of dimensions, and \(\lambda_i\) is the
eigenvalue of each dimension.

\begin{equation}
    diagonal = \sqrt{\sum_{i=1}^{k}|max(k_i) - min(k_i)|}
\end{equation}

Where \emph{k} is the number of dimensions.

\begin{equation}
    ranges = |max(k_i) - min(k_i)|
\end{equation}

Where \emph{k} is the number of dimensions.

\begin{equation}
    variances = \sigma^{2}{k_i}
\end{equation}

Where \emph{k} is the number of dimensions, and \(\sigma^{2}\) is their
variance.

\begin{equation}
    centroids = \sqrt{\sum_{i=1}^{n}{({k}_{n}-Centroid_{k})^2}}
\end{equation}

Where \emph{n} is each element in the ordinated space, \emph{k} is the
number of dimensions, and \(Centroid_{k}\) is their mean (or can be set
to another value).

\subsection{Using the different disparity
metrics}\label{using-the-different-disparity-metrics}

Here is a brief demonstration of the main metrics implemented in
\texttt{dispRity}. First, we will create a dummy/simulated ordinated
space using the \texttt{space.maker} utility function (more about that
\protect\hyperlink{space.maker}{here}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating a 10*5 normal space}
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{)}
\NormalTok{dummy_space <-}\StringTok{ }\KeywordTok{space.maker}\NormalTok{(}\DecValTok{10}\NormalTok{, }\DecValTok{5}\NormalTok{, rnorm)}
\end{Highlighting}
\end{Shaded}

We will use this simulated space to demonstrate the different metrics.

\subsubsection{Volumes and surface
metrics}\label{volumes-and-surface-metrics}

The functions \texttt{ellipse.volume}, \texttt{convhull.surface},
\texttt{convhull.volume} and \texttt{hyper.volume} all measure the
surface or the volume of the ordinated space occupied:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating the ellipsoid volume}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ ellipse.volume))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs
## 1          1 10 257.8
\end{verbatim}

\begin{quote}
Because there is only one subsample (i.e.~one matrix) in the dispRity
object, this operation is the equivalent of
\texttt{ellipse.volume(dummy\_space)} (with rounding).
\end{quote}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating the convex hull surface}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ convhull.surface))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs
## 1          1 10 11.91
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating the convex hull volume}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ convhull.volume))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs
## 1          1 10 1.031
\end{verbatim}

The convex hull functions make a (good) estimation of the
multidimensional properties of the ordinated space. For the full and
correct calculation of the volume of the ordinated space, it is possible
to use the \texttt{hyper.volume} function that has more options but
takes longer to calculate.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating the true multidimensional volume}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ hyper.volume))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in hypervolume::hypervolume(matrix, method = method, ...): Log number of observations (1.61) is less than or equal to the number of dimensions (4).
## You may not have enough data to accurately estimate a hypervolume with this dimensionality.
## Consider reducing the dimensionality of the analysis.
\end{verbatim}

\begin{verbatim}
## Warning in hypervolume::hypervolume(matrix, method = method, ...): Log number of observations (2.30) is less than or equal to the number of dimensions (5).
## You may not have enough data to accurately estimate a hypervolume with this dimensionality.
## Consider reducing the dimensionality of the analysis.
\end{verbatim}

\begin{verbatim}
##   subsamples  n   obs
## 1          1 10 408.2
\end{verbatim}

\begin{quote}
Cautionary note: measuring volumes in a high number of dimensions can be
strongly affected by the
\href{https://en.wikipedia.org/wiki/Curse_of_dimensionality}{curse of
dimensionality} that often results in near 0 disparity values.
\end{quote}

\subsubsection{Ranges, variances and
diagonal}\label{ranges-variances-and-diagonal}

The functions \texttt{ranges}, \texttt{variances} and \texttt{diagonal}
all measure properties of the ordinated space based on its dimensional
properties (they are also less affected by the ``curse of
dimensionality''):

\texttt{ranges} and \texttt{variances} both work on the same principle
and measure the range/variance of each dimension:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating the ranges of each dimension in the ordinated space}
\KeywordTok{ranges}\NormalTok{(dummy_space)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 2.430909 3.726481 2.908329 2.735739 1.588603
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating disparity as the distribution of these ranges}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ ranges))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n obs.median  2.5%   25%   75% 97.5%
## 1          1 10      2.736 1.673 2.431 2.908 3.645
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating disparity as the sum and the product of these ranges}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, ranges)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs
## 1          1 10 13.39
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(prod, ranges)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs
## 1          1 10 114.5
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating the variances of each dimension in the ordinated space}
\KeywordTok{variances}\NormalTok{(dummy_space)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 0.6093144 1.1438620 0.9131859 0.6537768 0.3549372
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating disparity as the distribution of these variances}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ variances))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n obs.median 2.5%   25%   75% 97.5%
## 1          1 10      0.654 0.38 0.609 0.913 1.121
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating disparity as the sum and the product of these variances}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, variances)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs
## 1          1 10 3.675
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(prod, variances)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs
## 1          1 10 0.148
\end{verbatim}

The \texttt{diagonal} function measures the multidimensional diagonal of
the whole space (i.e.~in our case the longest Euclidean distance in our
five dimensional space):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating the ordinated space's diagonal}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ diagonal))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs
## 1          1 10 3.659
\end{verbatim}

\begin{quote}
This metric is only a Euclidean diagonal (mathematically valid) if the
dimensions within the space are all orthogonal!
\end{quote}

\subsubsection{Centroids metric}\label{centroids-metric}

The \texttt{centroids} metric allows users to measure the position of
the different elements compared to a fixed point in the ordinated space.
By default, this function measures the distance between each element and
their centroid (centre point):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The distribution of the distances between each element and their centroid}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ centroids))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n obs.median  2.5%   25%   75% 97.5%
## 1          1 10      1.435 0.788 1.267 1.993 3.167
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Disparity as the median value of these distances}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(median, centroids)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs
## 1          1 10 1.435
\end{verbatim}

It is however possible to fix the coordinates of the centroid to a
specific point in the ordinated space, as long as it has the correct
number of dimensions:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The distance between each element and the origin of the ordinated space}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ centroids, }\DataTypeTok{centroid =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n obs.median  2.5% 25%   75% 97.5%
## 1          1 10      1.487 0.785 1.2 2.044 3.176
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Disparity as the distance between each element and a specific point in space}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ centroids, }\DataTypeTok{centroid =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{3}\NormalTok{,}\DecValTok{4}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n obs.median  2.5%   25%   75% 97.5%
## 1          1 10      5.489 4.293 5.032 6.155 6.957
\end{verbatim}

\section{Summarising dispRity data
(plots)}\label{summarising-disprity-data-plots}

Because of its architecture, printing \texttt{dispRity} objects only
summarises their content but does not print the disparity value measured
or associated analysis (more about this
\protect\hyperlink{manipulating-dispRity-objects}{here}). To actually
see what is in a dispRity object, one can either use the
\texttt{summary} function for visualising the data in a table or
\texttt{plot} to have a graphical representation of the results.

\subsection{\texorpdfstring{Summarising \texttt{dispRity}
data}{Summarising dispRity data}}\label{summarising-disprity-data}

This function is an S3 function (\texttt{summary.dispRity}) allowing
users to summarise the content of \texttt{dispRity} objects that contain
disparity calculations.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Example data from previous sections}
\NormalTok{crown_stem <-}\StringTok{ }\KeywordTok{custom.subsamples}\NormalTok{(BeckLee_mat50,}
                                \DataTypeTok{group =} \KeywordTok{list}\NormalTok{(}\StringTok{"crown"}\NormalTok{ =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{16}\NormalTok{, }\DecValTok{19}\OperatorTok{:}\DecValTok{41}\NormalTok{, }\DecValTok{45}\OperatorTok{:}\DecValTok{50}\NormalTok{), }
                                             \StringTok{"stem"}\NormalTok{ =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{15}\NormalTok{, }\DecValTok{17}\OperatorTok{:}\DecValTok{18}\NormalTok{, }\DecValTok{42}\OperatorTok{:}\DecValTok{44}\NormalTok{)))}
\NormalTok{## Bootstrapping and rarefying these groups}
\NormalTok{boot_crown_stem <-}\StringTok{ }\KeywordTok{boot.matrix}\NormalTok{(crown_stem, }\DataTypeTok{bootstraps =} \DecValTok{100}\NormalTok{, }\DataTypeTok{rarefaction =} \OtherTok{TRUE}\NormalTok{)}
\NormalTok{## Calculate disparity}
\NormalTok{disparity_crown_stem <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(boot_crown_stem, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, variances))}

\NormalTok{## Creating time slice subsamples}
\NormalTok{time_slices <-}\StringTok{ }\KeywordTok{time.subsamples}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat99, }\DataTypeTok{tree =}\NormalTok{ BeckLee_tree, }
    \DataTypeTok{method =} \StringTok{"continuous"}\NormalTok{, }\DataTypeTok{model =} \StringTok{"gradual"}\NormalTok{, }\DataTypeTok{time =} \KeywordTok{c}\NormalTok{(}\DecValTok{120}\NormalTok{, }\DecValTok{80}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{0}\NormalTok{),}
    \DataTypeTok{FADLAD =}\NormalTok{ BeckLee_ages)}
\NormalTok{## Bootstrapping the time slice subsamples}
\NormalTok{boot_time_slices <-}\StringTok{ }\KeywordTok{boot.matrix}\NormalTok{(time_slices, }\DataTypeTok{bootstraps =} \DecValTok{100}\NormalTok{)}
\NormalTok{## Calculate disparity}
\NormalTok{disparity_time_slices <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(boot_time_slices, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, variances))}

\NormalTok{## Creating time bin subsamples}
\NormalTok{time_bins <-}\StringTok{ }\KeywordTok{time.subsamples}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat99, }\DataTypeTok{tree =}\NormalTok{ BeckLee_tree, }
    \DataTypeTok{method =} \StringTok{"discrete"}\NormalTok{, }\DataTypeTok{time =} \KeywordTok{c}\NormalTok{(}\DecValTok{120}\NormalTok{, }\DecValTok{80}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{FADLAD =}\NormalTok{ BeckLee_ages,}
    \DataTypeTok{inc.nodes =} \OtherTok{TRUE}\NormalTok{)}
\NormalTok{## Bootstrapping the time bin subsamples}
\NormalTok{boot_time_bins <-}\StringTok{ }\KeywordTok{boot.matrix}\NormalTok{(time_bins, }\DataTypeTok{bootstraps =} \DecValTok{100}\NormalTok{)}
\NormalTok{## Calculate disparity}
\NormalTok{disparity_time_bins <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(boot_time_bins, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, variances))}
\end{Highlighting}
\end{Shaded}

These objects are easy to summarise as follows:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Default summary}
\KeywordTok{summary}\NormalTok{(disparity_time_slices)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs bs.median  2.5%   25%   75% 97.5%
## 1        120  5 2.823     2.295 1.292 2.039 2.598 2.829
## 2         80 19 3.233     3.079 2.834 2.996 3.138 3.278
## 3         40 15 3.421     3.211 2.909 3.103 3.309 3.518
## 4          0 10 4.055     3.684 3.247 3.557 3.783 3.940
\end{verbatim}

Information about the number of elements in each subsample and the
observed (i.e.~non-bootstrapped) disparity are also calculated. This is
specifically handy when rarefying the data for example:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{head}\NormalTok{(}\KeywordTok{summary}\NormalTok{(disparity_crown_stem))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs bs.median  2.5%   25%   75% 97.5%
## 1      crown 30 1.995     1.926 1.863 1.912 1.948 1.965
## 2      crown 29    NA     1.931 1.878 1.909 1.944 1.967
## 3      crown 28    NA     1.931 1.882 1.908 1.948 1.978
## 4      crown 27    NA     1.930 1.859 1.908 1.946 1.966
## 5      crown 26    NA     1.928 1.867 1.907 1.948 1.978
## 6      crown 25    NA     1.934 1.853 1.904 1.948 1.974
\end{verbatim}

The summary functions can also take various options such as:

\begin{itemize}
\tightlist
\item
  \texttt{quantile} values for the confidence interval levels (by
  default, the 50 and 95 quantiles are calculated)
\item
  \texttt{cent.tend} for the central tendency to use for summarising the
  results (default is \texttt{median})
\item
  rounding\texttt{option\ corresponding\ to\ the\ number\ of\ decimal\ places\ to\ print\ (default\ is}2`)
\item
  \texttt{recall} option for printing the call of the \texttt{dispRity}
  object as well (default is \texttt{FALSE})
\end{itemize}

These options can easily be changed from the defaults as follows:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Same as above but using the 88th quantile and the standard deviation as the summary }
\KeywordTok{summary}\NormalTok{(disparity_time_slices, }\DataTypeTok{quantile =} \DecValTok{88}\NormalTok{, }\DataTypeTok{cent.tend =}\NormalTok{ sd)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs bs.sd    6%   94%
## 1        120  5 2.823 0.398 1.658 2.705
## 2         80 19 3.233 0.116 2.896 3.243
## 3         40 15 3.421 0.161 2.967 3.485
## 4          0 10 4.055 0.184 3.342 3.907
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Printing the details of the object and rounding the values to the 5th decimal place}
\KeywordTok{summary}\NormalTok{(disparity_time_slices, }\DataTypeTok{recall =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{rounding =} \DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 4 continuous (gradual) time subsamples for 99 elements with 97 dimensions:
##     120, 80, 40, 0.
## Data was bootstrapped 100 times (method:"full").
## Disparity was calculated as: c(sum, variances).
\end{verbatim}

\begin{verbatim}
##   subsamples  n     obs bs.median    2.5%     25%     75%   97.5%
## 1        120  5 2.82292   2.29498 1.29247 2.03886 2.59781 2.82881
## 2         80 19 3.23312   3.07949 2.83426 2.99627 3.13758 3.27800
## 3         40 15 3.42091   3.21071 2.90858 3.10280 3.30911 3.51829
## 4          0 10 4.05457   3.68433 3.24678 3.55676 3.78275 3.93967
\end{verbatim}

Note that the summary table is a \texttt{data.frame}, hence it is as
easy to modify as any dataframe using \texttt{dplyr}. You can also
export it in \texttt{csv} format using \texttt{write.csv} or
\texttt{write\_csv} or even directly export into \texttt{LaTeX} format
using the following;

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Loading the xtable package}
\KeywordTok{require}\NormalTok{(xtable)}
\NormalTok{## Converting the table in LaTeX}
\KeywordTok{xtable}\NormalTok{(}\KeywordTok{summary}\NormalTok{(disparity_time_slices))}
\end{Highlighting}
\end{Shaded}

\subsection{\texorpdfstring{Plotting \texttt{dispRity}
data}{Plotting dispRity data}}\label{plotting-disprity-data}

An alternative (and more fun!) way to display the calculated disparity
is to plot the results using the S3 method \texttt{plot.dispRity}. This
function takes the same options as \texttt{summary.dispRity} along with
various graphical options described in the function help files (see
\texttt{?plot.dispRity}).

The plots can be of four different types:

\begin{itemize}
\tightlist
\item
  \texttt{continuous} for displaying continuous disparity curves
\item
  \texttt{box}, \texttt{lines}, and \texttt{polygons} to display
  discrete disparity results in respectively a boxplot, confidence
  interval lines, and confidence interval polygons.
\end{itemize}

\begin{quote}
This argument can be left empty. In this case, the algorithm will
automatically detect the type of subsamples from the \texttt{dispRity}
object and plot accordingly.
\end{quote}

It is also possible to display the number of elements in each subsample
(as a horizontal dotted line) using the option
\texttt{elements\ =\ TRUE}. Additionally, when the data is rarefied, one
can indicate which level of rarefaction to display (i.e.~only display
the results for a certain number of elements) by using the
\texttt{rarefaction} argument.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical parameters}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}

\NormalTok{## Plotting continuous disparity results}
\KeywordTok{plot}\NormalTok{(disparity_time_slices, }\DataTypeTok{type =} \StringTok{"continuous"}\NormalTok{)}

\NormalTok{## Plotting discrete disparity results}
\KeywordTok{plot}\NormalTok{(disparity_crown_stem, }\DataTypeTok{type =} \StringTok{"box"}\NormalTok{)}

\NormalTok{## As above but using lines for the rarefaction level of 20 elements only}
\KeywordTok{plot}\NormalTok{(disparity_crown_stem, }\DataTypeTok{type =} \StringTok{"line"}\NormalTok{, }\DataTypeTok{rarefaction =} \DecValTok{20}\NormalTok{)}

\NormalTok{## As above but using polygons while also displaying the number of elements}
\KeywordTok{plot}\NormalTok{(disparity_crown_stem, }\DataTypeTok{type =} \StringTok{"polygon"}\NormalTok{, }\DataTypeTok{elements =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{manual_files/figure-latex/unnamed-chunk-47-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Resetting graphical parameters}
\KeywordTok{par}\NormalTok{(op)}
\end{Highlighting}
\end{Shaded}

Since \texttt{plot.dispRity} uses the arguments from the generic
\texttt{plot} method, it is of course possible to change pretty much
everything using the regular plot arguments:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical options}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}

\NormalTok{## Plotting the results with some classic options from plot}
\KeywordTok{plot}\NormalTok{(disparity_time_slices, }\DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"blue"}\NormalTok{, }\StringTok{"orange"}\NormalTok{, }\StringTok{"green"}\NormalTok{),}
    \DataTypeTok{ylab =} \KeywordTok{c}\NormalTok{(}\StringTok{"Some measurement"}\NormalTok{), }\DataTypeTok{xlab =} \StringTok{"Some other measurement"}\NormalTok{,}
    \DataTypeTok{main =} \StringTok{"Many options..."}\NormalTok{, }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(}\DecValTok{4}\NormalTok{, }\DecValTok{0}\NormalTok{))}

\NormalTok{## Adding a legend}
\KeywordTok{legend}\NormalTok{(}\StringTok{"topleft"}\NormalTok{, }\DataTypeTok{legend =} \KeywordTok{c}\NormalTok{(}\StringTok{"Central tendency"}\NormalTok{,}
                             \StringTok{"Confidence interval 1"}\NormalTok{,}
                            \StringTok{"Confidence interval 2"}\NormalTok{),}
      \DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"blue"}\NormalTok{, }\StringTok{"orange"}\NormalTok{, }\StringTok{"green"}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{manual_files/figure-latex/unnamed-chunk-48-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Resetting graphical parameters}
\KeywordTok{par}\NormalTok{(op)}
\end{Highlighting}
\end{Shaded}

In addition to the classic \texttt{plot} arguments, the function can
also take arguments that are specific to \texttt{plot.dispRity} like
adding the number of elements or rarefaction level (as described above),
and also changing the values of the quantiles to plot as well as the
central tendency.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical options}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}

\NormalTok{## Plotting the results with some plot.dispRity arguments}
\KeywordTok{plot}\NormalTok{(disparity_time_slices, }\DataTypeTok{quantile =} \KeywordTok{c}\NormalTok{(}\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{10}\NormalTok{, }\DataTypeTok{to =} \DecValTok{100}\NormalTok{, }\DataTypeTok{by =} \DecValTok{10}\NormalTok{)),}
    \DataTypeTok{cent.tend =}\NormalTok{ sd, }\DataTypeTok{type =} \StringTok{"c"}\NormalTok{, }\DataTypeTok{elements =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"black"}\NormalTok{, }\KeywordTok{rainbow}\NormalTok{(}\DecValTok{10}\NormalTok{)),}
    \DataTypeTok{ylab =} \KeywordTok{c}\NormalTok{(}\StringTok{"Disparity"}\NormalTok{, }\StringTok{"Diversity"}\NormalTok{), }\DataTypeTok{time.subsamples =} \OtherTok{FALSE}\NormalTok{,}
    \DataTypeTok{xlab =} \StringTok{"Time (in in units from past to present)"}\NormalTok{, }\DataTypeTok{observed =} \OtherTok{TRUE}\NormalTok{,}
    \DataTypeTok{main =} \StringTok{"Many more options..."}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{manual_files/figure-latex/unnamed-chunk-49-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Resetting graphical parameters}
\KeywordTok{par}\NormalTok{(op)}
\end{Highlighting}
\end{Shaded}

\begin{quote}
Note that the argument \texttt{observed\ =\ TRUE} allows to plot the
disparity values calculated from the non-bootstrapped data as crosses on
the plot.
\end{quote}

For comparing results, it is also possible to add a plot to the existent
plot by using \texttt{add\ =\ TRUE}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical options}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}

\NormalTok{## Plotting the continuous disparity with a fixed y axis}
\KeywordTok{plot}\NormalTok{(disparity_time_slices, }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{4}\NormalTok{))}
\NormalTok{## Adding the discrete data}
\KeywordTok{plot}\NormalTok{(disparity_time_bins, }\DataTypeTok{type =} \StringTok{"line"}\NormalTok{, }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{4}\NormalTok{), }\DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{,}
    \DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{manual_files/figure-latex/unnamed-chunk-50-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Resetting graphical parameters}
\KeywordTok{par}\NormalTok{(op)}
\end{Highlighting}
\end{Shaded}

Finally, if your data has been fully rarefied, it is also possible to
easily look at rarefaction curves by using the
\texttt{rarefaction\ =\ TRUE} argument:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical options}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}

\NormalTok{## Plotting the rarefaction curves}
\KeywordTok{plot}\NormalTok{(disparity_crown_stem, }\DataTypeTok{rarefaction =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{manual_files/figure-latex/unnamed-chunk-51-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Resetting graphical parameters}
\KeywordTok{par}\NormalTok{(op)}
\end{Highlighting}
\end{Shaded}

\section{Testing disparity
hypotheses}\label{testing-disparity-hypotheses}

The \texttt{dispRity} package allows users to apply statistical tests to
the calculated disparity to test various hypotheses. The function
\texttt{test.dispRity} works in a similar way to the \texttt{dispRity}
function: it takes a \texttt{dispRity} object, a \texttt{test} and a
\texttt{comparisons} argument.

The \texttt{comparisons} argument indicates the way the test should be
applied to the data:

\begin{itemize}
\tightlist
\item
  \texttt{pairwise} (default): to compare each subsample in a pairwise
  manner
\item
  \texttt{referential}: to compare each subsample to the first subsample
\item
  \texttt{sequential}: to compare each subsample to the following
  subsample
\item
  \texttt{all}: to compare all the subsamples together (like in analysis
  of variance)
\end{itemize}

It is also possible to input a list of pairs of \texttt{numeric} values
or \texttt{characters} matching the subsample names to create
personalised tests. Some other tests implemented in \texttt{dispRity}
such as the \texttt{dispRity::null.test} have a specific way they are
applied to the data and therefore ignore the \texttt{comparisons}
argument.

The \texttt{test} argument can be any statistical or non-statistical
test to apply to the disparity object. It can be a common statistical
test function (e.g. \texttt{stats::t.test}), a function implemented in
\texttt{dispRity} (e.g.~see \texttt{?null.test}) or any function defined
by the user.

This function also allows users to correct for Type I error inflation
(false positives) when using multiple comparisons via the
\texttt{correction} argument. This argument can be empty (no correction
applied) or can contain one of the corrections from the
\texttt{stats::p.adjust} function (see \texttt{?p.adjust}).

Note that the \texttt{test.dispRity} algorithm deals with some classical
test outputs (\texttt{h.test}, \texttt{lm} and \texttt{numeric} vector)
and summarises the test output. It is, however, possible to get the full
detailed output by using the options \texttt{details\ =\ TRUE}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## T-test to test for a difference in disparity between crown and stem mammals}
\KeywordTok{test.dispRity}\NormalTok{(disparity_crown_stem, }\DataTypeTok{test =}\NormalTok{ t.test)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [[1]]
##              statistic
## crown : stem  55.60162
## 
## [[2]]
##              parameter
## crown : stem  164.6932
## 
## [[3]]
##                    p.value
## crown : stem 1.201122e-108
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Performing the same test but with the detailed t.test output}
\KeywordTok{test.dispRity}\NormalTok{(disparity_crown_stem, }\DataTypeTok{test =}\NormalTok{ t.test, }\DataTypeTok{details =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## $`crown : stem`
## $`crown : stem`[[1]]
## 
##  Welch Two Sample t-test
## 
## data:  dots[[1L]][[1L]] and dots[[2L]][[1L]]
## t = 55.602, df = 164.69, p-value < 2.2e-16
## alternative hypothesis: true difference in means is not equal to 0
## 95 percent confidence interval:
##  0.2849688 0.3059530
## sample estimates:
## mean of x mean of y 
##  1.926163  1.630702
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Wilcoxon test applied to time sliced disparity with sequential comparisons, with Bonferroni correction}
\KeywordTok{test.dispRity}\NormalTok{(disparity_time_slices, }\DataTypeTok{test =}\NormalTok{ wilcox.test,}
              \DataTypeTok{comparisons =} \StringTok{"sequential"}\NormalTok{, }\DataTypeTok{correction =} \StringTok{"bonferroni"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [[1]]
##          statistic
## 120 : 80        13
## 80 : 40       2497
## 40 : 0         426
## 
## [[2]]
##               p.value
## 120 : 80 1.130227e-33
## 80 : 40  2.904335e-09
## 40 : 0   1.624453e-28
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Measuring the overlap between distributions in the time bins (using the}
\NormalTok{## implemented Bhattacharyya Coefficient function - see ?bhatt.coeff)}
\KeywordTok{test.dispRity}\NormalTok{(disparity_time_bins, }\DataTypeTok{test =}\NormalTok{ bhatt.coeff)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                    bhatt.coeff
## 120 - 80 : 80 - 40   0.0000000
## 120 - 80 : 40 - 0    0.0000000
## 80 - 40 : 40 - 0     0.4195376
\end{verbatim}

It is also possible to apply some more \emph{complex} tests that have
their own output classes (like \texttt{stats::lm}).

The results can then be analysed as usual using the associated
\texttt{summary} S3 method:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Performing and linear model applied to the same data}
\NormalTok{(slice_lm <-}\StringTok{ }\KeywordTok{test.dispRity}\NormalTok{(disparity_time_slices, }\DataTypeTok{test =}\NormalTok{ lm,}
                            \DataTypeTok{comparisons =} \StringTok{"all"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## test(formula = data ~ subsamples, data = data)
## 
## Coefficients:
##   (Intercept)  subsamples120   subsamples40   subsamples80  
##        3.6571        -1.4046        -0.4481        -0.5863
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The output is a regular `lm` output}
\KeywordTok{class}\NormalTok{(slice_lm)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] "lm"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## This output can be summarised using summary}
\KeywordTok{summary}\NormalTok{(slice_lm)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## test(formula = data ~ subsamples, data = data)
## 
## Residuals:
##      Min       1Q   Median       3Q      Max 
## -1.26522 -0.10976  0.01515  0.13109  0.58161 
## 
## Coefficients:
##               Estimate Std. Error t value Pr(>|t|)    
## (Intercept)    3.65712    0.02406  151.99   <2e-16 ***
## subsamples120 -1.40459    0.03403  -41.28   <2e-16 ***
## subsamples40  -0.44812    0.03403  -13.17   <2e-16 ***
## subsamples80  -0.58634    0.03403  -17.23   <2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 0.2406 on 396 degrees of freedom
## Multiple R-squared:  0.818,  Adjusted R-squared:  0.8166 
## F-statistic: 593.1 on 3 and 396 DF,  p-value: < 2.2e-16
\end{verbatim}

Of course, due to the modular design of the package, tests can always be
made by the user (the same way disparity metrics can be user made). The
only condition is that the test can be applied to at least two
distributions. In practice, the \texttt{test.dispRity} function will
pass the calculated disparity data (distributions) to the provided
function in either pairs of distributions (if the \texttt{comparisons}
argument is set to \texttt{pairwise}, \texttt{referential} or
\texttt{sequential}) or a table containing all the distributions
(\texttt{comparisons\ =\ all}; this should be in the same format as data
passed to \texttt{lm} for example).

\hypertarget{disparity-as-a-distribution}{\section{Disparity as a
distribution}\label{disparity-as-a-distribution}}

Disparity is often regarded as a summary value of the position of the
all elements in the ordinated space. For example, the sum of variances,
the product of ranges or the median distance between the elements and
their centroid will summarise disparity as a single value. This value
can be pseudo-replicated (bootstrapped) to obtain a distribution of the
summary metric with estimated error. However, another way to perform
disparity analysis is to use the \emph{whole distribution} rather than
just a summary metric (e.g.~the variances or the ranges).

This is possible in the \texttt{dispRity} package by calculating
disparity as a dimension-level 2 metric only! Let's have a look using
our previous example of bootstrapped time slices but by measuring the
distances between each taxon and their centroid as disparity.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Measuring disparity as a whole distribution}
\NormalTok{disparity_centroids <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(boot_time_slices, }\DataTypeTok{metric =}\NormalTok{ centroids)}
\end{Highlighting}
\end{Shaded}

The resulting disparity object is of dimension-level 2, so it can easily
be transformed into a dimension-level 1 object by, for example,
measuring the median distance of all these distributions:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Measuring median disparity in each time slice}
\NormalTok{disparity_centroids_median <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(disparity_centroids, }\DataTypeTok{metric =}\NormalTok{ median)}
\end{Highlighting}
\end{Shaded}

And we can now compare the differences between these methods:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Summarising both disparity measurements:}
\NormalTok{## The distributions:}
\KeywordTok{summary}\NormalTok{(disparity_centroids)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n obs.median bs.median  2.5%   25%   75% 97.5%
## 1        120  5      1.508     1.324 0.536 1.069 1.562 1.868
## 2         80 19      1.790     1.702 1.393 1.597 1.815 1.943
## 3         40 15      1.719     1.703 1.387 1.581 1.857 2.076
## 4          0 10      1.910     1.807 1.366 1.700 1.971 2.085
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The summary of the distributions (as median)}
\KeywordTok{summary}\NormalTok{(disparity_centroids_median)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs bs.median  2.5%   25%   75% 97.5%
## 1        120  5 1.508     1.354 0.508 0.915 1.481 1.615
## 2         80 19 1.790     1.703 1.581 1.664 1.738 1.794
## 3         40 15 1.719     1.706 1.545 1.646 1.761 1.832
## 4          0 10 1.910     1.809 1.594 1.784 1.857 1.940
\end{verbatim}

We can see that the summary message for the distribution is slightly
different than before. Here \texttt{summary} also displays the observed
central tendency (i.e.~the central tendency of the measured
distributions). Note that, as expected, this central tendency is the
same in both metrics!

Another, maybe more intuitive way, to compare both approaches for
measuring disparity is to plot the distributions:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical parameters}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{, }\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{))}

\NormalTok{## Plotting both disparity measurements}
\KeywordTok{plot}\NormalTok{(disparity_centroids, }\DataTypeTok{ylab =} \StringTok{"Distribution of all the distances"}\NormalTok{)}
\KeywordTok{plot}\NormalTok{(disparity_centroids_median,}
     \DataTypeTok{ylab =} \StringTok{"Distribution of the medians of all the distances"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{manual_files/figure-latex/unnamed-chunk-57-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{par}\NormalTok{(op)}
\end{Highlighting}
\end{Shaded}

We can then test for differences in the resulting distributions using
\texttt{test.dispRity} and the \texttt{bhatt.coeff} test as described
above.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Probability of overlap in the distribution of medians}
\KeywordTok{test.dispRity}\NormalTok{(disparity_centroids_median, }\DataTypeTok{test =}\NormalTok{ bhatt.coeff)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##          bhatt.coeff
## 120 : 80  0.15380832
## 120 : 40  0.23690683
## 120 : 0   0.09196152
## 80 : 40   0.92367172
## 80 : 0    0.46576289
## 40 : 0    0.65992821
\end{verbatim}

In this case, we are looking at the probability of overlap of the
distribution of median distances from centroids among each pair of time
slices. In other words, we are measuring whether the medians from each
bootstrap pseudo-replicate for each time slice overlap. But of course,
we might be interested in the actual distribution of the distances from
the centroid rather than simply their central tendencies. This can be
problematic depending on the research question asked since we are
effectively comparing non-independent medians distributions (because of
the pseudo-replication).

One solution, therefore, is to look at the full distribution:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Probability of overlap for the full distributions}
\KeywordTok{test.dispRity}\NormalTok{(disparity_centroids, }\DataTypeTok{test =}\NormalTok{ bhatt.coeff)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##          bhatt.coeff
## 120 : 80   0.6401042
## 120 : 40   0.6402923
## 120 : 0    0.5787247
## 80 : 40    0.9423481
## 80 : 0     0.8444910
## 40 : 0     0.9392977
\end{verbatim}

These results show the actual overlap among all the measured distances
from centroids concatenated across all the bootstraps. For example, when
comparing the slices 120 and 80, we are effectively comparing the 5
\(\times\) 100 distances (the distances of the five elements in slice
120 bootstrapped 100 times) to the 19 \(\times\) 100 distances from
slice 80. However, this can also be problematic for some specific tests
since the \emph{n} \(\times\) 100 distances are also pseudo-replicates
and thus are still not independent.

A second solution is to compare the distributions to each other
\emph{for each replicate}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Boostrapped probability of overlap for the full distributions}
\KeywordTok{test.dispRity}\NormalTok{(disparity_centroids, }\DataTypeTok{test =}\NormalTok{ bhatt.coeff, }\DataTypeTok{concatenate =} \OtherTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##          bhatt.coeff      2.5%       25%       75%     97.5%
## 120 : 80   0.2534144 0.0000000 0.1450953 0.3554093 0.5615906
## 120 : 40   0.2726082 0.0000000 0.1632993 0.3942394 0.6435771
## 120 : 0    0.2172550 0.0000000 0.0000000 0.3464102 0.5913591
## 80 : 40    0.6016956 0.3097953 0.4779726 0.7208207 0.8306227
## 80 : 0     0.4770117 0.1382736 0.3964076 0.5772107 0.7169318
## 40 : 0     0.5327868 0.1724549 0.3887699 0.6761873 0.8781747
\end{verbatim}

These results show the median overlap among pairs of distributions in
the first column (\texttt{bhatt.coeff}) and then the distribution of
these overlaps among each pair of bootstraps. In other words, when two
distributions are compared, they are now compared for each bootstrap
pseudo-replicate, thus effectively creating a distribution of
probabilities of overlap. For example, when comparing the slices 120 and
80, we have a mean probability of overlap of 0.28 and a probability
between 0.18 and 0.43 in 50\% of the pseudo-replicates. Note that the
quantiles and central tendencies can be modified via the
\texttt{conc.quantiles} option.

\section{Ecological studies of
disparity}\label{ecological-studies-of-disparity}

Disparity analyses are not restricted to palaeobiology analysis. This
package is designed for dealing with ordinated matrices, no matter which
data was used in the ordination. Here is an short example of disparity
analysis using ecological data from Deirdre McClean (unpubl.).

The dataset contains an ordinated matrix of 20 dimensions (columns) for
40 elements (rows). The elements are different experimental plots with
discrete variations of nutrient enrichment, depth, and freshwater
benthic invertebrates.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Loading the data}
\KeywordTok{data}\NormalTok{(McClean_data)}

\NormalTok{## Ordinated matrix}
\NormalTok{ordinated_matrix <-}\StringTok{ }\NormalTok{McClean_data}\OperatorTok{$}\NormalTok{ordination}

\NormalTok{## Treatments and depths}
\NormalTok{treatments <-}\StringTok{ }\NormalTok{McClean_data}\OperatorTok{$}\NormalTok{treatment}
\NormalTok{depth <-}\StringTok{ }\NormalTok{McClean_data}\OperatorTok{$}\NormalTok{depth}
\end{Highlighting}
\end{Shaded}

\subsection{\texorpdfstring{Classic ecological ``disparity''
analysis}{Classic ecological disparity analysis}}\label{classic-ecological-disparity-analysis}

A classical way to represent this ordinated data would be to use two
dimensional plots to look at how the different experimental plots differ
in their values for nutrient enrichment, depth, and freshwater benthic
invertebrates.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Setting the colors that will represent the treatments}
\NormalTok{cols <-}\StringTok{ }\KeywordTok{sub}\NormalTok{(}\StringTok{"a"}\NormalTok{, }\StringTok{"red"}\NormalTok{, treatments)}
\NormalTok{cols <-}\StringTok{ }\KeywordTok{sub}\NormalTok{(}\StringTok{"b"}\NormalTok{, }\StringTok{"blue"}\NormalTok{, cols)}
\NormalTok{## Setting the symbols that will represent the depth}
\NormalTok{pchs <-}\StringTok{ }\KeywordTok{sub}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{16}\NormalTok{, depth)}
\NormalTok{pchs <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(}\KeywordTok{sub}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{17}\NormalTok{, pchs))}

\NormalTok{## Graphical option}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}
\NormalTok{## A classic 2D ordination plot}
\KeywordTok{plot}\NormalTok{(ordinated_matrix[, }\DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{], }\DataTypeTok{col =}\NormalTok{ cols, }\DataTypeTok{pch =}\NormalTok{ pchs,}
     \DataTypeTok{xlab =} \StringTok{"PC 1"}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{"PC 2"}\NormalTok{,}
     \DataTypeTok{xlim =} \KeywordTok{range}\NormalTok{(ordinated_matrix[, }\DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{]) }\OperatorTok{+}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{100}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\includegraphics{manual_files/figure-latex/unnamed-chunk-62-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{par}\NormalTok{(op)}
\end{Highlighting}
\end{Shaded}

This shows the distribution of the experimental plots along the two
first axis of variation of the ordinated distance matrix (i.e.~the first
two dimensions). At a first glance, it seems difficult to see a clear
effect of the treatments (blue or red) or the depth (rounds or
triangles) as the different experimental plots with the same parameters
don't seem to cluster together. However, this may be because this plot
ignores the 18 other dimensions of the ordination! Additionally, these
two represented dimensions do not represent a biological reality
\emph{per se}; i.e.~the values on the first dimension do not represent a
continuous trait (e.g.~depth), instead they just represent the
ordinations of correlations between the data and some factors.

Therefore, we might want to approach this problem without getting stuck
in only two dimensions and consider the whole dataset as a
\emph{n}-dimensional object.

\subsection{\texorpdfstring{Multidimensional analysis with
\texttt{dispRity}}{Multidimensional analysis with dispRity}}\label{multidimensional-analysis-with-disprity}

\subsubsection{Splitting the data with respect to the different
factors}\label{splitting-the-data-with-respect-to-the-different-factors}

The first step is to create different subsamples that represent
subsamples of the ordinated space (i.e.~sub-regions within the
\emph{n}-dimensional object). Each of these subsamples will contain a
certain number of elements (i.e.~a subset of the 40 experimental field
plots) that have some attributes in common. In our example, we are going
to group the elements according to their depth and treatment.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating a table that contain the elements and their attributes}
\NormalTok{factors <-}\StringTok{ }\KeywordTok{as.data.frame}\NormalTok{(}\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \KeywordTok{c}\NormalTok{(treatments, depth),}
          \DataTypeTok{nrow =} \KeywordTok{nrow}\NormalTok{(ordinated_matrix), }\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{, }\DataTypeTok{byrow =} \OtherTok{FALSE}\NormalTok{,}
          \DataTypeTok{dimnames =} \KeywordTok{list}\NormalTok{(}\KeywordTok{rownames}\NormalTok{(ordinated_matrix))))}
\KeywordTok{names}\NormalTok{(factors) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Treat"}\NormalTok{, }\StringTok{"Depth"}\NormalTok{)}
\KeywordTok{head}\NormalTok{(factors)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##     Treat Depth
## 1       a     1
## 1.1     a     2
## 2       b     1
## 2.1     b     2
## 3       a     1
## 3.1     a     1
\end{verbatim}

Second, let's split the data according by depth and treatment to create
the subsamples of the ordinated space by using the
\texttt{custom.subsamples} function:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Splitting the ordinated space into four subsamples}
\NormalTok{(customised_subsamples <-}\StringTok{ }\KeywordTok{custom.subsamples}\NormalTok{(ordinated_matrix, factors))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 4 customised subsamples for 40 elements:
##     Treat.a, Treat.b, Depth.1, Depth.2.
\end{verbatim}

\subsubsection{Calculating disparity}\label{calculating-disparity}

In this example, we are going to define disparity as the n-dimensional
ellipsoid volume (see \texttt{?hyper.volume}).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating disparity as the ellipsoid volume of each group}
\NormalTok{disparity <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(customised_subsamples, }\DataTypeTok{metric =}\NormalTok{ diagonal)}
\KeywordTok{summary}\NormalTok{(disparity)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs
## 1    Treat.a 21 36.89
## 2    Treat.b 19 40.99
## 3    Depth.1 23 36.38
## 4    Depth.2 17 40.74
\end{verbatim}

We can also bootstrap the data to test the robustness of the measured
disparity to outliers. Also, as we can see, each subsample has different
numbers of elements. It might also be interesting to rarefy the data to
have only subsamples with the same number of elements.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Bootstrapping the subsamples and using rarefaction}
\NormalTok{## (i.e. only re-sampling 17 elements each time)}
\NormalTok{bootstrapped_data <-}\StringTok{ }\KeywordTok{boot.matrix}\NormalTok{(customised_subsamples, }\DataTypeTok{bootstraps =} \DecValTok{100}\NormalTok{,}
                                 \DataTypeTok{rarefaction =} \DecValTok{17}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

We can now run a more robust disparity analysis using the bootstrapped
data:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating the bootstrapped disparity}
\NormalTok{disparity <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(bootstrapped_data, }\DataTypeTok{metric =}\NormalTok{ diagonal)}
\end{Highlighting}
\end{Shaded}

\subsubsection{Summarising and displaying the
results}\label{summarising-and-displaying-the-results}

Now let's summarise this data:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Summary table of the data}
\KeywordTok{summary}\NormalTok{(disparity)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs bs.median  2.5%   25%   75% 97.5%
## 1    Treat.a 21 36.89     35.01 31.32 32.88 35.44 36.11
## 2    Treat.a 17    NA     33.69 30.38 31.91 34.97 35.98
## 3    Treat.b 19 40.99     38.19 34.27 36.75 39.25 40.12
## 4    Treat.b 17    NA     37.52 33.38 36.08 38.85 39.80
## 5    Depth.1 23 36.38     34.44 31.86 33.63 35.01 35.52
## 6    Depth.1 17    NA     33.38 30.56 32.28 34.09 35.18
## 7    Depth.2 17 40.74     37.30 34.26 36.35 38.62 39.74
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical options}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =}\NormalTok{ (}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{)), }\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}

\NormalTok{## Plotting the bootstrapped disparity}
\KeywordTok{plot}\NormalTok{(disparity, }\DataTypeTok{main =} \StringTok{"Bootstrapped"}\NormalTok{, }\DataTypeTok{las =} \DecValTok{2}\NormalTok{, }\DataTypeTok{xlab =} \StringTok{""}\NormalTok{)}

\NormalTok{## Plotting the rarefied disparity}
\KeywordTok{plot}\NormalTok{(disparity, }\DataTypeTok{rarefaction =} \DecValTok{17}\NormalTok{, }\DataTypeTok{main =} \StringTok{"Rarefied"}\NormalTok{, }\DataTypeTok{las =} \DecValTok{2}\NormalTok{, }\DataTypeTok{xlab =} \StringTok{""}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{manual_files/figure-latex/unnamed-chunk-68-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{par}\NormalTok{(op)}
\end{Highlighting}
\end{Shaded}

As we can see, there seems to be no strong effect of the number of
experimental plots in each subsample (i.e.~the rarefied plot is really
similar to the bootstrapped plot) which is a good thing!

\subsubsection{Testing our hypotheses}\label{testing-our-hypotheses}

Finally, we can test our hypothesis (whether the water treatment at
certain depths alters invertebrate communities and composition in
natural habitats) by using the \texttt{test.dispRity} function.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Testing the effect of treatment and depth on the bootstrapped data}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{test.dispRity}\NormalTok{(disparity, }\DataTypeTok{test =}\NormalTok{ aov, }\DataTypeTok{comparisons =} \StringTok{"all"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##      Length Class Mode
## [1,] 13     aov   list
\end{verbatim}

\section{Simulating discrete morphological
data}\label{simulating-discrete-morphological-data}

The \texttt{dispRity} package also allows users to simulate discrete
morphological data matrices. In brief, the function \texttt{sim.morpho}
takes a phylogenetic tree, the number of required characters, the
evolutionary model, and a function from which to draw the rates. The
package also contains a function for quickly checking the matrix's
phylogenetic signal (as defined in systematics not phylogenetic
comparative methods) using parsimony. To understand these methods please
refer to the phylogenetics literature.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{3}\NormalTok{)}
\NormalTok{## Simulating a starting tree with 15 taxa as a random coalescent tree}
\NormalTok{my_tree <-}\StringTok{ }\KeywordTok{rcoal}\NormalTok{(}\DecValTok{15}\NormalTok{)}

\NormalTok{## Generating a matrix with 100 characters (85% binary and 15% three state) and}
\NormalTok{## an equal rates model with a gamma rate distribution (0.5, 1) with no }
\NormalTok{## invariant characters.}
\NormalTok{my_matrix <-}\StringTok{ }\KeywordTok{sim.morpho}\NormalTok{(}\DataTypeTok{tree =}\NormalTok{ my_tree, }\DataTypeTok{characters =} \DecValTok{100}\NormalTok{, }\DataTypeTok{states =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.85}\NormalTok{,}
    \FloatTok{0.15}\NormalTok{), }\DataTypeTok{rates =} \KeywordTok{c}\NormalTok{(rgamma, }\FloatTok{0.5}\NormalTok{, }\DecValTok{1}\NormalTok{), }\DataTypeTok{invariant =} \OtherTok{FALSE}\NormalTok{)}

\NormalTok{## The first few lines of the matrix}
\NormalTok{my_matrix[}\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{, }\DecValTok{1}\OperatorTok{:}\DecValTok{10}\NormalTok{]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##     [,1] [,2] [,3] [,4] [,5] [,6] [,7] [,8] [,9] [,10]
## t15 "1"  "1"  "0"  "1"  "1"  "2"  "1"  "1"  "0"  "0"  
## t12 "1"  "1"  "0"  "1"  "1"  "2"  "1"  "1"  "0"  "0"  
## t14 "1"  "1"  "0"  "1"  "1"  "2"  "1"  "1"  "0"  "0"  
## t6  "1"  "1"  "0"  "1"  "1"  "2"  "1"  "1"  "0"  "0"  
## t3  "0"  "0"  "1"  "0"  "1"  "0"  "0"  "2"  "1"  "0"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Checking the matrix properties with a quick Maximum Parsimony tree search}
\KeywordTok{check.morpho}\NormalTok{(my_matrix, my_tree)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                                     
## Maximum parsimony        139.0000000
## Consistency index          0.7625899
## Retention index            0.8881356
## Robinson-Foulds distance   0.0000000
\end{verbatim}

Note that this example produces a tree with a great consistency index
and an identical topology to the random coalescent tree! Nearly too good
to be true\ldots{}

\subsection{A more detailed
description}\label{a-more-detailed-description}

\texttt{sim.morpho} is really flexible and takes many different
arguments to allow users to simulate realistic matrices. It has three
implemented models: \texttt{"ER"} for Equal Rates (the M\emph{k} model);
\texttt{"HKY"}, the molecular HKY model but transforms pyrines in 0's
and pyrimidines in 1's; or the \texttt{"mixed"} model that randomly uses
an \texttt{"ER"} and/or \texttt{"HKY"} for the binary characters and
\texttt{"ER"} for the multistate (\textgreater{}2) characters. Both
models take specific distributions for their rate or substitution
models. These distributions should be passed to these arguments in the
format of \texttt{c(sampler\_function,\ distribution\_parameters)} where
the the sampler function is a the random generation function of that
distribution (e.g. \texttt{rnorm}, \texttt{runif}, etc\ldots{}) and the
parameters are any parameters to be passed to this function.

\texttt{check.morpho} runs a quick Maximum Parsimony tree search using
the \texttt{phangorn} parsimony algorithm. It quickly calculates the
parsimony score, the consistency and retention indices and, if a tree is
provided (e.g.~the tree used to generate the matrix) it calculates the
Robinson-Foulds distance between the most parsimonious tree and the
provided tree to determine how different they are.

\subsection{Parameters for a realistic(ish)
matrix}\label{parameters-for-a-realisticish-matrix}

There are many parameters that can create a
\texttt{realistic\textquotesingle{}\textquotesingle{}\ matrix\ (i.e.\ not\ too\ different\ from\ the\ input\ tree\ with\ a\ consistency\ and\ retention\ index\ close\ to\ what\ is\ seen\ in\ the\ literature)\ but\ because\ of\ the\ randomness\ of\ the\ matrix\ generation\ not\ all\ end\ up\ creating\ "good"\ matrices.\ The\ following\ parameters\ however,\ seem\ to\ generate\ fairly}realistic''
matrices with a starting coalescent tree, equal rates model with 0.85
binary characters and 0.15 three state characters, a gamma distribution
with a shape parameter (\(\alpha\)) of 5 and no scaling (\(\beta\) = 1)
with a rate of 100.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{0}\NormalTok{)}
\NormalTok{## tree}
\NormalTok{my_tree <-}\StringTok{ }\KeywordTok{rcoal}\NormalTok{(}\DecValTok{15}\NormalTok{)}
\NormalTok{## matrix}
\NormalTok{morpho_mat <-}\StringTok{ }\KeywordTok{sim.morpho}\NormalTok{(my_tree, }\DataTypeTok{characters =} \DecValTok{100}\NormalTok{, }\DataTypeTok{model =} \StringTok{"ER"}\NormalTok{,}
    \DataTypeTok{rates =} \KeywordTok{c}\NormalTok{(rgamma, }\DataTypeTok{rate =} \DecValTok{100}\NormalTok{, }\DataTypeTok{shape =} \DecValTok{5}\NormalTok{), }\DataTypeTok{invariant =} \OtherTok{FALSE}\NormalTok{)}
\KeywordTok{check.morpho}\NormalTok{(morpho_mat, my_tree)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                                     
## Maximum parsimony        104.0000000
## Consistency index          0.0000000
## Retention index            0.7886179
## Robinson-Foulds distance   0.0000000
\end{verbatim}

\hypertarget{space.maker}{\subsection{\texorpdfstring{\texttt{space.maker}}{space.maker}}\label{space.maker}}

Another way to simulate data is to directly simulate an ordinated space
with the \texttt{space.maker} function. This function allows users to
simulate multidimensional spaces with a certain number of properties. It
takes as arguments the number of elements (data points)~and dimensions
to create the space but also permits more fine tuning in the data
simulation: it is possible to give a specific distribution to each
dimension, provide a correlation matrix to link the dimensions or even
provide an \emph{a priori} distribution of the variance for each
distributions!

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical options}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =}\NormalTok{ (}\KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{)), }\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}
\NormalTok{## Visualising 3D spaces}
\KeywordTok{require}\NormalTok{(scatterplot3d)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Loading required package: scatterplot3d
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## A cube space}
\KeywordTok{scatterplot3d}\NormalTok{(}\KeywordTok{space.maker}\NormalTok{(}\DecValTok{2500}\NormalTok{, }\DecValTok{3}\NormalTok{, runif), }\DataTypeTok{pch =} \DecValTok{20}\NormalTok{,}
              \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{zlab =} \StringTok{""}\NormalTok{,}
              \DataTypeTok{main =} \StringTok{"Uniform cube space"}\NormalTok{)}

\NormalTok{## A plane space}
\KeywordTok{scatterplot3d}\NormalTok{(}\KeywordTok{space.maker}\NormalTok{(}\DecValTok{2500}\NormalTok{, }\DecValTok{3}\NormalTok{, }\KeywordTok{c}\NormalTok{(runif, runif, runif),}
              \DataTypeTok{arguments =} \KeywordTok{list}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{0}\NormalTok{), }\OtherTok{NULL}\NormalTok{, }\OtherTok{NULL}\NormalTok{)), }\DataTypeTok{pch =} \DecValTok{20}\NormalTok{,}
              \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{zlab =} \StringTok{""}\NormalTok{,}
              \DataTypeTok{main =} \StringTok{"Uniform plane space"}\NormalTok{)}

\NormalTok{## An ellipsoid space (=a spheric space with correlation)}
\NormalTok{cor_matrix <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\KeywordTok{cbind}\NormalTok{(}\DecValTok{1}\NormalTok{, }\FloatTok{0.8}\NormalTok{, }\FloatTok{0.2}\NormalTok{, }\FloatTok{0.8}\NormalTok{, }\DecValTok{1}\NormalTok{, }\FloatTok{0.7}\NormalTok{, }\FloatTok{0.2}\NormalTok{, }\FloatTok{0.7}\NormalTok{, }\DecValTok{1}\NormalTok{), }\DataTypeTok{nrow =} \DecValTok{3}\NormalTok{)}
\KeywordTok{scatterplot3d}\NormalTok{(}\KeywordTok{space.maker}\NormalTok{(}\DecValTok{2500}\NormalTok{, }\DecValTok{3}\NormalTok{, rnorm, }\DataTypeTok{cor.matrix =}\NormalTok{ cor_matrix), }\DataTypeTok{pch =} \DecValTok{20}\NormalTok{,}
              \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{zlab =} \StringTok{""}\NormalTok{,}
              \DataTypeTok{main =} \StringTok{"Normal ellipsoid space"}\NormalTok{)}

\NormalTok{## A cylindrical space with a decreasing variance per axis}
\KeywordTok{scatterplot3d}\NormalTok{(}\KeywordTok{space.maker}\NormalTok{(}\DecValTok{2500}\NormalTok{, }\DecValTok{3}\NormalTok{, }\KeywordTok{c}\NormalTok{(rnorm, rnorm, runif),}
              \DataTypeTok{scree =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.7}\NormalTok{, }\FloatTok{0.2}\NormalTok{, }\FloatTok{0.1}\NormalTok{)), }\DataTypeTok{pch =} \DecValTok{20}\NormalTok{,}
              \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{zlab =} \StringTok{""}\NormalTok{,}
              \DataTypeTok{main =} \StringTok{"Normal cylindrical space"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{manual_files/figure-latex/unnamed-chunk-72-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Resetting the graphic parameters}
\KeywordTok{par}\NormalTok{(op)}
\end{Highlighting}
\end{Shaded}

\chapter{The guts of the dispRity
package}\label{the-guts-of-the-disprity-package}

\section{\texorpdfstring{Manipulating \texttt{dispRity}
objects}{Manipulating dispRity objects}}\label{manipulating-disprity-objects}

Disparity analysis involves a lot of manipulation of many matrices
(especially when bootstrapping) which can be impractical to visualise
and will quickly overwhelm your \texttt{R} console. Even the simple Beck
and Lee 2014 example above produces an object with \textgreater{} 72
lines of lists of lists of matrices!

Therefore \texttt{dispRity} uses a specific class of object called a
\texttt{dispRity} object. These objects allow users to use S3 method
functions such as \texttt{summary.dispRity}, \texttt{plot.dispRity} and
\texttt{print.dispRity}. \texttt{dispRity} also contains various utility
functions that manipulate the \texttt{dispRity} object (e.g.
\texttt{sort.dispRity}, \texttt{extract.dispRity} see the full list in
the next section). These functions modify the \texttt{dispRity} object
without having to delve into its complex structure! The full structure
of a \texttt{dispRity} object is detailed
\href{https://github.com/TGuillerme/dispRity/blob/master/disparity_object.md}{here}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Loading the example data}
\KeywordTok{data}\NormalTok{(disparity)}

\NormalTok{## What is the class of the median_centroids object?}
\KeywordTok{class}\NormalTok{(disparity)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] "dispRity"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## What does the object contain?}
\KeywordTok{names}\NormalTok{(disparity)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] "matrix"     "call"       "subsamples" "disparity"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Summarising it using the S3 method print.dispRity}
\NormalTok{disparity}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 7 continuous (acctran) time subsamples for 99 elements with 97 dimensions:
##      90, 80, 70, 60, 50 ...
## Data was bootstrapped 100 times (method:"full") and rarefied to 20, 15, 10, 5 elements.
## Disparity was calculated as: c(median, centroids).
\end{verbatim}

Note that it is always possible to recall the full object using the
argument \texttt{all\ =\ TRUE} in \texttt{print.dispRity}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Display the full object}
\KeywordTok{print}\NormalTok{(disparity, }\DataTypeTok{all =} \OtherTok{TRUE}\NormalTok{)}
\NormalTok{## This is more nearly ~ 5000 lines on my 13 inch laptop screen!}
\end{Highlighting}
\end{Shaded}

\section{\texorpdfstring{\texttt{dispRity}
utilities}{dispRity utilities}}\label{disprity-utilities}

The package also provides some utility functions to facilitate
multidimensional analysis.

\subsection{\texorpdfstring{\texttt{dispRity} object utilities
}{dispRity object utilities }}\label{disprity-object-utilities}

The first set of utilities are functions for manipulating
\texttt{dispRity} objects:

\subsubsection{\texorpdfstring{\texttt{make.dispRity}}{make.dispRity}}\label{make.disprity}

This function creates empty \texttt{dispRity} objects.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating an empty dispRity object}
\KeywordTok{make.dispRity}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Empty dispRity object.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating an "empty" dispRity object with a matrix}
\NormalTok{(disparity_obj <-}\StringTok{ }\KeywordTok{make.dispRity}\NormalTok{(}\KeywordTok{matrix}\NormalTok{(}\KeywordTok{rnorm}\NormalTok{(}\DecValTok{20}\NormalTok{), }\DecValTok{5}\NormalTok{, }\DecValTok{4}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Empty dispRity object.
\end{verbatim}

\subsubsection{\texorpdfstring{\texttt{fill.dispRity}}{fill.dispRity}}\label{fill.disprity}

This function initialises a \texttt{dispRity} object and generates its
call properties.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The dispRity object's call is indeed empty}
\NormalTok{disparity_obj}\OperatorTok{$}\NormalTok{call}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## list()
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Filling an empty disparity object (that needs to contain at least a matrix)}
\NormalTok{(disparity_obj <-}\StringTok{ }\KeywordTok{fill.dispRity}\NormalTok{(disparity_obj))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 5 elements with 4 dimensions.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The dipRity object has now the correct minimal attributes}
\NormalTok{disparity_obj}\OperatorTok{$}\NormalTok{call}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## $dimensions
## [1] 4
\end{verbatim}

\subsubsection{\texorpdfstring{\texttt{matrix.dispRity}}{matrix.dispRity}}\label{matrix.disprity}

This function extracts a specific matrix from a disparity object. The
matrix can be one of the bootstrapped matrices or/and a rarefied matrix.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Extracting the matrix containing the coordinates of the elements at time 50}
\KeywordTok{str}\NormalTok{(}\KeywordTok{matrix.dispRity}\NormalTok{(disparity, }\StringTok{"50"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  num [1:18, 1:97] -0.1038 0.2844 0.2848 0.0927 0.1619 ...
##  - attr(*, "dimnames")=List of 2
##   ..$ : chr [1:18] "Leptictis" "Dasypodidae" "n24" "Potamogalinae" ...
##   ..$ : NULL
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Extracting the 3rd bootstrapped matrix with the 2nd rarefaction level}
\NormalTok{## (15 elements) from the second group (80 Mya)}
\KeywordTok{str}\NormalTok{(}\KeywordTok{matrix.dispRity}\NormalTok{(disparity, }\DataTypeTok{subsamples =} \DecValTok{1}\NormalTok{, }\DataTypeTok{bootstrap =} \DecValTok{3}\NormalTok{, }\DataTypeTok{rarefaction =} \DecValTok{2}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  num [1:15, 1:97] -0.7161 0.3496 -0.573 -0.0445 -0.1427 ...
##  - attr(*, "dimnames")=List of 2
##   ..$ : chr [1:15] "n7" "n34" "Maelestes" "n20" ...
##   ..$ : NULL
\end{verbatim}

\subsubsection{\texorpdfstring{\texttt{get.subsamples.dispRity}}{get.subsamples.dispRity}}\label{get.subsamples.disprity}

This function creates a dispRity object that contains only elements from
one specific subsamples.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Extracting all the data for the crown mammals}
\NormalTok{(crown_mammals <-}\StringTok{ }\KeywordTok{get.subsamples.dispRity}\NormalTok{(disp_crown_stemBS, }\StringTok{"Group.crown"}\NormalTok{))}

\NormalTok{## The object keeps the properties of the parent object but is composed of only one subsamples}
\KeywordTok{length}\NormalTok{(crown_mammals}\OperatorTok{$}\NormalTok{subsamples)}
\end{Highlighting}
\end{Shaded}

\subsubsection{\texorpdfstring{\texttt{extract.dispRity}}{extract.dispRity}}\label{extract.disprity}

This function extracts the calculated disparity values of a specific
matrix.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Extracting the observed disparity (default)}
\KeywordTok{extract.dispRity}\NormalTok{(disparity)}

\NormalTok{## Extracting the disparity from the bootstrapped values from the}
\NormalTok{## 10th rarefaction level from the second subsamples (80 Mya)}
\KeywordTok{extract.dispRity}\NormalTok{(disparity, }\DataTypeTok{observed =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{subsamples =} \DecValTok{2}\NormalTok{, }\DataTypeTok{rarefaction =} \DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\subsubsection{\texorpdfstring{\texttt{scale.dispRity}}{scale.dispRity}}\label{scale.disprity}

This is the S3 method of \texttt{scale} (scaling and/or centring) that
can be applied to the disparity data of a \texttt{dispRity} object.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Getting the disparity values of the time subsamples}
\KeywordTok{head}\NormalTok{(}\KeywordTok{summary}\NormalTok{(disparity))}

\NormalTok{## Scaling the same disparity values}
\KeywordTok{head}\NormalTok{(}\KeywordTok{summary}\NormalTok{(}\KeywordTok{scale}\NormalTok{(disparity, }\DataTypeTok{scale =} \OtherTok{TRUE}\NormalTok{)))}

\NormalTok{## Scaling and centering:}
\KeywordTok{head}\NormalTok{(}\KeywordTok{summary}\NormalTok{(}\KeywordTok{scale}\NormalTok{(disparity, }\DataTypeTok{scale =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{center =} \OtherTok{TRUE}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\subsubsection{\texorpdfstring{\texttt{sort.dispRity}}{sort.dispRity}}\label{sort.disprity}

This is the S3 method of \texttt{sort} for sorting the subsamples
alphabetically (default) or following a specific pattern.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Sorting the disparity subsamples in inverse alphabetic order}
\KeywordTok{head}\NormalTok{(}\KeywordTok{summary}\NormalTok{(}\KeywordTok{sort}\NormalTok{(disparity, }\DataTypeTok{decreasing =} \OtherTok{TRUE}\NormalTok{)))}

\NormalTok{## Customised sorting}
\KeywordTok{head}\NormalTok{(}\KeywordTok{summary}\NormalTok{(}\KeywordTok{sort}\NormalTok{(disparity, }\DataTypeTok{sort =} \KeywordTok{c}\NormalTok{(}\DecValTok{7}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{3}\NormalTok{, }\DecValTok{4}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{6}\NormalTok{))))}
\end{Highlighting}
\end{Shaded}

\chapter{Palaeobiology demo: disparity-through-time and within
groups}\label{palaeobiology-demo-disparity-through-time-and-within-groups}

This demo aims to give quick overview of the \texttt{dispRity} package
(v.0.3) for palaeobiology analyses of disparity, including disparity
through time analyses. Please refer to my
\href{https://github.com/TGuillerme/dispRity}{GitHub page} for other
vignettes, namely the
\href{http://htmlpreview.github.com/?https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-manual.html}{\texttt{dispRity}
manual} that explains the functions in more detail.

In brief, this package allows users to easily perform
\textbf{disparity-through-time} analysis. This type of analysis often
starts by ordinating \textbf{morphometric} or \textbf{cladistic} data
into a multidimensional object hereafter called the
\textbf{morphospace}. One might be interested in studying how the
occupancy of the morphospace has changed through time by measuring a
summary metric of this morphospace through time, called
\textbf{disparity}. Disparity can then be seen as a value or a
distribution of values summarising the diversity of morphologies.

This demo showcases a typical disparity-through-time analysis: we are
going to test whether the disparity changed through time in a subset of
eutherian mammals from the last 100 million years.

\section{Before starting}\label{before-starting}

\subsection{\texorpdfstring{Installing
\texttt{dispRity}}{Installing dispRity}}\label{installing-disprity}

You can install this package easily if you are using the latest version
of \texttt{R} and \texttt{devtools}.

\begin{Shaded}
\begin{Highlighting}[]
\ControlFlowTok{if}\NormalTok{(}\OperatorTok{!}\KeywordTok{require}\NormalTok{(devtools)) }\KeywordTok{install.packages}\NormalTok{(}\StringTok{"devtools"}\NormalTok{)}
\KeywordTok{install_github}\NormalTok{(}\StringTok{"TGuillerme/dispRity"}\NormalTok{, }\DataTypeTok{ref =} \StringTok{"release"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(dispRity)}
\end{Highlighting}
\end{Shaded}

\subsection{The morphospace}\label{the-morphospace}

In this example, we are going to use a subset of the data from
\citep{beckancient2014}. This dataset contains an ordinated matrix of 50
discrete characters from mammals (\texttt{BeckLee\_mat50}), another
matrix of the same 50 mammals and the estimated discrete data characters
of their descendants (thus 50 + 49 rows, \texttt{BeckLee\_mat99}), a
dataframe containing the ages of each taxon in the dataset
(\texttt{BeckLee\_ages}) and finally a phylogenetic tree with the
relationships among the 50 mammals (\texttt{BeckLee\_tree}). The
ordinated matrix will represent our full morphospace, i.e.~all the
mammalian morphologies that ever existed through time (for this
dataset).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Loading demo and the package data}
\KeywordTok{library}\NormalTok{(dispRity)}

\NormalTok{## Setting the random seed for repeatability}
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{123}\NormalTok{)}

\NormalTok{## Loading the ordinated matrix/morphospace:}
\KeywordTok{data}\NormalTok{(BeckLee_mat50)}
\KeywordTok{head}\NormalTok{(BeckLee_mat50[,}\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{])}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                    [,1]          [,2]        [,3]       [,4]      [,5]
## Cimolestes   -0.5319679  0.1117759259  0.09865194 -0.1933148 0.2035833
## Maelestes    -0.4087147  0.0139690317  0.26268300  0.2297096 0.1310953
## Batodon      -0.6923194  0.3308625215 -0.10175223 -0.1899656 0.1003108
## Bulaklestes  -0.6802291 -0.0134872777  0.11018009 -0.4103588 0.4326298
## Daulestes    -0.7386111  0.0009001369  0.12006449 -0.4978191 0.4741342
## Uchkudukodon -0.5105254 -0.2420633915  0.44170317 -0.1172972 0.3602273
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{dim}\NormalTok{(BeckLee_mat50)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 50 48
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The morphospace contains 50 taxa and has 48 dimensions (or axes)}

\NormalTok{## Showing a list of first and last occurrences data for some fossils}
\KeywordTok{data}\NormalTok{(BeckLee_ages)}
\KeywordTok{head}\NormalTok{(BeckLee_ages)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##             FAD  LAD
## Adapis     37.2 36.8
## Asioryctes 83.6 72.1
## Leptictis  33.9 33.3
## Miacis     49.0 46.7
## Mimotona   61.6 59.2
## Notharctus 50.2 47.0
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Plotting a phylogeny}
\KeywordTok{data}\NormalTok{(BeckLee_tree)}
\KeywordTok{plot}\NormalTok{(BeckLee_tree, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{)}
\KeywordTok{axisPhylo}\NormalTok{(}\DataTypeTok{root =} \DecValTok{140}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{manual_files/figure-latex/unnamed-chunk-83-1.pdf}

\begin{quote}
You can have an even nicer looking tree if you use the \texttt{strap}
package!
\end{quote}

\begin{Shaded}
\begin{Highlighting}[]
\ControlFlowTok{if}\NormalTok{(}\OperatorTok{!}\KeywordTok{require}\NormalTok{(strap)) }\KeywordTok{install.packages}\NormalTok{(}\StringTok{"strap"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(strap)}
\KeywordTok{geoscalePhylo}\NormalTok{(BeckLee_tree, }\DataTypeTok{cex.tip =} \FloatTok{0.7}\NormalTok{, }\DataTypeTok{cex.ts =} \FloatTok{0.6}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\section{A disparity-through-time
analysis}\label{a-disparity-through-time-analysis}

\subsection{Splitting the morphospace through
time}\label{splitting-the-morphospace-through-time}

One of the crucial steps in disparity-through-time analysis is to split
the full morphospace into smaller time subsamples that contain the total
number of morphologies at certain points in time (time-slicing) or
during certain periods in time (time-binning). Basically, the full
morphospace represents the total number of morphologies across all time
and will be greater than any of the time subsamples of the morphospace.

The \texttt{dispRity} package provides a \texttt{time.subsamples}
function that allows users to split the morphospace into time slices
(using \texttt{method\ =\ continuous}) or into time bins (using
\texttt{method\ =\ discrete}). In this example, we are going to split
the morphospace into five equal time bins of 20 million years long from
100 million years ago to the present. We will also provide to the
function a table containing the first and last occurrences dates for
some fossils to take into account that some fossils might occur in
several of our different time bins.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating the vector of time bins ages}
\NormalTok{(time_bins <-}\StringTok{ }\KeywordTok{rev}\NormalTok{(}\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{0}\NormalTok{, }\DataTypeTok{to =} \DecValTok{100}\NormalTok{, }\DataTypeTok{by =} \DecValTok{20}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 100  80  60  40  20   0
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Splitting the morphospace using the time.subsamples function}
\NormalTok{(binned_morphospace <-}\StringTok{ }\KeywordTok{time.subsamples}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat50, }\DataTypeTok{tree =}\NormalTok{ BeckLee_tree,}
    \DataTypeTok{method =} \StringTok{"discrete"}\NormalTok{, }\DataTypeTok{time =}\NormalTok{ time_bins, }\DataTypeTok{inc.nodes =} \OtherTok{FALSE}\NormalTok{,}
    \DataTypeTok{FADLAD =}\NormalTok{ BeckLee_ages))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 5 discrete time subsamples for 50 elements:
##     100 - 80, 80 - 60, 60 - 40, 40 - 20, 20 - 0.
\end{verbatim}

The output object is a \texttt{dispRity} object. For details about this
object class, please refer to the
\href{http://htmlpreview.github.com/?https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-manual.html}{\texttt{dispRity}
manual}. In brief, however, \texttt{dispRity} objects are lists of
different elements (i.e.~disparity results, morphospace time subsamples,
morphospace attributes, etc.) that display only a summary of the object
when calling the object to avoiding filling the \texttt{R} console with
superfluous output.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Printing the class of the object}
\KeywordTok{class}\NormalTok{(binned_morphospace)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] "dispRity"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Printing the content of the object}
\KeywordTok{str}\NormalTok{(binned_morphospace)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## List of 3
##  $ matrix    : num [1:50, 1:48] -0.532 -0.409 -0.692 -0.68 -0.739 ...
##   ..- attr(*, "dimnames")=List of 2
##   .. ..$ : chr [1:50] "Cimolestes" "Maelestes" "Batodon" "Bulaklestes" ...
##   .. ..$ : NULL
##  $ call      :List of 1
##   ..$ subsamples: chr "discrete"
##  $ subsamples:List of 5
##   ..$ 100 - 80:List of 1
##   .. ..$ elements: int [1:8, 1] 5 4 6 8 43 10 11 42
##   ..$ 80 - 60 :List of 1
##   .. ..$ elements: int [1:15, 1] 7 8 9 1 2 3 12 13 14 44 ...
##   ..$ 60 - 40 :List of 1
##   .. ..$ elements: int [1:13, 1] 41 49 24 25 26 27 28 21 22 19 ...
##   ..$ 40 - 20 :List of 1
##   .. ..$ elements: int [1:6, 1] 15 39 40 35 23 47
##   ..$ 20 - 0  :List of 1
##   .. ..$ elements: int [1:10, 1] 36 37 38 32 33 34 50 48 29 30
##  - attr(*, "class")= chr "dispRity"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{names}\NormalTok{(binned_morphospace)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] "matrix"     "call"       "subsamples"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Printing the object as a dispRity class}
\NormalTok{binned_morphospace}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 5 discrete time subsamples for 50 elements:
##     100 - 80, 80 - 60, 60 - 40, 40 - 20, 20 - 0.
\end{verbatim}

\begin{quote}
These objects will gradually contain more information when completing
the following steps in the disparity-through-time analysis.
\end{quote}

\subsection{Bootstrapping the data}\label{bootstrapping-the-data}

Once we obtain our different time subsamples, we can bootstrap and
rarefy them (i.e.~pseudo-replicating the data). The bootstrapping allows
us to make each subsample more robust to outliers and the rarefaction
allows us to compare subsamples with the same number of taxa to remove
sampling biases (i.e.~more taxa in one subsample than the others). The
\texttt{boot.matrix} function bootstraps the \texttt{dispRity} object
and the \texttt{rarefaction} option within performs rarefaction.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Bootstrapping each time subsample 100 times}
\NormalTok{(boot_bin_morphospace <-}\StringTok{ }\KeywordTok{boot.matrix}\NormalTok{(binned_morphospace, }\DataTypeTok{bootstraps =} \DecValTok{100}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 5 discrete time subsamples for 50 elements with 48 dimensions:
##     100 - 80, 80 - 60, 60 - 40, 40 - 20, 20 - 0.
## Data was bootstrapped 100 times (method:"full").
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Getting the minimum number of rows (i.e. taxa) in the time subsamples}
\KeywordTok{min}\NormalTok{(}\KeywordTok{unlist}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(boot_bin_morphospace}\OperatorTok{$}\NormalTok{subsamples, lapply, nrow)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 6
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Bootstrapping each time subsample 100 times and rarefying them }
\NormalTok{(rare_bin_morphospace <-}\StringTok{ }\KeywordTok{boot.matrix}\NormalTok{(binned_morphospace, }\DataTypeTok{bootstraps =} \DecValTok{100}\NormalTok{,}
    \DataTypeTok{rarefaction =} \DecValTok{6}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 5 discrete time subsamples for 50 elements with 48 dimensions:
##     100 - 80, 80 - 60, 60 - 40, 40 - 20, 20 - 0.
## Data was bootstrapped 100 times (method:"full") and rarefied to 6 elements.
\end{verbatim}

\subsection{Calculating disparity}\label{calculating-disparity-1}

We can now calculate the disparity within each time subsamples along
with some confidence intervals generated by the pseudoreplication step
above (bootstraps/rarefaction). Disparity can be calculated in many ways
and this package allows users to come up with their own disparity
metrics. For more details, please refer to the
\href{http://htmlpreview.github.com/?https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-metrics.html}{\texttt{dispRity}
metric vignette}.

In this example, we are going to calculate the spread of the data in
each time subsample by calculating disparity as the sum of the variance
of each dimension of the morphospace in each time subsample using the
\texttt{dispRity} function. Thus, in this example, disparity is defined
by the multi-dimensional variance of each time subsample (i.e.~the
spread of the taxa within the morphospace). Note that this metric comes
with a caveat (not solved here) since it ignores covariances among the
dimensions of the morphospace. We use this here because it is a standard
metric used in disparity-through-time analysis.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating disparity for the bootstrapped data}
\NormalTok{(boot_disparity <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(boot_bin_morphospace, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, variances)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 5 discrete time subsamples for 50 elements with 48 dimensions:
##     100 - 80, 80 - 60, 60 - 40, 40 - 20, 20 - 0.
## Data was bootstrapped 100 times (method:"full").
## Disparity was calculated as: c(sum, variances).
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating disparity for the rarefied data}
\NormalTok{(rare_disparity <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(rare_bin_morphospace, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, variances)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 5 discrete time subsamples for 50 elements with 48 dimensions:
##     100 - 80, 80 - 60, 60 - 40, 40 - 20, 20 - 0.
## Data was bootstrapped 100 times (method:"full") and rarefied to 6 elements.
## Disparity was calculated as: c(sum, variances).
\end{verbatim}

The \texttt{dispRity} function does not actually display the calculated
disparity values but rather only the properties of the disparity object
(size, subsamples, metric, etc.). To display the actual calculated
scores, we need to summarise the disparity object using the S3 method
\texttt{summary} that is applied to a \texttt{dispRity} object (see
\texttt{?summary.dispRity} for more details).

As for any \texttt{R} package, you can refer to the help files for each
individual function for more details.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Summarising the disparity results}
\KeywordTok{summary}\NormalTok{(boot_disparity)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs bs.median  2.5%   25%   75% 97.5%
## 1   100 - 80  8 1.675     1.488 1.087 1.389 1.568 1.648
## 2    80 - 60 15 1.782     1.679 1.538 1.631 1.728 1.792
## 3    60 - 40 13 1.913     1.772 1.607 1.734 1.826 1.886
## 4    40 - 20  6 2.022     1.707 1.212 1.537 1.822 1.942
## 5     20 - 0 10 1.971     1.794 1.598 1.716 1.842 1.890
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(rare_disparity)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsamples  n   obs bs.median  2.5%   25%   75% 97.5%
## 1   100 - 80  8 1.675     1.484 1.194 1.400 1.547 1.636
## 2   100 - 80  6    NA     1.477 0.993 1.361 1.569 1.698
## 3    80 - 60 15 1.782     1.674 1.517 1.600 1.725 1.793
## 4    80 - 60  6    NA     1.655 1.299 1.532 1.754 1.882
## 5    60 - 40 13 1.913     1.767 1.601 1.714 1.829 1.861
## 6    60 - 40  6    NA     1.787 1.314 1.672 1.879 1.984
## 7    40 - 20  6 2.022     1.736 1.281 1.603 1.822 1.948
## 8     20 - 0 10 1.971     1.807 1.595 1.729 1.856 1.917
## 9     20 - 0  6    NA     1.790 1.435 1.718 1.873 1.995
\end{verbatim}

\begin{quote}
The summary.dispRity function comes with many options on which values to
calculate (central tendency and quantiles) and on how many digits to
display. Refer to the function's manual for more details.
\end{quote}

\subsection{Plotting the results}\label{plotting-the-results}

It is sometimes easier to visualise the results in a plot than in a
table. For that we can use the \texttt{plot} S3 function to plot the
\texttt{dispRity} objects (see \texttt{?plot.dispRity} for more
details).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical options}
\KeywordTok{quartz}\NormalTok{(}\DataTypeTok{width =} \DecValTok{10}\NormalTok{, }\DataTypeTok{height =} \DecValTok{5}\NormalTok{) ; }\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =}\NormalTok{ (}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{)), }\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}

\NormalTok{## Plotting the bootstrapped and rarefied results}
\KeywordTok{plot}\NormalTok{(boot_disparity, }\DataTypeTok{type =} \StringTok{"continuous"}\NormalTok{, }\DataTypeTok{main =} \StringTok{"bootstrapped results"}\NormalTok{)}
\KeywordTok{plot}\NormalTok{(rare_disparity, }\DataTypeTok{type =} \StringTok{"continuous"}\NormalTok{, }\DataTypeTok{main =} \StringTok{"rarefied results"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{manual_files/figure-latex/unnamed-chunk-90-1.pdf}

\section{Testing differences}\label{testing-differences}

Finally, to draw some valid conclusions from these results, we can apply
some statistical tests. We can test, for example, if mammalian disparity
changed significantly through time over the last 100 million years. To
do so, we can compare the means of each time-bin in a sequential manner
to see whether the disparity in bin \emph{n} is equal to the disparity
in bin \emph{n+1}, and whether this is in turn equal to the disparity in
bin \emph{n+2}, etc. Because our data is temporally autocorrelated
(i.e.~what happens in bin \emph{n+1} depends on what happened in bin
\emph{n}) and pseudoreplicated (i.e.~each bootstrap draw creates
non-independent time subsamples because they are all based on the same
time subsamples), we apply a non-parametric mean comparison: the
\texttt{wilcox.test}. Also, we need to apply a p-value correction
(e.g.~Bonferroni correction) to correct for multiple testing (see
\texttt{?p.adjust} for more details).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Testing the differences between bins in the bootstrapped dataset.}
\KeywordTok{test.dispRity}\NormalTok{(boot_disparity, }\DataTypeTok{test =}\NormalTok{ wilcox.test, }\DataTypeTok{comparison =} \StringTok{"sequential"}\NormalTok{,}
    \DataTypeTok{correction =} \StringTok{"bonferroni"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [[1]]
##                    statistic
## 100 - 80 : 80 - 60       471
## 80 - 60 : 60 - 40       1562
## 60 - 40 : 40 - 20       6250
## 40 - 20 : 20 - 0        3725
## 
## [[2]]
##                         p.value
## 100 - 80 : 80 - 60 7.427563e-28
## 80 - 60 : 60 - 40  1.798899e-16
## 60 - 40 : 40 - 20  9.061511e-03
## 40 - 20 : 20 - 0   7.379715e-03
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Testing the differences between bins in the rarefied dataset.}
\KeywordTok{test.dispRity}\NormalTok{(rare_disparity, }\DataTypeTok{test =}\NormalTok{ wilcox.test, }\DataTypeTok{comparison =} \StringTok{"sequential"}\NormalTok{,}
    \DataTypeTok{correction =} \StringTok{"bonferroni"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [[1]]
##                    statistic
## 100 - 80 : 80 - 60       662
## 80 - 60 : 60 - 40       1814
## 60 - 40 : 40 - 20       5752
## 40 - 20 : 20 - 0        3621
## 
## [[2]]
##                         p.value
## 100 - 80 : 80 - 60 1.214988e-25
## 80 - 60 : 60 - 40  2.823697e-14
## 60 - 40 : 40 - 20  2.653018e-01
## 40 - 20 : 20 - 0   3.026079e-03
\end{verbatim}

Here our results show significant changes in disparity through time
between all time bins (all p-values \textless{} 0.05). However, when
looking at the rarefied results, there is no significant difference
between the time bins in the Palaeogene (60-40 to 40-20 Mya), suggesting
that the differences detected in the first test might just be due to the
differences in number of taxa sampled (13 or 6 taxa) in each time bin.

\subsection{References}\label{references}

\end{document}