### 2017-08-24 paper/CoverLetter.tex
Updated cover letter

\documentclass[11pt]{letter}
\usepackage[a4paper,left=2.5cm, right=2.5cm, top=1cm, bottom=1cm]{geometry}
\usepackage[osf]{mathpazo}
\signature{Thomas Guillerme}
\address{Imperial College London \\Silwood Park Campus \\Buckhurst Road \\Ascot SL57PY, United Kingdom \\guillert@tcd.ie}
\longindentation=0pt
\begin{document}

\begin{letter}{}
\opening{Dear Editors,}

Studying multivariate data as an alternative to univariate data has become increasingly popular in the last decades both in ecology and evolution.
This includes, for example, studying form and function (e.g. Diaz et al., 2016, \textit{Nature}) or morphological diversity (i.e. disparity - e.g. Brusatte et al. 2012, \textit{Science}).
These methods have in common to use a multidimensional space (through ordination or other techniques) representing the combination of all possible observed traits and then measuring some aspect of the space (e.g. its hyper-volume) for different groups and compare them between each other or to a null hypothesis.

There exist a multitude of ways to define a multidimensional space, a metric measuring a specific aspect thereof and a test for many hypothesis.
There are numerous implementation of such analysis pipeline in several R packages (e.g. \texttt{Claddis}, \texttt{geomorph}, \texttt{ade4}, \texttt{vegan}, etc.) or even in other languages (e.g. \texttt{GINGKO} in javascript).
However, each of these implementations are particular to the software/package author's definition of the multidimensional space, metric and test.
This leads researchers to often use these tool-kits by default, thus hampering them to use the best definitions of the space/metric/test fitted to their particular question.
At worst this leads to wrong statistic claims, and, at best to time consuming code retro-engineering.

In this application manuscript entitled ``dispRity: an R package for measuring disparity'', I present a complete and flexible R package allowing to run multidimensional analysis where each parameter can be easily defined by the user (the space, the metric and the test).
In fact, this package provides a tidy and modular environment where users can use their own multidimensional space defined as a matrix and apply to it any metric and any test (provided as functions).
This allows for highly modular multidimensional analysis and the tidy interface allows easily reproducible analysis.
Finally, this package comes with an in depth gitbook manual covering most functionalities and being updated regularly following users requests.

I look forward to hearing from you soon,

\closing{Yours sincerely,}

\end{letter}
\end{document}

%Editors:
%Sam Price, Daniele Silvestro

%Reviewers
%Graeme Lloyd, Dave Bapst, Emma Sherratt