### 2017-11-10 inst/gitbook/05_guts.Rmd

---
title: "The guts of the dispRity package"
author: "Thomas Guillerme (guillert@tcd.ie)"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

# The guts of the `dispRity` package

## Manipulating `dispRity` objects

Disparity analysis involves a lot of manipulation of many matrices (especially when bootstrapping) which can be impractical to visualise and will quickly overwhelm your `R` console.
Even the simple Beck and Lee 2014 example above produces an object with > 72 lines of lists of lists of matrices!

Therefore `dispRity` uses a specific class of object called a `dispRity` object.
These objects allow users to use S3 method functions such as `summary.dispRity`, `plot.dispRity` and `print.dispRity`. 
`dispRity` also contains various utility functions that manipulate the `dispRity` object (e.g. `sort.dispRity`, `extract.dispRity` see the full list in the next section).
These functions modify the `dispRity` object without having to delve into its complex structure!
The full structure of a `dispRity` object is detailed [here](https://github.com/TGuillerme/dispRity/blob/master/disparity_object.md).

```{r}
## Loading the example data
data(disparity)

## What is the class of the median_centroids object?
class(disparity)

## What does the object contain?
names(disparity)

## Summarising it using the S3 method print.dispRity
disparity
```

Note that it is always possible to recall the full object using the argument `all = TRUE` in `print.dispRity`:

```{r, eval=FALSE}
## Display the full object
print(disparity, all = TRUE)
## This is more nearly ~ 5000 lines on my 13 inch laptop screen!
```

## `dispRity` utilities

The package also provides some utility functions to facilitate multidimensional analysis.

### `dispRity` object utilities  <a name="dispRity.utilities"></a>
The first set of utilities are functions for manipulating `dispRity` objects:

#### `make.dispRity`
This function creates empty `dispRity` objects.

```{r}
## Creating an empty dispRity object
make.dispRity()

## Creating an "empty" dispRity object with a matrix
(disparity_obj <- make.dispRity(matrix(rnorm(20), 5, 4)))
```

####  `fill.dispRity`
This function initialises a `dispRity` object and generates its call properties.

```{r}
## The dispRity object's call is indeed empty
disparity_obj$call

## Filling an empty disparity object (that needs to contain at least a matrix)
(disparity_obj <- fill.dispRity(disparity_obj))

## The dipRity object has now the correct minimal attributes
disparity_obj$call
```

####  `matrix.dispRity`
This function extracts a specific matrix from a disparity object.
The matrix can be one of the bootstrapped matrices or/and a rarefied matrix.

```{r}
## Extracting the matrix containing the coordinates of the elements at time 50
str(matrix.dispRity(disparity, "50"))

## Extracting the 3rd bootstrapped matrix with the 2nd rarefaction level
## (15 elements) from the second group (80 Mya)
str(matrix.dispRity(disparity, subsamples = 1, bootstrap = 3, rarefaction = 2))
```

#### `get.subsamples.dispRity`
This function creates a dispRity object that contains only elements from one specific subsamples.

```{r, eval=FALSE}
## Extracting all the data for the crown mammals
(crown_mammals <- get.subsamples.dispRity(disp_crown_stemBS, "Group.crown"))

## The object keeps the properties of the parent object but is composed of only one subsamples
length(crown_mammals$subsamples)
```

#### `extract.dispRity`
This function extracts the calculated disparity values of a specific matrix.

```{r, eval=FALSE}
## Extracting the observed disparity (default)
extract.dispRity(disparity)

## Extracting the disparity from the bootstrapped values from the
## 10th rarefaction level from the second subsamples (80 Mya)
extract.dispRity(disparity, observed = FALSE, subsamples = 2, rarefaction = 10)
```

#### `scale.dispRity`
This is the S3 method of `scale` (scaling and/or centring) that can be applied to the disparity data of a `dispRity` object.

```{r, eval=FALSE}
## Getting the disparity values of the time subsamples
head(summary(disparity))

## Scaling the same disparity values
head(summary(scale(disparity, scale = TRUE)))

## Scaling and centering:
head(summary(scale(disparity, scale = TRUE, center = TRUE)))
```

#### `sort.dispRity`
This is the S3 method of `sort` for sorting the subsamples alphabetically (default) or following a specific pattern.

```{r, eval=FALSE}
## Sorting the disparity subsamples in inverse alphabetic order
head(summary(sort(disparity, decreasing = TRUE)))

## Customised sorting
head(summary(sort(disparity, sort = c(7, 1, 3, 4, 5, 2, 6))))
```

## The `dispRity` object content
The functions above are utilities to easily and safely access different elements in the `dispRity` object.
Alternatively, of course, each elements can be accessed manually.
Here is an explanation on how it works.
The `dispRity` object is a `list` of two to four elements, each of which are detailed below:

 * `$matrix`: an object of class `matrix`, the full multidimensional space.
 * `$call`: an object of class `list` containing information on the `dispRity` object content.
 * `$subsamples`: an object of class `list` containing the subsamples of the multidimensional space.
 * `$disparity`: an object of class `list` containing the disparity values.

The `dispRity` object is loosely based on `C` structure objects.
In fact, it is composed of one unique instance of a matrix (the multidimensional space) upon which the metric function is called via "pointers" to only a certain number of elements and/or dimensions of this matrix.
This allows for: (1) faster and easily tractable execution time: the metric functions are called through apply family function and can be parallelised; and (2) a really low memory footprint:  at any time, only one matrix is present in the `R` environment rather than multiple copies of it for each subsample.

### `$matrix`
This is the multidimensional space, stored in the `R` environment as a `matrix` object.
It requires row names but not column names.
By default, if the row names are missing, `dispRity` function will arbitrarily generate them in numeric order (i.e. `rownames(matrix) <- 1:nrow(matrix)`).
This element of the `dispRity` object is never modified.

### `$call`
This element contains the information on the `dispRity` object content.
It is a `list` that can contain the following:

 * `$call$subsamples`: a vector of `character` with information on the subsamples type (either `"continuous"`, `"discrete"` or `"custom"`) and their eventual model (`"acctran"`, `"deltran"`, `"punctuated"`, `"gradual"`). This element generated only once via `time.subsamples()` and `custom.subsamples()`.
 * `$call$dimensions`: either a single `numeric` value indicating how many dimensions to use or a vector of `numeric` values indicating which specific dimensions to use. This element is by default the number of columns in `$matrix` but can be modified through `boot.matrix()` or `dispRity()`.
 * `$call$bootstrap`: this is a `list` containing three elements:
    * `[[1]]`: the number of bootstrap replicates (`numeric`)
    * `[[2]]`: the bootstrap method (`character`)
    * `[[3]]`: the rarefaction levels (`numeric` vector)
 * `$call$disparity`: this is a `list` containing one element, `$metric`, that is a `list` containing the different functions passed to the `metric` argument in `dispRity`. These are `call` elements and get modified each time the `dispRity` function is used (the first element is the first metric(s), the second, the second metric(s), etc.).

### `$subsamples`
This element contain the eventual subsamples of the multidimensional space.
It is a `list` of subsample names.
Each subsample name is in turn a `list` of at least one element called `elements` which is in turn a `matrix`.
This `elements` matrix is the raw (observed) elements in the subsamples.
The `elements` matrix is composed of `numeric` values in one column and _n_ rows (the number of elements in the subsample).
Each of these values are a "pointer" (`C` inspired) to the element of the `$matrix`.
For example, lets assume a `dispRity` object called `disparity`, composed of at least one subsamples called `sub1`:

```
 disparity$subsamples$sub1$elements
      [,1]
 [1,]    5
 [2,]    4
 [3,]    6
 [4,]    7
```

The values in the matrix "point" to the elements in `$matrix`: here, the multidimensional space with only the 4th, 5th, 6th and 7th elements.
The following elements in `diparity$subsamples$sub1` will correspond to the same "pointers" but drawn from the bootstrap replicates.
The columns will correspond to different bootstrap replicates.
For example:

```
 disparity$subsamples$sub1[[2]]
      [,1] [,2] [,3] [,4]
 [1,]   57   43   70    4
 [2,]   43   44    4    4
 [3,]   42   84   44    1
 [4,]   84    7    2   10
```
This signifies that we have four bootstrap pseudo-replicates pointing each time to four elements in `$matrix`.
The next element (`[[3]]`) will be the same for the eventual first rarefaction level (i.e. the resulting bootstrap matrix will have _m_ rows where _m_ is the number of elements for this rarefaction level).
The next element after that (`[[4]]`) will be the same for with an other rarefaction level and so forth...

### `$disparity`
The `$disparity` element is identical to the `$subsamples` element structure (a list of list(s) containing matrices) but the matrices don't contain "pointers" to `$matrix` but the disparity result of the disparity metric applied to the "pointers".
For example, in our first example (`$elements`) from above, if the disparity metric is of dimensions level 1, we would have:

```
 disparity$disparity$sub1$elements
      [,1]
 [1,]    1.82
```

This is the observed disparity (1.82) for the subsample called `sub1`.
If the disparity metric is of dimension level 2 (say the function `range` that outputs two values), we would have:

```
 disparity$disparity$sub1$elements
      [,1]
 [1,]    0.82
 [2,]    2.82
```

The following elements in the list follow the same logic as before: rows are disparity values (one row for a dimension level 1 metric, multiple for a dimensions level 2 metric) and columns are the bootstrap replicates (the bootstrap with all elements followed by the eventual rarefaction levels).
For example for the bootstrap without rarefaction (second element of the list):

```
 disparity$disparity$sub1[[2]]
         [,1]     [,2]     [,3]     [,4]
[1,] 1.744668 1.777418 1.781624 1.739679 
```

<!-- ## Modularity

### Making your own metrics

### Making your own tests

## Where is this going?
 -->
<!-- ## Running `dispRity` in parallel

The computationally intensive function `dispRity` have a `parallel` option to speed up their calculations.

This option requires the package `snow` and takes arguments that are to be passed to  -->