### 2018-07-12 inst/gitbook/index.Rmd
added cites to the manual

--- 
title: "dispRity manual"
author: "Thomas Guillerme (guillert@tcd.ie) and Natalie Cooper (natalie.cooper@nhm.ac.uk)"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [References.bib, packages.bib]
link-citations: yes
github-repo: TGuillerme/dispRity
description: "dispRity R package vignette"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval = TRUE, echo = FALSE}
## master version
version_raw <- RCurl::getURL("https://raw.githubusercontent.com/TGuillerme/dispRity/master/DESCRIPTION")
version_master <- strsplit(strsplit(version_raw, split = "Version: ")[[1]][2], split = "\nDate:")[[1]][1]

## release version
version_raw <- RCurl::getURL("https://raw.githubusercontent.com/TGuillerme/dispRity/release/DESCRIPTION")
version_release <- strsplit(strsplit(version_raw, split = "Version: ")[[1]][2], split = "\nDate:")[[1]][1]

## CRAN version
version_raw <- RCurl::getURL("https://cran.r-project.org/web/packages/dispRity/index.html")
version_CRAN <- strsplit(strsplit(version_raw, split = "\n<td>Version:</td>\n<td>")[[1]][2], split = "</td>\n</tr>\n<tr>\n<td>Depends:")[[1]][1]
```

# `dispRity`

This is a package for measuring disparity in `R`.
It allows users to summarise matrices as representations as multidimensional spaces into a single value or distribution describing a specific aspect of this multidimensional space (the disparity).
Multidimensional spaces can be ordinated matrices from MDS, PCA, PCO, PCoA but the package is *not* restricted to any type of matrices!
This manual is based on the version `r version_release`.

## What is `dispRity`?
This is a modular package for measuring disparity in `R`.
It allows users to summarise ordinated matrices (e.g. MDS, PCA, PCO, PCoA) to perform some multidimensional analysis.
Typically, these analysis are used in palaeobiology and evolutionary biology to study the changes in morphology through time.
However, there are many more applications in ecology, evolution and beyond.

### Modular?
Because their exist a multitude of ways to measure disparity, each adapted to every specific question, this package uses an easy to modify modular architecture.
In coding, each module is simply a function or a modification of a function that can be passed to the main functions of the package to tweak it to your proper needs!
In practice, you will notice throughout this manual that some function can take other functions as arguments: the modular architecture of this package allows you to use any function for these arguments (with some restrictions explained for each specific cases).
This will allow you to finely tune your multidimensional analysis to the needs of your specific question!

## Installing and running the package

You can install this package easily, directly from the CRAN:

```{r, eval = FALSE}
install.packages("dispRity")
```

Alternatively, for the most up to data version and some functionalities not compatible with the CRAN, you can use the package through GitHub using `devtool` (see [to CRAN or not to CRAN?](#noCRAN) for more details):

```{r, eval=FALSE}
## Checking if devtools is already installed
if(!require(devtools)) install.packages("devtools")

## Installing the latest released version directly from GitHub
install_github("TGuillerme/dispRity", ref = "release")
```

Anyways to upload it, you can use

```{r, eval = TRUE, echo = FALSE, message = FALSE}
## loading the package and setting up the start seed.
library(dispRity)
## Setting a random seed for repeatability
set.seed(123)
```

Note this uses the `release` branch (`r version_release`).
For the piping-hot (but potentially unstable) version, you can change the argument `ref = release` to `ref = master`.
`dispRity` depends mainly on the `ape` package and uses functions from several other packages (`ade4`, `geometry`, `grDevices`, `hypervolume`, `paleotree`, `snow`, `Claddis`, `geomorph` and `RCurl`).

## To CRAN or not to CRAN? {#noCRAN}

This package can be installed via [the CRAN]() or via [GitHub](https://github.com/TGuillerme/dispRity).

 * The CRAN version is a more "strict" version of the package that does not include some functionalities (such as [`Claddis.ordination`](#Claddis-ordination) or this gitbook vignette) because of compatibility issues with the CRAN requirement.

 * The GitHub version is a more "reactive" version (fed by user's suggestions and comments) and can contain more functionalities (see list below). This version exists on a `"release"` branch which is tested as strictly as the CRAN version with Travis and codecov. The `"master"` branch contains always the latest "piping hot", not yet release version. It might (though rarely) be bugged.

<!-- badges out start -->
[![Build Status](https://travis-ci.org/TGuillerme/dispRity.svg?branch=release)](https://travis-ci.org/TGuillerme/dispRity)
[![codecov](https://codecov.io/gh/TGuillerme/dispRity/branch/release/graph/badge.svg)](https://codecov.io/gh/TGuillerme/dispRity)
[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
<!-- badges out end   -->

Note however, that apart from the functions listed below, both versions of the package have the same tested and implemented functions.

Functions not available on the CRAN version (`r version_CRAN`):

 * [`Claddis.ordination`](#Claddis-ordination)

## Help

If you need help with the package, hopefully the following manual will be useful.
However, parts of this package are still in development and some other parts are probably not covered.
Thus if you have suggestions or comments on on what has already been developed or will be developed, please send me an email (<guillert@tcd.ie>) or if you are a GitHub user, directly create an issue on the [GitHub page](https://github.com/TGuillerme/dispRity).

## Citations

To cite the package, this manual or some specific functionalities, you can use the following references:

The package main paper:

> Guillerme T. dispRity: A modular R package for measuring disparity. Methods Ecol Evol. 2018;9:1755–1763. [doi.org/10.1111/2041-210X.13022](https://doi.org/10.1111/2041-210X.13022).

The package manual (regularly updated!):

> Guillerme, T. & Cooper, N. (2018): dispRity manual. figshare. Preprint. [10.6084/m9.figshare.6187337.v1](https://doi.org/10.6084/m9.figshare.6187337.v1).

The time-slicing method implemented in `chrono.subsets` (unfortunately not Open Access, but you can still get a [free copy from here](https://tguillerme.github.io/papers/Guillerme&Cooper-2018-Palaeontology.pdf)):

> Guillerme, T. and Cooper, N. (2018), Time for a rethink: time sub-sampling methods in disparity-through-time analyses. Palaeontology, 61: 481-493. [doi:10.1111/pala.12364](https://scholar.google.co.uk/scholar?hl=en&as_sdt=0%2C5&q=Guillerme%2C+T.+and+Cooper%2C+N.+%282018%29%2C+Time+for+a+rethink%3A+time+sub-sampling+methods+in+disparity-through-time+analyses.+Palaeontology%2C&btnG=).

<!-- ## Associated papers -->

<!-- Here are some papers that have used the `dispRity` package: -->