### 2018-09-27 inst/gitbook/dispRity_manual.tex

\documentclass[]{book}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={dispRity manual},
            pdfauthor={Thomas Guillerme (guillert@tcd.ie), Mark Puttick (marknputtick@gmail.com) and Natalie Cooper (natalie.cooper@nhm.ac.uk)},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{dispRity manual}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Thomas Guillerme
(\href{mailto:guillert@tcd.ie}{\nolinkurl{guillert@tcd.ie}}), Mark
Puttick
(\href{mailto:marknputtick@gmail.com}{\nolinkurl{marknputtick@gmail.com}})
and Natalie Cooper
(\href{mailto:natalie.cooper@nhm.ac.uk}{\nolinkurl{natalie.cooper@nhm.ac.uk}})}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{2018-09-27}

\usepackage{booktabs}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{lemma}{Lemma}[chapter]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[chapter]
\newtheorem{corollary}{Corollary}[chapter]
\newtheorem{proposition}{Proposition}[chapter]
\theoremstyle{definition}
\newtheorem{example}{Example}[chapter]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[chapter]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}
\maketitle

{
\setcounter{tocdepth}{1}
\tableofcontents
}
\chapter{\texorpdfstring{\texttt{dispRity}}{dispRity}}\label{disprity}

This is a package for measuring disparity in \texttt{R}. It allows users
to summarise matrices as representations as multidimensional spaces into
a single value or distribution describing a specific aspect of this
multidimensional space (the disparity). Multidimensional spaces can be
ordinated matrices from MDS, PCA, PCO, PCoA but the package is
\emph{not} restricted to any type of matrices! This manual is based on
the version 1.2.

\section{\texorpdfstring{What is
\texttt{dispRity}?}{What is dispRity?}}\label{what-is-disprity}

This is a modular package for measuring disparity in \texttt{R}. It
allows users to summarise ordinated matrices (e.g.~MDS, PCA, PCO, PCoA)
to perform some multidimensional analysis. Typically, these analysis are
used in palaeobiology and evolutionary biology to study the changes in
morphology through time. However, there are many more applications in
ecology, evolution and beyond.

\subsection{Modular?}\label{modular}

Because their exist a multitude of ways to measure disparity, each
adapted to every specific question, this package uses an easy to modify
modular architecture. In coding, each module is simply a function or a
modification of a function that can be passed to the main functions of
the package to tweak it to your proper needs! In practice, you will
notice throughout this manual that some function can take other
functions as arguments: the modular architecture of this package allows
you to use any function for these arguments (with some restrictions
explained for each specific cases). This will allow you to finely tune
your multidimensional analysis to the needs of your specific question!

\section{Installing and running the
package}\label{installing-and-running-the-package}

You can install this package easily, directly from the CRAN:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{install.packages}\NormalTok{(}\StringTok{"dispRity"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Alternatively, for the most up to data version and some functionalities
not compatible with the CRAN, you can use the package through GitHub
using \texttt{devtool} (see \protect\hyperlink{noCRAN}{to CRAN or not to
CRAN?} for more details):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Checking if devtools is already installed}
\ControlFlowTok{if}\NormalTok{(}\OperatorTok{!}\KeywordTok{require}\NormalTok{(devtools)) }\KeywordTok{install.packages}\NormalTok{(}\StringTok{"devtools"}\NormalTok{)}

\NormalTok{## Installing the latest released version directly from GitHub}
\KeywordTok{install_github}\NormalTok{(}\StringTok{"TGuillerme/dispRity"}\NormalTok{, }\DataTypeTok{ref =} \StringTok{"release"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Anyways to upload it, you can use

Note this uses the \texttt{release} branch (1.2). For the piping-hot
(but potentially unstable) version, you can change the argument
\texttt{ref\ =\ release} to \texttt{ref\ =\ master}. \texttt{dispRity}
depends mainly on the \texttt{ape} package and uses functions from
several other packages (\texttt{ade4}, \texttt{geometry},
\texttt{grDevices}, \texttt{hypervolume}, \texttt{paleotree},
\texttt{snow}, \texttt{Claddis}, \texttt{geomorph} and \texttt{RCurl}).

\hypertarget{noCRAN}{\section{To CRAN or not to CRAN?}\label{noCRAN}}

This package can be installed via
\href{https://cran.r-project.org/package=dispRity}{the CRAN} or via
\href{https://github.com/TGuillerme/dispRity}{GitHub}.

\begin{itemize}
\item
  The CRAN version is a more ``strict'' version of the package that does
  not include some functionalities (such as
  \protect\hyperlink{Claddis-ordination}{\texttt{Claddis.ordination}} or
  this gitbook vignette) because of compatibility issues with the CRAN
  requirement.
\item
  The GitHub version is a more ``reactive'' version (fed by user's
  suggestions and comments) and can contain more functionalities (see
  list below). This version exists on a \texttt{"release"} branch which
  is tested as strictly as the CRAN version with Travis and codecov. The
  \texttt{"master"} branch contains always the latest ``piping hot'',
  not yet release version. It might (though rarely) be bugged.
\end{itemize}

Note however, that apart from the functions listed below, both versions
of the package have the same tested and implemented functions.

Functions not available on the CRAN version (1.2):

\begin{itemize}
\tightlist
\item
  \protect\hyperlink{Claddis-ordination}{\texttt{Claddis.ordination}}
\end{itemize}

\subsection{So which version do I choose?}\label{version}

There are always three version of the package available:

\begin{itemize}
\tightlist
\item
  The CRAN one
\item
  The GitHub \texttt{release} one
\item
  The GitHub \texttt{master} one
\end{itemize}

The differences between the CRAN one and the GitHub \texttt{release} or
\texttt{master} ones is explained just above. For the the GitHub
version, the differences are that the \texttt{release} one is more
stable (i.e.~more rarely modified) and the \texttt{master} one is more
live one (i.e.~bug fixes and new functionalities are added as they
come).

If you want the latest-latest version of the package I suggest using the
GitHub \texttt{master} one, especially if you recently emailed me
reporting a minor bug or wanting a new functionality! Note however that
\emph{it can happen} that the \texttt{master} version can sometimes be
bugged (especially when there are major \texttt{R} and \texttt{R}
packages updates), however, the status of the package state on both the
\texttt{release} and the \texttt{master} version is constantly displayed
on the \href{https://github.com/TGuillerme/dispRity/}{\texttt{README}
page} of the package with the nice badges displaying these different
(and constantly tested) information

\begin{itemize}
\tightlist
\item
  The build status:
  \href{https://travis-ci.org/TGuillerme/dispRity}{\includegraphics{https://travis-ci.org/TGuillerme/dispRity.svg?branch=master}}
  which tests whether the package compiles (i.e.~can be installed on
  your computer).
\item
  The code coverage:
  \href{https://codecov.io/gh/TGuillerme/dispRity}{\includegraphics{https://codecov.io/gh/TGuillerme/dispRity/branch/master/graph/badge.svg}}
  which indicates the percentage of lines in the code that have been
  tested for expected behaviour.
\item
  The project status:
  \href{http://www.repostatus.org/\#active}{\includegraphics{http://www.repostatus.org/badges/latest/active.svg}}
  which indicates whether this package is maintained regularly or
  forgotten in the abyss of internet.
\item
  The development version:
  \href{https://github.com/TGuillerme/dispRity}{\includegraphics{https://img.shields.io/badge/devel\%20version-1.2.1-green.svg?style=flat}}
  which indicates which version of the package you are about to install.
\end{itemize}

Note that when we (the \texttt{dispRity} package collaborators!) are
developing some more major versions, including some more complex
features for example, they will never be released on either the the
versions described above. Therefore you can rest assure that you will
never use a badly bugged version of \texttt{dispRity}! Enjoy!

\section{Help}\label{help}

If you need help with the package, hopefully the following manual will
be useful. However, parts of this package are still in development and
some other parts are probably not covered. Thus if you have suggestions
or comments on on what has already been developed or will be developed,
please send me an email
(\href{mailto:guillert@tcd.ie}{\nolinkurl{guillert@tcd.ie}}) or if you
are a GitHub user, directly create an issue on the
\href{https://github.com/TGuillerme/dispRity}{GitHub page}.

\section{Citations}\label{citations}

To cite the package, this manual or some specific functionalities, you
can use the following references:

The package main paper:

\begin{quote}
Guillerme T. dispRity: A modular R package for measuring disparity.
Methods Ecol Evol. 2018;9:1755--1763.
\href{https://doi.org/10.1111/2041-210X.13022}{doi.org/10.1111/2041-210X.13022}.
\end{quote}

The package manual (regularly updated!):

\begin{quote}
Guillerme, T. \& Cooper, N. (2018): dispRity manual. figshare. Preprint.
\href{https://doi.org/10.6084/m9.figshare.6187337.v1}{10.6084/m9.figshare.6187337.v1}.
\end{quote}

The time-slicing method implemented in \texttt{chrono.subsets}
(unfortunately not Open Access, but you can still get a
\href{https://tguillerme.github.io/papers/Guillerme\&Cooper-2018-Palaeontology.pdf}{free
copy from here}):

\begin{quote}
Guillerme, T. and Cooper, N. (2018), Time for a rethink: time
sub-sampling methods in disparity-through-time analyses. Palaeontology,
61: 481-493.
\href{https://onlinelibrary.wiley.com/doi/abs/10.1111/pala.12364}{doi:10.1111/pala.12364}.
\end{quote}

\chapter{Glossary}\label{glossary}

\begin{itemize}
\item
  \textbf{Multidimensional space} (or just space). The mathematical
  multidimensional object that will be analysed with this package. In
  morphometrics, this is often referred to as the morphospace. However
  it may also be referred to as the cladisto-space for cladistic data or
  the eco-space for ecological data etc. In practice, this term
  designates a matrix where the columns represent the dimensions of the
  space (often -- but not necessarily - \textgreater{} 3!) and the rows
  represent the elements within this space.
\item
  \textbf{Elements}. The rows of the multidimensional space matrix.
  Elements can be taxa, field sites, countries etc.
\item
  \textbf{Dimensions}. The columns of the multidimensional space matrix.
  The dimensions can be referred to as axes of variation, or principal
  components, for ordinated spaces obtained from a PCA for example.
\item
  \textbf{Subsets}. Subsets of the multidimensional space. A subset (or
  subsets) contains the same number of dimensions as the space but may
  contain a smaller subset of elements. For example, if our space is
  composed of birds and mammals (the elements) and 50 principal
  components of variation (the dimensions), we can create two subsets
  containing just mammals or birds, but with the same 50 dimensions, to
  compare disparity in the two clades.
\item
  \textbf{Disparity}. A metric expressing the
  similarities/dissimilarities of the elements within the space or a
  summarising the space dimensions. For example the pairwise distances
  between elements or the range of each dimensions.
\end{itemize}

\section{Glossary equivalences in palaeobiology and
ecology}\label{glossary-equivalences-in-palaeobiology-and-ecology}

\begin{longtable}[]{@{}llll@{}}
\toprule
\begin{minipage}[b]{0.20\columnwidth}\raggedright\strut
In this manual\strut
\end{minipage} & \begin{minipage}[b]{0.20\columnwidth}\raggedright\strut
In \texttt{dispRity}\strut
\end{minipage} & \begin{minipage}[b]{0.23\columnwidth}\raggedright\strut
E.g. in palaeobiology\strut
\end{minipage} & \begin{minipage}[b]{0.17\columnwidth}\raggedright\strut
E.g. in ecology\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
the multidimensional space\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
a \texttt{matrix} object (\(n\times k\))\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
a morphospace\strut
\end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
a function-space\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
elements\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
rows (\(n\))\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
taxa\strut
\end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
field experiments\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
dimensions\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
columns (\(k\))\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
morphological characters\strut
\end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
communities' compositions\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
subsets\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
a \texttt{matrix} (\(m \times k\), with \(m \leq n\))\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
time series\strut
\end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
experimental treatments\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
disparity\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
a \texttt{function}\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
sum of variances\strut
\end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright\strut
ellipsoid volume\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\chapter{\texorpdfstring{Getting started with
\texttt{dispRity}}{Getting started with dispRity}}\label{getting-started-with-disprity}

\section{\texorpdfstring{What sort of data does \texttt{dispRity} work
with?}{What sort of data does dispRity work with?}}\label{what-sort-of-data-does-disprity-work-with}

\textbf{Any matrix object in \texttt{R}.} Disparity can be estimated
from pretty much any matrix as long as rows represent the elements and
columns the dimensions. These matrices can be observations, pairwise
differences between elements, ordinations, etc\ldots{}

\begin{quote}
Note that whether the matrix is an ordinated one or of any other type,
it is always considered as the multidimensional space to analysis and no
corrections are applied to it (e.g.~if ordinated and with negative eigen
values, no correction is used). Always make sure the input is the right
multidimensional space to analyse!
\end{quote}

\section{Ordinated matrices}\label{ordinated-matrices}

Classically, when a high number of variables is used, disparity is
calculated from ordinated matrices. These can be any type of ordinations
(PCO, PCA, PCoA, MDS, etc.) as long as elements are the rows (taxa,
countries, field experiments) and the dimensions are the columns.

\subsection{\texorpdfstring{Ordination matrices from
\texttt{geomorph}}{Ordination matrices from geomorph}}\label{ordination-matrices-from-geomorph}

You can also easily use data from \texttt{geomorph} using the
\texttt{geomorph.ordination} function. This function simply takes
Procrustes aligned data and performs an ordination:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{require}\NormalTok{(geomorph)}

\NormalTok{## Loading the plethodon dataset}
\KeywordTok{data}\NormalTok{(plethodon)}

\NormalTok{## Performing a Procrustes transform on the landmarks}
\NormalTok{procrustes <-}\StringTok{ }\KeywordTok{gpagen}\NormalTok{(plethodon}\OperatorTok{$}\NormalTok{land, }\DataTypeTok{PrinAxes =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{print.progress =} \OtherTok{FALSE}\NormalTok{)}

\NormalTok{## Ordinating this data}
\KeywordTok{geomorph.ordination}\NormalTok{(procrustes)[}\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{,}\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                PC1        PC2           PC3          PC4          PC5
## [1,] -0.0369931363 0.05118247 -0.0016971082 -0.003128809 -0.010936371
## [2,] -0.0007493738 0.05942082  0.0001371715 -0.002768680 -0.008117383
## [3,]  0.0056004654 0.07419599 -0.0052612103 -0.005034566 -0.002746592
## [4,] -0.0134808572 0.06463959 -0.0458436015 -0.007887369  0.009816827
## [5,] -0.0334696244 0.06863518  0.0136292041  0.007359409  0.022347225
\end{verbatim}

Options for the ordination (from \texttt{?prcomp}) can be directly
passed to this function to perform customised ordinations. Additionally
you can give the function a \texttt{geomorph.data.frame} object. If the
latter contains sorting information (i.e.~factors), they can be directly
used to make a customised \texttt{dispRity} object
\protect\hyperlink{customised-subsets}{customised \texttt{dispRity}
object}!

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Using a geomorph.data.frame}
\NormalTok{geomorph_df <-}\StringTok{ }\KeywordTok{geomorph.data.frame}\NormalTok{(procrustes,}
     \DataTypeTok{species =}\NormalTok{ plethodon}\OperatorTok{$}\NormalTok{species, }\DataTypeTok{site =}\NormalTok{ plethodon}\OperatorTok{$}\NormalTok{site)}

\NormalTok{## Ordinating this data and making a dispRity object}
\KeywordTok{geomorph.ordination}\NormalTok{(geomorph_df)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 4 customised subsets for 40 elements:
##     species.Jord, species.Teyah, site.Allo, site.Symp.
\end{verbatim}

More about these \texttt{dispRity} objects below!

\hypertarget{Claddis-ordination}{\subsection{\texorpdfstring{Ordination
matrices from
\texttt{Claddis}}{Ordination matrices from Claddis}}\label{Claddis-ordination}}

\begin{quote}
This functionality is not available on the CRAN version of the package
(1.2) due to CRAN compatibility issues see \protect\hyperlink{noCRAN}{to
CRAN or not to CRAN?} for more information and on how to install the
GitHub (more complete) version of the package (1.2).
\end{quote}

\texttt{dispRity} package can easily take data from \texttt{Claddis}
using the \texttt{Claddis.ordination} function. For this, simply input a
matrix in the \texttt{Claddis} format to the function and it will
automatically calculate and ordinate the distances among taxa:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{require}\NormalTok{(Claddis)}

\NormalTok{## Ordinating the example data from Claddis}
\KeywordTok{Claddis.ordination}\NormalTok{(Michaux1989) }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                      [,1]          [,2]       [,3]
## Ancilla      0.000000e+00  4.154578e-01  0.2534942
## Turrancilla -5.106645e-01  1.956921e-16 -0.2534942
## Ancillista   5.106645e-01  6.523071e-17 -0.2534942
## Amalda      -2.004476e-16 -4.154578e-01  0.2534942
\end{verbatim}

Note that several options are available, namely which type of distance
should be computed. See more info in the function manual
(\texttt{?Claddis.ordination}). Alternatively, it is of course also
possible to manual calculate the ordination matrix using the functions
\texttt{Claddis::MorphDistMatrix} and \texttt{stats::cmdscale}.

\subsection{Other kinds of ordination
matrices}\label{other-kinds-of-ordination-matrices}

If you are not using the packages mentioned above (\texttt{Claddis} and
\texttt{geomorph}) you can easily make your own ordination matrices by
using the following functions from the \texttt{stats} package. Here is
how to do it for the following types of matrices:

\begin{itemize}
\tightlist
\item
  Multivariate matrices (principal components analysis; PCA)
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## A multivariate matrix}
\KeywordTok{head}\NormalTok{(USArrests)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##            Murder Assault UrbanPop Rape
## Alabama      13.2     236       58 21.2
## Alaska       10.0     263       48 44.5
## Arizona       8.1     294       80 31.0
## Arkansas      8.8     190       50 19.5
## California    9.0     276       91 40.6
## Colorado      7.9     204       78 38.7
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Ordinating the matrix using `prcomp` }
\NormalTok{ordination <-}\StringTok{ }\KeywordTok{prcomp}\NormalTok{(USArrests)}

\NormalTok{## Selecting the ordinated matrix}
\NormalTok{ordinated_matrix <-}\StringTok{ }\NormalTok{ordination}\OperatorTok{$}\NormalTok{x}
\KeywordTok{head}\NormalTok{(ordinated_matrix)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                  PC1        PC2        PC3        PC4
## Alabama     64.80216 -11.448007 -2.4949328 -2.4079009
## Alaska      92.82745 -17.982943 20.1265749  4.0940470
## Arizona    124.06822   8.830403 -1.6874484  4.3536852
## Arkansas    18.34004 -16.703911  0.2101894  0.5209936
## California 107.42295  22.520070  6.7458730  2.8118259
## Colorado    34.97599  13.719584 12.2793628  1.7214637
\end{verbatim}

This results in a ordinated matrix with US states as elements and four
dimensions (PC 1 to 4). For an alternative method, see the
\texttt{?princomp} function.

\begin{itemize}
\tightlist
\item
  Distance matrices (classical multidimensional scaling; MDS)
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## A matrix of distances between cities}
\KeywordTok{str}\NormalTok{(eurodist)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  'dist' num [1:210] 3313 2963 3175 3339 2762 ...
##  - attr(*, "Size")= num 21
##  - attr(*, "Labels")= chr [1:21] "Athens" "Barcelona" "Brussels" "Calais" ...
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Ordinating the matrix using cmdscale() with k = 5 dimensions }
\NormalTok{ordinated_matrix <-}\StringTok{ }\KeywordTok{cmdscale}\NormalTok{(eurodist, }\DataTypeTok{k =} \DecValTok{5}\NormalTok{)}
\KeywordTok{head}\NormalTok{(ordinated_matrix)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                 [,1]      [,2]       [,3]       [,4]       [,5]
## Athens    2290.27468 1798.8029   53.79314 -103.82696 -156.95511
## Barcelona -825.38279  546.8115 -113.85842   84.58583  291.44076
## Brussels    59.18334 -367.0814  177.55291   38.79751  -95.62045
## Calais     -82.84597 -429.9147  300.19274  106.35369 -180.44614
## Cherbourg -352.49943 -290.9084  457.35294  111.44915 -417.49668
## Cologne    293.68963 -405.3119  360.09323 -636.20238  159.39266
\end{verbatim}

This results in a ordinated matrix with European cities as elements and
five dimensions.

Of course any other method for creating the ordination matrix is totally
valid, you can also not use any ordination at all! The only requirements
for the \texttt{dispRity} functions is that the input is a matrix with
elements as rows and dimensions as columns.

\section{Performing a simple dispRity analysis}\label{simpleanalysis}

Two \texttt{dispRity} functions allow users to run an analysis pipeline
simply by inputting an ordination matrix. These functions allow users to
either calculate the disparity through time
(\texttt{dispRity.through.time}) or the disparity of user-defined groups
(\texttt{dispRity.per.group}).

\textbf{IMPORTANT}

Note that \texttt{disparity.through.time} and
\texttt{disparity.per.group} are wrapper functions (i.e.~they
incorporate lots of other functions) that allow users to run a basic
disparity-through-time, or disparity among groups, analysis without too
much effort. As such they use a lot of default options. These are
described in the help files for the functions that are used to make the
wrapper functions, and not described in the help files for
\texttt{disparity.through.time} and \texttt{disparity.per.group}. These
defaults are good enough for \textbf{data exploration}, but for a proper
analysis you should consider the \textbf{best parameters for your
question and data}. For example, which metric should you use? How many
bootstraps do you require? What model of evolution is most appropriate
if you are time slicing? Should you rarefy the data? See
\protect\hyperlink{time-slicing}{\texttt{chrono.subsets}},
\protect\hyperlink{customised-subsets}{\texttt{custom.subsets}},
\protect\hyperlink{bootstraps-and-rarefactions}{\texttt{boot.matrix}}
and \protect\hyperlink{disparity-metrics}{\texttt{dispRity.metric}} for
more details of the defaults used in each of these functions. Note that
any of these default arguments can be changed within the
\texttt{disparity.through.time} or \texttt{disparity.per.group}
functions.

\hypertarget{example-data}{\subsection{Example
data}\label{example-data}}

To illustrate these functions, we will use data from
\citet{beckancient2014}. This dataset contains an ordinated matrix of 50
discrete characters from mammals (\texttt{BeckLee\_mat50}), another
matrix of the same 50 mammals and the estimated discrete data characters
of their descendants (thus 50 + 49 rows, \texttt{BeckLee\_mat99}), a
dataframe containing the ages of each taxon in the dataset
(\texttt{BeckLee\_ages}) and finally a phylogenetic tree with the
relationships among the 50 mammals (\texttt{BeckLee\_tree}).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Loading the ordinated matrices}
\KeywordTok{data}\NormalTok{(BeckLee_mat50)}
\KeywordTok{data}\NormalTok{(BeckLee_mat99)}

\NormalTok{## The first five taxa and dimensions of the 50 taxa matrix}
\KeywordTok{head}\NormalTok{(BeckLee_mat50[, }\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{])}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                    [,1]          [,2]        [,3]       [,4]      [,5]
## Cimolestes   -0.5319679  0.1117759259  0.09865194 -0.1933148 0.2035833
## Maelestes    -0.4087147  0.0139690317  0.26268300  0.2297096 0.1310953
## Batodon      -0.6923194  0.3308625215 -0.10175223 -0.1899656 0.1003108
## Bulaklestes  -0.6802291 -0.0134872777  0.11018009 -0.4103588 0.4326298
## Daulestes    -0.7386111  0.0009001369  0.12006449 -0.4978191 0.4741342
## Uchkudukodon -0.5105254 -0.2420633915  0.44170317 -0.1172972 0.3602273
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The first five taxa and dimensions of the 99 taxa + ancestors matrix}
\NormalTok{BeckLee_mat99[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{98}\NormalTok{, }\DecValTok{99}\NormalTok{), }\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                   [,1]       [,2]       [,3]        [,4]        [,5]
## Cimolestes -0.60824375 -0.0323683 0.08458885 -0.43384481 -0.30536875
## Maelestes  -0.57302058 -0.2840361 0.01308847 -0.12588477  0.06123611
## n48        -0.05529018  0.4799330 0.04118477  0.04944912 -0.35588301
## n49        -0.13067785  0.4478168 0.11956268  0.13800340 -0.32227852
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Loading a list of first and last occurrence dates for the fossils}
\KeywordTok{data}\NormalTok{(BeckLee_ages)}
\KeywordTok{head}\NormalTok{(BeckLee_ages)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##             FAD  LAD
## Adapis     37.2 36.8
## Asioryctes 83.6 72.1
## Leptictis  33.9 33.3
## Miacis     49.0 46.7
## Mimotona   61.6 59.2
## Notharctus 50.2 47.0
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Loading and plotting the phylogeny}
\KeywordTok{data}\NormalTok{(BeckLee_tree)}
\KeywordTok{plot}\NormalTok{(BeckLee_tree, }\DataTypeTok{cex =} \FloatTok{0.8}\NormalTok{) }
\KeywordTok{axisPhylo}\NormalTok{(}\DataTypeTok{root =} \DecValTok{140}\NormalTok{)}
\KeywordTok{nodelabels}\NormalTok{(}\DataTypeTok{cex =} \FloatTok{0.5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-12-1.pdf}

Of course you can use your own data as detailed in the
\protect\hyperlink{What-sort-of-data-does-dispRity-work-with}{previous
chapter}.

\subsection{Disparity through time}\label{disparity-through-time}

The \texttt{dispRity.through.time} function calculates disparity through
time, a common analysis in palaeontology. This function (and the
following one) uses an analysis pipeline with a lot of default
parameters to make the analysis as simple as possible. Of course all the
defaults can be changed if required, more on this later.

For a disparity through time analysis, you will need:

\begin{itemize}
\tightlist
\item
  An ordinated matrix (we covered that above)
\item
  A phylogenetic tree: this must be a \texttt{phylo} object (from the
  \texttt{ape} package) and needs a \texttt{root.time} element. To give
  your tree a root time (i.e.~an age for the root), you can simply
  do\textbackslash{}
  \texttt{my\_tree\$root.time\ \textless{}-\ my\_age}.
\item
  The required number of time subsets (here \texttt{time\ =\ 3})
\item
  Your favourite disparity metric (here the sum of variances)
\end{itemize}

Using the Beck and Lee (2014) data described
\protect\hyperlink{example-data}{above}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Measuring disparity through time}
\NormalTok{disparity_data <-}\StringTok{ }\KeywordTok{dispRity.through.time}\NormalTok{(BeckLee_mat50, BeckLee_tree,}
                                        \DataTypeTok{time =} \DecValTok{3}\NormalTok{, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, variances))}
\end{Highlighting}
\end{Shaded}

This generates a \texttt{dispRity} object (see
\protect\hyperlink{guts}{here} for technical details). When displayed,
these \texttt{dispRity} objects provide us with information on the
operations done to the matrix:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Print the disparity_data object}
\NormalTok{disparity_data}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 3 discrete time subsets for 50 elements with 48 dimensions:
##     133.51104 - 89.00736, 89.00736 - 44.50368, 44.50368 - 0.
## Data was bootstrapped 100 times (method:"full").
## Disparity was calculated as: metric.
\end{verbatim}

We asked for three subsets (evenly spread across the age of the tree),
the data was bootstrapped 100 times (default) and the metric used was
the sum of variances.

We can now summarise or plot the \texttt{disparity\_data} object, or
perform statistical tests on it (e.g.~a simple \texttt{lm}):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Summarising disparity through time}
\KeywordTok{summary}\NormalTok{(disparity_data)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                subsets  n   obs bs.median  2.5%   25%   75% 97.5%
## 1 133.51104 - 89.00736  5 1.575     1.305 0.729 1.118 1.420 1.509
## 2  89.00736 - 44.50368 29 1.922     1.867 1.775 1.830 1.889 1.922
## 3         44.50368 - 0 16 1.990     1.871 1.716 1.831 1.914 1.942
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Plotting the results}
\KeywordTok{plot}\NormalTok{(disparity_data, }\DataTypeTok{type =} \StringTok{"continuous"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-15-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Testing for an difference among the time bins}
\NormalTok{disp_lm <-}\StringTok{ }\KeywordTok{test.dispRity}\NormalTok{(disparity_data, }\DataTypeTok{test =}\NormalTok{ lm, }\DataTypeTok{comparisons =} \StringTok{"all"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in test.dispRity(disparity_data, test = lm, comparisons = "all"): Multiple p-values will be calculated without adjustment!
## This can inflate Type I error!
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(disp_lm)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## test(formula = data ~ subsets, data = data)
## 
## Residuals:
##      Min       1Q   Median       3Q      Max 
## -0.56623 -0.04160  0.01049  0.05507  0.31886 
## 
## Coefficients:
##                            Estimate Std. Error t value Pr(>|t|)    
## (Intercept)                 1.25647    0.01270   98.97   <2e-16 ***
## subsets44.50368 - 0         0.60863    0.01795   33.90   <2e-16 ***
## subsets89.00736 - 44.50368  0.60169    0.01795   33.51   <2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 0.127 on 297 degrees of freedom
## Multiple R-squared:  0.8361, Adjusted R-squared:  0.835 
## F-statistic: 757.5 on 2 and 297 DF,  p-value: < 2.2e-16
\end{verbatim}

Please refer to the \protect\hyperlink{specific-tutorial}{specific
tutorials} for (much!) more information on the nuts and bolts of the
package. You can also directly explore the specific function help files
within R and navigate to related functions.

\hypertarget{disparity-among-groups}{\subsection{Disparity among
groups}\label{disparity-among-groups}}

The \texttt{dispRity.per.group} function is used if you are interested
in looking at disparity among groups rather than through time. For
example, you could ask if there is a difference in disparity between two
groups?

To perform such an analysis, you will need:

\begin{itemize}
\tightlist
\item
  An matrix with rows as elements and columns as dimensions (always!)
\item
  A list of group members: this list should be a list of numeric vectors
  or names corresponding to the row names in the matrix. For example
  \texttt{list("a"\ =\ c(1,2),\ "b"\ =\ c(3,4))} will create a group
  \emph{a} containing elements 1 and 2 from the matrix and a group
  \emph{b} containing elements 3 and 4. Note that elements can be
  present in multiple groups at once.
\item
  Your favourite disparity metric (here the sum of variances)
\end{itemize}

Using the \citet{beckancient2014} data described
\protect\hyperlink{example-data}{above}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating the two groups (crown versus stem) as a list}
\NormalTok{mammal_groups <-}\StringTok{ }\KeywordTok{crown.stem}\NormalTok{(BeckLee_tree, }\DataTypeTok{inc.nodes =} \OtherTok{FALSE}\NormalTok{)}

\NormalTok{## Measuring disparity for each group}
\NormalTok{disparity_data <-}\StringTok{ }\KeywordTok{dispRity.per.group}\NormalTok{(BeckLee_mat50, }\DataTypeTok{group =}\NormalTok{ mammal_groups,}
                                     \DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, variances))}
\end{Highlighting}
\end{Shaded}

We can display the disparity of both groups by simply looking at the
output variable (\texttt{disparity\_data}) and then summarising the
\texttt{disparity\_data} object and plotting it, and/or by performing a
statistical test to compare disparity across the groups (here a Wilcoxon
test).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Print the disparity_data object}
\NormalTok{disparity_data}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 2 customised subsets for 50 elements with 48 dimensions:
##     crown, stem.
## Data was bootstrapped 100 times (method:"full").
## Disparity was calculated as: metric.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Summarising disparity in the different groups}
\KeywordTok{summary}\NormalTok{(disparity_data)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs bs.median  2.5%   25%   75% 97.5%
## 1   crown 30 1.995     1.933 1.883 1.917 1.951 1.983
## 2    stem 20 1.715     1.635 1.535 1.610 1.664 1.691
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Plotting the results}
\KeywordTok{plot}\NormalTok{(disparity_data)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-17-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Testing for a difference between the groups}
\KeywordTok{test.dispRity}\NormalTok{(disparity_data, }\DataTypeTok{test =}\NormalTok{ wilcox.test, }\DataTypeTok{details =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## $`crown : stem`
## $`crown : stem`[[1]]
## 
##  Wilcoxon rank sum test with continuity correction
## 
## data:  dots[[1L]][[1L]] and dots[[2L]][[1L]]
## W = 10000, p-value < 2.2e-16
## alternative hypothesis: true location shift is not equal to 0
\end{verbatim}

\chapter{Details of specific
functions}\label{details-of-specific-functions}

The following section contains information specific to some functions.
If any of your questions are not covered in these sections, please refer
to the function help files in R, send me an email
(\href{mailto:guillert@tcd.ie}{\nolinkurl{guillert@tcd.ie}}), or raise
an issue on GitHub. The several tutorials below describe specific
functionalities of certain functions; please always refer to the
function help files for the full function documentation!

Before each section, make sure you loaded the \citet{beckancient2014}
data (see \protect\hyperlink{example-data}{example data} for more
details).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Loading the data}
\KeywordTok{data}\NormalTok{(BeckLee_mat50)}
\KeywordTok{data}\NormalTok{(BeckLee_mat99)}
\KeywordTok{data}\NormalTok{(BeckLee_tree)}
\KeywordTok{data}\NormalTok{(BeckLee_ages)}
\end{Highlighting}
\end{Shaded}

\hypertarget{time-slicing}{\section{Time slicing}\label{time-slicing}}

The function \texttt{chrono.subsets} allows users to divide the matrix
into different time subsets or slices given a dated phylogeny that
contains all the elements (i.e.~taxa) from the matrix. Each subset
generated by this function will then contain all the elements present at
a specific point in time or during a specific period in time.

Two types of time subsets can be performed by using the \texttt{method}
option:

\begin{itemize}
\tightlist
\item
  Discrete time subsets (or time-binning) using
  \texttt{method\ =\ discrete}
\item
  Continuous time subsets (or time-slicing) using
  \texttt{method\ =\ continuous}
\end{itemize}

For the time-slicing method details see \citet{time-slice}. For both
methods, the function takes the \texttt{time} argument which can be a
vector of \texttt{numeric} values for:

\begin{itemize}
\tightlist
\item
  Defining the boundaries of the time bins (when
  \texttt{method\ =\ discrete})
\item
  Defining the time slices (when \texttt{method\ =\ continuous})
\end{itemize}

Otherwise, the \texttt{time} argument can be set as a single
\texttt{numeric} value for automatically generating a given number of
equidistant time-bins/slices. Additionally, it is also possible to input
a dataframe containing the first and last occurrence data (FAD/LAD) for
taxa that span over a longer time than the given tips/nodes age, so taxa
can appear in more than one time bin/slice.

Here is an example for \texttt{method\ =\ discrete}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Generating three time bins containing the taxa present every 40 Ma}
\KeywordTok{chrono.subsets}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat50, }\DataTypeTok{tree =}\NormalTok{ BeckLee_tree, }\DataTypeTok{method =} \StringTok{"discrete"}\NormalTok{,}
                \DataTypeTok{time =} \KeywordTok{c}\NormalTok{(}\DecValTok{120}\NormalTok{, }\DecValTok{80}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{0}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 3 discrete time subsets for 50 elements:
##     120 - 80, 80 - 40, 40 - 0.
\end{verbatim}

Note that we can also generate equivalent results by just telling the
function that we want three time-bins as follow:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Automatically generate three equal length bins:}
\KeywordTok{chrono.subsets}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat50, }\DataTypeTok{tree =}\NormalTok{ BeckLee_tree, }\DataTypeTok{method =} \StringTok{"discrete"}\NormalTok{,}
                \DataTypeTok{time =} \DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 3 discrete time subsets for 50 elements:
##     133.51104 - 89.00736, 89.00736 - 44.50368, 44.50368 - 0.
\end{verbatim}

In this example, the taxa were split inside each time-bin according to
their age. However, the taxa here are considered as single points in
time. It is totally possible that some taxa could have had longer
longevity and that they exist in multiple time bins. In this case, it is
possible to include them in more than one bin by providing a table of
first and last occurrence dates (FAD/LAD). This table should have the
taxa names as row names and two columns for respectively the first and
last occurrence age:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Displaying the table of first and last occurrence dates for each taxa}
\KeywordTok{head}\NormalTok{(BeckLee_ages)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##             FAD  LAD
## Adapis     37.2 36.8
## Asioryctes 83.6 72.1
## Leptictis  33.9 33.3
## Miacis     49.0 46.7
## Mimotona   61.6 59.2
## Notharctus 50.2 47.0
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Generating time bins including taxa that might span between them}
\KeywordTok{chrono.subsets}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat50, }\DataTypeTok{tree =}\NormalTok{ BeckLee_tree, }\DataTypeTok{method =} \StringTok{"discrete"}\NormalTok{,}
                \DataTypeTok{time =} \KeywordTok{c}\NormalTok{(}\DecValTok{120}\NormalTok{, }\DecValTok{80}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{FADLAD =}\NormalTok{ BeckLee_ages)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 3 discrete time subsets for 50 elements:
##     120 - 80, 80 - 40, 40 - 0.
\end{verbatim}

When using this method, the oldest boundary of the first bin (or the
first slice, see below) is automatically generated as the root age plus
1\% of the tree length, as long as at least three elements/taxa are
present at that point in time. The algorithm adds an extra 1\% tree
length until reaching the required minimum of three elements. It is also
possible to include nodes in each bin by using
\texttt{inc.nodes\ =\ TRUE} and providing a matrix that contains the
ordinated distance among tips \emph{and} nodes.

For the time-slicing method (\texttt{method\ =\ continuous}), the idea
is fairly similar. This option, however, requires a matrix that contains
the ordinated distance among taxa \emph{and} nodes and an extra argument
describing the assumed evolutionary model (via the \texttt{model}
argument). This model argument is used when the time slice occurs along
a branch of the tree rather than on a tip or a node, meaning that a
decision must be made about what the value for the branch should be. The
model can be one of the following:

\begin{itemize}
\item
  \textbf{Punctuated models}
\item
  \texttt{acctran} where the data chosen along the branch is always the
  one of the descendant
\item
  \texttt{deltran} where the data chosen along the branch is always the
  one of the ancestor
\item
  \texttt{random} where the data chosen along the branch is randomly
  chosen between the descendant or the ancestor
\item
  \texttt{proximity} where the data chosen along the branch is either
  the descendant or the ancestor depending on branch length
\item
  \textbf{Gradual models}
\item
  \texttt{equal.split} where the data chosen along the branch is both
  the descendant and the ancestor with an even probability
\item
  \texttt{gradual.split} where the data chosen along the branch is both
  the descendant and the ancestor with a probability depending on branch
  length
\end{itemize}

\begin{quote}
Note that the four first models are a proxy for punctuated evolution:
the selected data is always either the one of the descendant or the
ancestor. In other words, changes along the branches always occur at
either ends of it. The two last models are a proxy for gradual
evolution: the data from both the descendant and the ancestor is used
with an associate probability. These later models perform better when
bootstrapped, effectively approximating the ``intermediate'' state
between and the ancestor and the descendants.
\end{quote}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Generating four time slices every 40 million years under a model of proximity evolution}
\KeywordTok{chrono.subsets}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat99, }\DataTypeTok{tree =}\NormalTok{ BeckLee_tree, }
    \DataTypeTok{method =} \StringTok{"continuous"}\NormalTok{, }\DataTypeTok{model =} \StringTok{"proximity"}\NormalTok{, }\DataTypeTok{time =} \KeywordTok{c}\NormalTok{(}\DecValTok{120}\NormalTok{, }\DecValTok{80}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{0}\NormalTok{),}
    \DataTypeTok{FADLAD =}\NormalTok{ BeckLee_ages)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 4 continuous (proximity) time subsets for 99 elements:
##     120, 80, 40, 0.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Generating four time slices automatically}
\KeywordTok{chrono.subsets}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat99, }\DataTypeTok{tree =}\NormalTok{ BeckLee_tree,}
    \DataTypeTok{method =} \StringTok{"continuous"}\NormalTok{, }\DataTypeTok{model =} \StringTok{"proximity"}\NormalTok{, }\DataTypeTok{time =} \DecValTok{4}\NormalTok{, }\DataTypeTok{FADLAD =}\NormalTok{ BeckLee_ages)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 4 continuous (proximity) time subsets for 99 elements:
##     133.51104, 89.00736, 44.50368, 0.
\end{verbatim}

If you want to generate time subsets based on stratigraphy, the package
proposes a useful functions to do it for you: \texttt{get.bin.ages}
(check out the function's manual in \texttt{R})!

\hypertarget{customised-subsets}{\section{Customised
subsets}\label{customised-subsets}}

Another way of separating elements into different categories is to use
customised subsets as briefly explained
\protect\hyperlink{disparity-among-groups}{above}. This function simply
takes the list of elements to put in each group (whether they are the
actual element names or their position in the matrix).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating the two groups (crown and stems)}
\NormalTok{mammal_groups <-}\StringTok{ }\KeywordTok{crown.stem}\NormalTok{(BeckLee_tree, }\DataTypeTok{inc.nodes =} \OtherTok{FALSE}\NormalTok{)}

\NormalTok{## Separating the dataset into two different groups}
\KeywordTok{custom.subsets}\NormalTok{(BeckLee_mat50, }\DataTypeTok{group =}\NormalTok{ mammal_groups)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 2 customised subsets for 50 elements:
##     crown, stem.
\end{verbatim}

Like in this example, you can use the utility function
\texttt{crown.stem} that allows to automatically separate the crown and
stems taxa given a phylogenetic tree. Also, elements can easily be
assigned to different groups if necessary!

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating the three groups as a list}
\NormalTok{weird_groups <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}\StringTok{"even"}\NormalTok{ =}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{1}\NormalTok{, }\DataTypeTok{to =} \DecValTok{49}\NormalTok{, }\DataTypeTok{by =} \DecValTok{2}\NormalTok{),}
                      \StringTok{"odd"}\NormalTok{ =}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{2}\NormalTok{, }\DataTypeTok{to =} \DecValTok{50}\NormalTok{, }\DataTypeTok{by =} \DecValTok{2}\NormalTok{),}
                      \StringTok{"all"}\NormalTok{ =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{50}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

The \texttt{custom.subsets} function can also take a phylogeny (as a
\texttt{phylo} object) as an argument to create groups as clades:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating groups as clades}
\KeywordTok{custom.subsets}\NormalTok{(BeckLee_mat50, }\DataTypeTok{group =}\NormalTok{ BeckLee_tree)}
\end{Highlighting}
\end{Shaded}

This automatically creates 49 (the number of nodes) groups containing
between two and 50 (the number of tips) elements.

\hypertarget{bootstraps-and-rarefactions}{\section{Bootstraps and
rarefactions}\label{bootstraps-and-rarefactions}}

One important step in analysing ordinated matrices is to
pseudo-replicate the data to see how robust the results are, and how
sensitive they are to outliers in the dataset. This can be achieved
using the function \texttt{boot.matrix} to bootstrap and/or rarefy the
data. The default options will bootstrap the matrix 100 times without
rarefaction using the ``full'' bootstrap method (see below):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Default bootstrapping}
\KeywordTok{boot.matrix}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat50)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 50 elements with 48 dimensions.
## Data was bootstrapped 100 times (method:"full").
\end{verbatim}

The number of bootstrap replicates can be defined using the
\texttt{bootstraps} option. The method can be modified by controlling
which bootstrap algorithm to use through the \texttt{boot.type}
argument. Currently two algorithms are implemented:

\begin{itemize}
\tightlist
\item
  \texttt{full} where the bootstrapping is entirely stochastic (\emph{n}
  elements are replaced by any \emph{m} elements drawn from the data)
\item
  \texttt{single} where only one random element is replaced by one other
  random element for each pseudo-replicate
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Bootstrapping with the single bootstrap method}
\KeywordTok{boot.matrix}\NormalTok{(BeckLee_mat50, }\DataTypeTok{boot.type =} \StringTok{"single"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 50 elements with 48 dimensions.
## Data was bootstrapped 100 times (method:"single").
\end{verbatim}

This function also allows users to rarefy the data using the
\texttt{rarefaction} argument. Rarefaction allows users to limit the
number of elements to be drawn at each bootstrap replication. This is
useful if, for example, one is interested in looking at the effect of
reducing the number of elements on the results of an analysis.

This can be achieved by using the \texttt{rarefaction} option that draws
only \emph{n-x} at each bootstrap replicate (where \emph{x} is the
number of elements not sampled). The default argument is \texttt{FALSE}
but it can be set to \texttt{TRUE} to fully rarefy the data (i.e.~remove
\emph{x} elements for the number of pseudo-replicates, where \emph{x}
varies from the maximum number of elements present in each subset to a
minimum of three elements). It can also be set to one or more
\texttt{numeric} values to only rarefy to the corresponding number of
elements.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Bootstrapping with the full rarefaction}
\KeywordTok{boot.matrix}\NormalTok{(BeckLee_mat50, }\DataTypeTok{bootstraps =} \DecValTok{20}\NormalTok{, }\DataTypeTok{rarefaction =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 50 elements with 48 dimensions.
## Data was bootstrapped 20 times (method:"full") and fully rarefied.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Or with a set number of rarefaction levels}
\KeywordTok{boot.matrix}\NormalTok{(BeckLee_mat50, }\DataTypeTok{bootstraps =} \DecValTok{20}\NormalTok{, }\DataTypeTok{rarefaction =} \KeywordTok{c}\NormalTok{(}\DecValTok{6}\OperatorTok{:}\DecValTok{8}\NormalTok{, }\DecValTok{3}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 50 elements with 48 dimensions.
## Data was bootstrapped 20 times (method:"full") and rarefied to 6, 7, 8, 3 elements.
\end{verbatim}

One other argument is \texttt{dimensions} that specifies how many
dimensions from the matrix should be used for further analysis. When
missing, all dimensions from the ordinated matrix are used.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Using the first 50% of the dimensions}
\KeywordTok{boot.matrix}\NormalTok{(BeckLee_mat50, }\DataTypeTok{dimensions =} \FloatTok{0.5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 50 elements with 24 dimensions.
## Data was bootstrapped 100 times (method:"full").
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Using the first 10 dimensions}
\KeywordTok{boot.matrix}\NormalTok{(BeckLee_mat50, }\DataTypeTok{dimensions =} \DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 50 elements with 10 dimensions.
## Data was bootstrapped 100 times (method:"full").
\end{verbatim}

It is also possible to specify the sampling probability in the bootstrap
for each elements. This can be useful for weighting analysis for example
(i.e.~giving more importance to specific elements). These probabilities
can be passed to the \texttt{prob} argument individually with a vector
with the elements names or with a matrix with the rownames as elements
names. The elements with no specified probability will be assigned a
probability of 1 (or 1/maximum weight if the argument is weights rather
than probabilities).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Attributing a weight of 0 to Cimolestes and 10 to Maelestes}
\KeywordTok{boot.matrix}\NormalTok{(BeckLee_mat50, }\DataTypeTok{prob =} \KeywordTok{c}\NormalTok{(}\StringTok{"Cimolestes"}\NormalTok{ =}\StringTok{ }\DecValTok{0}\NormalTok{, }\StringTok{"Maelestes"}\NormalTok{ =}\StringTok{ }\DecValTok{10}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 50 elements with 48 dimensions.
## Data was bootstrapped 100 times (method:"full").
\end{verbatim}

Of course, one could directly supply the subsets generated above (using
\texttt{chrono.subsets} or \texttt{custom.subsets}) to this function.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating subsets of crown and stem mammals}
\NormalTok{crown_stem <-}\StringTok{ }\KeywordTok{custom.subsets}\NormalTok{(BeckLee_mat50,}
                                \DataTypeTok{group =} \KeywordTok{list}\NormalTok{(}\StringTok{"crown"}\NormalTok{ =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{16}\NormalTok{, }\DecValTok{19}\OperatorTok{:}\DecValTok{41}\NormalTok{, }\DecValTok{45}\OperatorTok{:}\DecValTok{50}\NormalTok{), }
                                             \StringTok{"stem"}\NormalTok{ =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{15}\NormalTok{, }\DecValTok{17}\OperatorTok{:}\DecValTok{18}\NormalTok{, }\DecValTok{42}\OperatorTok{:}\DecValTok{44}\NormalTok{)))}
\NormalTok{## Bootstrapping and rarefying these groups}
\KeywordTok{boot.matrix}\NormalTok{(crown_stem, }\DataTypeTok{bootstraps =} \DecValTok{200}\NormalTok{, }\DataTypeTok{rarefaction =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 2 customised subsets for 50 elements with 48 dimensions:
##     crown, stem.
## Data was bootstrapped 200 times (method:"full") and fully rarefied.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating time slice subsets}
\NormalTok{time_slices <-}\StringTok{ }\KeywordTok{chrono.subsets}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat99, }\DataTypeTok{tree =}\NormalTok{ BeckLee_tree, }
                               \DataTypeTok{method =} \StringTok{"continuous"}\NormalTok{, }\DataTypeTok{model =} \StringTok{"proximity"}\NormalTok{, }
                               \DataTypeTok{time =} \KeywordTok{c}\NormalTok{(}\DecValTok{120}\NormalTok{, }\DecValTok{80}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{0}\NormalTok{),}
                               \DataTypeTok{FADLAD =}\NormalTok{ BeckLee_ages)}

\NormalTok{## Bootstrapping the time slice subsets}
\KeywordTok{boot.matrix}\NormalTok{(time_slices, }\DataTypeTok{bootstraps =} \DecValTok{100}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 4 continuous (proximity) time subsets for 99 elements with 97 dimensions:
##     120, 80, 40, 0.
## Data was bootstrapped 100 times (method:"full").
\end{verbatim}

\hypertarget{disparity-metrics}{\section{Disparity
metrics}\label{disparity-metrics}}

There are many ways of measuring disparity! In brief, disparity is a
summary metric that will represent an aspect of an ordinated space
(e.g.~a MDS, PCA, PCO, PCoA). For example, one can look at ellipsoid
hyper-volume of the ordinated space (Donohue \emph{et al.} 2013), the
sum and the product of the ranges and variances (Wills \emph{et al.}
1994) or the median position of the elements relative to their centroid
(Wills \emph{et al.} 1994). Of course, there are many more examples of
metrics one can use for describing some aspect of the ordinated space,
with some performing better than other ones at particular descriptive
tasks, and some being more generalist.

Because of this great diversity of metrics, the package
\texttt{dispRity} does not have one way to measure disparity but rather
proposes to facilitate users in defining their own disparity metric that
will best suit their particular analysis. In fact, the core function of
the package, \texttt{dispRity}, allows the user to define any metric
with the \texttt{metric} argument. However the \texttt{metric} argument
has to follow certain rules:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  It must be composed from one to three \texttt{function} objects;
\item
  The function(s) must take as a first argument a \texttt{matrix} or a
  \texttt{vector};
\item
  The function(s) must be of one of the three dimension-levels described
  below;
\item
  At least one of the functions must be of dimension-level 1 or 2 (see
  below).
\end{enumerate}

\subsection{The function
dimension-levels}\label{the-function-dimension-levels}

The metric function dimension-levels determine the ``dimensionality of
decomposition'' of the input matrix. In other words, each
dimension-level designates the dimensions of the output, i.e.~either
three (a \texttt{matrix}); two (a \texttt{vector}); or one (a single
\texttt{numeric} value) dimension.

\begin{figure}
\centering
\includegraphics{dispRity_fun.png}
\caption{Illustration of the different dimension-levels of functions
with an input \texttt{matrix}}
\end{figure}

\subsubsection{Dimension-level 1
functions}\label{dimension-level-1-functions}

A dimension-level 1 function will decompose a \texttt{matrix} or a
\texttt{vector} into a single value:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating a dummy matrix}
\NormalTok{dummy_matrix <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\KeywordTok{rnorm}\NormalTok{(}\DecValTok{12}\NormalTok{), }\DecValTok{4}\NormalTok{, }\DecValTok{3}\NormalTok{)}

\NormalTok{## Example of dimension-level 1 functions}
\KeywordTok{mean}\NormalTok{(dummy_matrix)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 0.4719271
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{median}\NormalTok{(dummy_matrix)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 0.4816107
\end{verbatim}

Any summary metric such as mean or median are good examples of
dimension-level 1 functions as they reduce the matrix to a single
dimension (i.e.~one value).

\subsubsection{Dimension-level 2
functions}\label{dimension-level-2-functions}

A dimension-level 2 function will decompose a \texttt{matrix} into a
\texttt{vector}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Defining the function as the product of rows}
\NormalTok{prod.rows <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(matrix) }\KeywordTok{apply}\NormalTok{(matrix, }\DecValTok{1}\NormalTok{, prod)}

\NormalTok{## A dimension-level 2 metric}
\KeywordTok{prod.rows}\NormalTok{(dummy_matrix)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] -0.1693259  0.1151639  0.4023232  0.2070852
\end{verbatim}

Several dimension-level 2 functions are implemented in \texttt{dispRity}
(see \texttt{?dispRity.metric}) such as the \texttt{variances} or
\texttt{ranges} functions that calculate the variance or the range of
each dimension of the ordinated matrix respectively.

\subsubsection{Dimension-level 3
functions}\label{dimension-level-3-functions}

Finally a dimension-level 3 function will transform the matrix into
another matrix. Note that the dimension of the output matrix doesn't
need to match the the input matrix:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## A dimension-level 3 metric}
\KeywordTok{var}\NormalTok{(dummy_matrix)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##           [,1]      [,2]      [,3]
## [1,] 0.2275789 0.2927228 0.2720648
## [2,] 0.2927228 0.6394468 0.6796798
## [3,] 0.2720648 0.6796798 1.9455929
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## A dimension-level 3 metric with a forced matrix output}
\KeywordTok{as.matrix}\NormalTok{(}\KeywordTok{dist}\NormalTok{(dummy_matrix))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##          1        2         3         4
## 1 0.000000 2.284299 3.1469895 3.2820238
## 2 2.284299 0.000000 1.8652681 1.9829745
## 3 3.146990 1.865268 0.0000000 0.6684051
## 4 3.282024 1.982975 0.6684051 0.0000000
\end{verbatim}

\subsection{\texorpdfstring{\texttt{make.metric}}{make.metric}}\label{make.metric}

Of course, functions can be more complex and involve multiple operations
such as the \texttt{centroids} function (see \texttt{?dispRity.metric})
that calculates the Euclidean distance between each element and the
centroid of the ordinated space. The \texttt{make.metric} function
implemented in \texttt{dispRity} is designed to help test and find the
dimension-level of the functions. This function tests:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  If your function can deal with a \texttt{matrix} or a \texttt{vector}
  as an input;
\item
  Your function's dimension-level according to its output
  (dimension-level 1, 2 or 3, see above);
\item
  Whether the function can be implemented in the \texttt{dispRity}
  function (the function is fed into a \texttt{lapply} loop).
\end{enumerate}

For example, let's see if the functions described above are the right
dimension-levels:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Which dimension-level is the mean function? And can it be used in dispRity?}
\KeywordTok{make.metric}\NormalTok{(mean)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## mean outputs a single value.
## mean is detected as being a dimension-level 1 function.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Which dimension-level is the prod.rows function? And can it be used in dispRity?}
\KeywordTok{make.metric}\NormalTok{(prod.rows)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## prod.rows outputs a matrix object.
## prod.rows is detected as being a dimension-level 2 function.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Which dimension-level is the var function? And can it be used in dispRity?}
\KeywordTok{make.metric}\NormalTok{(var)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## var outputs a matrix object.
## var is detected as being a dimension-level 3 function.
## Additional dimension-level 2 and/or 1 function(s) will be needed.
\end{verbatim}

A non verbose version of the function is also available. This can be
done using the option \texttt{silent\ =\ TRUE} and will simply output
the dimension-level of the metric.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Testing whether mean is dimension-level 1}
\ControlFlowTok{if}\NormalTok{(}\KeywordTok{make.metric}\NormalTok{(mean, }\DataTypeTok{silent =} \OtherTok{TRUE}\NormalTok{) }\OperatorTok{!=}\StringTok{ "level1"}\NormalTok{) \{}
    \KeywordTok{message}\NormalTok{(}\StringTok{"The metric is not dimension-level 1."}\NormalTok{)}
\NormalTok{\}}
\NormalTok{## Testing whether var is dimension-level 1}
\ControlFlowTok{if}\NormalTok{(}\KeywordTok{make.metric}\NormalTok{(var, }\DataTypeTok{silent =} \OtherTok{TRUE}\NormalTok{) }\OperatorTok{!=}\StringTok{ "level1"}\NormalTok{) \{}
    \KeywordTok{message}\NormalTok{(}\StringTok{"The metric is not dimension-level 1."}\NormalTok{)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## The metric is not dimension-level 1.
\end{verbatim}

\subsection{\texorpdfstring{Metrics in the \texttt{dispRity}
function}{Metrics in the dispRity function}}\label{metrics-in-the-disprity-function}

Using this metric structure, we can easily use any disparity metric in
the \texttt{dispRity} function as follows:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Measuring disparity as the standard deviation of all the values of the}
\NormalTok{## ordinated matrix (dimension-level 1 function).}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(BeckLee_mat50, }\DataTypeTok{metric =}\NormalTok{ sd))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs
## 1       1 50 0.201
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Measuring disparity as the standard deviation of the variance of each axis of}
\NormalTok{## the ordinated matrix (dimension-level 1 and 2 functions).}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(BeckLee_mat50, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sd, variances)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs
## 1       1 50 0.028
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Measuring disparity as the standard deviation of the variance of each axis of}
\NormalTok{## the variance covariance matrix (dimension-level 1, 2 and 3 functions).}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(BeckLee_mat50, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sd, variances, var)), }\DataTypeTok{round =} \DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n obs
## 1       1 50   0
\end{verbatim}

Note that the order of each function in the metric argument does not
matter, the \texttt{dispRity} function will automatically detect the
function dimension-levels (using \texttt{make.metric}) and apply them to
the data in decreasing order (dimension-level 3 \textgreater{} 2
\textgreater{} 1).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Disparity as the standard deviation of the variance of each axis of the}
\NormalTok{## variance covariance matrix:}
\NormalTok{disparity1 <-}\StringTok{ }\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(BeckLee_mat50, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sd, variances, var)),}
                      \DataTypeTok{round =} \DecValTok{10}\NormalTok{)}

\NormalTok{## Same as above but using a different function order for the metric argument}
\NormalTok{disparity2 <-}\StringTok{ }\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(BeckLee_mat50, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(variances, sd, var)),}
                      \DataTypeTok{round =} \DecValTok{10}\NormalTok{)}

\NormalTok{## Both ways output the same disparity values:}
\NormalTok{disparity1 }\OperatorTok{==}\StringTok{ }\NormalTok{disparity2}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##      subsets    n  obs
## [1,]    TRUE TRUE TRUE
\end{verbatim}

In these examples, we considered disparity to be a single value. For
example, in the previous example, we defined disparity as the standard
deviation of the variances of each column of the variance/covariance
matrix (\texttt{metric\ =\ c(variances,\ sd,\ var)}). It is, however,
possible to calculate
\protect\hyperlink{disparity-as-a-distribution}{disparity as a
distribution}.

\subsection{\texorpdfstring{Metrics implemented in
\texttt{dispRity}}{Metrics implemented in dispRity}}\label{metrics-implemented-in-disprity}

Several disparity metrics are implemented in the \texttt{dispRity}
package. The detailed list can be found in \texttt{?dispRity.metric}
along with some description of each metric.

\begin{longtable}[]{@{}llll@{}}
\toprule
\begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
Level\strut
\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
Name\strut
\end{minipage} & \begin{minipage}[b]{0.61\columnwidth}\raggedright\strut
Description\strut
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Source\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
2\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{ancestral.dist}\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The distance between an element and its ancestor\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\texttt{dispRity}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
2\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{centroids}1\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The distance between each element and the centroid of the ordinated
space\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\texttt{dispRity}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
1\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{convhull.surface}\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The surface of the convex hull formed by all the elements\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\href{https://cran.r-project.org/web/packages/geometry/index.html}{\texttt{geometry}}\texttt{::convhulln\$area}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
1\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{convhull.volume}\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The volume of the convex hull formed by all the elements\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\href{https://cran.r-project.org/web/packages/geometry/index.html}{\texttt{geometry}}\texttt{::convhulln\$vol}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
1\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{diagonal}\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The longest distance in the ordinated space (like the diagonal in two
dimensions)\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\texttt{dispRity}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
1\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{ellipse.volume}1\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The volume of the ellipsoid of the space\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Donohue \emph{et al.} (2013)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
1\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{mode.val}\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The modal value\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\texttt{dispRity}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
1\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{n.ball.volume}\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The hyper-spherical (\emph{n}-ball) volume\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\texttt{dispRity}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
2\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{pairwise.dist}\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The pairwise distances between elements\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\href{https://cran.r-project.org/web/packages/vegan/index.html}{\texttt{vegan}}\texttt{::vegist}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
2\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{radius}\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The radius of each dimensions\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\texttt{dispRity}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
2\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{ranges}\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The range of each dimension\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\texttt{dispRity}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
1\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{span.tree.length}\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The minimal spanning tree length\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\href{https://cran.r-project.org/web/packages/vegan/index.html}{\texttt{vegan}}\texttt{::spantree}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
2\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\texttt{variances}\strut
\end{minipage} & \begin{minipage}[t]{0.61\columnwidth}\raggedright\strut
The variance of each dimension\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\texttt{dispRity}\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

1: Note that by default, the centroid is the centroid of the elements.
It can, however, be fixed to a different value by using the
\texttt{centroid} argument
\texttt{centroids(space,\ centroid\ =\ rep(0,\ ncol(space)))}, for
example the origin of the ordinated space.

2: This function uses an estimation of the eigenvalue that only works
for MDS or PCoA ordinations (\emph{not} PCA).

\subsection{Equations and
implementations}\label{equations-and-implementations}

Some of the functions described below are implemented in the
\texttt{dispRity} package and do not require any other packages to
calculate
(\href{https://github.com/TGuillerme/dispRity/blob/master/R/dispRity.metric.R}{see
implementation here}).

\begin{equation}
    ancestral.dist = \sqrt{\sum_{i=1}^{n}{({k}_{n}-Ancestor_{n})^2}}
\end{equation}

\begin{equation}
    centroids = \sqrt{\sum_{i=1}^{n}{({k}_{n}-Centroid_{k})^2}}
\end{equation}

\begin{equation}
    diagonal = \sqrt{\sum_{i=1}^{k}|max(k_i) - min(k_i)|}
\end{equation}

\begin{equation}
    ellipse.volume = \frac{\pi^{k/2}}{\Gamma(\frac{k}{2}+1)}\displaystyle\prod_{i=1}^{k} (\lambda_{i}^{0.5})
\end{equation}

\begin{equation}
    n.ball.volume = \frac{\pi^{k/2}}{\Gamma(\frac{k}{2}+1)}\displaystyle\prod_{i=1}^{k} R
\end{equation}

\begin{equation}
    radius = |\frac{\sum_{i=1}^{n}k_i}{n} - f(\mathbf{v}k)|
\end{equation}

\begin{equation}
    ranges = |max(k_i) - min(k_i)|
\end{equation}

\begin{equation}
    variances = \sigma^{2}{k_i}
\end{equation}

\begin{equation}
    span.tree.length = \sum(\mathrm{branch\ length})
\end{equation}

Where \emph{k} is the number of dimensions, \emph{n} the number of
elements, \(\Gamma\) is the Gamma distribution, \(\lambda_i\) is the
eigenvalue of each dimensions, \(\sigma^{2}\) is their variance and
\(Centroid_{k}\) is their mean, \(Ancestor_{n}\) is the coordinates of
the ancestor of element \(n\), \(f(\mathbf{v}k)\) is function to select
one value from the vector \(\mathbf{v}\) of the dimension \(k\)
(e.g.~it's maximum, minimum, mean, etc.), \emph{R} is the radius of the
sphere or the product of the radii of each dimensions
(\(\displaystyle\prod_{i=1}^{k}R_{i}\) - for a hyper-ellipsoid).

\subsection{Using the different disparity
metrics}\label{using-the-different-disparity-metrics}

Here is a brief demonstration of the main metrics implemented in
\texttt{dispRity}. First, we will create a dummy/simulated ordinated
space using the \texttt{space.maker} utility function (more about that
\protect\hyperlink{space.maker}{here}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating a 10*5 normal space}
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{)}
\NormalTok{dummy_space <-}\StringTok{ }\KeywordTok{space.maker}\NormalTok{(}\DecValTok{10}\NormalTok{, }\DecValTok{5}\NormalTok{, rnorm)}
\end{Highlighting}
\end{Shaded}

We will use this simulated space to demonstrate the different metrics.

\subsubsection{Volumes and surface
metrics}\label{volumes-and-surface-metrics}

The functions \texttt{ellipse.volume}, \texttt{convhull.surface},
\texttt{convhull.volume} and \texttt{n.ball.volume} all measure the
surface or the volume of the ordinated space occupied:

Because there is only one subset (i.e.~one matrix) in the dispRity
object, the operations below are the equivalent of
\texttt{metric(dummy\_space)} (with rounding).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating the ellipsoid volume}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ ellipse.volume))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs
## 1       1 10 257.8
\end{verbatim}

\begin{quote}
WARNING: in such dummy space, this gives the estimation of the ellipsoid
volume, not the real ellipsoid volume! See the cautionary note in
\texttt{?ellipse.volume}.
\end{quote}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating the convex hull surface}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ convhull.surface))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs
## 1       1 10 11.91
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating the convex hull volume}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ convhull.volume))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs
## 1       1 10 1.031
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating the convex hull volume}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ n.ball.volume))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n  obs
## 1       1 10 4.43
\end{verbatim}

The convex hull based functions are a call to the
\texttt{geometry::convhulln} function with the \texttt{"FA"} option
(computes total area and volume). Also note that they are really
sensitive to the size of the dataset.

\begin{quote}
Cautionary note: measuring volumes in a high number of dimensions can be
strongly affected by the
\href{https://en.wikipedia.org/wiki/Curse_of_dimensionality}{curse of
dimensionality} that often results in near 0 disparity values. I
strongly recommend reading
\href{https://beta.observablehq.com/@tophtucker/theres-plenty-of-room-in-the-corners}{this
really intuitive explanation} from
\href{https://github.com/tophtucker}{Toph Tucker}.
\end{quote}

\subsubsection{Ranges, variances, radius, pairwise distance, modal value
and
diagonal}\label{ranges-variances-radius-pairwise-distance-modal-value-and-diagonal}

The functions \texttt{ranges}, \texttt{variances} \texttt{radius},
\texttt{pairwise.dist}, \texttt{mode.val} and \texttt{diagonal} all
measure properties of the ordinated space based on its dimensional
properties (they are also less affected by the ``curse of
dimensionality''):

\texttt{ranges}, \texttt{variances} and \texttt{radius} work on the same
principle and measure the range/variance/radius of each dimension:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating the ranges of each dimension in the ordinated space}
\KeywordTok{ranges}\NormalTok{(dummy_space)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 2.430909 3.726481 2.908329 2.735739 1.588603
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating disparity as the distribution of these ranges}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ ranges))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n obs.median  2.5%   25%   75% 97.5%
## 1       1 10      2.431 1.673 2.431 2.908 3.645
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating disparity as the sum and the product of these ranges}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, ranges)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs
## 1       1 10 13.39
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(prod, ranges)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs
## 1       1 10 114.5
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating the variances of each dimension in the ordinated space}
\KeywordTok{variances}\NormalTok{(dummy_space)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 0.6093144 1.1438620 0.9131859 0.6537768 0.3549372
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating disparity as the distribution of these variances}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ variances))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n obs.median 2.5%   25%   75% 97.5%
## 1       1 10      0.609 0.38 0.609 0.913 1.121
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating disparity as the sum and the product of these variances}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, variances)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs
## 1       1 10 3.675
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(prod, variances)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs
## 1       1 10 0.148
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating the radius of each dimension in the ordinated space}
\KeywordTok{radius}\NormalTok{(dummy_space)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 1.4630780 2.4635449 1.8556785 1.4977898 0.8416318
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## By default the radius is the maximum distance from the centre of}
\NormalTok{## the dimension. It can however be changed to any function:}
\KeywordTok{radius}\NormalTok{(dummy_space, }\DataTypeTok{type =}\NormalTok{ min)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 0.05144054 0.14099827 0.02212226 0.17453525 0.23044528
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{radius}\NormalTok{(dummy_space, }\DataTypeTok{type =}\NormalTok{ mean)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 0.6233501 0.7784888 0.7118713 0.6253263 0.5194332
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating disparity as the mean average radius}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(mean, radius), }\DataTypeTok{type =}\NormalTok{ mean))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs
## 1       1 10 0.652
\end{verbatim}

The pairwise distances uses the function \texttt{vegan::vegdist} and can
take the normal \texttt{vegdist} options:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The average pairwise euclidean distance}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(mean, pairwise.dist)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs
## 1       1 10 2.539
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The distribution of the Manhattan distances}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ pairwise.dist, }\DataTypeTok{method =} \StringTok{"manhattan"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n obs.median  2.5%   25%   75% 97.5%
## 1       1 10      3.619 2.566 3.335 5.672  9.63
\end{verbatim}

Note that this function is a direct call to
\texttt{vegan::vegdist(matrix,\ method\ =\ method,\ diag\ =\ FALSE,\ upper\ =\ FALSE,\ ...)}.

The \texttt{diagonal} function measures the multidimensional diagonal of
the whole space (i.e.~in our case the longest Euclidean distance in our
five dimensional space). The \texttt{mode.val} function measures the
modal value of the matrix:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating the ordinated space's diagonal}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ diagonal))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs
## 1       1 10 3.659
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating the modal value of the matrix}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ mode.val))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs
## 1       1 10 -2.21
\end{verbatim}

\begin{quote}
This metric is only a Euclidean diagonal (mathematically valid) if the
dimensions within the space are all orthogonal!
\end{quote}

\subsubsection{Centroids and ancestral distance
metrics}\label{centroids-and-ancestral-distance-metrics}

The \texttt{centroids} metric allows users to measure the position of
the different elements compared to a fixed point in the ordinated space.
By default, this function measures the distance between each element and
their centroid (centre point):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The distribution of the distances between each element and their centroid}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ centroids))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n obs.median  2.5%   25%   75% 97.5%
## 1       1 10      2.214 0.788 1.267 1.993 3.167
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Disparity as the median value of these distances}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(median, centroids)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs
## 1       1 10 1.435
\end{verbatim}

It is however possible to fix the coordinates of the centroid to a
specific point in the ordinated space, as long as it has the correct
number of dimensions:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The distance between each element and the origin of the ordinated space}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ centroids, }\DataTypeTok{centroid =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n obs.median  2.5% 25%   75% 97.5%
## 1       1 10      2.323 0.785 1.2 2.044 3.176
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Disparity as the distance between each element and a specific point in space}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ centroids, }\DataTypeTok{centroid =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{3}\NormalTok{,}\DecValTok{4}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n obs.median  2.5%   25%   75% 97.5%
## 1       1 10      4.675 4.293 5.032 6.155 6.957
\end{verbatim}

The \texttt{ancestral.dist} metric works on a similar principle as the
\texttt{centroids} function but changes the centroid to be the
coordinates of each element's ancestor. Therefore this functions needs a
tree and node coordinates as additional arguments:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## A generating a random tree with node labels}
\NormalTok{tree <-}\StringTok{ }\KeywordTok{rtree}\NormalTok{(}\DecValTok{5}\NormalTok{) ; tree}\OperatorTok{$}\NormalTok{node.label <-}\StringTok{ }\KeywordTok{paste0}\NormalTok{(}\StringTok{"n"}\NormalTok{, }\DecValTok{1}\OperatorTok{:}\DecValTok{4}\NormalTok{)}
\NormalTok{## Adding the tip and node names to the matrix}
\NormalTok{dummy_space2 <-}\StringTok{ }\NormalTok{dummy_space[}\OperatorTok{-}\DecValTok{1}\NormalTok{,]}
\KeywordTok{rownames}\NormalTok{(dummy_space2) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(tree}\OperatorTok{$}\NormalTok{tip.label, tree}\OperatorTok{$}\NormalTok{node.label)}

\NormalTok{## Calculating all the ancestral nodes}
\NormalTok{all_anc_centroids <-}\StringTok{ }\KeywordTok{nodes.coordinates}\NormalTok{(dummy_space2, tree, }\DataTypeTok{full =} \OtherTok{TRUE}\NormalTok{)}

\NormalTok{## Calculating the distances from the ancestral nodes}
\NormalTok{ancestral_dist <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(dummy_space2, }\DataTypeTok{metric =}\NormalTok{ ancestral.dist,}
                           \DataTypeTok{nodes.coords =}\NormalTok{ all_anc_centroids)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in fun(matrix, ...): Missing tree and full argument for nodes.coordinates.
## See ?nodes.coordinates manual.
## The centroids function was applied instead.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating disparity as the sum of the distances from all the ancestral nodes}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(ancestral_dist, }\DataTypeTok{metric =}\NormalTok{ sum))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets n   obs
## 1       1 9 36.56
\end{verbatim}

\subsubsection{Minimal spanning tree
length}\label{minimal-spanning-tree-length}

The \texttt{span.tree.length} uses the \texttt{vegan::spantree} function
to heuristically calculate the minimum spanning tree (the shortest
multidimensional tree connecting each elements) and calculates its
length as the sum of every branch lengths.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The length of the minimal spanning tree}
\KeywordTok{summary}\NormalTok{(}\KeywordTok{dispRity}\NormalTok{(dummy_space, }\DataTypeTok{metric =}\NormalTok{ span.tree.length))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n  obs
## 1       1 10 15.4
\end{verbatim}

Note that because the solution is heuristic, this metric can take a long
time to compute for big matrices.

\section{Summarising dispRity data
(plots)}\label{summarising-disprity-data-plots}

Because of its architecture, printing \texttt{dispRity} objects only
summarises their content but does not print the disparity value measured
or associated analysis (more about this
\protect\hyperlink{manipulating-dispRity-objects}{here}). To actually
see what is in a dispRity object, one can either use the
\texttt{summary} function for visualising the data in a table or
\texttt{plot} to have a graphical representation of the results.

\subsection{\texorpdfstring{Summarising \texttt{dispRity}
data}{Summarising dispRity data}}\label{summarising-disprity-data}

This function is an S3 function (\texttt{summary.dispRity}) allowing
users to summarise the content of \texttt{dispRity} objects that contain
disparity calculations.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Example data from previous sections}
\NormalTok{crown_stem <-}\StringTok{ }\KeywordTok{custom.subsets}\NormalTok{(BeckLee_mat50,}
                                \DataTypeTok{group =} \KeywordTok{list}\NormalTok{(}\StringTok{"crown"}\NormalTok{ =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{16}\NormalTok{, }\DecValTok{19}\OperatorTok{:}\DecValTok{41}\NormalTok{, }\DecValTok{45}\OperatorTok{:}\DecValTok{50}\NormalTok{), }
                                             \StringTok{"stem"}\NormalTok{ =}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{15}\NormalTok{, }\DecValTok{17}\OperatorTok{:}\DecValTok{18}\NormalTok{, }\DecValTok{42}\OperatorTok{:}\DecValTok{44}\NormalTok{)))}
\NormalTok{## Bootstrapping and rarefying these groups}
\NormalTok{boot_crown_stem <-}\StringTok{ }\KeywordTok{boot.matrix}\NormalTok{(crown_stem, }\DataTypeTok{bootstraps =} \DecValTok{100}\NormalTok{, }\DataTypeTok{rarefaction =} \OtherTok{TRUE}\NormalTok{)}
\NormalTok{## Calculate disparity}
\NormalTok{disparity_crown_stem <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(boot_crown_stem, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, variances))}

\NormalTok{## Creating time slice subsets}
\NormalTok{time_slices <-}\StringTok{ }\KeywordTok{chrono.subsets}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat99, }\DataTypeTok{tree =}\NormalTok{ BeckLee_tree, }
    \DataTypeTok{method =} \StringTok{"continuous"}\NormalTok{, }\DataTypeTok{model =} \StringTok{"proximity"}\NormalTok{, }\DataTypeTok{time =} \KeywordTok{c}\NormalTok{(}\DecValTok{120}\NormalTok{, }\DecValTok{80}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{0}\NormalTok{),}
    \DataTypeTok{FADLAD =}\NormalTok{ BeckLee_ages)}
\NormalTok{## Bootstrapping the time slice subsets}
\NormalTok{boot_time_slices <-}\StringTok{ }\KeywordTok{boot.matrix}\NormalTok{(time_slices, }\DataTypeTok{bootstraps =} \DecValTok{100}\NormalTok{)}
\NormalTok{## Calculate disparity}
\NormalTok{disparity_time_slices <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(boot_time_slices, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, variances))}

\NormalTok{## Creating time bin subsets}
\NormalTok{time_bins <-}\StringTok{ }\KeywordTok{chrono.subsets}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat99, }\DataTypeTok{tree =}\NormalTok{ BeckLee_tree, }
    \DataTypeTok{method =} \StringTok{"discrete"}\NormalTok{, }\DataTypeTok{time =} \KeywordTok{c}\NormalTok{(}\DecValTok{120}\NormalTok{, }\DecValTok{80}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{FADLAD =}\NormalTok{ BeckLee_ages,}
    \DataTypeTok{inc.nodes =} \OtherTok{TRUE}\NormalTok{)}
\NormalTok{## Bootstrapping the time bin subsets}
\NormalTok{boot_time_bins <-}\StringTok{ }\KeywordTok{boot.matrix}\NormalTok{(time_bins, }\DataTypeTok{bootstraps =} \DecValTok{100}\NormalTok{)}
\NormalTok{## Calculate disparity}
\NormalTok{disparity_time_bins <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(boot_time_bins, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, variances))}
\end{Highlighting}
\end{Shaded}

These objects are easy to summarise as follows:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Default summary}
\KeywordTok{summary}\NormalTok{(disparity_time_slices)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs bs.median  2.5%   25%   75% 97.5%
## 1     120  5 2.823     2.295 1.398 2.037 2.563 2.767
## 2      80 19 3.233     3.065 2.785 2.973 3.133 3.266
## 3      40 15 3.359     3.149 2.764 2.996 3.285 3.443
## 4       0 10 4.055     3.685 3.169 3.453 3.760 3.961
\end{verbatim}

Information about the number of elements in each subset and the observed
(i.e.~non-bootstrapped) disparity are also calculated. This is
specifically handy when rarefying the data for example:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{head}\NormalTok{(}\KeywordTok{summary}\NormalTok{(disparity_crown_stem))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs bs.median  2.5%   25%   75% 97.5%
## 1   crown 30 1.995     1.933 1.873 1.915 1.943 1.972
## 2   crown 29    NA     1.935 1.861 1.906 1.953 1.976
## 3   crown 28    NA     1.929 1.872 1.910 1.945 1.970
## 4   crown 27    NA     1.933 1.852 1.905 1.947 1.981
## 5   crown 26    NA     1.932 1.867 1.910 1.948 1.971
## 6   crown 25    NA     1.929 1.847 1.913 1.949 1.975
\end{verbatim}

The summary functions can also take various options such as:

\begin{itemize}
\tightlist
\item
  \texttt{quantile} values for the confidence interval levels (by
  default, the 50 and 95 quantiles are calculated)
\item
  \texttt{cent.tend} for the central tendency to use for summarising the
  results (default is \texttt{median})
\item
  digits\texttt{option\ corresponding\ to\ the\ number\ of\ decimal\ places\ to\ print\ (default\ is}2`)
\item
  \texttt{recall} option for printing the call of the \texttt{dispRity}
  object as well (default is \texttt{FALSE})
\end{itemize}

These options can easily be changed from the defaults as follows:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Same as above but using the 88th quantile and the standard deviation as the summary }
\KeywordTok{summary}\NormalTok{(disparity_time_slices, }\DataTypeTok{quantile =} \DecValTok{88}\NormalTok{, }\DataTypeTok{cent.tend =}\NormalTok{ sd)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs bs.sd  2.5%   25%   75% 97.5%
## 1     120  5 2.823 0.380 1.398 2.037 2.563 2.767
## 2      80 19 3.233 0.127 2.785 2.973 3.133 3.266
## 3      40 15 3.359 0.189 2.764 2.996 3.285 3.443
## 4       0 10 4.055 0.205 3.169 3.453 3.760 3.961
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Printing the details of the object and digits the values to the 5th decimal place}
\KeywordTok{summary}\NormalTok{(disparity_time_slices, }\DataTypeTok{recall =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{digits =} \DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 4 continuous (proximity) time subsets for 99 elements with 97 dimensions:
##     120, 80, 40, 0.
## Data was bootstrapped 100 times (method:"full").
## Disparity was calculated as: c(sum, variances).
\end{verbatim}

\begin{verbatim}
##   subsets  n     obs bs.median    2.5%     25%     75%   97.5%
## 1     120  5 2.82292   2.29516 1.39758 2.03734 2.56317 2.76690
## 2      80 19 3.23312   3.06472 2.78542 2.97332 3.13308 3.26617
## 3      40 15 3.35947   3.14918 2.76395 2.99563 3.28541 3.44267
## 4       0 10 4.05457   3.68545 3.16864 3.45263 3.76034 3.96114
\end{verbatim}

Note that the summary table is a \texttt{data.frame}, hence it is as
easy to modify as any dataframe using \texttt{dplyr}. You can also
export it in \texttt{csv} format using \texttt{write.csv} or
\texttt{write\_csv} or even directly export into \texttt{LaTeX} format
using the following;

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Loading the xtable package}
\KeywordTok{require}\NormalTok{(xtable)}
\NormalTok{## Converting the table in LaTeX}
\KeywordTok{xtable}\NormalTok{(}\KeywordTok{summary}\NormalTok{(disparity_time_slices))}
\end{Highlighting}
\end{Shaded}

\subsection{\texorpdfstring{Plotting \texttt{dispRity}
data}{Plotting dispRity data}}\label{plotting-disprity-data}

An alternative (and more fun!) way to display the calculated disparity
is to plot the results using the S3 method \texttt{plot.dispRity}. This
function takes the same options as \texttt{summary.dispRity} along with
various graphical options described in the function help files (see
\texttt{?plot.dispRity}).

The plots can be of four different types:

\begin{itemize}
\tightlist
\item
  \texttt{continuous} for displaying continuous disparity curves
\item
  \texttt{box}, \texttt{lines}, and \texttt{polygons} to display
  discrete disparity results in respectively a boxplot, confidence
  interval lines, and confidence interval polygons.
\end{itemize}

\begin{quote}
This argument can be left empty. In this case, the algorithm will
automatically detect the type of subsets from the \texttt{dispRity}
object and plot accordingly.
\end{quote}

It is also possible to display the number of elements in each subset (as
a horizontal dotted line) using the option \texttt{elements\ =\ TRUE}.
Additionally, when the data is rarefied, one can indicate which level of
rarefaction to display (i.e.~only display the results for a certain
number of elements) by using the \texttt{rarefaction} argument.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical parameters}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}

\NormalTok{## Plotting continuous disparity results}
\KeywordTok{plot}\NormalTok{(disparity_time_slices, }\DataTypeTok{type =} \StringTok{"continuous"}\NormalTok{)}

\NormalTok{## Plotting discrete disparity results}
\KeywordTok{plot}\NormalTok{(disparity_crown_stem, }\DataTypeTok{type =} \StringTok{"box"}\NormalTok{)}

\NormalTok{## As above but using lines for the rarefaction level of 20 elements only}
\KeywordTok{plot}\NormalTok{(disparity_crown_stem, }\DataTypeTok{type =} \StringTok{"line"}\NormalTok{, }\DataTypeTok{rarefaction =} \DecValTok{20}\NormalTok{)}

\NormalTok{## As above but using polygons while also displaying the number of elements}
\KeywordTok{plot}\NormalTok{(disparity_crown_stem, }\DataTypeTok{type =} \StringTok{"polygon"}\NormalTok{, }\DataTypeTok{elements =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-54-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Resetting graphical parameters}
\KeywordTok{par}\NormalTok{(op)}
\end{Highlighting}
\end{Shaded}

Since \texttt{plot.dispRity} uses the arguments from the generic
\texttt{plot} method, it is of course possible to change pretty much
everything using the regular plot arguments:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical options}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}

\NormalTok{## Plotting the results with some classic options from plot}
\KeywordTok{plot}\NormalTok{(disparity_time_slices, }\DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"blue"}\NormalTok{, }\StringTok{"orange"}\NormalTok{, }\StringTok{"green"}\NormalTok{),}
    \DataTypeTok{ylab =} \KeywordTok{c}\NormalTok{(}\StringTok{"Some measurement"}\NormalTok{), }\DataTypeTok{xlab =} \StringTok{"Some other measurement"}\NormalTok{,}
    \DataTypeTok{main =} \StringTok{"Many options..."}\NormalTok{, }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(}\DecValTok{4}\NormalTok{, }\DecValTok{0}\NormalTok{))}

\NormalTok{## Adding a legend}
\KeywordTok{legend}\NormalTok{(}\StringTok{"topleft"}\NormalTok{, }\DataTypeTok{legend =} \KeywordTok{c}\NormalTok{(}\StringTok{"Central tendency"}\NormalTok{,}
                             \StringTok{"Confidence interval 1"}\NormalTok{,}
                            \StringTok{"Confidence interval 2"}\NormalTok{),}
      \DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"blue"}\NormalTok{, }\StringTok{"orange"}\NormalTok{, }\StringTok{"green"}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-55-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Resetting graphical parameters}
\KeywordTok{par}\NormalTok{(op)}
\end{Highlighting}
\end{Shaded}

In addition to the classic \texttt{plot} arguments, the function can
also take arguments that are specific to \texttt{plot.dispRity} like
adding the number of elements or rarefaction level (as described above),
and also changing the values of the quantiles to plot as well as the
central tendency.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical options}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}

\NormalTok{## Plotting the results with some plot.dispRity arguments}
\KeywordTok{plot}\NormalTok{(disparity_time_slices, }\DataTypeTok{quantile =} \KeywordTok{c}\NormalTok{(}\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{10}\NormalTok{, }\DataTypeTok{to =} \DecValTok{100}\NormalTok{, }\DataTypeTok{by =} \DecValTok{10}\NormalTok{)),}
    \DataTypeTok{cent.tend =}\NormalTok{ sd, }\DataTypeTok{type =} \StringTok{"c"}\NormalTok{, }\DataTypeTok{elements =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"black"}\NormalTok{, }\KeywordTok{rainbow}\NormalTok{(}\DecValTok{10}\NormalTok{)),}
    \DataTypeTok{ylab =} \KeywordTok{c}\NormalTok{(}\StringTok{"Disparity"}\NormalTok{, }\StringTok{"Diversity"}\NormalTok{), }\DataTypeTok{chrono.subsets =} \OtherTok{FALSE}\NormalTok{,}
    \DataTypeTok{xlab =} \StringTok{"Time (in in units from past to present)"}\NormalTok{, }\DataTypeTok{observed =} \OtherTok{TRUE}\NormalTok{,}
    \DataTypeTok{main =} \StringTok{"Many more options..."}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in plot.window(...): "quantile" is not a graphical parameter
\end{verbatim}

\begin{verbatim}
## Warning in plot.xy(xy, type, ...): "quantile" is not a graphical parameter
\end{verbatim}

\begin{verbatim}
## Warning in axis(side = side, at = at, labels = labels, ...): "quantile" is
## not a graphical parameter

## Warning in axis(side = side, at = at, labels = labels, ...): "quantile" is
## not a graphical parameter
\end{verbatim}

\begin{verbatim}
## Warning in box(...): "quantile" is not a graphical parameter
\end{verbatim}

\begin{verbatim}
## Warning in title(...): "quantile" is not a graphical parameter
\end{verbatim}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-56-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Resetting graphical parameters}
\KeywordTok{par}\NormalTok{(op)}
\end{Highlighting}
\end{Shaded}

\begin{quote}
Note that the argument \texttt{observed\ =\ TRUE} allows to plot the
disparity values calculated from the non-bootstrapped data as crosses on
the plot.
\end{quote}

For comparing results, it is also possible to add a plot to the existent
plot by using \texttt{add\ =\ TRUE}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical options}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}

\NormalTok{## Plotting the continuous disparity with a fixed y axis}
\KeywordTok{plot}\NormalTok{(disparity_time_slices, }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{4}\NormalTok{))}
\NormalTok{## Adding the discrete data}
\KeywordTok{plot}\NormalTok{(disparity_time_bins, }\DataTypeTok{type =} \StringTok{"line"}\NormalTok{, }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{4}\NormalTok{), }\DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{,}
    \DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-57-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Resetting graphical parameters}
\KeywordTok{par}\NormalTok{(op)}
\end{Highlighting}
\end{Shaded}

Finally, if your data has been fully rarefied, it is also possible to
easily look at rarefaction curves by using the
\texttt{rarefaction\ =\ TRUE} argument:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical options}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}

\NormalTok{## Plotting the rarefaction curves}
\KeywordTok{plot}\NormalTok{(disparity_crown_stem, }\DataTypeTok{rarefaction =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-58-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Resetting graphical parameters}
\KeywordTok{par}\NormalTok{(op)}
\end{Highlighting}
\end{Shaded}

\section{Testing disparity
hypotheses}\label{testing-disparity-hypotheses}

The \texttt{dispRity} package allows users to apply statistical tests to
the calculated disparity to test various hypotheses. The function
\texttt{test.dispRity} works in a similar way to the \texttt{dispRity}
function: it takes a \texttt{dispRity} object, a \texttt{test} and a
\texttt{comparisons} argument.

The \texttt{comparisons} argument indicates the way the test should be
applied to the data:

\begin{itemize}
\tightlist
\item
  \texttt{pairwise} (default): to compare each subset in a pairwise
  manner
\item
  \texttt{referential}: to compare each subset to the first subset
\item
  \texttt{sequential}: to compare each subset to the following subset
\item
  \texttt{all}: to compare all the subsets together (like in analysis of
  variance)
\end{itemize}

It is also possible to input a list of pairs of \texttt{numeric} values
or \texttt{characters} matching the subset names to create personalised
tests. Some other tests implemented in \texttt{dispRity} such as the
\texttt{dispRity::null.test} have a specific way they are applied to the
data and therefore ignore the \texttt{comparisons} argument.

The \texttt{test} argument can be any statistical or non-statistical
test to apply to the disparity object. It can be a common statistical
test function (e.g. \texttt{stats::t.test}), a function implemented in
\texttt{dispRity} (e.g.~see \texttt{?null.test}) or any function defined
by the user.

This function also allows users to correct for Type I error inflation
(false positives) when using multiple comparisons via the
\texttt{correction} argument. This argument can be empty (no correction
applied) or can contain one of the corrections from the
\texttt{stats::p.adjust} function (see \texttt{?p.adjust}).

Note that the \texttt{test.dispRity} algorithm deals with some classical
test outputs and summarises the test output. It is, however, possible to
get the full detailed output by using the options
\texttt{details\ =\ TRUE}.

Here we are using the variables generated in the
\protect\hyperlink{summarising-dispRity-data-plots}{section above}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## T-test to test for a difference in disparity between crown and stem mammals}
\KeywordTok{test.dispRity}\NormalTok{(disparity_crown_stem, }\DataTypeTok{test =}\NormalTok{ t.test)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [[1]]
##              statistic: t
## crown : stem     66.67693
## 
## [[2]]
##              parameter: df
## crown : stem      169.1446
## 
## [[3]]
##                    p.value
## crown : stem 2.275095e-123
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Performing the same test but with the detailed t.test output}
\KeywordTok{test.dispRity}\NormalTok{(disparity_crown_stem, }\DataTypeTok{test =}\NormalTok{ t.test, }\DataTypeTok{details =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## $`crown : stem`
## $`crown : stem`[[1]]
## 
##  Welch Two Sample t-test
## 
## data:  dots[[1L]][[1L]] and dots[[2L]][[1L]]
## t = 66.677, df = 169.14, p-value < 2.2e-16
## alternative hypothesis: true difference in means is not equal to 0
## 95 percent confidence interval:
##  0.2894967 0.3071619
## sample estimates:
## mean of x mean of y 
##  1.928128  1.629799
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Wilcoxon test applied to time sliced disparity with sequential comparisons,}
\NormalTok{## with Bonferroni correction}
\KeywordTok{test.dispRity}\NormalTok{(disparity_time_slices, }\DataTypeTok{test =}\NormalTok{ wilcox.test,}
              \DataTypeTok{comparisons =} \StringTok{"sequential"}\NormalTok{, }\DataTypeTok{correction =} \StringTok{"bonferroni"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [[1]]
##          statistic: W
## 120 : 80           27
## 80 : 40          3466
## 40 : 0            455
## 
## [[2]]
##               p.value
## 120 : 80 1.720545e-33
## 80 : 40  5.370214e-04
## 40 : 0   3.599349e-28
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Measuring the overlap between distributions in the time bins (using the}
\NormalTok{## implemented Bhattacharyya Coefficient function - see ?bhatt.coeff)}
\KeywordTok{test.dispRity}\NormalTok{(disparity_time_bins, }\DataTypeTok{test =}\NormalTok{ bhatt.coeff)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in test.dispRity(disparity_time_bins, test = bhatt.coeff): Multiple p-values will be calculated without adjustment!
## This can inflate Type I error!
\end{verbatim}

\begin{verbatim}
##                    bhatt.coeff
## 120 - 80 : 80 - 40    0.000000
## 120 - 80 : 40 - 0     0.000000
## 80 - 40 : 40 - 0      0.490448
\end{verbatim}

Because of the modular design of the package, tests can always be made
by the user (the same way disparity metrics can be user made). The only
condition is that the test can be applied to at least two distributions.
In practice, the \texttt{test.dispRity} function will pass the
calculated disparity data (distributions) to the provided function in
either pairs of distributions (if the \texttt{comparisons} argument is
set to \texttt{pairwise}, \texttt{referential} or \texttt{sequential})
or a table containing all the distributions
(\texttt{comparisons\ =\ all}; this should be in the same format as data
passed to \texttt{lm}-type functions for example).

\subsection{\texorpdfstring{NPMANOVA in
\texttt{dispRity}}{NPMANOVA in dispRity}}\label{npmanova-in-disprity}

One often useful test to apply to multidimensional data is the
permutational multivariate analysis of variance based on distance
matrices \texttt{vegan::adonis}. This can be done on \texttt{dispRity}
objects using the \texttt{adonis.dispRity} wrapper function. Basically,
this function takes the exact same arguments as \texttt{adonis} and a
\texttt{dispRity} object for data and performs a PERMANOVA based on the
distance matrix of the multidimensional space (unless the
multidimensional space was already defined as a distance matrix). The
\texttt{adonis.dispRity} function uses the information from the
\texttt{dispRity} object to generate default formulas:

\begin{itemize}
\tightlist
\item
  If the object contains customised subsets, it applies the default
  formula \texttt{matrix\ \textasciitilde{}~group} testing the effect of
  \texttt{group} as a predictor on \texttt{matrix} (called from the
  \texttt{dispRity} object as \texttt{data\$matrix} see
  \protect\hyperlink{The-dispRity-object-content}{\texttt{dispRitu}
  object details})
\item
  If the object contains time subsets, it applies the default formula
  \texttt{matrix\ \textasciitilde{}\ time} testing the effect of
  \texttt{time} as a predictor (were the different levels of
  \texttt{time} are the different time slices/bins)
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{)}
\NormalTok{## Generating a random character matrix}
\NormalTok{character_matrix <-}\StringTok{ }\KeywordTok{sim.morpho}\NormalTok{(}\KeywordTok{rtree}\NormalTok{(}\DecValTok{20}\NormalTok{), }\DecValTok{50}\NormalTok{, }\DataTypeTok{rates =} \KeywordTok{c}\NormalTok{(rnorm, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{))}

\NormalTok{## Calculating the distance matrix}
\NormalTok{distance_matrix <-}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(}\KeywordTok{dist}\NormalTok{(character_matrix))}

\NormalTok{## Creating two groups}
\NormalTok{random_groups <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}\StringTok{"group1"}\NormalTok{ =}\StringTok{ }\DecValTok{1}\OperatorTok{:}\DecValTok{10}\NormalTok{, }\StringTok{"group2"}\NormalTok{ =}\StringTok{ }\DecValTok{11}\OperatorTok{:}\DecValTok{20}\NormalTok{)}

\NormalTok{## Generating a dispRity object}
\NormalTok{random_disparity <-}\StringTok{ }\KeywordTok{custom.subsets}\NormalTok{(distance_matrix, random_groups)}

\NormalTok{## Running a default NPMANOVA}
\KeywordTok{adonis.dispRity}\NormalTok{(random_disparity)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## vegan::adonis(formula = matrix ~ group, data = random_disparity,      method = "euclidean") 
## 
## Permutation: free
## Number of permutations: 999
## 
## Terms added sequentially (first to last)
## 
##           Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)  
## group      1      16.5  16.500  1.3904 0.07171  0.095 .
## Residuals 18     213.6  11.867         0.92829         
## Total     19     230.1                 1.00000         
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
\end{verbatim}

Of course, it is possible to pass customised formulas if the disparity
object contains more more groups. In that case the predictors must
correspond to the names of the groups explained data must be set as
\texttt{matrix}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating two groups with two states each}
\NormalTok{groups <-}\StringTok{ }\KeywordTok{as.data.frame}\NormalTok{(}\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{10}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{10}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{), }\DecValTok{10}\NormalTok{)),}
         \DataTypeTok{nrow =} \DecValTok{20}\NormalTok{, }\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{, }\DataTypeTok{dimnames =} \KeywordTok{list}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(}\StringTok{"t"}\NormalTok{, }\DecValTok{1}\OperatorTok{:}\DecValTok{20}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\StringTok{"g1"}\NormalTok{, }\StringTok{"g2"}\NormalTok{))))}

\NormalTok{## Creating the dispRity object}
\NormalTok{multi_groups <-}\StringTok{ }\KeywordTok{custom.subsets}\NormalTok{(distance_matrix, groups)}

\NormalTok{## Running the NPMANOVA}
\KeywordTok{adonis.dispRity}\NormalTok{(multi_groups, matrix }\OperatorTok{~}\StringTok{ }\NormalTok{g1 }\OperatorTok{+}\StringTok{ }\NormalTok{g2)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## vegan::adonis(formula = matrix ~ g1 + g2, data = multi_groups,      method = "euclidean") 
## 
## Permutation: free
## Number of permutations: 999
## 
## Terms added sequentially (first to last)
## 
##           Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)
## g1         1      16.5  16.500 1.34921 0.07171  0.111
## g2         1       5.7   5.700 0.46609 0.02477  0.998
## Residuals 17     207.9  12.229         0.90352       
## Total     19     230.1                 1.00000
\end{verbatim}

Finally, it is possible to use objects generated by
\texttt{chrono.subsets}. In this case, \texttt{adonis.dispRity} will
applied the \texttt{matrix\ \textasciitilde{}\ time} formula by default:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating time series}
\NormalTok{time_subsets <-}\StringTok{ }\KeywordTok{chrono.subsets}\NormalTok{(BeckLee_mat50, BeckLee_tree, }
     \DataTypeTok{method =} \StringTok{"discrete"}\NormalTok{, }\DataTypeTok{inc.nodes =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{time =} \KeywordTok{c}\NormalTok{(}\DecValTok{100}\NormalTok{, }\DecValTok{85}\NormalTok{, }\DecValTok{65}\NormalTok{, }\DecValTok{0}\NormalTok{),}
     \DataTypeTok{FADLAD =}\NormalTok{ BeckLee_ages)}

\NormalTok{## Running the NPMANOVA with time as a predictor}
\KeywordTok{adonis.dispRity}\NormalTok{(time_subsets)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in adonis.dispRity(time_subsets): The input data for adonis.dispRity was not a distance matrix.
## The results are thus based on the distance matrix for the input data (i.e. dist(data$matrix)).
## Make sure that this is the desired methodological approach!
\end{verbatim}

\begin{verbatim}
## 
## Call:
## vegan::adonis(formula = dist(matrix) ~ time, data = time_subsets,      method = "euclidean") 
## 
## Permutation: free
## Number of permutations: 999
## 
## Terms added sequentially (first to last)
## 
##           Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)    
## time       2     8.049  4.0245  2.1303 0.08311  0.001 ***
## Residuals 47    88.792  1.8892         0.91689           
## Total     49    96.841                 1.00000           
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
\end{verbatim}

Note that the function warns you that the input data was transformed
into a distance matrix. This is reflected in the Call part of the output
(\texttt{formula\ =\ dist(matrix)\ \textasciitilde{}\ time}).

To use each time subset as a separate predictor, you can use the
\texttt{matrix\ \textasciitilde{}\ chrono.subsets} formula; this is
equivalent to
\texttt{matrix\ \textasciitilde{}\ first\_time\_subset\ +\ second\_time\_subset\ +\ ...}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Running the NPMANOVA with each time bin as a predictor}
\KeywordTok{adonis.dispRity}\NormalTok{(time_subsets, matrix }\OperatorTok{~}\StringTok{ }\NormalTok{chrono.subsets)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in adonis.dispRity(time_subsets, matrix ~ chrono.subsets): The input data for adonis.dispRity was not a distance matrix.
## The results are thus based on the distance matrix for the input data (i.e. dist(data$matrix)).
## Make sure that this is the desired methodological approach!
\end{verbatim}

\begin{verbatim}
## 
## Call:
## vegan::adonis(formula = dist(matrix) ~ chrono.subsets, data = time_subsets,      method = "euclidean") 
## 
## Permutation: free
## Number of permutations: 999
## 
## Terms added sequentially (first to last)
## 
##           Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)    
## t100to85   1     3.090  3.0897  1.6354 0.03190  0.002 ** 
## t85to65    1     4.959  4.9593  2.6251 0.05121  0.001 ***
## Residuals 47    88.792  1.8892         0.91689           
## Total     49    96.841                 1.00000           
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
\end{verbatim}

\subsection{\texorpdfstring{\texttt{geiger::dtt} model fitting in
\texttt{dispRity}}{geiger::dtt model fitting in dispRity}}\label{geigerdtt-model-fitting-in-disprity}

The \texttt{dtt} function from the \texttt{geiger} package is also often
used to compare a trait's disparity observed in living taxa to the
disparity of a simulated trait based on a given phylogeny. The
\texttt{dispRity} package proposes a wrapper function for
\texttt{geiger::dtt}, \texttt{dtt.dispRity} that allows the use of any
disparity metric. Unfortunately, this implementation is slower that
\texttt{geiger::dtt} (so if you're using the metrics implemented in
\texttt{geiger} prefer the original version) and, as the original
function, is limited to ultrametric trees (only living taxa!)\ldots{}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{require}\NormalTok{(geiger)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Loading required package: geiger
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{geiger_data <-}\StringTok{ }\KeywordTok{get}\NormalTok{(}\KeywordTok{data}\NormalTok{(geospiza))}

\NormalTok{## Calculate the disparity of the dataset using the sum of variance}
\NormalTok{dispRity_dtt <-}\StringTok{ }\KeywordTok{dtt.dispRity}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ geiger_data}\OperatorTok{$}\NormalTok{dat, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, variances),}
                             \DataTypeTok{tree =}\NormalTok{ geiger_data}\OperatorTok{$}\NormalTok{phy, }\DataTypeTok{nsim =} \DecValTok{100}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in dtt.dispRity(data = geiger_data$dat, metric = c(sum,
## variances), : The following tip(s) was not present in the data: olivacea.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Plotting the results}
\KeywordTok{plot}\NormalTok{(dispRity_dtt, }\DataTypeTok{fig.width=}\DecValTok{8}\NormalTok{, }\DataTypeTok{fig.height=}\DecValTok{8}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-64-1.pdf}

Note that, like in the original \texttt{dtt} function, it is possible to
change the evolutionary model (see \texttt{?geiger::sim.char}
documentation).

\subsection{\texorpdfstring{null morphospace testing with
\texttt{null.test}}{null morphospace testing with null.test}}\label{null-morphospace-testing-with-null.test}

This test is equivalent to the test performed in \citet{diaz2016global}.
It compares the disparity measured in the observed space to the
disparity measured in a set of simulated spaces. These simulated spaces
can be built with based on the hypothesis assumptions: for example, we
can test whether our space is normal.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{123}\NormalTok{)}
\NormalTok{## A "normal" multidimensional space with 50 dimensions and 10 elements}
\NormalTok{normal_space <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\KeywordTok{rnorm}\NormalTok{(}\DecValTok{1000}\NormalTok{), }\DataTypeTok{ncol =} \DecValTok{50}\NormalTok{)}

\NormalTok{## Calculating the disparity as the average pairwise distances}
\NormalTok{obs_disparity <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(normal_space, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(mean, pairwise.dist))}

\NormalTok{## Testing against 100 randomly generated normal spaces}
\NormalTok{(results <-}\StringTok{ }\KeywordTok{null.test}\NormalTok{(obs_disparity, }\DataTypeTok{replicates =} \DecValTok{100}\NormalTok{, }\DataTypeTok{null.distrib =}\NormalTok{ rnorm))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Monte-Carlo test
## Call: [1] "dispRity::null.test"
## 
## Observation: 9.910536 
## 
## Based on 100 replicates
## Simulated p-value: 0.7524752 
## Alternative hypothesis: two-sided 
## 
##     Std.Obs Expectation    Variance 
## -0.24885893  9.96420000  0.04650127
\end{verbatim}

Here the results show that disparity measured in our observed space is
not significantly different than the one measured in a normal space. We
can then propose that our observed space is normal!

These results have an attributed \texttt{dispRity} and \texttt{randtest}
class and can be plotted as \texttt{randtest} objects using the
\texttt{dispRity} S3 \texttt{plot} method:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Plotting the results}
\KeywordTok{plot}\NormalTok{(results, }\DataTypeTok{main =} \StringTok{"Is this space normal?"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-66-1.pdf}

For more details on generating spaces see the
\protect\hyperlink{Simulating-multidimensional-spaces}{\texttt{space.maker}}
function tutorial.

\section{Fitting modes of evolution to disparity
data}\label{fitting-modes-of-evolution-to-disparity-data}

The code used for these models is based on those developed by Gene Hunt
\citep{hunt2006fitting, hunt2012measuring, hunt2015simple}. So we
acknowledge and thank Gene Hunt for developing these models and writing
the original R code that served as inspiration for these models.

\subsection{Simple modes of disparity change through
time}\label{simple-modes-of-disparity-change-through-time}

\subsubsection{\texorpdfstring{\texttt{model.test}}{model.test}}\label{model.test}

Changes in disparity-through-time can follow a range of models, such as
random walks, stasis, constrained evolution, trends, or an early burst
model of evolution. We will start with by fitting the simplest modes of
evolution to our data. For example we may have a null expectation of
time-invariant change in disparity in which values fluctuate with a
variance around the mean - this would be best describe by a Stasis
model:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Loading premade disparity data}
\KeywordTok{data}\NormalTok{(BeckLee_disparity)}
\NormalTok{disp_time <-}\StringTok{ }\KeywordTok{model.test}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_disparity, }\DataTypeTok{model =} \StringTok{"Stasis"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Evidence of equal variance (Bartlett's test of equal variances p = 0).
## Variance is not pooled.
## Running Stasis model...Done. Log-likelihood = -59.501
\end{verbatim}

We can see the standard output from \texttt{model.test}. The first
output message tells us it has tested for equal variances in each
sample. The model uses Bartlett's test of equal variances to assess if
variances are equal, so if p \textgreater{} 0.05 then variance is
treated as the same for all samples, but if (p \textless{} 0.05) then
each bin variance is unique. Here we have p \textless{} 0.05, so
variance is not pooled between samples.

By default \texttt{model.test} will use Bartlett's test to assess for
homogeneity of variances, and then use this to decide to pool variances
or not. This is ignored if the argument \texttt{pool.variance} in
\texttt{model.test} is changed from the default \texttt{NULL} to
\texttt{TRUE} or \texttt{FALSE}. For example, to ignore Bartlett's test
and pool variances manually we would do the following:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{disp_time_pooled <-}\StringTok{ }\KeywordTok{model.test}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_disparity, }\DataTypeTok{model =} \StringTok{"Stasis"}\NormalTok{, }\DataTypeTok{pool.variance =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Running Stasis model...Done. Log-likelihood = -58.233
\end{verbatim}

However, unless you have good reason to choose otherwise it is
recommended to use the default of \texttt{pool.variance\ =\ NULL}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{disp_time <-}\StringTok{ }\KeywordTok{model.test}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_disparity, }\DataTypeTok{model =} \StringTok{"Stasis"}\NormalTok{, }\DataTypeTok{pool.variance =} \OtherTok{NULL}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Evidence of equal variance (Bartlett's test of equal variances p = 0).
## Variance is not pooled.
## Running Stasis model...Done. Log-likelihood = -59.501
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{disp_time}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Disparity evolution model fitting:
## Call: model.test(data = BeckLee_disparity, model = "Stasis", pool.variance = NULL) 
## 
##            aicc delta_aicc weight_aicc
## Stasis 123.1027          0           1
## 
## Use x$full.details for displaying the models details
## or summary(x) for summarising them.
\end{verbatim}

The remaining output gives us the log-likelihood of the Stasis model of
-59.501 (you may notice this change when we pooled variances above). The
output also gives us the small sample Akaike Information Criterion
(AICc), the delta AICc (the distance from the best fitting model), and
the AICc weights (\textasciitilde{}the relative support of this model
compared to all models, scaled to one).

These are all metrics of relative fit, so when we test a single model
they are not useful. By using the function summary in \texttt{dispRity}
we can see the maximum likelihood estimates of the model parameters:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(disp_time)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##         aicc delta_aicc weight_aicc log.lik param theta.1 omega
## Stasis 123.1          0           1   -59.5     2     3.4   0.1
\end{verbatim}

So we again see the AICc, delta AICc, AICc weight, and the
log-likelihood we saw previously. We now also see the number of
parameters from the model (2: theta and omega), and their estimates so
the variance (omega = 0.01) and the mean (theta.1 = 3.4).

The \texttt{model.test} function is designed to test relative model fit,
so we need to test more than one model to make relative comparisons. So
let's compare to the fit of the Stasis model to another model with two
parameters: the Brownian motion. Brownian motion assumes a constant mean
that is equal to the ancestral estimate of the sequence, and the
variance around this mean increases linearly with time. The easier way
to compare these models is to simply add \texttt{"BM"} to the
\texttt{models} vector argument:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{disp_time <-}\StringTok{ }\KeywordTok{model.test}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_disparity, }\DataTypeTok{model =} \KeywordTok{c}\NormalTok{(}\StringTok{"Stasis"}\NormalTok{, }\StringTok{"BM"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Evidence of equal variance (Bartlett's test of equal variances p = 0).
## Variance is not pooled.
## Running Stasis model...Done. Log-likelihood = -59.501
## Running BM model...Done. Log-likelihood = 123.938
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{disp_time}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Disparity evolution model fitting:
## Call: model.test(data = BeckLee_disparity, model = c("Stasis", "BM")) 
## 
##             aicc delta_aicc  weight_aicc
## Stasis  123.1027   366.8774 2.155677e-80
## BM     -243.7747     0.0000 1.000000e+00
## 
## Use x$full.details for displaying the models details
## or summary(x) for summarising them.
\end{verbatim}

Et voilà! Here we can see by the log-likelihood, AICc, delta AICc, and
AICc weight Brownian motion has a much better relative fit to these data
than the Stasis model. Brownian motion has a relative AICc fit 366 units
better than Stasis, and virtually has a AICc weight of 1.

We can also all the information about the relative fit of models
alongside the maximum likelihood estimates of model parameters using the
summary function

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(disp_time)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##        aicc delta_aicc weight_aicc log.lik param theta.1 omega
## Stasis  123      366.9           0   -59.5     2   3.403  0.15
## BM     -244        0.0           1   123.9     2      NA    NA
##        ancestral state sigma squared
## Stasis              NA            NA
## BM               2.858         0.003
\end{verbatim}

Not that because the parameters per models differ, the summary includes
\texttt{NA} for inapplicable parameters per models (e.g.~the theta and
omega parameters from the Stasis models are inapplicable for a Brownian
motion model).

We can plot the relative fit of our models using the \texttt{plot}
function

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(disp_time)}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics{dispRity_manual_files/figure-latex/plot1-1}

}

\caption{relative fit (AICc weight) of Stasis and Brownian models of disparity through time}\label{fig:plot1}
\end{figure}

Here we see and overwhelming support for the Brownian motion model.

Alternatively, we could test all available models single modes: Stasis,
Brownian motion, Ornstein-Uhlenbeck (evolution constrained to an
optima), Trend (increasing or decreasing mean through time), and Early
Burst (exponentially decreasing rate through time)

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{disp_time <-}\StringTok{ }\KeywordTok{model.test}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_disparity, }\DataTypeTok{model =} \KeywordTok{c}\NormalTok{(}\StringTok{"Stasis"}\NormalTok{, }\StringTok{"BM"}\NormalTok{, }\StringTok{"OU"}\NormalTok{, }\StringTok{"Trend"}\NormalTok{, }\StringTok{"EB"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Evidence of equal variance (Bartlett's test of equal variances p = 0).
## Variance is not pooled.
## Running Stasis model...Done. Log-likelihood = -59.501
## Running BM model...Done. Log-likelihood = 123.938
## Running OU model...Done. Log-likelihood = 126.431
## Running Trend model...Done. Log-likelihood = 126.361
## Running EB model...Done. Log-likelihood = 113.081
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(disp_time)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##        aicc delta_aicc weight_aicc log.lik param theta.1 omega
## Stasis  123      369.6       0.000   -59.5     2   3.403  0.15
## BM     -244        2.7       0.157   123.9     2      NA    NA
## OU     -245        2.0       0.227   126.4     4      NA    NA
## Trend  -247        0.0       0.617   126.4     3      NA    NA
## EB     -220       26.6       0.000   113.1     3      NA    NA
##        ancestral state sigma squared alpha optima.1 trend     eb
## Stasis              NA            NA    NA       NA    NA     NA
## BM               2.858         0.003    NA       NA    NA     NA
## OU               2.835         0.002 0.004    5.707    NA     NA
## Trend            2.839         0.002    NA       NA  0.01     NA
## EB               4.055         0.002    NA       NA    NA -0.014
\end{verbatim}

These models indicate support for a Trend model, and we can plot the
relative support of all model AICc weights

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(disp_time)}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics{dispRity_manual_files/figure-latex/plot2-1}

}

\caption{relative fit (AICc weight) of various modes of evolution}\label{fig:plot2}
\end{figure}

Is this a trend of increasing or decreasing disparity through time? One
way to find out is to look at the summary function for the Trend model:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(disp_time)[}\StringTok{"Trend"}\NormalTok{,]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##            aicc      delta_aicc     weight_aicc         log.lik 
##        -247.000           0.000           0.617         126.400 
##           param         theta.1           omega ancestral state 
##           3.000              NA              NA           2.839 
##   sigma squared           alpha        optima.1           trend 
##           0.002              NA              NA           0.010 
##              eb 
##              NA
\end{verbatim}

This show a positive trend (0.01) of increasing disparity through time.

\subsection{Plot and run simulation tests in a single
step}\label{plot-and-run-simulation-tests-in-a-single-step}

\subsubsection{\texorpdfstring{\texttt{model.test.wrapper}}{model.test.wrapper}}\label{model.test.wrapper}

Patterns of evolution can be fit using \texttt{model.test}, but the
\texttt{model.test.wrapper} fits the same models as \texttt{model.test}
as well as running predictive tests and plots.

The predictive tests use the maximum likelihood estimates of model
parameters to simulate a number of datasets (default = 1000), and
analyse whether this is significantly different to the empirical input
data using the Rank Envelope test \citep{murrell2018global}. Finally we
can plot the empirical data, simulated data, and the Rank Envelope test
p values. This can all be done using the function
\texttt{model.test.wrapper}, and we will set the argument
\texttt{show.p\ =\ TRUE} so \emph{p} values from the Rank Envelope test
are printed on the plot:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{disp_time <-}\StringTok{ }\KeywordTok{model.test.wrapper}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_disparity, }\DataTypeTok{model =} \KeywordTok{c}\NormalTok{(}\StringTok{"Stasis"}\NormalTok{, }\StringTok{"BM"}\NormalTok{, }\StringTok{"OU"}\NormalTok{, }\StringTok{"Trend"}\NormalTok{, }\StringTok{"EB"}\NormalTok{),}
                                \DataTypeTok{show.p =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Evidence of equal variance (Bartlett's test of equal variances p = 0).
## Variance is not pooled.
## Running Stasis model...Done. Log-likelihood = -59.501
## Running BM model...Done. Log-likelihood = 123.938
## Running OU model...Done. Log-likelihood = 126.431
## Running Trend model...Done. Log-likelihood = 126.361
## Running EB model...Done. Log-likelihood = 113.081
\end{verbatim}

\begin{figure}

{\centering \includegraphics{dispRity_manual_files/figure-latex/plot3-1}

}

\caption{Empirical disparity through time (pink), simulate data based on estimated model parameters (grey), delta AICc, and range of p values from the Rank Envelope test for Trend, OU, BM, EB, and Stasis models}\label{fig:plot3}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{disp_time}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##        aicc delta_aicc weight_aicc log.lik param theta.1 omega
## Trend  -247        0.0       0.617   126.4     3      NA    NA
## OU     -245        2.0       0.227   126.4     4      NA    NA
## BM     -244        2.7       0.157   123.9     2      NA    NA
## EB     -220       26.6       0.000   113.1     3      NA    NA
## Stasis  123      369.6       0.000   -59.5     2   3.403  0.15
##        ancestral state sigma squared alpha optima.1 trend     eb
## Trend            2.839         0.002    NA       NA  0.01     NA
## OU               2.835         0.002 0.004    5.707    NA     NA
## BM               2.858         0.003    NA       NA    NA     NA
## EB               4.055         0.002    NA       NA    NA -0.014
## Stasis              NA            NA    NA       NA    NA     NA
##        median p value lower p value upper p value
## Trend      0.97902098   0.978021978     0.9800200
## OU         0.93406593   0.934065934     0.9340659
## BM         0.33266733   0.322677323     0.3426573
## EB         0.06293706   0.000999001     0.1248751
## Stasis     1.00000000   1.000000000     1.0000000
\end{verbatim}

From this plot we can see the empirical estimates of disparity through
time (pink) compared to the predictive data based upon the simulations
using the estimated parameters from each model. There is no significant
differences between the empirical data and simulated data, except for
the Early Burst model.

Trend is the best-fitting model but the plot suggests the OU model also
follows a trend-like pattern. This is because the optima for the OU
model (5.7) is different to the ancestral state 2.835 and outside the
observed value. This is potentially unrealistic, and one way to
alleviate this issue is to set the optima of the OU model to equal the
ancestral estimate - this is the normal practice for OU models in
comparative phylogenetics. To set the optima to the ancestral value we
change the argument \texttt{fixed.optima\ =\ TRUE}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{disp_time <-}\StringTok{ }\KeywordTok{model.test.wrapper}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_disparity, }\DataTypeTok{model =} \KeywordTok{c}\NormalTok{(}\StringTok{"Stasis"}\NormalTok{, }\StringTok{"BM"}\NormalTok{, }\StringTok{"OU"}\NormalTok{, }\StringTok{"Trend"}\NormalTok{, }\StringTok{"EB"}\NormalTok{),}
                                \DataTypeTok{show.p =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{fixed.optima =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Evidence of equal variance (Bartlett's test of equal variances p = 0).
## Variance is not pooled.
## Running Stasis model...Done. Log-likelihood = -59.501
## Running BM model...Done. Log-likelihood = 123.938
## Running OU model...Done. Log-likelihood = 123.938
## Running Trend model...Done. Log-likelihood = 126.361
## Running EB model...Done. Log-likelihood = 113.081
\end{verbatim}

\begin{figure}

{\centering \includegraphics{dispRity_manual_files/figure-latex/plot4-1}

}

\caption{Empirical disparity through time (pink), simulate data based on estimated model parameters (grey), delta AICc, and range of p values from the Rank Envelope test for Trend, OU, BM, EB, and Stasis models with the optima of the OU model set to equal the ancestral value}\label{fig:plot4}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{disp_time}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##        aicc delta_aicc weight_aicc log.lik param theta.1 omega
## Trend  -247        0.0       0.745   126.4     3      NA    NA
## BM     -244        2.7       0.189   123.9     2      NA    NA
## OU     -242        4.8       0.066   123.9     3      NA    NA
## EB     -220       26.6       0.000   113.1     3      NA    NA
## Stasis  123      369.6       0.000   -59.5     2   3.403  0.15
##        ancestral state sigma squared alpha trend     eb median p value
## Trend            2.839         0.002    NA  0.01     NA     0.97402597
## BM               2.858         0.003    NA    NA     NA     0.29870130
## OU               2.858         0.003     0    NA     NA     0.35864136
## EB               4.055         0.002    NA    NA -0.014     0.06393606
## Stasis              NA            NA    NA    NA     NA     1.00000000
##        lower p value upper p value
## Trend    0.973026973     0.9750250
## BM       0.285714286     0.3116883
## OU       0.347652348     0.3696304
## EB       0.000999001     0.1268731
## Stasis   1.000000000     1.0000000
\end{verbatim}

The relative fit of the OU model is decreased by constraining the fit of
the optima to equal the ancestral state value. In fact as the OU
attraction parameter (alpha) is zero, the model is equal to a Brownian
motion model but is penalised by having an extra parameter. Note that
indeed, the plots of the BM model and the OU model look nearly
identical.

\subsection{Multiple modes of evolution (time
shifts)}\label{multiple-modes-of-evolution-time-shifts}

As well as fitting a single model to a sequence of disparity values we
can also allow for the mode of evolution to shift at a single or
multiple points in time. The timing of a shift in mode can be based on
an a prior expectation, such as a mass extinction event, or the model
can test multiple points to allow to find time shift point with the
highest likelihood.

Models can be fit using \texttt{model.test} but it can be more
convenient to use \texttt{model.test.wrapper}. Here we will compare the
relative fit of Brownian motion, Trend, Ornstein-Uhlenbeck and a
multi-mode Ornstein Uhlenbck model in which the optima changes at 66
million years ago, the Cretaceous-Palaeogene boundary.

For example, we could be testing the hypothesis that the extinction of
non-avian dinosaurs allowed mammals to go from scurrying in the
undergrowth (low optima/low disparity) to dominating all habitats (high
optima/high disparity). We will constrain the optima of OU model in the
first time begin (i.e, pre-66 Mya) to equal the ancestral value:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{disp_time <-}\StringTok{ }\KeywordTok{model.test.wrapper}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_disparity, }\DataTypeTok{model =} \KeywordTok{c}\NormalTok{(}\StringTok{"BM"}\NormalTok{, }\StringTok{"Trend"}\NormalTok{, }\StringTok{"OU"}\NormalTok{, }\StringTok{"multi.OU"}\NormalTok{),}
                                \DataTypeTok{time.split =} \DecValTok{66}\NormalTok{, }\DataTypeTok{pool.variance =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{show.p =} \OtherTok{TRUE}\NormalTok{,}
                                \DataTypeTok{fixed.optima =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Evidence of equal variance (Bartlett's test of equal variances p = 0).
## Variance is not pooled.
## Running BM model...Done. Log-likelihood = 123.938
## Running Trend model...Done. Log-likelihood = 126.361
## Running OU model...Done. Log-likelihood = 123.938
## Running multi.OU model...Done. Log-likelihood = 126.469
\end{verbatim}

\begin{figure}

{\centering \includegraphics{dispRity_manual_files/figure-latex/plot5-1}

}

\caption{Empirical disparity through time (pink), simulate data based on estimated model parameters (grey), delta AICc, and range of p values from the Rank Envelope test for BM, Trend, OU, and multi OU models with a shift in optima allowed at 66 Ma}\label{fig:plot5}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{disp_time}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##          aicc delta_aicc weight_aicc log.lik param ancestral state
## Trend    -247      0.000       0.580   126.4     3           2.839
## multi.OU -245      1.922       0.222   126.5     4           2.836
## BM       -244      2.742       0.147   123.9     2           2.858
## OU       -242      4.845       0.051   123.9     3           2.858
##          sigma squared trend alpha optima.2 median p value lower p value
## Trend            0.002  0.01    NA       NA      0.9920080     0.9920080
## multi.OU         0.002    NA 0.005    5.526      0.6308691     0.6263736
## BM               0.003    NA    NA       NA      0.3506494     0.3396603
## OU               0.003    NA 0.000       NA      0.3666334     0.3576424
##          upper p value
## Trend        0.9920080
## multi.OU     0.6353646
## BM           0.3616384
## OU           0.3756244
\end{verbatim}

The multi-OU model shows an increase an optima at the
Cretaceous-Palaeogene boundary, indicating a shift in disparity.
However, this model does not fit as well as a model in which there is an
increasing trend through time. We can also fit a model in which the we
specify a heterogeneous model but we do not give a \texttt{time.split}.
In this instance the model will test all splits that have at least 10
time slices on either side of the split. That's 102 potential time
shifts in this example dataset so be warned, the following code will
estimate 105 models!

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## An example of a time split model in which all potential splits are tested}
\NormalTok{## }\AlertTok{WARNING}\NormalTok{: this will take between 20 minutes and half and hour to run!}
\NormalTok{disp_time <-}\StringTok{ }\KeywordTok{model.test.wrapper}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_disparity, }\DataTypeTok{model =} \KeywordTok{c}\NormalTok{(}\StringTok{"BM"}\NormalTok{, }\StringTok{"Trend"}\NormalTok{, }\StringTok{"OU"}\NormalTok{, }\StringTok{"multi.OU"}\NormalTok{),}
                                \DataTypeTok{show.p =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{fixed.optima =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

As well as specifying a multi-OU model we can run any combination of
models. For example we could fit a model at the Cretaceous-Palaeogene
boundary that goes from an OU to a BM model, a Trend to an OU model, a
Stasis to a Trend model or any combination you want to use. The only
model that can't be used in combination is a multi-OU model.

These can be introduced by changing the input for the models into a
list, and supplying a vector with the two models. This is easier to see
with an example:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The models to test}
\NormalTok{my_models <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"BM"}\NormalTok{, }\StringTok{"OU"}\NormalTok{),}
                  \KeywordTok{c}\NormalTok{(}\StringTok{"Stasis"}\NormalTok{, }\StringTok{"OU"}\NormalTok{),}
                  \KeywordTok{c}\NormalTok{(}\StringTok{"BM"}\NormalTok{, }\StringTok{"Stasis"}\NormalTok{),}
                  \KeywordTok{c}\NormalTok{(}\StringTok{"OU"}\NormalTok{, }\StringTok{"Trend"}\NormalTok{),}
                  \KeywordTok{c}\NormalTok{(}\StringTok{"Stasis"}\NormalTok{, }\StringTok{"BM"}\NormalTok{))}

\NormalTok{## Testing the models}
\NormalTok{disp_time <-}\StringTok{ }\KeywordTok{model.test.wrapper}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_disparity, }\DataTypeTok{model =}\NormalTok{ my_models, }\DataTypeTok{time.split =} \DecValTok{66}\NormalTok{,}
                                \DataTypeTok{show.p =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{fixed.optima =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Evidence of equal variance (Bartlett's test of equal variances p = 0).
## Variance is not pooled.
## Running BM:OU model...Done. Log-likelihood = 115.367
## Running Stasis:OU model...Done. Log-likelihood = 80.16
## Running BM:Stasis model...Done. Log-likelihood = 34.278
## Running OU:Trend model...Done. Log-likelihood = 118.166
## Running Stasis:BM model...Done. Log-likelihood = 80.16
\end{verbatim}

\begin{figure}

{\centering \includegraphics{dispRity_manual_files/figure-latex/plot6-1}

}

\caption{Empirical disparity through time (pink), simulate data based on estimated model parameters (grey), delta AICc, and range of p values from the Rank Envelope test for a variety of models with a shift in optima allowed at 66 Ma}\label{fig:plot6}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{disp_time}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##           aicc delta_aicc weight_aicc log.lik param ancestral state
## OU:Trend  -228        0.0       0.943   118.2     4           3.051
## BM:OU     -222        5.6       0.057   115.4     4           3.052
## Stasis:BM -154       73.9       0.000    80.2     3              NA
## Stasis:OU -150       78.2       0.000    80.2     5              NA
## BM:Stasis  -60      167.8       0.000    34.3     4           2.858
##           sigma squared alpha optima.1 theta.1 omega trend median p value
## OU:Trend          0.003 0.039       NA      NA    NA 0.015      0.4840160
## BM:OU             0.003 0.000    4.055      NA    NA    NA      0.7857143
## Stasis:BM         0.004    NA       NA   3.109 0.025    NA      0.9860140
## Stasis:OU         0.004 0.000    4.055   3.109 0.025    NA      0.9635365
## BM:Stasis         0.002    NA       NA   3.648 0.113    NA      0.9615385
##           lower p value upper p value
## OU:Trend      0.4685315     0.4995005
## BM:OU         0.7792208     0.7922078
## Stasis:BM     0.9850150     0.9870130
## Stasis:OU     0.9610390     0.9660340
## BM:Stasis     0.9560440     0.9670330
\end{verbatim}

\subsection{\texorpdfstring{\texttt{model.test.sim}}{model.test.sim}}\label{model.test.sim}

Note that all the models above where run using the
\texttt{model.test.wrapper} function that is a\ldots{} wrapping
function! In practice, this function runs two main functions from the
\texttt{dispRity} package and then plots the results:

\begin{itemize}
\tightlist
\item
  \texttt{model.test} and
\item
  \texttt{model.test.sim}
\end{itemize}

The \texttt{model.test.sim} allows to simulate disparity evolution given
a \texttt{dispRity} object input (as in \texttt{model.test.wrapper}) or
given a model and its specification. For example, it is possible to
simulate a simple Brownian motion model (or any of the other models or
models combination described above):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## A simple BM model}
\NormalTok{model_simulation <-}\StringTok{ }\KeywordTok{model.test.sim}\NormalTok{(}\DataTypeTok{sim =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{model =} \StringTok{"BM"}\NormalTok{, }\DataTypeTok{time.span =} \DecValTok{50}\NormalTok{, }\DataTypeTok{variance =} \FloatTok{0.1}\NormalTok{,}
                                   \DataTypeTok{sample.size =} \DecValTok{100}\NormalTok{, }\DataTypeTok{parameters =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{ancestral.state =} \DecValTok{0}\NormalTok{))}
\NormalTok{model_simulation}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Disparity evolution model simulation:
## Call: model.test.sim(sim = 1000, model = "BM", time.span = 50, variance = 0.1, sample.size = 100, parameters = list(ancestral.state = 0)) 
## 
## Model simulated (1000 times):
## [1] "BM"
\end{verbatim}

This will simulate 1000 Brownian motions for 50 units of time with 100
sampled elements, a variance of 0.1 and an ancestral state of 0. We can
also pass multiple models in the same way we did it for
\texttt{model.test} This model can then be summarised and plotted as
most \texttt{dispRity} objects:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Displaying the 5 first rows of the summary}
\KeywordTok{head}\NormalTok{(}\KeywordTok{summary}\NormalTok{(model_simulation))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets   n var      median      2.5%       25%       75%    97.5%
## 1      50 100 0.1 -0.09312806 -1.718481 -0.712478 0.5854576 1.772700
## 2      49 100 0.1 -0.01881396 -2.697017 -0.967495 0.9105681 2.712452
## 3      48 100 0.1 -0.06914146 -3.443629 -1.205211 1.1171606 3.185465
## 4      47 100 0.1 -0.01584249 -3.897544 -1.474240 1.3762412 3.639836
## 5      46 100 0.1 -0.04118287 -4.402526 -1.496096 1.5347916 4.239536
## 6      45 100 0.1 -0.17175251 -4.764863 -1.697076 1.5894314 4.349500
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Plotting the simulations}
\KeywordTok{plot}\NormalTok{(model_simulation)}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics{dispRity_manual_files/figure-latex/plot7-1}

}

\caption{A simulated Brownian motion}\label{fig:plot7}
\end{figure}

Note that these functions can take all the arguments that can be passed
to \texttt{plot}, \texttt{summary}, \texttt{plot.dispRity} and
\texttt{summary.dispRity}.

\subsubsection{Simulating tested models}\label{simulating-tested-models}

Maybe more interestingly though, it is possible to pass the output of
\texttt{model.test} directly to \texttt{model.test.sim} to simulate the
models that fits the data the best and calculate the Rank Envelope test
\emph{p} value. Let's see that using the simple example from the start:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Fitting multiple models on the data set}
\NormalTok{disp_time <-}\StringTok{ }\KeywordTok{model.test}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_disparity, }\DataTypeTok{model =} \KeywordTok{c}\NormalTok{(}\StringTok{"Stasis"}\NormalTok{, }\StringTok{"BM"}\NormalTok{, }\StringTok{"OU"}\NormalTok{, }\StringTok{"Trend"}\NormalTok{, }\StringTok{"EB"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Evidence of equal variance (Bartlett's test of equal variances p = 0).
## Variance is not pooled.
## Running Stasis model...Done. Log-likelihood = -59.501
## Running BM model...Done. Log-likelihood = 123.938
## Running OU model...Done. Log-likelihood = 126.431
## Running Trend model...Done. Log-likelihood = 126.361
## Running EB model...Done. Log-likelihood = 113.081
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(disp_time)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##        aicc delta_aicc weight_aicc log.lik param theta.1 omega
## Stasis  123      369.6       0.000   -59.5     2   3.403  0.15
## BM     -244        2.7       0.157   123.9     2      NA    NA
## OU     -245        2.0       0.227   126.4     4      NA    NA
## Trend  -247        0.0       0.617   126.4     3      NA    NA
## EB     -220       26.6       0.000   113.1     3      NA    NA
##        ancestral state sigma squared alpha optima.1 trend     eb
## Stasis              NA            NA    NA       NA    NA     NA
## BM               2.858         0.003    NA       NA    NA     NA
## OU               2.835         0.002 0.004    5.707    NA     NA
## Trend            2.839         0.002    NA       NA  0.01     NA
## EB               4.055         0.002    NA       NA    NA -0.014
\end{verbatim}

As seen before, the Trend model fitted this dataset the best. To
simulate what 1000 Trend models would look like using the same
parameters as the ones estimated with \texttt{model.test} (here the
ancestral state being 2.839, the sigma squared beeing 0.002 and the
trend of 0.01), we can simply pass this model to
\texttt{model.test.sim}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Simulating 1000 Trend model with the observed parameters}
\NormalTok{sim_trend <-}\StringTok{ }\KeywordTok{model.test.sim}\NormalTok{(}\DataTypeTok{sim =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{model =}\NormalTok{ disp_time)}
\NormalTok{sim_trend}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Disparity evolution model simulation:
## Call: model.test.sim(sim = 1000, model = disp_time) 
## 
## Model simulated (1000 times):
##       aicc log.lik param ancestral state sigma squared trend
## Trend -247   126.4     3           2.839         0.002  0.01
## 
## Rank envelope test
##  p-value of the test: 0.987013 (ties method: midrank)
##  p-interval         : (0.987013, 0.987013)
\end{verbatim}

By default, the model simulated is the one with the lowest AICc
(\texttt{model.rank\ =\ 1}) but it is possible to choose any ranked
model, for example, the OU (second one):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Simulating 1000 OU model with the observed parameters}
\NormalTok{sim_OU <-}\StringTok{ }\KeywordTok{model.test.sim}\NormalTok{(}\DataTypeTok{sim =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{model =}\NormalTok{ disp_time, }\DataTypeTok{model.rank =} \DecValTok{2}\NormalTok{)}
\NormalTok{sim_OU}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Disparity evolution model simulation:
## Call: model.test.sim(sim = 1000, model = disp_time, model.rank = 2) 
## 
## Model simulated (1000 times):
##    aicc log.lik param ancestral state sigma squared alpha optima.1
## OU -245   126.4     4           2.835         0.002 0.004    5.707
## 
## Rank envelope test
##  p-value of the test: 0.8971029 (ties method: midrank)
##  p-interval         : (0.8941059, 0.9000999)
\end{verbatim}

And as the example above, the simulated data can be plotted or
summarised:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{head}\NormalTok{(}\KeywordTok{summary}\NormalTok{(sim_trend))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets n        var   median     2.5%      25%      75%    97.5%
## 1     120 5 0.06056490 2.837869 2.614855 2.760862 2.908493 3.059927
## 2     119 5 0.07453663 2.858233 2.598562 2.756114 2.945448 3.106885
## 3     118 6 0.07556947 2.856971 2.589279 2.769942 2.941288 3.129721
## 4     117 6 0.07556947 2.870782 2.584536 2.777017 2.965582 3.147397
## 5     116 6 0.07556947 2.869144 2.589171 2.769528 2.978955 3.173132
## 6     115 7 0.06590243 2.883672 2.593479 2.790683 2.978769 3.179178
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{head}\NormalTok{(}\KeywordTok{summary}\NormalTok{(sim_OU))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets n        var   median     2.5%      25%      75%    97.5%
## 1     120 5 0.06056490 2.832068 2.618984 2.759091 2.908017 3.051097
## 2     119 5 0.07453663 2.849006 2.592719 2.759753 2.940149 3.115460
## 3     118 6 0.07556947 2.858304 2.579423 2.767239 2.946567 3.125425
## 4     117 6 0.07556947 2.872844 2.585720 2.782702 2.971421 3.142925
## 5     116 6 0.07556947 2.881609 2.599897 2.779863 2.985630 3.168541
## 6     115 7 0.06590243 2.897410 2.602577 2.790346 2.998301 3.182402
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The trend model with some graphical options}
\KeywordTok{plot}\NormalTok{(sim_trend, }\DataTypeTok{xlab =} \StringTok{"Time (Mya)"}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{"sum of variances"}\NormalTok{,}
    \DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"#F65205"}\NormalTok{, }\StringTok{"#F38336"}\NormalTok{, }\StringTok{"#F7B27E"}\NormalTok{))}

\NormalTok{## Adding the observed disparity through time}
\KeywordTok{plot}\NormalTok{(BeckLee_disparity, }\DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"#3E9CBA"}\NormalTok{, }\StringTok{"#98D4CF90"}\NormalTok{, }\StringTok{"#BFE4E390"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics{dispRity_manual_files/figure-latex/plot8-1}

}

\caption{The best fitted model (Trend) and the observed disparity through time}\label{fig:plot8}
\end{figure}

\hypertarget{disparity-as-a-distribution}{\section{Disparity as a
distribution}\label{disparity-as-a-distribution}}

Disparity is often regarded as a summary value of the position of the
all elements in the ordinated space. For example, the sum of variances,
the product of ranges or the median distance between the elements and
their centroid will summarise disparity as a single value. This value
can be pseudo-replicated (bootstrapped) to obtain a distribution of the
summary metric with estimated error. However, another way to perform
disparity analysis is to use the \emph{whole distribution} rather than
just a summary metric (e.g.~the variances or the ranges).

This is possible in the \texttt{dispRity} package by calculating
disparity as a dimension-level 2 metric only! Let's have a look using
our \protect\hyperlink{summarising-dispRity-data-plots}{previous
example} of bootstrapped time slices but by measuring the distances
between each taxon and their centroid as disparity.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Measuring disparity as a whole distribution}
\NormalTok{disparity_centroids <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(boot_time_slices, }\DataTypeTok{metric =}\NormalTok{ centroids)}
\end{Highlighting}
\end{Shaded}

The resulting disparity object is of dimension-level 2, so it can easily
be transformed into a dimension-level 1 object by, for example,
measuring the median distance of all these distributions:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Measuring median disparity in each time slice}
\NormalTok{disparity_centroids_median <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(disparity_centroids, }\DataTypeTok{metric =}\NormalTok{ median)}
\end{Highlighting}
\end{Shaded}

And we can now compare the differences between these methods:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Summarising both disparity measurements:}
\NormalTok{## The distributions:}
\KeywordTok{summary}\NormalTok{(disparity_centroids)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n obs.median bs.median  2.5%   25%   75% 97.5%
## 1     120  5      1.539     1.287 0.536 1.092 1.553 1.864
## 2      80 19      1.846     1.688 1.409 1.582 1.814 1.945
## 3      40 15      1.892     1.685 1.327 1.563 1.839 2.062
## 4       0 10      1.855     1.801 1.313 1.680 1.962 2.104
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The summary of the distributions (as median)}
\KeywordTok{summary}\NormalTok{(disparity_centroids_median)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   subsets  n   obs bs.median  2.5%   25%   75% 97.5%
## 1     120  5 1.508     1.287 0.535 0.943 1.477 1.508
## 2      80 19 1.790     1.682 1.534 1.645 1.716 1.791
## 3      40 15 1.689     1.678 1.552 1.635 1.717 1.822
## 4       0 10 1.910     1.804 1.577 1.761 1.855 1.909
\end{verbatim}

We can see that the summary message for the distribution is slightly
different than before. Here \texttt{summary} also displays the observed
central tendency (i.e.~the central tendency of the measured
distributions). Note that, as expected, this central tendency is the
same in both metrics!

Another, maybe more intuitive way, to compare both approaches for
measuring disparity is to plot the distributions:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical parameters}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{, }\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{))}

\NormalTok{## Plotting both disparity measurements}
\KeywordTok{plot}\NormalTok{(disparity_centroids, }\DataTypeTok{ylab =} \StringTok{"Distribution of all the distances"}\NormalTok{)}
\KeywordTok{plot}\NormalTok{(disparity_centroids_median,}
     \DataTypeTok{ylab =} \StringTok{"Distribution of the medians of all the distances"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-84-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{par}\NormalTok{(op)}
\end{Highlighting}
\end{Shaded}

We can then test for differences in the resulting distributions using
\texttt{test.dispRity} and the \texttt{bhatt.coeff} test as described
above.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Probability of overlap in the distribution of medians}
\KeywordTok{test.dispRity}\NormalTok{(disparity_centroids_median, }\DataTypeTok{test =}\NormalTok{ bhatt.coeff)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in test.dispRity(disparity_centroids_median, test = bhatt.coeff): Multiple p-values will be calculated without adjustment!
## This can inflate Type I error!
\end{verbatim}

\begin{verbatim}
##          bhatt.coeff
## 120 : 80   0.1905968
## 120 : 40   0.1936492
## 120 : 0    0.1381682
## 80 : 40    0.9416139
## 80 : 0     0.6076888
## 40 : 0     0.7124871
\end{verbatim}

In this case, we are looking at the probability of overlap of the
distribution of median distances from centroids among each pair of time
slices. In other words, we are measuring whether the medians from each
bootstrap pseudo-replicate for each time slice overlap. But of course,
we might be interested in the actual distribution of the distances from
the centroid rather than simply their central tendencies. This can be
problematic depending on the research question asked since we are
effectively comparing non-independent medians distributions (because of
the pseudo-replication).

One solution, therefore, is to look at the full distribution:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Probability of overlap for the full distributions}
\KeywordTok{test.dispRity}\NormalTok{(disparity_centroids, }\DataTypeTok{test =}\NormalTok{ bhatt.coeff)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in test.dispRity(disparity_centroids, test = bhatt.coeff): Multiple p-values will be calculated without adjustment!
## This can inflate Type I error!
\end{verbatim}

\begin{verbatim}
##          bhatt.coeff
## 120 : 80   0.6286805
## 120 : 40   0.6631398
## 120 : 0    0.5819930
## 80 : 40    0.9397416
## 80 : 0     0.8578102
## 40 : 0     0.9361860
\end{verbatim}

These results show the actual overlap among all the measured distances
from centroids concatenated across all the bootstraps. For example, when
comparing the slices 120 and 80, we are effectively comparing the 5
\(\times\) 100 distances (the distances of the five elements in slice
120 bootstrapped 100 times) to the 19 \(\times\) 100 distances from
slice 80. However, this can also be problematic for some specific tests
since the \emph{n} \(\times\) 100 distances are also pseudo-replicates
and thus are still not independent.

A second solution is to compare the distributions to each other
\emph{for each replicate}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Boostrapped probability of overlap for the full distributions}
\KeywordTok{test.dispRity}\NormalTok{(disparity_centroids, }\DataTypeTok{test =}\NormalTok{ bhatt.coeff, }\DataTypeTok{concatenate =} \OtherTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in test.dispRity(disparity_centroids, test = bhatt.coeff, concatenate = FALSE): Multiple p-values will be calculated without adjustment!
## This can inflate Type I error!
\end{verbatim}

\begin{verbatim}
##          bhatt.coeff      2.5%       25%       75%     97.5%
## 120 : 80   0.2507287 0.0000000 0.1450953 0.3829003 0.5431424
## 120 : 40   0.3184407 0.0000000 0.2000000 0.4312174 0.7651213
## 120 : 0    0.2368001 0.0000000 0.0000000 0.3464102 0.5791641
## 80 : 40    0.5739203 0.2252672 0.4678055 0.6965587 0.8184853
## 80 : 0     0.4649973 0.1382736 0.3688643 0.5757552 0.7292754
## 40 : 0     0.5418625 0.1954287 0.4385665 0.6767986 0.8082483
\end{verbatim}

These results show the median overlap among pairs of distributions in
the first column (\texttt{bhatt.coeff}) and then the distribution of
these overlaps among each pair of bootstraps. In other words, when two
distributions are compared, they are now compared for each bootstrap
pseudo-replicate, thus effectively creating a distribution of
probabilities of overlap. For example, when comparing the slices 120 and
80, we have a mean probability of overlap of 0.28 and a probability
between 0.18 and 0.43 in 50\% of the pseudo-replicates. Note that the
quantiles and central tendencies can be modified via the
\texttt{conc.quantiles} option.

\section{Disparity from other
matrices}\label{disparity-from-other-matrices}

In the example so far, disparity was measured from an ordinated
multidimensional space (i.e.~a PCO of the distances between taxa based
on discrete morphological characters). This is a common approach in
palaeobiology, morphometrics or ecology but ordinated matrices are not
mandatory for the \texttt{dispRity} package! It is totally possible to
perform the same analysis detailed above using other types of matrices
as long as your elements are rows in your matrix.

For example, we can use the data set \texttt{eurodist}, an \texttt{R}
inbuilt dataset that contains the distances (in km) between European
cities. We can check for example, if Northern European cities are closer
to each other than Southern ones:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Making the eurodist data set into a matrix (rather than "dist" object)}
\NormalTok{eurodist <-}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(eurodist)}
\NormalTok{eurodist[}\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{, }\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##           Athens Barcelona Brussels Calais Cherbourg
## Athens         0      3313     2963   3175      3339
## Barcelona   3313         0     1318   1326      1294
## Brussels    2963      1318        0    204       583
## Calais      3175      1326      204      0       460
## Cherbourg   3339      1294      583    460         0
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The two groups of cities}
\NormalTok{Northern <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Brussels"}\NormalTok{, }\StringTok{"Calais"}\NormalTok{, }\StringTok{"Cherbourg"}\NormalTok{, }\StringTok{"Cologne"}\NormalTok{, }\StringTok{"Copenhagen"}\NormalTok{,}
              \StringTok{"Hamburg"}\NormalTok{, }\StringTok{"Hook of Holland"}\NormalTok{, }\StringTok{"Paris"}\NormalTok{, }\StringTok{"Stockholm"}\NormalTok{)}
\NormalTok{Southern <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Athens"}\NormalTok{, }\StringTok{"Barcelona"}\NormalTok{, }\StringTok{"Geneva"}\NormalTok{, }\StringTok{"Gibraltar"}\NormalTok{, }\StringTok{"Lisbon"}\NormalTok{, }\StringTok{"Lyons"}\NormalTok{,}
              \StringTok{"Madrid"}\NormalTok{, }\StringTok{"Marseilles"}\NormalTok{, }\StringTok{"Milan"}\NormalTok{, }\StringTok{"Munich"}\NormalTok{, }\StringTok{"Rome"}\NormalTok{, }\StringTok{"Vienna"}\NormalTok{)}

\NormalTok{## Creating the subset dispRity object}
\NormalTok{eurodist_subsets <-}\StringTok{ }\KeywordTok{custom.subsets}\NormalTok{(eurodist, }\DataTypeTok{group =} \KeywordTok{list}\NormalTok{(}\StringTok{"Northern"}\NormalTok{ =}\StringTok{ }\NormalTok{Northern,}
                                                        \StringTok{"Southern"}\NormalTok{ =}\StringTok{ }\NormalTok{Southern))}

\NormalTok{## Bootstrapping and rarefying to 9 elements (the number of Northern cities)}
\NormalTok{eurodist_bs <-}\StringTok{ }\KeywordTok{boot.matrix}\NormalTok{(eurodist_subsets, }\DataTypeTok{rarefaction =} \DecValTok{9}\NormalTok{)}

\NormalTok{## Measuring disparity as the median distance from group's centroid}
\NormalTok{euro_disp <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(eurodist_bs, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(median, centroids))}

\NormalTok{## Testing the differences using a simple wilcox.test}
\NormalTok{euro_diff <-}\StringTok{ }\KeywordTok{test.dispRity}\NormalTok{(euro_disp, }\DataTypeTok{test =}\NormalTok{ wilcox.test)}
\NormalTok{euro_diff_rar <-}\StringTok{ }\KeywordTok{test.dispRity}\NormalTok{(euro_disp, }\DataTypeTok{test =}\NormalTok{ wilcox.test, }\DataTypeTok{rarefaction =} \DecValTok{9}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

We can compare this approach to an ordination one:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Ordinating the eurodist matrix}
\NormalTok{euro_ord <-}\StringTok{ }\KeywordTok{cmdscale}\NormalTok{(eurodist, }\DataTypeTok{k =} \KeywordTok{nrow}\NormalTok{(eurodist) }\OperatorTok{-}\StringTok{ }\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in cmdscale(eurodist, k = nrow(eurodist) - 2): only 11 of the first
## 19 eigenvalues are > 0
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating disparity on the bootstrapped and rarefied subset data}
\NormalTok{euro_ord_disp <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(}\KeywordTok{boot.matrix}\NormalTok{(}\KeywordTok{custom.subsets}\NormalTok{(euro_ord, }\DataTypeTok{group =}
        \KeywordTok{list}\NormalTok{(}\StringTok{"Northern"}\NormalTok{ =}\StringTok{ }\NormalTok{Northern, }\StringTok{"Southern"}\NormalTok{ =}\StringTok{ }\NormalTok{Southern)), }\DataTypeTok{rarefaction =} \DecValTok{9}\NormalTok{),}
        \DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(median, centroids))}

\NormalTok{## Testing the differences using a simple wilcox.test}
\NormalTok{euro_ord_diff <-}\StringTok{ }\KeywordTok{test.dispRity}\NormalTok{(euro_ord_disp, }\DataTypeTok{test =}\NormalTok{ wilcox.test)}
\NormalTok{euro_ord_diff_rar <-}\StringTok{ }\KeywordTok{test.dispRity}\NormalTok{(euro_ord_disp, }\DataTypeTok{test =}\NormalTok{ wilcox.test, }\DataTypeTok{rarefaction =} \DecValTok{9}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

And visualise the differences:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Plotting the differences}
\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{), }\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}
\NormalTok{## Plotting the normal disparity}
\KeywordTok{plot}\NormalTok{(euro_disp, }\DataTypeTok{main =} \StringTok{"Distance differences"}\NormalTok{)}
\NormalTok{## Adding the p-value}
\KeywordTok{text}\NormalTok{(}\FloatTok{1.5}\NormalTok{, }\DecValTok{4000}\NormalTok{, }\KeywordTok{paste0}\NormalTok{(}\StringTok{"p="}\NormalTok{,}\KeywordTok{round}\NormalTok{(euro_diff[[}\DecValTok{2}\NormalTok{]][[}\DecValTok{1}\NormalTok{]], }\DataTypeTok{digit =} \DecValTok{5}\NormalTok{)))}
\NormalTok{## Plotting the rarefied disparity}
\KeywordTok{plot}\NormalTok{(euro_disp, }\DataTypeTok{rarefaction =} \DecValTok{9}\NormalTok{, }\DataTypeTok{main =} \StringTok{"Distance differences (rarefied)"}\NormalTok{)}
\NormalTok{## Adding the p-value}
\KeywordTok{text}\NormalTok{(}\FloatTok{1.5}\NormalTok{, }\DecValTok{4000}\NormalTok{, }\KeywordTok{paste0}\NormalTok{(}\StringTok{"p="}\NormalTok{,}\KeywordTok{round}\NormalTok{(euro_diff_rar[[}\DecValTok{2}\NormalTok{]][[}\DecValTok{1}\NormalTok{]], }\DataTypeTok{digit =} \DecValTok{5}\NormalTok{)))}

\NormalTok{## Plotting the ordinated disparity}
\KeywordTok{plot}\NormalTok{(euro_ord_disp, }\DataTypeTok{main =} \StringTok{"Ordinated differences"}\NormalTok{)}
\NormalTok{## Adding the p-value}
\KeywordTok{text}\NormalTok{(}\FloatTok{1.5}\NormalTok{, }\DecValTok{1400}\NormalTok{, }\KeywordTok{paste0}\NormalTok{(}\StringTok{"p="}\NormalTok{,}\KeywordTok{round}\NormalTok{(euro_ord_diff[[}\DecValTok{2}\NormalTok{]][[}\DecValTok{1}\NormalTok{]], }\DataTypeTok{digit =} \DecValTok{5}\NormalTok{) ))}
\NormalTok{## Plotting the rarefied disparity}
\KeywordTok{plot}\NormalTok{(euro_ord_disp, }\DataTypeTok{rarefaction =} \DecValTok{9}\NormalTok{, }\DataTypeTok{main =} \StringTok{"Ordinated differences (rarefied)"}\NormalTok{)}
\NormalTok{## Adding the p-value}
\KeywordTok{text}\NormalTok{(}\FloatTok{1.5}\NormalTok{, }\DecValTok{1400}\NormalTok{, }\KeywordTok{paste0}\NormalTok{(}\StringTok{"p="}\NormalTok{,}\KeywordTok{round}\NormalTok{(euro_ord_diff_rar[[}\DecValTok{2}\NormalTok{]][[}\DecValTok{1}\NormalTok{]], }\DataTypeTok{digit =} \DecValTok{5}\NormalTok{) ))}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-90-1.pdf}

As expected, the results are pretty similar in pattern but different in
terms of scale. The median centroids distance is expressed in km in the
``Distance differences'' plots and in Euclidean units of variation in
the ``Ordinated differences'' plots.

\chapter{Making stuff up!}\label{making-stuff-up}

The \texttt{dispRity} package also offers some advanced data simulation
features to allow to test hypothesis, explore multidimensional spaces,
metrics properties or simply playing around with data! All the following
functions are based on the same modular architecture of the package and
therefore can be used with most of the functions of the package.

\section{Simulating discrete morphological
data}\label{simulating-discrete-morphological-data}

The function \texttt{sim.morpho} allows to simulate discrete
morphological data matrices (sometimes referred to as ``cladistic''
matrices). It allows to evolve multiple discrete characters on a given
phylogenetic tree, given different evolutionary models, rates, and
states. It even allows to include proper inapplicable data to make
datasets as messy as in real life!

In brief, the function \texttt{sim.morpho} takes a phylogenetic tree,
the number of required characters, the evolutionary model and a function
from which to draw the rates. The package also contains a function for
quickly checking the matrix's phylogenetic signal (as defined in
systematics not phylogenetic comparative methods) using parsimony. The
methods are described in details below

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{3}\NormalTok{)}
\NormalTok{## Simulating a starting tree with 15 taxa as a random coalescent tree}
\NormalTok{my_tree <-}\StringTok{ }\KeywordTok{rcoal}\NormalTok{(}\DecValTok{15}\NormalTok{)}

\NormalTok{## Generating a matrix with 100 characters (85% binary and 15% three state) and}
\NormalTok{## an equal rates model with a gamma rate distribution (0.5, 1) with no }
\NormalTok{## invariant characters.}
\NormalTok{my_matrix <-}\StringTok{ }\KeywordTok{sim.morpho}\NormalTok{(}\DataTypeTok{tree =}\NormalTok{ my_tree, }\DataTypeTok{characters =} \DecValTok{100}\NormalTok{, }\DataTypeTok{states =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.85}\NormalTok{,}
    \FloatTok{0.15}\NormalTok{), }\DataTypeTok{rates =} \KeywordTok{c}\NormalTok{(rgamma, }\DataTypeTok{shape =} \FloatTok{0.5}\NormalTok{, }\DataTypeTok{rate =} \DecValTok{1}\NormalTok{), }\DataTypeTok{invariant =} \OtherTok{FALSE}\NormalTok{)}

\NormalTok{## The first few lines of the matrix}
\NormalTok{my_matrix[}\DecValTok{1}\OperatorTok{:}\DecValTok{10}\NormalTok{, }\DecValTok{1}\OperatorTok{:}\DecValTok{10}\NormalTok{]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##     [,1] [,2] [,3] [,4] [,5] [,6] [,7] [,8] [,9] [,10]
## t15 "1"  "1"  "0"  "1"  "1"  "2"  "1"  "1"  "0"  "0"  
## t12 "1"  "1"  "0"  "1"  "1"  "2"  "1"  "1"  "0"  "0"  
## t14 "1"  "1"  "0"  "1"  "1"  "2"  "1"  "1"  "0"  "0"  
## t6  "1"  "1"  "0"  "1"  "1"  "2"  "1"  "1"  "0"  "0"  
## t3  "0"  "0"  "1"  "0"  "1"  "0"  "0"  "2"  "1"  "0"  
## t2  "0"  "0"  "1"  "0"  "1"  "0"  "0"  "2"  "1"  "0"  
## t11 "0"  "0"  "1"  "0"  "1"  "0"  "0"  "2"  "1"  "1"  
## t7  "0"  "0"  "1"  "0"  "1"  "0"  "0"  "2"  "1"  "0"  
## t1  "0"  "0"  "1"  "0"  "0"  "0"  "0"  "2"  "0"  "0"  
## t5  "0"  "0"  "1"  "0"  "0"  "0"  "0"  "2"  "0"  "0"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Checking the matrix properties with a quick Maximum Parsimony tree search}
\KeywordTok{check.morpho}\NormalTok{(my_matrix, my_tree)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                                     
## Maximum parsimony        139.0000000
## Consistency index          0.7625899
## Retention index            0.8881356
## Robinson-Foulds distance   0.0000000
\end{verbatim}

Note that this example produces a tree with a great consistency index
and an identical topology to the random coalescent tree!

\subsection{A more detailed
description}\label{a-more-detailed-description}

The protocol implemented here to generate discrete morphological
matrices is based on the ones developed in \citet{GuillermeCooper}
\citet{OReilly2016} \citet{puttick2017uncertain} and
\citet{OReilly2017}.

\begin{itemize}
\tightlist
\item
  The first \texttt{tree} argument will be the tree on which to
  ``evolve'' the characters and therefore requires branch length. You
  can generate quick and easy random Yule trees using
  \texttt{ape::rtree(number\_of\_taxa)} but I would advise to use more
  realistic trees for more realistic simulations based on more realistic
  models (really realistic then) using the function \texttt{tree.bd}
  from the
  \href{http://www.zoology.ubc.ca/prog/diversitree/}{\texttt{diversitree}}
  package \citep{fitzjohndiversitree2012} for example.
\item
  The second argument, \texttt{character} is the number of characters.
\item
  The third, \texttt{states} is the proportion of characters states
  above two. This argument intakes the proportion of \emph{n}-states
  characters, for example \texttt{states\ =\ c(0.5,0.3,0.2)} will
  generate 50\% of binary-state characters, 30\% of three-state
  characters and 20\% of four-state characters. There is no limit in the
  number of state characters proportion as long as the total makes up
  100\%.
\item
  The fourth, \texttt{model} is the evolutionary model for generating
  the character(s). More about this below.
\item
  The fifth and sixth, \texttt{rates} and \texttt{substitution} are the
  model parameters described below as well.
\item
  Finally, the two logical arguments, are self explanatory:
  \texttt{invariant} whether to allow invariant characters
  (i.e.~characters that don't change) and \texttt{verbose} whether to
  print the simulation progress on your console.
\end{itemize}

\subsubsection{Available evolutionary
models}\label{available-evolutionary-models}

There are currently three evolutionary models implemented in
\texttt{sim.morpho} but more will come in the future. Note also that
they allow fine tuning parameters making them pretty plastic!

\begin{itemize}
\tightlist
\item
  \texttt{"ER"}: this model allows any number of character states and is
  based on the Mk model \citep{lewisa2001}. It assumes a unique overall
  evolutionary rate equal substitution rate between character states.
  This model is based on the \texttt{ape::rTraitDisc} function.
\item
  \texttt{"HKY"}: this is binary state character model based on the
  molecular HKY model \citep{HKY85}. It uses the four molecular states
  (A,C,G,T) with a unique overall evolutionary rate and a biased
  substitution rate towards transitions (A \textless{}-\textgreater{} G
  or C \textless{}-\textgreater{} T) against transvertions (A
  \textless{}-\textgreater{} C and G \textless{}-\textgreater{} T).
  After evolving the nucleotides, this model transforms them into binary
  states by converting the purines (A and G) into state 0 and the
  pyrimidines (C and T) into state 1. This method is based on the
  \texttt{phyclust::seq.gen.HKY} function and was first proposed by
  \citet{OReilly2016}.
\item
  \texttt{"MIXED"}: this model uses a random (uniform) mix between both
  the \texttt{"ER"} and the \texttt{"HKY"} models.
\end{itemize}

The models can take the following parameters: (1) \texttt{rates} is the
evolutionary rate (i.e.~the rate of changes along a branch; the
evolutionary speed) and (2) \texttt{substitution} is the frequency of
changes between one state or another. For example if a character can
have high probability of changing (the \emph{evolutionary} rate) with,
each time a change occurs a probability of changing from state \emph{X}
to state \emph{Y} (the \emph{substitution} rate). Note that in the
\texttt{"ER"} model, the substitution rate is ignore because by
definition this rate is equal!

The parameters arguments \texttt{rates} and \texttt{substitution} takes
a distributions from which to draw the parameters values for each
character. For example, if you want an \texttt{"HKY"} model with an
evolutionary rate (i.e.~speed) drawn from a uniform distribution bounded
between 0.001 and 0.005, you can define it as
\texttt{rates\ =\ c(runif,\ min\ =\ 0.001,\ max\ =\ 0.005)},
\texttt{runif} being the function for random draws from a uniform
distribution and \texttt{max} and \texttt{min} being the distribution
parameters. These distributions should always be passed in the format
\texttt{c(random\_distribution\_function,\ distribution\_parameters)}
with the names of the distribution parameters arguments.

\subsubsection{Checking the results}\label{checking-the-results}

An additional function, \texttt{check.morpho} runs a quick Maximum
Parsimony tree search using the \texttt{phangorn} parsimony algorithm.
It quickly calculates the parsimony score, the consistency and retention
indices and, if a tree is provided (e.g.~the tree used to generate the
matrix) it calculates the Robinson-Foulds distance between the most
parsimonious tree and the provided tree to determine how different they
are.

\subsubsection{Adding inapplicable
characters}\label{adding-inapplicable-characters}

Once a matrix is generated, it is possible to apply inapplicable
characters to it for increasing realism! Inapplicable characters are
commonly designated as \texttt{NA} or simply \texttt{-}. They differ
from missing characters \texttt{?} in their nature by being inapplicable
rather than unknown. For example, considering a binary character defined
as ``colour of the tail'' with the following states ``blue'' and
``red''; on a taxa with no tail, the character should be coded as
inapplicable (``\texttt{-}'') since the state of the character ``colour
of tail'' is \emph{known}: it's neither ``blue'' or ``red'', it's just
not there! It contrasts with coding it as missing (``\texttt{?}'' - also
called as ambiguous) where the state is \emph{unknown}, for example, the
taxon of interest is a fossil where the tail has no colour preserved or
is not present at all due to bad conservation!

This type of characters can be added to the simulated matrices using the
\texttt{apply.NA} function. It takes, as arguments, the \texttt{matrix},
the source of inapplicability (\texttt{NAs} - more below), the
\texttt{tree} used to generate the matrix and the two same
\texttt{invariant} and \texttt{verbose} arguments as defined above. The
\texttt{NAs} argument allows two types of sources of inapplicability:

\begin{itemize}
\tightlist
\item
  \texttt{"character"} where the inapplicability is due to the character
  (e.g.~coding a character tail for species with no tail). In practice,
  the algorithm chooses a character \emph{X} as the underlying character
  (e.g. ``presence and absence of tail''), arbitrarily chooses one of
  the states as ``absent'' (e.g.~0 = absent) and changes in the next
  character \emph{Y} any state next to character \emph{X} state 0 into
  an inapplicable token (``\texttt{-}''). This simulates the
  inapplicability induced by coding the characters (i.e.~not always
  biological).
\item
  \texttt{"clade"} where the inapplicability is due to evolutionary
  history (e.g.~a clade loosing its tail). In practice, the algorithm
  chooses a random clade in the tree and a random character \emph{Z} and
  replaces the state of the taxa present in the clade by the
  inapplicable token (``\texttt{-}''). This simulates the
  inapplicability induced by evolutionary biology (e.g.~the lose of a
  feature in a clade).
\end{itemize}

To apply these sources of inapplicability, simply repeat the number of
inapplicable sources for the desired number of characters with
inapplicable data.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Generating 5 "character" NAs and 10 "clade" NAs}
\NormalTok{my_matrix_NA <-}\StringTok{ }\KeywordTok{apply.NA}\NormalTok{(my_matrix, }\DataTypeTok{tree =}\NormalTok{ my_tree,}
                         \DataTypeTok{NAs =} \KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\StringTok{"character"}\NormalTok{, }\DecValTok{5}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\StringTok{"clade"}\NormalTok{, }\DecValTok{10}\NormalTok{)))}

\NormalTok{## The first few lines of the resulting matrix}
\NormalTok{my_matrix_NA[}\DecValTok{1}\OperatorTok{:}\DecValTok{10}\NormalTok{, }\DecValTok{90}\OperatorTok{:}\DecValTok{100}\NormalTok{]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##     [,1] [,2] [,3] [,4] [,5] [,6] [,7] [,8] [,9] [,10] [,11]
## t15 "0"  "1"  "0"  "1"  "0"  "0"  "0"  "1"  "0"  "1"   "0"  
## t12 "1"  "1"  "0"  "1"  "0"  "0"  "0"  "1"  "0"  "1"   "0"  
## t14 "0"  "0"  "0"  "1"  "0"  "0"  "0"  "1"  "0"  "1"   "0"  
## t6  "0"  "0"  "0"  "1"  "0"  "0"  "0"  "1"  "0"  "1"   "0"  
## t3  "0"  "0"  "1"  "0"  "-"  "0"  "0"  "0"  "-"  "0"   "1"  
## t2  "0"  "0"  "1"  "0"  "-"  "0"  "0"  "0"  "-"  "0"   "1"  
## t11 "0"  "0"  "1"  "0"  "-"  "0"  "0"  "0"  "-"  "0"   "1"  
## t7  "0"  "0"  "1"  "0"  "-"  "0"  "0"  "0"  "-"  "0"   "1"  
## t1  "0"  "0"  "0"  "0"  "-"  "0"  "0"  "0"  "-"  "2"   "1"  
## t5  "0"  "0"  "0"  "0"  "-"  "0"  "0"  "0"  "-"  "0"   "1"
\end{verbatim}

\subsection{Parameters for a realistic(ish)
matrix}\label{parameters-for-a-realisticish-matrix}

There are many parameters that can create a ``realistic'' matrix
(i.e.~not too different from the input tree with a consistency and
retention index close to what is seen in the literature) but because of
the randomness of the matrix generation not all parameters combination
end up creating ``good'' matrices. The following parameters however,
seem to generate fairly ``realist'' matrices with a starting coalescent
tree, equal rates model with 0.85 binary characters and 0.15 three state
characters, a gamma distribution with a shape parameter (\(\alpha\)) of
10 and no scaling (\(\beta\) = 1) with a rate of 50.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{0}\NormalTok{)}
\NormalTok{## tree}
\NormalTok{my_tree <-}\StringTok{ }\KeywordTok{rcoal}\NormalTok{(}\DecValTok{30}\NormalTok{)}
\NormalTok{## matrix}
\NormalTok{morpho_mat <-}\StringTok{ }\KeywordTok{sim.morpho}\NormalTok{(}\DataTypeTok{tree =}\NormalTok{ my_tree, }\DataTypeTok{characters =} \DecValTok{100}\NormalTok{, }\DataTypeTok{states =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.85}\NormalTok{,}
    \FloatTok{0.15}\NormalTok{), }\DataTypeTok{rates =} \KeywordTok{c}\NormalTok{(rgamma, }\DataTypeTok{shape =} \DecValTok{10}\NormalTok{, }\DataTypeTok{rate =} \DecValTok{50}\NormalTok{), }\DataTypeTok{invariant =} \OtherTok{FALSE}\NormalTok{)}

\KeywordTok{check.morpho}\NormalTok{(morpho_mat, my_tree)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                                     
## Maximum parsimony        133.0000000
## Consistency index          0.7819549
## Retention index            0.9597222
## Robinson-Foulds distance   4.0000000
\end{verbatim}

\section{Simulating multidimensional
spaces}\label{simulating-multidimensional-spaces}

Another way to simulate data is to directly simulate a multidimensional
space with the \texttt{space.maker} function. This function allows users
to simulate spaces with a certain number of properties. For example, it
is possible to design a space with a specific distribution on each axis,
a correlation between the axes and a specific cumulative variance per
axis. This can be useful for creating ordinated spaces for null
hypothesis, for example if you're using the function \texttt{null.test}
\citep{diaz2016global}.

This function takes as arguments the number of elements (data points -
\texttt{elements} argument)~and dimensions (\texttt{dimensions}
argument) to create the space and the distribution functions to be used
for each axis. The distributions are passed through the
\texttt{distribution} argument as modular functions. You can either pass
a single distribution function for all the axes (for example
\texttt{distribution\ =\ runif} for all the axis being uniform) or a
specific distribution function for each specific axis (for example
\texttt{distribution\ =\ c(runif,\ rnorm,\ rgamma))} for the first axis
being uniform, the second normal and the third gamma). You can of course
use your very own functions or use the ones implemented in
\texttt{dispRity} for more complex ones (see below). Specific optional
arguments for each of these distributions can be passed as a list via
the \texttt{arguments} argument.

Furthermore, it is possible to add a correlation matrix to add a
correlation between the axis via the \texttt{cor.matrix} argument or
even a vector of proportion of variance to be bear by each axis via the
\texttt{scree} argument to simulate realistic ordinated spaces.

Here is a simple two dimensional example:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical options}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}

\NormalTok{## A square space}
\NormalTok{square_space <-}\StringTok{ }\KeywordTok{space.maker}\NormalTok{(}\DecValTok{100}\NormalTok{, }\DecValTok{2}\NormalTok{, runif)}

\NormalTok{## The resulting 2D matrix}
\KeywordTok{head}\NormalTok{(square_space)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##           [,1]       [,2]
## [1,] 0.3322704 0.73741888
## [2,] 0.5157255 0.33160759
## [3,] 0.3716847 0.62284731
## [4,] 0.4367013 0.76535328
## [5,] 0.7424290 0.07764732
## [6,] 0.0111013 0.65066885
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Visualising the space}
\KeywordTok{plot}\NormalTok{(square_space, }\DataTypeTok{pch =} \DecValTok{20}\NormalTok{, }\DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{main =} \StringTok{"Uniform 2D space"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-94-1.pdf}

Of course, more complex spaces can be created by changing the
distributions, their arguments or adding a correlation matrix or a
cumulative variance vector:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## A plane space: uniform with one dimensions equal to 0}
\NormalTok{plane_space <-}\StringTok{ }\KeywordTok{space.maker}\NormalTok{(}\DecValTok{2500}\NormalTok{, }\DecValTok{3}\NormalTok{, }\KeywordTok{c}\NormalTok{(runif, runif, runif),}
                           \DataTypeTok{arguments =} \KeywordTok{list}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{0}\NormalTok{), }\OtherTok{NULL}\NormalTok{, }\OtherTok{NULL}\NormalTok{))}
\KeywordTok{head}\NormalTok{(plane_space)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##      [,1]       [,2]       [,3]
## [1,]    0 0.18294640 0.32682604
## [2,]    0 0.67243526 0.77182827
## [3,]    0 0.56028603 0.06603386
## [4,]    0 0.80409259 0.19401581
## [5,]    0 0.31732158 0.16460302
## [6,]    0 0.09244429 0.99873117
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Correlation matrix for a 3D space}
\NormalTok{(cor_matrix <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\KeywordTok{cbind}\NormalTok{(}\DecValTok{1}\NormalTok{, }\FloatTok{0.8}\NormalTok{, }\FloatTok{0.2}\NormalTok{, }\FloatTok{0.8}\NormalTok{, }\DecValTok{1}\NormalTok{, }\FloatTok{0.7}\NormalTok{, }\FloatTok{0.2}\NormalTok{, }\FloatTok{0.7}\NormalTok{, }\DecValTok{1}\NormalTok{), }\DataTypeTok{nrow =} \DecValTok{3}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##      [,1] [,2] [,3]
## [1,]  1.0  0.8  0.2
## [2,]  0.8  1.0  0.7
## [3,]  0.2  0.7  1.0
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## An ellipsoid space (normal space with correlation)}
\NormalTok{ellipse_space <-}\StringTok{ }\KeywordTok{space.maker}\NormalTok{(}\DecValTok{2500}\NormalTok{, }\DecValTok{3}\NormalTok{, rnorm, }\DataTypeTok{cor.matrix =}\NormalTok{ cor_matrix)}
\KeywordTok{head}\NormalTok{(ellipse_space)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##            [,1]       [,2]       [,3]
## [1,] -0.5080285 -0.7014059 -1.0286500
## [2,]  0.7250674  0.9498248  0.7220638
## [3,]  0.4476787  0.2102857 -0.3835125
## [4,] -1.3252827 -0.5060518  0.2211104
## [5,] -0.9193155 -1.3861888 -1.6151720
## [6,] -0.7217653 -0.1967257  0.2672965
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## A cylindrical space with decreasing axes variance}
\NormalTok{cylindrical_space <-}\StringTok{ }\KeywordTok{space.maker}\NormalTok{(}\DecValTok{2500}\NormalTok{, }\DecValTok{3}\NormalTok{, }\KeywordTok{c}\NormalTok{(rnorm, rnorm, runif),}
                                 \DataTypeTok{scree =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.7}\NormalTok{, }\FloatTok{0.2}\NormalTok{, }\FloatTok{0.1}\NormalTok{))}
\KeywordTok{head}\NormalTok{(cylindrical_space)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##             [,1]        [,2]         [,3]
## [1,] -0.07935545  0.10932383 0.0017309562
## [2,] -0.20619425 -0.14809591 0.0008416656
## [3,] -0.20133258  0.05766707 0.0041888464
## [4,]  0.69412491 -0.47106729 0.0036967387
## [5,]  0.41300602  0.23924098 0.0077537936
## [6,]  0.53369953  0.02882589 0.0073476805
\end{verbatim}

See below for the visualisation of the spaces.

\subsection{Personalised dimensions
distributions}\label{personalised-dimensions-distributions}

Following the modular architecture of the package, it is of course
possible to pass home made distribution functions to the
\texttt{distribution} argument. For example, the \texttt{random.circle}
function is a personalised one implemented in \texttt{dispRity}. This
function allows to create circles based on trigonometry allowing two
axis to covary to produce circle coordinates. By default, this function
generates two sets of coordinates with a \texttt{distribution} argument
and a minimum and maximum boundary (\texttt{inner} and \texttt{outer}
respectively) to create nice sharp edges to the circle. The maximum
boundary is equivalent to the radius of the circle (it removes
coordinates beyond the circle radius) and the minimum is equivalent to
the radius of a smaller circle with no data (it removes coordinates
below this inner circle radius).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical options}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}

\NormalTok{## Generating coordinates for a normal circle with a upper boundary of 1}
\NormalTok{circle <-}\StringTok{ }\KeywordTok{random.circle}\NormalTok{(}\DecValTok{1000}\NormalTok{, rnorm, }\DataTypeTok{inner =} \DecValTok{0}\NormalTok{, }\DataTypeTok{outer =} \DecValTok{1}\NormalTok{)}

\NormalTok{## Plotting the circle}
\KeywordTok{plot}\NormalTok{(circle, }\DataTypeTok{xlab =} \StringTok{"x"}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{"y"}\NormalTok{, }\DataTypeTok{main =} \StringTok{"A normal circle"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-96-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating doughnut space (a spherical space with a hole)}
\NormalTok{doughnut_space <-}\StringTok{ }\KeywordTok{space.maker}\NormalTok{(}\DecValTok{5000}\NormalTok{, }\DecValTok{3}\NormalTok{, }\KeywordTok{c}\NormalTok{(rnorm, random.circle),}
     \DataTypeTok{arguments =} \KeywordTok{list}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{mean =} \DecValTok{0}\NormalTok{), }\KeywordTok{list}\NormalTok{(runif, }\DataTypeTok{inner =} \FloatTok{0.5}\NormalTok{, }\DataTypeTok{outer =} \DecValTok{1}\NormalTok{)))}
\KeywordTok{head}\NormalTok{(doughnut_space)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##            [,1]        [,2]       [,3]
## [1,]  1.0692065 -0.34286592 -0.4077478
## [2,]  0.3898210 -0.06393936  0.5212298
## [3,] -0.9878641  0.36223496  0.4519167
## [4,]  0.1031183  0.08781562  0.9176893
## [5,]  1.1351951 -0.34076407  0.8712020
## [6,]  0.1769964 -0.77904629  0.5770467
\end{verbatim}

See below for the visualisation of the ``doughnut'' space.

\subsection{Visualising the space}\label{visualising-the-space}

I suggest using the excellent \texttt{scatterplot3d} package to play
around and visualise the simulated spaces:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical options}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =}\NormalTok{ (}\KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{)), }\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}
\NormalTok{## Visualising 3D spaces}
\KeywordTok{require}\NormalTok{(scatterplot3d)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Loading required package: scatterplot3d
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The plane space}
\KeywordTok{scatterplot3d}\NormalTok{(plane_space, }\DataTypeTok{pch =} \DecValTok{20}\NormalTok{, }\DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{zlab =} \StringTok{""}\NormalTok{,}
              \DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{), }\DataTypeTok{main =} \StringTok{"Plane space"}\NormalTok{)}

\NormalTok{## The ellipsoid space}
\KeywordTok{scatterplot3d}\NormalTok{(ellipse_space, }\DataTypeTok{pch =} \DecValTok{20}\NormalTok{, }\DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{zlab =} \StringTok{""}\NormalTok{,}
              \DataTypeTok{main =} \StringTok{"Normal ellipsoid space"}\NormalTok{)}

\NormalTok{## A cylindrical space with a decreasing variance per axis}
\KeywordTok{scatterplot3d}\NormalTok{(cylindrical_space, }\DataTypeTok{pch =} \DecValTok{20}\NormalTok{, }\DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{zlab =} \StringTok{""}\NormalTok{,}
              \DataTypeTok{main =} \StringTok{"Normal cylindrical space"}\NormalTok{)}
\NormalTok{## Axes have different orders of magnitude}

\NormalTok{## Plotting the doughnut space}
\KeywordTok{scatterplot3d}\NormalTok{(doughnut_space[,}\KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{1}\NormalTok{,}\DecValTok{3}\NormalTok{)], }\DataTypeTok{pch =} \DecValTok{20}\NormalTok{, }\DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{,}
              \DataTypeTok{zlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{main =} \StringTok{"Doughnut space"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-97-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{par}\NormalTok{(op)}
\end{Highlighting}
\end{Shaded}

\subsection{Generating realistic
spaces}\label{generating-realistic-spaces}

It is possible to generate ``realistic'' spaces by simply extracting the
parameters of an existing space and scaling it up to the simulated
space. For example, we can extract the parameters of the
\texttt{BeckLee\_mat50} ordinated space and simulate a similar space.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Loading the data}
\KeywordTok{data}\NormalTok{(BeckLee_mat50)}

\NormalTok{## Number of dimensions}
\NormalTok{obs_dim <-}\StringTok{ }\KeywordTok{ncol}\NormalTok{(BeckLee_mat50)}

\NormalTok{## Observed correlation between the dimensions}
\NormalTok{obs_correlations <-}\StringTok{ }\KeywordTok{cor}\NormalTok{(BeckLee_mat50)}

\NormalTok{## Observed mean and standard deviation per axis}
\NormalTok{obs_mu_sd_axis <-}\StringTok{ }\KeywordTok{mapply}\NormalTok{(}\ControlFlowTok{function}\NormalTok{(x,y) }\KeywordTok{list}\NormalTok{(}\StringTok{"mean"}\NormalTok{ =}\StringTok{ }\NormalTok{x, }\StringTok{"sd"}\NormalTok{ =}\StringTok{ }\NormalTok{y),}
                         \KeywordTok{as.list}\NormalTok{(}\KeywordTok{apply}\NormalTok{(BeckLee_mat50, }\DecValTok{2}\NormalTok{, mean)),}
                         \KeywordTok{as.list}\NormalTok{(}\KeywordTok{apply}\NormalTok{(BeckLee_mat50, }\DecValTok{2}\NormalTok{, sd)), }\DataTypeTok{SIMPLIFY =} \OtherTok{FALSE}\NormalTok{)}

\NormalTok{## Observed overall mean and standard deviation}
\NormalTok{obs_mu_sd_glob <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}\StringTok{"mean"}\NormalTok{ =}\StringTok{ }\KeywordTok{mean}\NormalTok{(BeckLee_mat50), }\StringTok{"sd"}\NormalTok{ =}\StringTok{ }\KeywordTok{sd}\NormalTok{(BeckLee_mat50))}

\NormalTok{## Scaled observed variance per axis (scree plot)}
\NormalTok{obs_scree <-}\StringTok{ }\KeywordTok{variances}\NormalTok{(BeckLee_mat50)}\OperatorTok{/}\KeywordTok{sum}\NormalTok{(}\KeywordTok{variances}\NormalTok{(BeckLee_mat50))}

\NormalTok{## Generating our simulated space}
\NormalTok{simulated_space <-}\StringTok{ }\KeywordTok{space.maker}\NormalTok{(}\DecValTok{1000}\NormalTok{, }\DataTypeTok{dimensions =}\NormalTok{ obs_dim, }
                               \DataTypeTok{distribution =} \KeywordTok{rep}\NormalTok{(}\KeywordTok{list}\NormalTok{(rnorm), obs_dim),}
                               \DataTypeTok{arguments =}\NormalTok{ obs_mu_sd_axis,}
                               \DataTypeTok{cor.matrix =}\NormalTok{ obs_correlations)}

\NormalTok{## Visualising the fit of our data in the space (in the two first dimensions)}
\KeywordTok{plot}\NormalTok{(simulated_space[,}\DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{], }\DataTypeTok{xlab =} \StringTok{"PC1"}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{"PC2"}\NormalTok{)}
\KeywordTok{points}\NormalTok{(BeckLee_mat50[,}\DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{], }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{20}\NormalTok{)}
\KeywordTok{legend}\NormalTok{(}\StringTok{"topleft"}\NormalTok{, }\DataTypeTok{legend =} \KeywordTok{c}\NormalTok{(}\StringTok{"observed"}\NormalTok{, }\StringTok{"simulated"}\NormalTok{),}
        \DataTypeTok{pch =} \KeywordTok{c}\NormalTok{(}\DecValTok{20}\NormalTok{,}\DecValTok{21}\NormalTok{), }\DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"red"}\NormalTok{, }\StringTok{"black"}\NormalTok{), }\DataTypeTok{bg =} \StringTok{"white"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-98-1.pdf}

It is now possible to simulate a space using these observed arguments to
test several hypothesis:

\begin{itemize}
\tightlist
\item
  Is the space uniform or normal?
\item
  If the space is normal, is the mean and variance global or specific
  for each axis?
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Measuring disparity as the sum of variance}
\NormalTok{observed_disp <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(BeckLee_mat50, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(median, centroids))}

\NormalTok{## Is the space uniform?}
\NormalTok{test_unif <-}\StringTok{ }\KeywordTok{null.test}\NormalTok{(observed_disp, }\DataTypeTok{null.distrib =}\NormalTok{ runif)}

\NormalTok{## Is the space normal with a mean of 0 and a sd of 1?}
\NormalTok{test_norm1 <-}\StringTok{ }\KeywordTok{null.test}\NormalTok{(observed_disp, }\DataTypeTok{null.distrib =}\NormalTok{ rnorm)}

\NormalTok{## Is the space normal with the observed mean and sd and cumulative variance}
\NormalTok{test_norm2 <-}\StringTok{ }\KeywordTok{null.test}\NormalTok{(observed_disp, }\DataTypeTok{null.distrib =} \KeywordTok{rep}\NormalTok{(}\KeywordTok{list}\NormalTok{(rnorm), obs_dim),}
                        \DataTypeTok{null.args =} \KeywordTok{rep}\NormalTok{(}\KeywordTok{list}\NormalTok{(obs_mu_sd_glob), obs_dim),}
                        \DataTypeTok{null.scree =}\NormalTok{ obs_scree)}

\NormalTok{## Is the space multiple normal with multiple means and sds and a correlation?}
\NormalTok{test_norm3 <-}\StringTok{ }\KeywordTok{null.test}\NormalTok{(observed_disp, }\DataTypeTok{null.distrib =} \KeywordTok{rep}\NormalTok{(}\KeywordTok{list}\NormalTok{(rnorm), obs_dim),}
                        \DataTypeTok{null.args =}\NormalTok{ obs_mu_sd_axis, }\DataTypeTok{null.cor =}\NormalTok{ obs_correlations)}

\NormalTok{## Graphical options}
\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =}\NormalTok{ (}\KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{)), }\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}
\NormalTok{## Plotting the results}
\KeywordTok{plot}\NormalTok{(test_unif, }\DataTypeTok{main =} \StringTok{"Uniform (0,1)"}\NormalTok{)}
\KeywordTok{plot}\NormalTok{(test_norm1, }\DataTypeTok{main =} \StringTok{"Normal (0,1)"}\NormalTok{)}
\KeywordTok{plot}\NormalTok{(test_norm2, }\DataTypeTok{main =} \StringTok{"Normal (global observed}\CharTok{\textbackslash{}n}\StringTok{mean and sd)"}\NormalTok{)}
\KeywordTok{plot}\NormalTok{(test_norm3, }\DataTypeTok{main =} \StringTok{"Normal (variable mean}\CharTok{\textbackslash{}n}\StringTok{and sd with correlation)"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-99-1.pdf}

If we measure disparity as the median distance from the morphospace
centroid, we can explain the distribution of the data as normal with the
variable observed mean and standard deviation and with a correlation
between the dimensions.

\chapter{\texorpdfstring{The guts of the \texttt{dispRity}
package}{The guts of the dispRity package}}\label{the-guts-of-the-disprity-package}

\section{\texorpdfstring{Manipulating \texttt{dispRity}
objects}{Manipulating dispRity objects}}\label{manipulating-disprity-objects}

Disparity analysis involves a lot of manipulation of many matrices
(especially when bootstrapping) which can be impractical to visualise
and will quickly overwhelm your \texttt{R} console. Even the simple Beck
and Lee 2014 example above produces an object with \textgreater{} 72
lines of lists of lists of matrices!

Therefore \texttt{dispRity} uses a specific class of object called a
\texttt{dispRity} object. These objects allow users to use S3 method
functions such as \texttt{summary.dispRity}, \texttt{plot.dispRity} and
\texttt{print.dispRity}. \texttt{dispRity} also contains various utility
functions that manipulate the \texttt{dispRity} object (e.g.
\texttt{sort.dispRity}, \texttt{extract.dispRity} see the full list in
the next section). These functions modify the \texttt{dispRity} object
without having to delve into its complex structure! The full structure
of a \texttt{dispRity} object is detailed
\href{https://github.com/TGuillerme/dispRity/blob/master/disparity_object.md}{here}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Loading the example data}
\KeywordTok{data}\NormalTok{(disparity)}

\NormalTok{## What is the class of the median_centroids object?}
\KeywordTok{class}\NormalTok{(disparity)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] "dispRity"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## What does the object contain?}
\KeywordTok{names}\NormalTok{(disparity)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] "matrix"    "call"      "subsets"   "disparity"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Summarising it using the S3 method print.dispRity}
\NormalTok{disparity}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 7 continuous (acctran) time subsets for 99 elements with 97 dimensions:
##      90, 80, 70, 60, 50 ...
## Data was bootstrapped 100 times (method:"full") and rarefied to 20, 15, 10, 5 elements.
## Disparity was calculated as: c(median, centroids).
\end{verbatim}

Note that it is always possible to recall the full object using the
argument \texttt{all\ =\ TRUE} in \texttt{print.dispRity}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Display the full object}
\KeywordTok{print}\NormalTok{(disparity, }\DataTypeTok{all =} \OtherTok{TRUE}\NormalTok{)}
\NormalTok{## This is more nearly ~ 5000 lines on my 13 inch laptop screen!}
\end{Highlighting}
\end{Shaded}

\section{\texorpdfstring{\texttt{dispRity}
utilities}{dispRity utilities}}\label{disprity-utilities}

The package also provides some utility functions to facilitate
multidimensional analysis.

\subsection{\texorpdfstring{\texttt{dispRity} object utilities
}{dispRity object utilities }}\label{disprity-object-utilities}

The first set of utilities are functions for manipulating
\texttt{dispRity} objects:

\subsubsection{\texorpdfstring{\texttt{make.dispRity}}{make.dispRity}}\label{make.disprity}

This function creates empty \texttt{dispRity} objects.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating an empty dispRity object}
\KeywordTok{make.dispRity}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Empty dispRity object.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating an "empty" dispRity object with a matrix}
\NormalTok{(disparity_obj <-}\StringTok{ }\KeywordTok{make.dispRity}\NormalTok{(}\KeywordTok{matrix}\NormalTok{(}\KeywordTok{rnorm}\NormalTok{(}\DecValTok{20}\NormalTok{), }\DecValTok{5}\NormalTok{, }\DecValTok{4}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## Contains only a matrix 5x4.
\end{verbatim}

\subsubsection{\texorpdfstring{\texttt{fill.dispRity}}{fill.dispRity}}\label{fill.disprity}

This function initialises a \texttt{dispRity} object and generates its
call properties.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The dispRity object's call is indeed empty}
\NormalTok{disparity_obj}\OperatorTok{$}\NormalTok{call}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## list()
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Filling an empty disparity object (that needs to contain at least a matrix)}
\NormalTok{(disparity_obj <-}\StringTok{ }\KeywordTok{fill.dispRity}\NormalTok{(disparity_obj))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 5 elements with 4 dimensions.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The dipRity object has now the correct minimal attributes}
\NormalTok{disparity_obj}\OperatorTok{$}\NormalTok{call}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## $dimensions
## [1] 4
\end{verbatim}

\subsubsection{\texorpdfstring{\texttt{matrix.dispRity}}{matrix.dispRity}}\label{matrix.disprity}

This function extracts a specific matrix from a disparity object. The
matrix can be one of the bootstrapped matrices or/and a rarefied matrix.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Extracting the matrix containing the coordinates of the elements at time 50}
\KeywordTok{str}\NormalTok{(}\KeywordTok{matrix.dispRity}\NormalTok{(disparity, }\StringTok{"50"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  num [1:18, 1:97] -0.1038 0.2844 0.2848 0.0927 0.1619 ...
##  - attr(*, "dimnames")=List of 2
##   ..$ : chr [1:18] "Leptictis" "Dasypodidae" "n24" "Potamogalinae" ...
##   ..$ : NULL
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Extracting the 3rd bootstrapped matrix with the 2nd rarefaction level}
\NormalTok{## (15 elements) from the second group (80 Mya)}
\KeywordTok{str}\NormalTok{(}\KeywordTok{matrix.dispRity}\NormalTok{(disparity, }\DataTypeTok{subsets =} \DecValTok{1}\NormalTok{, }\DataTypeTok{bootstrap =} \DecValTok{3}\NormalTok{, }\DataTypeTok{rarefaction =} \DecValTok{2}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  num [1:15, 1:97] -0.7161 0.3496 -0.573 -0.0445 -0.1427 ...
##  - attr(*, "dimnames")=List of 2
##   ..$ : chr [1:15] "n7" "n34" "Maelestes" "n20" ...
##   ..$ : NULL
\end{verbatim}

\subsubsection{\texorpdfstring{\texttt{get.subsets}}{get.subsets}}\label{get.subsets}

This function creates a dispRity object that contains only elements from
one specific subsets.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Extracting all the data for the crown mammals}
\NormalTok{(crown_mammals <-}\StringTok{ }\KeywordTok{get.subsets}\NormalTok{(disp_crown_stemBS, }\StringTok{"Group.crown"}\NormalTok{))}

\NormalTok{## The object keeps the properties of the parent object but is composed of only one subsets}
\KeywordTok{length}\NormalTok{(crown_mammals}\OperatorTok{$}\NormalTok{subsets)}
\end{Highlighting}
\end{Shaded}

\subsubsection{\texorpdfstring{\texttt{combine.subsets}}{combine.subsets}}\label{combine.subsets}

This function allows to merge different subsets. If the

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Combine the two first subsets in the dispRity data example}
\KeywordTok{combine.subsets}\NormalTok{(disparity, }\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

Note that the computed values (bootstrapped data + disparity metric) are
not merge.

\subsubsection{\texorpdfstring{\texttt{extract.dispRity}}{extract.dispRity}}\label{extract.disprity}

This function extracts the calculated disparity values of a specific
matrix.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Extracting the observed disparity (default)}
\KeywordTok{extract.dispRity}\NormalTok{(disparity)}

\NormalTok{## Extracting the disparity from the bootstrapped values from the}
\NormalTok{## 10th rarefaction level from the second subsets (80 Mya)}
\KeywordTok{extract.dispRity}\NormalTok{(disparity, }\DataTypeTok{observed =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{subsets =} \DecValTok{2}\NormalTok{, }\DataTypeTok{rarefaction =} \DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\subsubsection{\texorpdfstring{\texttt{rescale.dispRity}}{rescale.dispRity}}\label{rescale.disprity}

This is the modified S3 method for \texttt{scale} (scaling and/or
centring) that can be applied to the disparity data of a
\texttt{dispRity} object and can take optional arguments (for example
the rescaling by dividing by a maximum value).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Getting the disparity values of the time subsets}
\KeywordTok{head}\NormalTok{(}\KeywordTok{summary}\NormalTok{(disparity))}

\NormalTok{## Scaling the same disparity values}
\KeywordTok{head}\NormalTok{(}\KeywordTok{summary}\NormalTok{(}\KeywordTok{rescale.dispRity}\NormalTok{(disparity, }\DataTypeTok{scale =} \OtherTok{TRUE}\NormalTok{)))}

\NormalTok{## Scaling and centering:}
\KeywordTok{head}\NormalTok{(}\KeywordTok{summary}\NormalTok{(}\KeywordTok{rescale.dispRity}\NormalTok{(disparity, }\DataTypeTok{scale =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{center =} \OtherTok{TRUE}\NormalTok{)))}

\NormalTok{## Rescaling the value by dividing by a maximum value}
\KeywordTok{head}\NormalTok{(}\KeywordTok{summary}\NormalTok{(}\KeywordTok{rescale.dispRity}\NormalTok{(disparity, }\DataTypeTok{max =} \DecValTok{10}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\subsubsection{\texorpdfstring{\texttt{sort.dispRity}}{sort.dispRity}}\label{sort.disprity}

This is the S3 method of \texttt{sort} for sorting the subsets
alphabetically (default) or following a specific pattern.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Sorting the disparity subsets in inverse alphabetic order}
\KeywordTok{head}\NormalTok{(}\KeywordTok{summary}\NormalTok{(}\KeywordTok{sort}\NormalTok{(disparity, }\DataTypeTok{decreasing =} \OtherTok{TRUE}\NormalTok{)))}

\NormalTok{## Customised sorting}
\KeywordTok{head}\NormalTok{(}\KeywordTok{summary}\NormalTok{(}\KeywordTok{sort}\NormalTok{(disparity, }\DataTypeTok{sort =} \KeywordTok{c}\NormalTok{(}\DecValTok{7}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{3}\NormalTok{, }\DecValTok{4}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{6}\NormalTok{))))}
\end{Highlighting}
\end{Shaded}

\section{\texorpdfstring{The \texttt{dispRity} object
content}{The dispRity object content}}\label{the-disprity-object-content}

The functions above are utilities to easily and safely access different
elements in the \texttt{dispRity} object. Alternatively, of course, each
elements can be accessed manually. Here is an explanation on how it
works. The \texttt{dispRity} object is a \texttt{list} of two to four
elements, each of which are detailed below:

\begin{itemize}
\tightlist
\item
  \texttt{\$matrix}: an object of class \texttt{matrix}, the full
  multidimensional space.
\item
  \texttt{\$call}: an object of class \texttt{list} containing
  information on the \texttt{dispRity} object content.
\item
  \texttt{\$subsets}: an object of class \texttt{list} containing the
  subsets of the multidimensional space.
\item
  \texttt{\$disparity}: an object of class \texttt{list} containing the
  disparity values.
\end{itemize}

The \texttt{dispRity} object is loosely based on \texttt{C} structure
objects. In fact, it is composed of one unique instance of a matrix (the
multidimensional space) upon which the metric function is called via
``pointers'' to only a certain number of elements and/or dimensions of
this matrix. This allows for: (1) faster and easily tractable execution
time: the metric functions are called through apply family function and
can be parallelised; and (2) a really low memory footprint: at any time,
only one matrix is present in the \texttt{R} environment rather than
multiple copies of it for each subset.

\subsection{\texorpdfstring{\texttt{\$matrix}}{\$matrix}}\label{matrix}

This is the multidimensional space, stored in the \texttt{R} environment
as a \texttt{matrix} object. It requires row names but not column names.
By default, if the row names are missing, \texttt{dispRity} function
will arbitrarily generate them in numeric order (i.e.
\texttt{rownames(matrix)\ \textless{}-\ 1:nrow(matrix)}). This element
of the \texttt{dispRity} object is never modified.

\subsection{\texorpdfstring{\texttt{\$call}}{\$call}}\label{call}

This element contains the information on the \texttt{dispRity} object
content. It is a \texttt{list} that can contain the following:

\begin{itemize}
\tightlist
\item
  \texttt{\$call\$subsets}: a vector of \texttt{character} with
  information on the subsets type (either \texttt{"continuous"},
  \texttt{"discrete"} or \texttt{"custom"}) and their eventual model
  (\texttt{"acctran"}, \texttt{"deltran"}, \texttt{"random"},
  \texttt{"proximity"}, \texttt{"equal.split"},
  \texttt{"gradual.split"}). This element generated only once via
  \texttt{chrono.subsets()} and \texttt{custom.subsets()}.
\item
  \texttt{\$call\$dimensions}: either a single \texttt{numeric} value
  indicating how many dimensions to use or a vector of \texttt{numeric}
  values indicating which specific dimensions to use. This element is by
  default the number of columns in \texttt{\$matrix} but can be modified
  through \texttt{boot.matrix()} or \texttt{dispRity()}.
\item
  \texttt{\$call\$bootstrap}: this is a \texttt{list} containing three
  elements:

  \begin{itemize}
  \tightlist
  \item
    \texttt{{[}{[}1{]}{]}}: the number of bootstrap replicates
    (\texttt{numeric})
  \item
    \texttt{{[}{[}2{]}{]}}: the bootstrap method (\texttt{character})
  \item
    \texttt{{[}{[}3{]}{]}}: the rarefaction levels (\texttt{numeric}
    vector)
  \end{itemize}
\item
  \texttt{\$call\$disparity}: this is a \texttt{list} containing one
  element, \texttt{\$metric}, that is a \texttt{list} containing the
  different functions passed to the \texttt{metric} argument in
  \texttt{dispRity}. These are \texttt{call} elements and get modified
  each time the \texttt{dispRity} function is used (the first element is
  the first metric(s), the second, the second metric(s), etc.).
\end{itemize}

\subsection{\texorpdfstring{\texttt{\$subsets}}{\$subsets}}\label{subsets}

This element contain the eventual subsets of the multidimensional space.
It is a \texttt{list} of subset names. Each subset name is in turn a
\texttt{list} of at least one element called \texttt{elements} which is
in turn a \texttt{matrix}. This \texttt{elements} matrix is the raw
(observed) elements in the subsets. The \texttt{elements} matrix is
composed of \texttt{numeric} values in one column and \emph{n} rows (the
number of elements in the subset). Each of these values are a
``pointer'' (\texttt{C} inspired) to the element of the
\texttt{\$matrix}. For example, lets assume a \texttt{dispRity} object
called \texttt{disparity}, composed of at least one subsets called
\texttt{sub1}:

\begin{verbatim}
 disparity$subsets$sub1$elements
      [,1]
 [1,]    5
 [2,]    4
 [3,]    6
 [4,]    7
\end{verbatim}

The values in the matrix ``point'' to the elements in \texttt{\$matrix}:
here, the multidimensional space with only the 4th, 5th, 6th and 7th
elements. The following elements in \texttt{diparity\$subsets\$sub1}
will correspond to the same ``pointers'' but drawn from the bootstrap
replicates. The columns will correspond to different bootstrap
replicates. For example:

\begin{verbatim}
 disparity$subsets$sub1[[2]]
      [,1] [,2] [,3] [,4]
 [1,]   57   43   70    4
 [2,]   43   44    4    4
 [3,]   42   84   44    1
 [4,]   84    7    2   10
\end{verbatim}

This signifies that we have four bootstrap pseudo-replicates pointing
each time to four elements in \texttt{\$matrix}. The next element
(\texttt{{[}{[}3{]}{]}}) will be the same for the eventual first
rarefaction level (i.e.~the resulting bootstrap matrix will have
\emph{m} rows where \emph{m} is the number of elements for this
rarefaction level). The next element after that (\texttt{{[}{[}4{]}{]}})
will be the same for with an other rarefaction level and so
forth\ldots{}

\subsection{\texorpdfstring{\texttt{\$disparity}}{\$disparity}}\label{disparity}

The \texttt{\$disparity} element is identical to the \texttt{\$subsets}
element structure (a list of list(s) containing matrices) but the
matrices don't contain ``pointers'' to \texttt{\$matrix} but the
disparity result of the disparity metric applied to the ``pointers''.
For example, in our first example (\texttt{\$elements}) from above, if
the disparity metric is of dimensions level 1, we would have:

\begin{verbatim}
 disparity$disparity$sub1$elements
      [,1]
 [1,]    1.82
\end{verbatim}

This is the observed disparity (1.82) for the subset called
\texttt{sub1}. If the disparity metric is of dimension level 2 (say the
function \texttt{range} that outputs two values), we would have:

\begin{verbatim}
 disparity$disparity$sub1$elements
      [,1]
 [1,]    0.82
 [2,]    2.82
\end{verbatim}

The following elements in the list follow the same logic as before: rows
are disparity values (one row for a dimension level 1 metric, multiple
for a dimensions level 2 metric) and columns are the bootstrap
replicates (the bootstrap with all elements followed by the eventual
rarefaction levels). For example for the bootstrap without rarefaction
(second element of the list):

\begin{verbatim}
 disparity$disparity$sub1[[2]]
         [,1]     [,2]     [,3]     [,4]
[1,] 1.744668 1.777418 1.781624 1.739679 
\end{verbatim}

\chapter{Palaeobiology demo: disparity-through-time and within
groups}\label{palaeobiology-demo-disparity-through-time-and-within-groups}

This demo aims to give quick overview of the \texttt{dispRity} package
(v.1.2) for palaeobiology analyses of disparity, including disparity
through time analyses.

This demo showcases a typical disparity-through-time analysis: we are
going to test whether the disparity changed through time in a subset of
eutherian mammals from the last 100 million years using a dataset from
Beck and Lee (2014).

\section{Before starting}\label{before-starting}

\subsection{The morphospace}\label{the-morphospace}

In this example, we are going to use a subset of the data from
\citet{beckancient2014}. See the
\protect\hyperlink{example-data}{example data} description for more
details. Briefly, this dataset contains an ordinated matrix of 50
discrete characters from mammals (\texttt{BeckLee\_mat50}), another
matrix of the same 50 mammals and the estimated discrete data characters
of their descendants (thus 50 + 49 rows, \texttt{BeckLee\_mat99}), a
dataframe containing the ages of each taxon in the dataset
(\texttt{BeckLee\_ages}) and finally a phylogenetic tree with the
relationships among the 50 mammals (\texttt{BeckLee\_tree}). The
ordinated matrix will represent our full morphospace, i.e.~all the
mammalian morphologies that ever existed through time (for this
dataset).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Loading demo and the package data}
\KeywordTok{library}\NormalTok{(dispRity)}

\NormalTok{## Setting the random seed for repeatability}
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{123}\NormalTok{)}

\NormalTok{## Loading the ordinated matrix/morphospace:}
\KeywordTok{data}\NormalTok{(BeckLee_mat50)}
\KeywordTok{head}\NormalTok{(BeckLee_mat50[,}\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{])}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                    [,1]          [,2]        [,3]       [,4]      [,5]
## Cimolestes   -0.5319679  0.1117759259  0.09865194 -0.1933148 0.2035833
## Maelestes    -0.4087147  0.0139690317  0.26268300  0.2297096 0.1310953
## Batodon      -0.6923194  0.3308625215 -0.10175223 -0.1899656 0.1003108
## Bulaklestes  -0.6802291 -0.0134872777  0.11018009 -0.4103588 0.4326298
## Daulestes    -0.7386111  0.0009001369  0.12006449 -0.4978191 0.4741342
## Uchkudukodon -0.5105254 -0.2420633915  0.44170317 -0.1172972 0.3602273
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{dim}\NormalTok{(BeckLee_mat50)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 50 48
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## The morphospace contains 50 taxa and has 48 dimensions (or axes)}

\NormalTok{## Showing a list of first and last occurrences data for some fossils}
\KeywordTok{data}\NormalTok{(BeckLee_ages)}
\KeywordTok{head}\NormalTok{(BeckLee_ages)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##             FAD  LAD
## Adapis     37.2 36.8
## Asioryctes 83.6 72.1
## Leptictis  33.9 33.3
## Miacis     49.0 46.7
## Mimotona   61.6 59.2
## Notharctus 50.2 47.0
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Plotting a phylogeny}
\KeywordTok{data}\NormalTok{(BeckLee_tree)}
\KeywordTok{plot}\NormalTok{(BeckLee_tree, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{)}
\KeywordTok{axisPhylo}\NormalTok{(}\DataTypeTok{root =} \DecValTok{140}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-110-1.pdf}

\begin{quote}
You can have an even nicer looking tree if you use the \texttt{strap}
package!
\end{quote}

\begin{Shaded}
\begin{Highlighting}[]
\ControlFlowTok{if}\NormalTok{(}\OperatorTok{!}\KeywordTok{require}\NormalTok{(strap)) }\KeywordTok{install.packages}\NormalTok{(}\StringTok{"strap"}\NormalTok{)}
\NormalTok{strap}\OperatorTok{::}\KeywordTok{geoscalePhylo}\NormalTok{(BeckLee_tree, }\DataTypeTok{cex.tip =} \FloatTok{0.7}\NormalTok{, }\DataTypeTok{cex.ts =} \FloatTok{0.6}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-111-1.pdf}

\section{A disparity-through-time
analysis}\label{a-disparity-through-time-analysis}

\subsection{Splitting the morphospace through
time}\label{splitting-the-morphospace-through-time}

One of the crucial steps in disparity-through-time analysis is to split
the full morphospace into smaller time subsets that contain the total
number of morphologies at certain points in time (time-slicing) or
during certain periods in time (time-binning). Basically, the full
morphospace represents the total number of morphologies across all time
and will be greater than any of the time subsets of the morphospace.

The \texttt{dispRity} package provides a \texttt{chrono.subsets}
function that allows users to split the morphospace into time slices
(using \texttt{method\ =\ continuous}) or into time bins (using
\texttt{method\ =\ discrete}). In this example, we are going to split
the morphospace into five equal time bins of 20 million years long from
100 million years ago to the present. We will also provide to the
function a table containing the first and last occurrences dates for
some fossils to take into account that some fossils might occur in
several of our different time bins.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating the vector of time bins ages}
\NormalTok{(time_bins <-}\StringTok{ }\KeywordTok{rev}\NormalTok{(}\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{0}\NormalTok{, }\DataTypeTok{to =} \DecValTok{100}\NormalTok{, }\DataTypeTok{by =} \DecValTok{20}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 100  80  60  40  20   0
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Splitting the morphospace using the chrono.subsets function}
\NormalTok{(binned_morphospace <-}\StringTok{ }\KeywordTok{chrono.subsets}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ BeckLee_mat50, }\DataTypeTok{tree =}\NormalTok{ BeckLee_tree,}
    \DataTypeTok{method =} \StringTok{"discrete"}\NormalTok{, }\DataTypeTok{time =}\NormalTok{ time_bins, }\DataTypeTok{inc.nodes =} \OtherTok{FALSE}\NormalTok{,}
    \DataTypeTok{FADLAD =}\NormalTok{ BeckLee_ages))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 5 discrete time subsets for 50 elements:
##     100 - 80, 80 - 60, 60 - 40, 40 - 20, 20 - 0.
\end{verbatim}

The output object is a \texttt{dispRity} object (see more about that
\protect\hyperlink{The-guts-of-the-dispRity-package}{here}. In brief,
however, \texttt{dispRity} objects are lists of different elements
(i.e.~disparity results, morphospace time subsets, morphospace
attributes, etc.) that display only a summary of the object when calling
the object to avoiding filling the \texttt{R} console with superfluous
output.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Printing the class of the object}
\KeywordTok{class}\NormalTok{(binned_morphospace)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] "dispRity"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Printing the content of the object}
\KeywordTok{str}\NormalTok{(binned_morphospace)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## List of 3
##  $ matrix : num [1:50, 1:48] -0.532 -0.409 -0.692 -0.68 -0.739 ...
##   ..- attr(*, "dimnames")=List of 2
##   .. ..$ : chr [1:50] "Cimolestes" "Maelestes" "Batodon" "Bulaklestes" ...
##   .. ..$ : NULL
##  $ call   :List of 1
##   ..$ subsets: chr "discrete"
##  $ subsets:List of 5
##   ..$ 100 - 80:List of 1
##   .. ..$ elements: int [1:8, 1] 5 4 6 8 43 10 11 42
##   ..$ 80 - 60 :List of 1
##   .. ..$ elements: int [1:15, 1] 7 8 9 1 2 3 12 13 14 44 ...
##   ..$ 60 - 40 :List of 1
##   .. ..$ elements: int [1:13, 1] 41 49 24 25 26 27 28 21 22 19 ...
##   ..$ 40 - 20 :List of 1
##   .. ..$ elements: int [1:6, 1] 15 39 40 35 23 47
##   ..$ 20 - 0  :List of 1
##   .. ..$ elements: int [1:10, 1] 36 37 38 32 33 34 50 48 29 30
##  - attr(*, "class")= chr "dispRity"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{names}\NormalTok{(binned_morphospace)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] "matrix"  "call"    "subsets"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Printing the object as a dispRity class}
\NormalTok{binned_morphospace}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 5 discrete time subsets for 50 elements:
##     100 - 80, 80 - 60, 60 - 40, 40 - 20, 20 - 0.
\end{verbatim}

\begin{quote}
These objects will gradual.splitly contain more information when
completing the following steps in the disparity-through-time analysis.
\end{quote}

\subsection{Bootstrapping the data}\label{bootstrapping-the-data}

Once we obtain our different time subsets, we can bootstrap and rarefy
them (i.e.~pseudo-replicating the data). The bootstrapping allows us to
make each subset more robust to outliers and the rarefaction allows us
to compare subsets with the same number of taxa to remove sampling
biases (i.e.~more taxa in one subset than the others). The
\texttt{boot.matrix} function bootstraps the \texttt{dispRity} object
and the \texttt{rarefaction} option within performs rarefaction.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Bootstrapping each time subset 100 times (default)}
\NormalTok{(boot_bin_morphospace <-}\StringTok{ }\KeywordTok{boot.matrix}\NormalTok{(binned_morphospace))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 5 discrete time subsets for 50 elements with 48 dimensions:
##     100 - 80, 80 - 60, 60 - 40, 40 - 20, 20 - 0.
## Data was bootstrapped 100 times (method:"full").
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Getting the minimum number of rows (i.e. taxa) in the time subsets}
\KeywordTok{min}\NormalTok{(}\KeywordTok{size.subsets}\NormalTok{(boot_bin_morphospace))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 6
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Bootstrapping each time subset 100 times and rarefying them }
\NormalTok{(rare_bin_morphospace <-}\StringTok{ }\KeywordTok{boot.matrix}\NormalTok{(binned_morphospace, }\DataTypeTok{bootstraps =} \DecValTok{100}\NormalTok{,}
    \DataTypeTok{rarefaction =} \DecValTok{6}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 5 discrete time subsets for 50 elements with 48 dimensions:
##     100 - 80, 80 - 60, 60 - 40, 40 - 20, 20 - 0.
## Data was bootstrapped 100 times (method:"full") and rarefied to 6 elements.
\end{verbatim}

\subsection{Calculating disparity}\label{calculating-disparity}

We can now calculate the disparity within each time subsets along with
some confidence intervals generated by the pseudoreplication step above
(bootstraps/rarefaction). Disparity can be calculated in many ways and
this package allows users to come up with their own disparity metrics.
For more details, please refer to the
\protect\hyperlink{disparity-metrics}{\texttt{dispRity} metric section}.

In this example, we are going to calculate the spread of the data in
each time subset by calculating disparity as the sum of the variance of
each dimension of the morphospace in each time subset using the
\texttt{dispRity} function. Thus, in this example, disparity is defined
by the multi-dimensional variance of each time subset (i.e.~the spread
of the taxa within the morphospace). Note that this metric comes with a
caveat (not solved here) since it ignores covariances among the
dimensions of the morphospace. We use this here because it is a standard
metric used in disparity-through-time analysis \citep{Wills1994}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating disparity for the bootstrapped data}
\NormalTok{(boot_disparity <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(boot_bin_morphospace, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, variances)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 5 discrete time subsets for 50 elements with 48 dimensions:
##     100 - 80, 80 - 60, 60 - 40, 40 - 20, 20 - 0.
## Data was bootstrapped 100 times (method:"full").
## Disparity was calculated as: c(sum, variances).
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating disparity for the rarefied data}
\NormalTok{(rare_disparity <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(rare_bin_morphospace, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(sum, variances)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 5 discrete time subsets for 50 elements with 48 dimensions:
##     100 - 80, 80 - 60, 60 - 40, 40 - 20, 20 - 0.
## Data was bootstrapped 100 times (method:"full") and rarefied to 6 elements.
## Disparity was calculated as: c(sum, variances).
\end{verbatim}

The \texttt{dispRity} function does not actually display the calculated
disparity values but rather only the properties of the disparity object
(size, subsets, metric, etc.). To display the actual calculated scores,
we need to summarise the disparity object using the S3 method
\texttt{summary} that is applied to a \texttt{dispRity} object (see
\texttt{?summary.dispRity} for more details).

As for any \texttt{R} package, you can refer to the help files for each
individual function for more details.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Summarising the disparity results}
\KeywordTok{summary}\NormalTok{(boot_disparity)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##    subsets  n   obs bs.median  2.5%   25%   75% 97.5%
## 1 100 - 80  8 1.675     1.488 1.087 1.389 1.568 1.648
## 2  80 - 60 15 1.782     1.679 1.538 1.631 1.728 1.792
## 3  60 - 40 13 1.913     1.772 1.607 1.734 1.826 1.886
## 4  40 - 20  6 2.022     1.707 1.212 1.537 1.822 1.942
## 5   20 - 0 10 1.971     1.794 1.598 1.716 1.842 1.890
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(rare_disparity)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##    subsets  n   obs bs.median  2.5%   25%   75% 97.5%
## 1 100 - 80  8 1.675     1.484 1.194 1.400 1.547 1.636
## 2 100 - 80  6    NA     1.477 0.993 1.361 1.569 1.698
## 3  80 - 60 15 1.782     1.674 1.517 1.600 1.725 1.793
## 4  80 - 60  6    NA     1.655 1.299 1.532 1.754 1.882
## 5  60 - 40 13 1.913     1.767 1.601 1.714 1.829 1.861
## 6  60 - 40  6    NA     1.787 1.314 1.672 1.879 1.984
## 7  40 - 20  6 2.022     1.736 1.281 1.603 1.822 1.948
## 8   20 - 0 10 1.971     1.807 1.595 1.729 1.856 1.917
## 9   20 - 0  6    NA     1.790 1.435 1.718 1.873 1.995
\end{verbatim}

\begin{quote}
The summary.dispRity function comes with many options on which values to
calculate (central tendency and quantiles) and on how many digits to
display. Refer to the function's manual for more details.
\end{quote}

\subsection{Plotting the results}\label{plotting-the-results}

It is sometimes easier to visualise the results in a plot than in a
table. For that we can use the \texttt{plot} S3 function to plot the
\texttt{dispRity} objects (see \texttt{?plot.dispRity} for more
details).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical options}
\KeywordTok{quartz}\NormalTok{(}\DataTypeTok{width =} \DecValTok{10}\NormalTok{, }\DataTypeTok{height =} \DecValTok{5}\NormalTok{) ; }\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =}\NormalTok{ (}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{)), }\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}

\NormalTok{## Plotting the bootstrapped and rarefied results}
\KeywordTok{plot}\NormalTok{(boot_disparity, }\DataTypeTok{type =} \StringTok{"continuous"}\NormalTok{, }\DataTypeTok{main =} \StringTok{"bootstrapped results"}\NormalTok{)}
\KeywordTok{plot}\NormalTok{(rare_disparity, }\DataTypeTok{type =} \StringTok{"continuous"}\NormalTok{, }\DataTypeTok{main =} \StringTok{"rarefied results"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-117-1.pdf}

\section{Testing differences}\label{testing-differences}

Finally, to draw some valid conclusions from these results, we can apply
some statistical tests. We can test, for example, if mammalian disparity
changed significantly through time over the last 100 million years. To
do so, we can compare the means of each time-bin in a sequential manner
to see whether the disparity in bin \emph{n} is equal to the disparity
in bin \emph{n+1}, and whether this is in turn equal to the disparity in
bin \emph{n+2}, etc. Because our data is temporally autocorrelated
(i.e.~what happens in bin \emph{n+1} depends on what happened in bin
\emph{n}) and pseudoreplicated (i.e.~each bootstrap draw creates
non-independent time subsets because they are all based on the same time
subsets), we apply a non-parametric mean comparison: the
\texttt{wilcox.test}. Also, we need to apply a p-value correction
(e.g.~Bonferroni correction) to correct for multiple testing (see
\texttt{?p.adjust} for more details).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Testing the differences between bins in the bootstrapped dataset.}
\KeywordTok{test.dispRity}\NormalTok{(boot_disparity, }\DataTypeTok{test =}\NormalTok{ wilcox.test, }\DataTypeTok{comparison =} \StringTok{"sequential"}\NormalTok{,}
    \DataTypeTok{correction =} \StringTok{"bonferroni"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [[1]]
##                    statistic: W
## 100 - 80 : 80 - 60          471
## 80 - 60 : 60 - 40          1562
## 60 - 40 : 40 - 20          6250
## 40 - 20 : 20 - 0           3725
## 
## [[2]]
##                         p.value
## 100 - 80 : 80 - 60 7.427563e-28
## 80 - 60 : 60 - 40  1.798899e-16
## 60 - 40 : 40 - 20  9.061511e-03
## 40 - 20 : 20 - 0   7.379715e-03
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Testing the differences between bins in the rarefied dataset.}
\KeywordTok{test.dispRity}\NormalTok{(rare_disparity, }\DataTypeTok{test =}\NormalTok{ wilcox.test, }\DataTypeTok{comparison =} \StringTok{"sequential"}\NormalTok{,}
    \DataTypeTok{correction =} \StringTok{"bonferroni"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [[1]]
##                    statistic: W
## 100 - 80 : 80 - 60          662
## 80 - 60 : 60 - 40          1814
## 60 - 40 : 40 - 20          5752
## 40 - 20 : 20 - 0           3621
## 
## [[2]]
##                         p.value
## 100 - 80 : 80 - 60 1.214988e-25
## 80 - 60 : 60 - 40  2.823697e-14
## 60 - 40 : 40 - 20  2.653018e-01
## 40 - 20 : 20 - 0   3.026079e-03
\end{verbatim}

Here our results show significant changes in disparity through time
between all time bins (all p-values \textless{} 0.05). However, when
looking at the rarefied results, there is no significant difference
between the time bins in the Palaeogene (60-40 to 40-20 Mya), suggesting
that the differences detected in the first test might just be due to the
differences in number of taxa sampled (13 or 6 taxa) in each time bin.

\chapter{Ecology demo}\label{ecology-demo}

This is an example of typical disparity analysis that can be performed
in ecology.

\section{Data}\label{data}

For this example, we will use the famous \texttt{iris} inbuilt data set

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{data}\NormalTok{(iris)}
\end{Highlighting}
\end{Shaded}

This data contains petal and sepal length for 150 individual plants
sorted into three species.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Separating the species}
\NormalTok{species <-}\StringTok{ }\NormalTok{iris[,}\DecValTok{5}\NormalTok{]}
\NormalTok{## Which species?}
\KeywordTok{unique}\NormalTok{(species)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] setosa     versicolor virginica 
## Levels: setosa versicolor virginica
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Separating the petal/sepal length}
\NormalTok{measurements <-}\StringTok{ }\NormalTok{iris[,}\DecValTok{1}\OperatorTok{:}\DecValTok{4}\NormalTok{]}
\KeywordTok{head}\NormalTok{(measurements)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   Sepal.Length Sepal.Width Petal.Length Petal.Width
## 1          5.1         3.5          1.4         0.2
## 2          4.9         3.0          1.4         0.2
## 3          4.7         3.2          1.3         0.2
## 4          4.6         3.1          1.5         0.2
## 5          5.0         3.6          1.4         0.2
## 6          5.4         3.9          1.7         0.4
\end{verbatim}

We can then ordinate the data using a PCA (\texttt{prcomp} function)
thus defining our four dimensional space as the poetically named
petal-space.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Ordinating the data}
\NormalTok{ordination <-}\StringTok{ }\KeywordTok{prcomp}\NormalTok{(measurements)}

\NormalTok{## The petal-space}
\NormalTok{petal_space <-}\StringTok{ }\NormalTok{ordination}\OperatorTok{$}\NormalTok{x}

\NormalTok{## Adding the elements names to the petal-space (the individuals IDs)}
\KeywordTok{rownames}\NormalTok{(petal_space) <-}\StringTok{ }\DecValTok{1}\OperatorTok{:}\KeywordTok{nrow}\NormalTok{(petal_space)}
\end{Highlighting}
\end{Shaded}

\section{Classic analysis}\label{classic-analysis}

A classical way to represent this ordinated data would be to use two
dimensional plots to look at how the different species are distributed
in the petal-space.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Measuring the variance on each axis}
\NormalTok{variances <-}\StringTok{ }\KeywordTok{apply}\NormalTok{(petal_space, }\DecValTok{2}\NormalTok{, var)}
\NormalTok{variances <-}\StringTok{ }\NormalTok{variances}\OperatorTok{/}\KeywordTok{sum}\NormalTok{(variances)}

\NormalTok{## Graphical option}
\KeywordTok{par}\NormalTok{(}\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}

\NormalTok{## A classic 2D ordination plot}
\KeywordTok{plot}\NormalTok{(petal_space[, }\DecValTok{1}\NormalTok{], petal_space[, }\DecValTok{2}\NormalTok{], }\DataTypeTok{col =}\NormalTok{ species,}
    \DataTypeTok{xlab =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"PC 1 ("}\NormalTok{, }\KeywordTok{round}\NormalTok{(variances[}\DecValTok{1}\NormalTok{], }\DecValTok{2}\NormalTok{), }\StringTok{")"}\NormalTok{),}
    \DataTypeTok{ylab =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"PC 2 ("}\NormalTok{, }\KeywordTok{round}\NormalTok{(variances[}\DecValTok{2}\NormalTok{], }\DecValTok{2}\NormalTok{), }\StringTok{")"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-122-1.pdf}

This shows the distribution of the different species in the petal-space
along the two first axis of variation. This is a pretty standard way to
visualise the multidimensional space and further analysis might be
necessary to test wether the groups are different such as a linear
discriminant analysis (LDA). However, in this case we are ignoring the
two other dimensions of the ordination! If we look at the two other axis
we see a totally different result:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Plotting the two second axis of the petal-space}
\KeywordTok{plot}\NormalTok{(petal_space[, }\DecValTok{3}\NormalTok{], petal_space[, }\DecValTok{4}\NormalTok{], }\DataTypeTok{col =}\NormalTok{ species,}
    \DataTypeTok{xlab =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"PC 3 ("}\NormalTok{, }\KeywordTok{round}\NormalTok{(variances[}\DecValTok{3}\NormalTok{], }\DecValTok{2}\NormalTok{), }\StringTok{")"}\NormalTok{),}
    \DataTypeTok{ylab =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"PC 4 ("}\NormalTok{, }\KeywordTok{round}\NormalTok{(variances[}\DecValTok{4}\NormalTok{], }\DecValTok{2}\NormalTok{), }\StringTok{")"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-123-1.pdf}

Additionally, these two represented dimensions do not represent a
biological reality \emph{per se}; i.e.~the values on the first dimension
do not represent a continuous trait (e.g.~petal length), instead they
just represent the ordinations of correlations between the data and some
factors.

Therefore, we might want to approach this problem without getting stuck
in only two dimensions and consider the whole dataset as a
\emph{n}-dimensional object.

\section{\texorpdfstring{A multidimensional approach with
\texttt{dispRity}}{A multidimensional approach with dispRity}}\label{a-multidimensional-approach-with-disprity}

The first step is to create different subsets that represent subsets of
the ordinated space (i.e.~sub-regions within the \emph{n}-dimensional
object). Each of these subsets will contain only the individuals of a
specific species.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Creating the table that contain the elements and their attributes}
\NormalTok{petal_subsets <-}\StringTok{ }\KeywordTok{custom.subsets}\NormalTok{(petal_space, }\DataTypeTok{group =} \KeywordTok{list}\NormalTok{(}
                                \StringTok{"setosa"}\NormalTok{ =}\StringTok{ }\KeywordTok{which}\NormalTok{(species }\OperatorTok{==}\StringTok{ "setosa"}\NormalTok{),}
                                \StringTok{"versicolor"}\NormalTok{ =}\StringTok{ }\KeywordTok{which}\NormalTok{(species }\OperatorTok{==}\StringTok{ "versicolor"}\NormalTok{),}
                                \StringTok{"virginica"}\NormalTok{ =}\StringTok{ }\KeywordTok{which}\NormalTok{(species }\OperatorTok{==}\StringTok{ "virginica"}\NormalTok{)))}

\NormalTok{## Visualising the dispRity object content}
\NormalTok{petal_subsets}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 3 customised subsets for 150 elements:
##     setosa, versicolor, virginica.
\end{verbatim}

This created a \texttt{dispRity} object (more about that
\protect\hyperlink{guts}{here}) with three subsets corresponding to each
subspecies.

\subsection{Bootstrapping the data}\label{bootstrapping-the-data-1}

We can the bootstrap the subsets to be able test the robustness of the
measured disparity to outliers. We can do that using the default options
of \texttt{boot.matrix} (more about that
\protect\hyperlink{bootstraps-and-rarefactions}{here}):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Bootstrapping the data}
\NormalTok{(petal_bootstrapped <-}\StringTok{ }\KeywordTok{boot.matrix}\NormalTok{(petal_subsets))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 3 customised subsets for 150 elements with 4 dimensions:
##     setosa, versicolor, virginica.
## Data was bootstrapped 100 times (method:"full").
\end{verbatim}

\subsection{Calculating disparity}\label{calculating-disparity-1}

Disparity can be calculated in many ways, therefore the
\texttt{dispRity} function allows users to define their own measure of
disparity. For more details on measuring disparity, see the
\protect\hyperlink{disparity-metrics}{dispRity metrics section}.

In this example, we are going to define disparity as the median distance
between the different individuals and the centroid of the ordinated
space. High values of disparity will indicate a generally high spread of
points from this centroid (i.e.~on average, the individuals are far
apart in the ordinated space). We can define the metrics easily in the
\texttt{dispRity} function by feeding them to the \texttt{metric}
argument. Here we are going to feed the functions \texttt{stats::median}
and \texttt{dispRity::centroids} which calculates distances between
elements and their centroid.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Calculating disparity as the median distance between each elements and}
\NormalTok{## the centroid of the petal-space}
\NormalTok{(petal_disparity <-}\StringTok{ }\KeywordTok{dispRity}\NormalTok{(petal_bootstrapped, }\DataTypeTok{metric =} \KeywordTok{c}\NormalTok{(median, centroids)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  ---- dispRity object ---- 
## 3 customised subsets for 150 elements with 4 dimensions:
##     setosa, versicolor, virginica.
## Data was bootstrapped 100 times (method:"full").
## Disparity was calculated as: c(median, centroids).
\end{verbatim}

\subsection{Summarising the results
(plot)}\label{summarising-the-results-plot}

Similarly to the \texttt{custom.subsets} and \texttt{boot.matrix}
function, \texttt{dispRity} displays a \texttt{dispRity} object. But we
are definitely more interested in actually look at the calculated
values.

First we can summarise the data in a table by simply using
\texttt{summary}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Displaying the summary of the calculated disparity}
\KeywordTok{summary}\NormalTok{(petal_disparity)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##      subsets  n   obs bs.median  2.5%   25%   75% 97.5%
## 1     setosa 50 0.421     0.419 0.342 0.401 0.445 0.503
## 2 versicolor 50 0.693     0.657 0.530 0.622 0.692 0.733
## 3  virginica 50 0.785     0.745 0.577 0.690 0.797 0.880
\end{verbatim}

We can also plot the results in a similar way:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Graphical options}
\KeywordTok{par}\NormalTok{(}\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}

\NormalTok{## Plotting the disparity in the petal_space}
\KeywordTok{plot}\NormalTok{(petal_disparity)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-128-1.pdf}

Now contrary to simply plotting the two first axis of the PCA where we
saw that the species have a different position in the two first
petal-space, we can now also see that they occupy this space clearly
differently!

\subsection{Testing hypothesis}\label{testing-hypothesis}

Finally we can test our hypothesis that we guessed from the disparity
plot (that some groups occupy different volume of the petal-space) by
using the \texttt{test.dispRity} option.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Running a PERMANOVA}
\KeywordTok{test.dispRity}\NormalTok{(petal_disparity, }\DataTypeTok{test =}\NormalTok{ adonis.dispRity)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in test.dispRity(petal_disparity, test = adonis.dispRity): adonis.dispRity test will be applied to the data matrix, not to the calculated disparity.
## See ?adonis.dispRity for more details.
\end{verbatim}

\begin{verbatim}
## Warning in adonis.dispRity(data, ...): The input data for adonis.dispRity was not a distance matrix.
## The results are thus based on the distance matrix for the input data (i.e. dist(data$matrix)).
## Make sure that this is the desired methodological approach!
\end{verbatim}

\begin{verbatim}
## 
## Call:
## vegan::adonis(formula = dist(matrix) ~ group, data = data, method = "euclidean") 
## 
## Permutation: free
## Number of permutations: 999
## 
## Terms added sequentially (first to last)
## 
##            Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)    
## group       2    592.07 296.037  487.33 0.86894  0.001 ***
## Residuals 147     89.30   0.607         0.13106           
## Total     149    681.37                 1.00000           
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Post-hoc testing of the differences between species (corrected for multiple tests)}
\KeywordTok{test.dispRity}\NormalTok{(petal_disparity, }\DataTypeTok{test =}\NormalTok{ t.test, }\DataTypeTok{correction =} \StringTok{"bonferroni"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [[1]]
##                        statistic: t
## setosa : versicolor      -33.714480
## setosa : virginica       -34.257797
## versicolor : virginica    -8.595829
## 
## [[2]]
##                        parameter: df
## setosa : versicolor         189.0217
## setosa : virginica          150.1371
## versicolor : virginica      171.2387
## 
## [[3]]
##                             p.value
## setosa : versicolor    2.107278e-81
## setosa : virginica     2.252270e-72
## versicolor : virginica 1.506037e-14
\end{verbatim}

We can now see that there is a significant difference in petal-space
occupancy between all species of iris.

\subsubsection{Setting up a multidimensional
null-hypothesis}\label{setting-up-a-multidimensional-null-hypothesis}

One other series of test can be done on the shape of the petal-space.
Using a MCMC permutation test we can simulate a petal-space with
specific properties and see if our observed petal-space matches these
properties (similarly to \citet{diaz2016global}):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Testing against a uniform distribution}
\NormalTok{disparity_uniform <-}\StringTok{ }\KeywordTok{null.test}\NormalTok{(petal_disparity, }\DataTypeTok{replicates =} \DecValTok{200}\NormalTok{,}
    \DataTypeTok{null.distrib =}\NormalTok{ runif, }\DataTypeTok{scale =} \OtherTok{FALSE}\NormalTok{)}
\KeywordTok{plot}\NormalTok{(disparity_uniform)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-130-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Testing against a normal distribution}
\NormalTok{disparity_normal <-}\StringTok{ }\KeywordTok{null.test}\NormalTok{(petal_disparity, }\DataTypeTok{replicates =} \DecValTok{200}\NormalTok{,}
    \DataTypeTok{null.distrib =}\NormalTok{ rnorm, }\DataTypeTok{scale =} \OtherTok{TRUE}\NormalTok{)}
\KeywordTok{plot}\NormalTok{(disparity_normal)}
\end{Highlighting}
\end{Shaded}

\includegraphics{dispRity_manual_files/figure-latex/unnamed-chunk-131-1.pdf}

In both cases we can see that our petal-space is not entirely normal or
uniform. This is expected because of the simplicity of these parameters.

\chapter{Future directions}\label{future-directions}

\section{More tests!}\label{more-tests}

Some more tests are being developed such as a \texttt{sequential.test}
to run sequential linear models for testing hypothesis in disparity
through time or a \texttt{model.test} developed in collaboration with
\href{https://puttickbiology.wordpress.com/}{Mark Puttick} to fit modes
of evolution to disparity curves. Stay tuned!

\section{More rarefactions!}\label{more-rarefactions}

I am working on including several ``clever'' bootstrap and rarefaction
algorithms for palaeobiology studies based on geological sampling
(following:and Butler et al. 2012 \emph{Evolution} and Kotric and Knoll
2015 \emph{Palaeobiology}).

\section{Faster disparity
calculations}\label{faster-disparity-calculations}

I am slowly implementing parallel disparity calculation as well as C
implementations of some disparity metrics to increase significantly
improve the speed of the \texttt{dispRity} function.

\section{More modularity}\label{more-modularity}

I am equally slowly developing functions to allow more of the options in
the package to be modular (in the same way as the \texttt{metric}
argument in \texttt{dispRity}). The next arguments to benefit this
increased modularity will be \texttt{chrono.subsets}'s \texttt{model}
argument and \texttt{boot.matrix}'s \texttt{type} argument.

\section{More suggestions?}\label{more-suggestions}

More generally, I welcome any suggestions or collaborations for
developing some additional specific ideas. Feel free to contact me via
email (\href{mailto:guillert@tcd.ie}{\nolinkurl{guillert@tcd.ie}}) or if
you are a GitHub user, directly create an issue on the
\href{https://github.com/TGuillerme/dispRity}{GitHub page}.

\bibliography{../References.bib,../packages.bib}

\end{document}