### 2019-07-10 inst/gitbook/09_future_directions.Rmd

---
title: "Future directions and references"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: [../References.bib, ../packages.bib]
bst: sysbio.bst
vignette: >
  %\VignetteIndexEntry{future directions}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Developing `dispRity`

Everyone is more than welcome to help developing this package, whether it is through developing functions (metrics, tests, plots, etc...), the manual (more examples, typos, etc...) or completely new functionalities!

## More functions

## More manuals

<!-- 
* Beck, R. M., & Lee, M. S. (2014). Ancient dates or accelerated rates? Morphological clocks and the antiquity of placental mammals. Proceedings of the Royal Society of London B: Biological Sciences, 281(1793), 20141278.
* Cooper, N., & Guillerme, T. (in prep.). Coming soonish!.
* Diaz, S., Kattge, J., Cornelissen, J.H., Wright, I.J., Lavorel, S., Dray, S., Reu, B., Kleyer, M., Wirth, C., Prentice, I.C. and Garnier, E. (2016). The global spectrum of plant form and function. Nature, 529(7585), 167.
* Donohue, I., Petchey, O.L., Montoya, J.M., Jackson, A.L., McNally, L., Viana, M., Healy, K., Lurgi, M., O'Connor, N.E. and Emmerson, M.C, (2013). On the dimensionality of ecological stability. Ecology letters, 16(4), 421-429.
* Wills, M. A., Briggs, D. E., & Fortey, R. A. (1994). Disparity as an evolutionary index: a comparison of Cambrian and Recent arthropods. Paleobiology, 20(2), 93-130. -->