### 2020-02-27 inst/gitbook/05_other-functionalities.Rmd
Added `char.diff` to the manual

---
title: "other functionalities"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: [../References.bib, ../packages.bib]
bst: sysbio.bst
vignette: >
  %\VignetteIndexEntry{other functionalities}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Other functionalities

The `dispRity` package also contains several other functions that are not specific to multidimensional analysis but that are often used by `dispRity` internal functions.
However, we decided to make these functions also available at a user level since they can be handy for certain specific operations!
You'll find a brief description of each of them (alphabetically) here:

## `char.diff`

This is yet another function for calculating distance matrices.
There are many functions for calculating pairwise distance matrices in `R` (`stats::dist`, `vegan::vegdist`, `cluster::daisy` or `Claddis::MorphDistMatrix`) but this one is the `dispRity` one.
It is slightly different to the ones mentioned above (though not that dissimilar from `Claddis::MorphDistMatrix`) in the fact that it focuses on comparing discrete morphological characters and tries to solve all the problems linked to these kind of matrices (especially dealing with special tokens).

The function intakes a matrix with either `numeric` or `integer` (`NA` included) or matrices with `character` that are indeed `integers` (e.g.`"0"` and `"1"`).
It then uses a [bitwise operations](https://en.wikipedia.org/wiki/Bitwise_operations_in_C) architecture implemented in `C` that renders the function pretty fast and pretty modular.
This bitwise operations translates the character states into binary values.
This way, `0` becomes `1`, `1` becomes `2`, `2` becomes `4`, `3` becomes `8`, etc...
Specifically it can handle any rules specific to special tokens (i.e. symbols) for discrete morphological characters. For example, should you treat missing values `"?"` as `NA` (ignoring them) or as any possible character state (e.g. `c("0", "1")`?)? And how to treat characters with a ampersand (`"&"`)? `char.diff` can answer to all these questions!

Let's start by a basic binary matrix 4\*3 with random `integer`:

```{r}
## A random binary matrix
matrix_binary <- matrix(sample(c(0,1), 12, replace = TRUE), ncol = 4,
                        dimnames = list(letters[1:3], LETTERS[1:4]))
```

By default, `char.diff` measures the hamming distance between characters:

```{r}
## The hamming distance between characters
(differences <- char.diff(matrix_binary))
```

Note that the results is just a pairwise distance (dissimilarity) matrix with some special dual class `matrix` and `char.diff`.
This means it can easily be plotted via the disparity package:

```{r}
## Visualising the matrix
plot(differences)
```

You can check all the numerous plotting options in the `?plot.char.diff` manual (it won't be developed here).

The `char.diff` function has much more options however (see all of them in the `?char.diff` manual) for example to measure different differences (via `method`) or making the comparison work per row (for a distance matrix between the rows):

```{r}
## Euclidean distance between rows
char.diff(matrix_binary, by.col = FALSE, method = "euclidean")
```

We can however make it more interesting by playing with the different rules to play with different tokens.
First let's create a matrix with morphological characters as numeric `character`s:

```{r}
## A random character matrix
(matrix_character <- matrix(sample(c("0","1","2"), 30, replace = TRUE), ncol = 5,
                           dimnames = list(letters[1:6], LETTERS[1:5])))

## The hamming difference between columns
char.diff(matrix_character)
```

Here the characters are automatically converted into bitwise integers to be compared efficiently. We can now add some more special tokens like `"?"` or `"0/1"` for uncertainties between state `"0"` and `"1"` but not `"2"`:

```{r}
## Adding uncertain characters
matrix_character[sample(1:30, 8)] <- "0/1"

## Adding missing data
(matrix_character[sample(1:30, 5)] <- "?")

## The hamming difference between columns including the special characters
char.diff(matrix_character)
```

Note here that it detected the default behaviours for the special tokens `"?"` and `"/"`: `"?"` are treated as `NA` (not compared) and `"/"` are treated as both states (e.g. `"0/1"` is treated as `"0"` and as `"1"`).
We can specify both the special tokens and the special behaviours to consider via `special.tokens` and `special.behaviours`.
The `special.tokens` are `missing = "?"`, `inapplicable = "-"`, `uncertainty = "\"` and `polymorphism = "&"` meaning we don't have to modify them for now.
However, say we want to change the behaviour for `"?"` and treat them as all possible characters and treat `"/"` as only the character `"0"` (as an integer) we can specify them giving a behaviour function:

```{r}
## Specifying some special behaviours
my_special_behaviours <- list(missing = function(x,y) return(y),
                              uncertainty = function(x,y) return(as.integer(0)))

## Passing these special behaviours to the char.diff function
char.diff(matrix_character, special.behaviour = my_special_behaviours)
```

The results are quiet different as before! Note that you can also specify some really specific behaviours for any type of special token.

```{r}
## Adding weird tokens to the matrix
matrix_character[sample(1:30, 8)] <- "%"

## Specify the new token and the new behaviour
char.diff(matrix_character, special.tokens = c(weird_one = "%"),
                            special.behaviours = list(
                                weird_one = function(x,y) return(as.integer(42)))
                            )
```

Of course the results can be quiet surprising then... But that's the essence of the modularity. You can see more options in the function manual `?char.diff`!

## `clean.data`

This is a rather useful function that allows matching a `matrix` or a `data.frame` to a tree (`phylo`) or a distribution of trees (`multiPhylo`).
This function outputs the cleaned data and trees (if cleaning was needed) and a list of dropped rows and tips.

```{r}
## Generating a trees with labels from a to e
dummy_tree <- rtree(5, tip.label = LETTERS[1:5])

## Generating a matrix with rows from b to f
dummy_data <- matrix(1, 5, 2, dimnames = list(LETTERS[2:6], c("var1", "var2")))

##Cleaning the trees and the data
(cleaned <- clean.data(data = dummy_data, tree = dummy_tree))
```

## `crown.stem`

This function quiet handily separates tips from a phylogeny between crown members (the living taxa and their descendants) and their stem members (the fossil taxa without any living relatives).

```{r}
data(BeckLee_tree)
## Diving both crow and stem species
(crown.stem(BeckLee_tree, inc.nodes = FALSE))
```

Note that it is possible to include or exclude nodes from the output.
To see a more applied example: this function is used in [chapter 03: specific tutorials](#custom-subsets).

## `get.bin.ages`

This function is similar than the `crown.stem` one as it is based on a tree but this one outputs the stratigraphic bins ages that the tree is covering.
This can be useful to generate precise bin ages for the `chrono.subsets` function:

```{r}
get.bin.ages(BeckLee_tree)
```

Note that this function outputs the stratigraphic age limits by default but this can be customisable by specifying the `type` of data (e.g. `type = "Eon"` for eons).
The function also intakes several optional arguments such as whether to output the startm end, range or midpoint of the stratigraphy or the year of reference of the International Commission of Stratigraphy.
To see a more applied example: this function is used in [chapter 03: specific tutorials](#chrono-subsets).

## `pair.plot`

This utility function allows to plot a matrix image of pairwise comparisons.
This can be useful when getting pairwise comparisons and if you'd like to see at a glance which pairs of comparisons have high or low values.

```{r}
## Random data
data <- matrix(data = runif(42), ncol = 2)

## Plotting the first column as a pairwise comparisons
pair.plot(data, what = 1, col = c("orange", "blue"), legend = TRUE, diag = 1)
```

Here blue squares are ones that have a high value and orange ones the ones that have low values.
Note that the values plotted correspond the first column of the data as designated by `what = 1`.

It is also possible to add some tokens or symbols to quickly highlight to specific cells, for example which elements in the data are below a certain value:

```{r}
## The same plot as before without the diagonal being the maximal observed value
pair.plot(data, what = 1, col = c("orange", "blue"), legend = TRUE, diag = "max")
## Highlighting with an asterisk which squares have a value below 0.2
pair.plot(data, what = 1, binary = 0.2, add = "*", cex = 2)
```

This function can also be used as a binary display when running a series of pairwise t-tests.
For example, the following script runs a wilcoxon test between the time-slices from the `disparity` example dataset and displays in black which pairs of slices have a p-value below 0.05:

```{r}
## Loading disparity data
data(disparity)

## Testing the pairwise difference between slices
tests <- test.dispRity(disparity, test = wilcox.test, correction = "bonferroni")

## Plotting the significance
pair.plot(as.data.frame(tests), what = "p.value", binary = 0.05)
```

## `reduce.matrix`

This function allows to reduce columns or rows of a matrix to make sure that there is enough overlap for further analysis.
This is particularly useful if you are going to use distance matrices since it uses the `vegan::vegdist` function to test whether distances can be calculated or not.

For example, if we have a patchy matrix like so (where the black squares represent available data):

```{r}
set.seed(1)
## A 10*5 matrix
na_matrix <- matrix(rnorm(50), 10, 5)
## Making sure some rows don't overlap
na_matrix[1, 1:2] <- NA
na_matrix[2, 3:5] <- NA
## Adding 50% NAs
na_matrix[sample(1:50, 25)] <- NA
## Illustrating the gappy matrix
image(t(na_matrix), col = "black")
```

We can use the `reduce.matrix` to double check whether any rows cannot be compared.
The functions needs as an input the type of distance that will be used, say a `"gower"` distance:

```{r}
## Reducing the matrix by row
(reduction <- reduce.matrix(na_matrix, distance = "gower"))
```

We can not remove the rows 1 and 9 and see if that improved the overlap:

```{r}
image(t(na_matrix[-as.numeric(reduction$rows.to.remove), ]), col = "black")
```

## `slice.tree`

This function is a modification of the `paleotree::timeSliceTree` function that allows to make slices through a phylogenetic tree.
Compared to the `paleotree::timeSliceTree`, this function allows a model to decide which tip or node to use when slicing through a branch (whereas `paleotree::timeSliceTree` always choose the first available tip alphabetically).
The models for choosing which tip or node are the same as the ones used in the `chrono.subsets` and are described in [chapter 03: specific tutorials](#chrono-subsets).

The function works by using at least a tree, a slice age and a model:

```{r}
set.seed(1)
## Generate a random ultrametric tree
tree <- rcoal(20)
## Add some node labels
tree$node.label <- letters[1:19]
## Add its root time
tree$root.time <- max(tree.age(tree)$ages)

## Slicing the tree at age 0.75
tree_75 <- slice.tree(tree, age = 0.75, "acctran")

## Showing both trees
par(mfrow = c(1,2))
plot(tree, main = "original tree")
axisPhylo() ; nodelabels(tree$node.label, cex = 0.8)
abline(v = (max(tree.age(tree)$ages) - 0.75), col = "red")
plot(tree_75, main = "sliced tree")
```

## `slide.nodes` and `remove.zero.brlen`

This function allows to slide nodes along a tree!
In other words it allows to change the branch length leading to a node without modifying the overall tree shape.
This can be useful to add some value to 0 branch lengths for example.

The function works by taking a node (or a list of nodes), a tree and a sliding value.
The node will be moved "up" (towards the tips) for the given sliding value.
You can move the node "down" (towards the roots) using a negative value.

```{r}
set.seed(42)
## Generating simple coalescent tree
tree <- rcoal(5)

## Sliding node 8 up and down
tree_slide_up <- slide.nodes(8, tree, slide = 0.075)
tree_slide_down <- slide.nodes(8, tree, slide = -0.075)

## Display the results
par(mfrow = c(3,1))
plot(tree, main = "original tree") ; axisPhylo() ; nodelabels()
plot(tree_slide_up, main = "slide up!") ; axisPhylo() ; nodelabels()
plot(tree_slide_down, main = "slide down!") ; axisPhylo() ; nodelabels()
```

The `remove.zero.brlen` is a "clever" wrapping function that uses the `slide.nodes` function to stochastically remove zero branch lengths across a whole tree.
This function will slide nodes up or down in successive postorder traversals (i.e. going down the tree clade by clade) in order to minimise the number of nodes to slide while making sure there are no silly negative branch lengths produced!
By default it is trying to slide the nodes using 1% of the minimum branch length to avoid changing the topology too much.

```{r}
set.seed(42)
## Generating a tree
tree <- rtree(20)

## Adding some zero branch lengths (5)
tree$edge.length[sample(1:Nedge(tree), 5)] <- 0

## And now removing these zero branch lengths!
tree_no_zero <- remove.zero.brlen(tree)

## Exaggerating the removal (to make it visible)
tree_exaggerated <- remove.zero.brlen(tree, slide = 1)

## Check the differences
any(tree$edge.length == 0)
any(tree_no_zero$edge.length == 0)
any(tree_exaggerated$edge.length == 0)

## Display the results
par(mfrow = c(3,1))
plot(tree, main = "with zero edges")
plot(tree_no_zero, main = "without zero edges!")
plot(tree_exaggerated, main = "with longer edges")
```

## `tree.age`

This function allows to quickly calculate the ages of each tips and nodes present in a tree.

```{r}
set.seed(1)
tree <- rtree(10)
## The tree age from a 10 tip tree
tree.age(tree)
```

It also allows to set the age of the root of the tree:

```{r}
## The ages starting from -100 units
tree.age(tree, age = 100)
```

Usually tree age is calculated from the present to the past (e.g. in million years ago) but it is possible to reverse it using the `order = present` option:

```{r}
## The ages in terms of tip/node height
tree.age(tree, order = "present")
```