### 2020-04-24 inst/gitbook/03_specific-tutorials.Rmd
book update (up until section 4)

---
title: "Details of specific functions"
author: "Thomas Guillerme (guillert@tcd.ie), Mark Puttick (marknputtick@gmail.com) and  Natalie Cooper (natalie.cooper@nhm.ac.uk)"
date: "`r Sys.Date()`"
bibliography: [../References.bib, ../packages.bib]
output:
  html_document: default
  pdf_document: default
---

# Details of specific functions

The following section contains information specific to some functions.
If any of your questions are not covered in these sections, please refer to the function help files in `R`, send me an email (<guillert@tcd.ie>), or raise an issue on [GitHub](https://github.com/TGuillerme/dispRity/issues).
The several tutorials below describe specific functionalities of certain functions; please always refer to the function help files for the full function documentation!

Before each section, make sure you loaded the @beckancient2014 data (see [example data](#example-data) for more details).

```{r}
## Loading the data
data(BeckLee_mat50)
data(BeckLee_mat99)
data(BeckLee_tree)
data(BeckLee_ages)
```

## Time slicing {#chrono-subsets}

The function `chrono.subsets` allows users to divide the matrix into different time subsets or slices given a dated phylogeny that contains all the elements (i.e. taxa) from the matrix.
Each subset generated by this function will then contain all the elements present at a specific point in time or during a specific period in time.

Two types of time subsets can be performed by using the `method` option:

 *  Discrete time subsets (or time-binning) using `method = discrete`
 *  Continuous time subsets (or time-slicing) using `method = continuous`

For the time-slicing method details see @time-slice.
For both methods, the function takes the `time` argument which can be a vector of `numeric` values for:

 *  Defining the boundaries of the time bins (when `method = discrete`)
 *  Defining the time slices (when `method = continuous`)

Otherwise, the `time` argument can be set as a single `numeric` value for automatically generating a given number of equidistant time-bins/slices.
Additionally, it is also possible to input a dataframe containing the first and last occurrence data (FAD/LAD) for taxa that span over a longer time than the given tips/nodes age, so taxa can appear in more than one time bin/slice.

### Time-binning

Here is an example for the time binning method (`method = discrete`):

```{r, eval=TRUE}
## Generating three time bins containing the taxa present every 40 Ma
chrono.subsets(data = BeckLee_mat50, tree = BeckLee_tree, method = "discrete",
                time = c(120, 80, 40, 0))
```

Note that we can also generate equivalent results by just telling the function that we want three time-bins as follow:

```{r, eval=TRUE}
## Automatically generate three equal length bins:
chrono.subsets(data = BeckLee_mat50, tree = BeckLee_tree, method = "discrete",
                time = 3)
```

In this example, the taxa were split inside each time-bin according to their age.
However, the taxa here are considered as single points in time.
It is totally possible that some taxa could have had longer longevity and that they exist in multiple time bins.
In this case, it is possible to include them in more than one bin by providing a table of first and last occurrence dates (FAD/LAD).
This table should have the taxa names as row names and two columns for respectively the first and last occurrence age:

```{r, eval=TRUE}
## Displaying the table of first and last occurrence dates for each taxa
head(BeckLee_ages)

## Generating time bins including taxa that might span between them
chrono.subsets(data = BeckLee_mat50, tree = BeckLee_tree, method = "discrete",
                time = c(120, 80, 40, 0), FADLAD = BeckLee_ages)
```

When using this method, the oldest boundary of the first bin (or the first slice, see below) is automatically generated as the root age plus 1\% of the tree length, as long as at least three elements/taxa are present at that point in time.
The algorithm adds an extra 1\% tree length until reaching the required minimum of three elements.
It is also possible to include nodes in each bin by using `inc.nodes = TRUE` and providing a matrix that contains the ordinated distance among tips *and* nodes.

If you want to generate time subsets based on stratigraphy, the package proposes a useful functions to do it for you: `get.bin.ages` (check out the function's manual in `R`)!

### Time-slicing

For the time-slicing method (`method = continuous`), the idea is fairly similar.
This option, however, requires a matrix that contains the ordinated distance among taxa *and* nodes and an extra argument describing the assumed evolutionary model (via the `model` argument).
This model argument is used when the time slice occurs along a branch of the tree rather than on a tip or a node, meaning that a decision must be made about what the value for the branch should be.
The model can be one of the following:

 * **Punctuated models**
   *  `acctran` where the data chosen along the branch is always the one of the descendant
   *  `deltran` where the data chosen along the branch is always the one of the ancestor
   *  `random` where the data chosen along the branch is randomly chosen between the descendant or the ancestor
   *  `proximity` where the data chosen along the branch is either the descendant or the ancestor depending on branch length

 * **Gradual models**
   *  `equal.split` where the data chosen along the branch is both the descendant and the ancestor with an even probability
   *  `gradual.split` where the data chosen along the branch is both the descendant and the ancestor with a probability depending on branch length

> Note that the four first models are a proxy for punctuated evolution: the selected data is always either the one of the descendant or the ancestor.
In other words, changes along the branches always occur at either ends of it.
The two last models are a proxy for gradual evolution: the data from both the descendant and the ancestor is used with an associate probability.
These later models perform better when bootstrapped, effectively approximating the "intermediate" state between and the ancestor and the descendants.

More details about the differences between these methods can be found in @time-slice.

```{r, eval=TRUE}
## Generating four time slices every 40 million years under a model of proximity evolution
chrono.subsets(data = BeckLee_mat99, tree = BeckLee_tree, 
    method = "continuous", model = "proximity", time = c(120, 80, 40, 0),
    FADLAD = BeckLee_ages)

## Generating four time slices automatically
chrono.subsets(data = BeckLee_mat99, tree = BeckLee_tree,
    method = "continuous", model = "proximity", time = 4, FADLAD = BeckLee_ages)
```

## Customised subsets {#custom-subsets}

Another way of separating elements into different categories is to use customised subsets as briefly explained [above](#disparity-among-groups).
This function simply takes the list of elements to put in each group (whether they are the actual element names or their position in the matrix).

```{r, eval=TRUE}
## Creating the two groups (crown and stems)
mammal_groups <- crown.stem(BeckLee_tree, inc.nodes = FALSE)

## Separating the dataset into two different groups
custom.subsets(BeckLee_mat50, group = mammal_groups)
```

Like in this example, you can use the utility function `crown.stem` that allows to automatically separate the crown and stems taxa given a phylogenetic tree.
Also, elements can easily be assigned to different groups if necessary!

```{r, eval=FALSE}
## Creating the three groups as a list
weird_groups <- list("even" = seq(from = 1, to = 49, by = 2),
                      "odd" = seq(from = 2, to = 50, by = 2),
                      "all" = c(1:50))
```

The `custom.subsets` function can also take a phylogeny (as a `phylo` object) as an argument to create groups as clades:

```{r, eval=FALSE}
## Creating groups as clades
custom.subsets(BeckLee_mat50, group = BeckLee_tree)
```

This automatically creates 49 (the number of nodes) groups containing between two and 50 (the number of tips) elements.

## Bootstraps and rarefactions

One important step in analysing ordinated matrices is to pseudo-replicate the data to see how robust the results are, and how sensitive they are to outliers in the dataset.
This can be achieved using the function `boot.matrix` to bootstrap and/or rarefy the data.
The default options will bootstrap the matrix 100 times without rarefaction using the "full" bootstrap method (see below):

```{r, eval=TRUE}
## Default bootstrapping
boot.matrix(data = BeckLee_mat50)
```

The number of bootstrap replicates can be defined using the `bootstraps` option.
The method can be modified by controlling which bootstrap algorithm to use through the `boot.type` argument.
Currently two algorithms are implemented:

 * `full` where the bootstrapping is entirely stochastic (*n* elements are replaced by any *m* elements drawn from the data)
 * `single` where only one random element is replaced by one other random element for each pseudo-replicate

```{r, eval=TRUE}
## Bootstrapping with the single bootstrap method
boot.matrix(BeckLee_mat50, boot.type = "single")
```

This function also allows users to rarefy the data using the `rarefaction` argument.
Rarefaction allows users to limit the number of elements to be drawn at each bootstrap replication.
This is useful if, for example, one is interested in looking at the effect of reducing the number of elements on the results of an analysis.

This can be achieved by using the `rarefaction` option that draws only *n-x* at each bootstrap replicate (where *x* is the number of elements not sampled).
The default argument is `FALSE` but it can be set to `TRUE` to fully rarefy the data (i.e. remove *x* elements for the number of pseudo-replicates, where *x* varies from the maximum number of elements present in each subset to a minimum of three elements).
It can also be set to one or more `numeric` values to only rarefy to the corresponding number of elements.

```{r, eval=TRUE}
## Bootstrapping with the full rarefaction
boot.matrix(BeckLee_mat50, bootstraps = 20, rarefaction = TRUE)

## Or with a set number of rarefaction levels
boot.matrix(BeckLee_mat50, bootstraps = 20, rarefaction = c(6:8, 3))
```

One other argument is `dimensions` that specifies how many dimensions from the matrix should be used for further analysis.
When missing, all dimensions from the ordinated matrix are used.

```{r, eval=TRUE}
## Using the first 50% of the dimensions
boot.matrix(BeckLee_mat50, dimensions = 0.5)

## Using the first 10 dimensions
boot.matrix(BeckLee_mat50, dimensions = 10)
```

It is also possible to specify the sampling probability in the bootstrap for each elements.
This can be useful for weighting analysis for example (i.e. giving more importance to specific elements).
These probabilities can be passed to the `prob` argument individually with a vector with the elements names or with a matrix with the rownames as elements names.
The elements with no specified probability will be assigned a probability of 1 (or 1/maximum weight if the argument is weights rather than probabilities).

```{r, eval=TRUE}
## Attributing a weight of 0 to Cimolestes and 10 to Maelestes
boot.matrix(BeckLee_mat50, prob = c("Cimolestes" = 0, "Maelestes" = 10))
```

Of course, one could directly supply the subsets generated above (using `chrono.subsets` or `custom.subsets`) to this function.

```{r, eval=TRUE}
## Creating subsets of crown and stem mammals
crown_stem <- custom.subsets(BeckLee_mat50,
                                group = crown.stem(BeckLee_tree, inc.nodes = FALSE))
## Bootstrapping and rarefying these groups
boot.matrix(crown_stem, bootstraps = 200, rarefaction = TRUE)

## Creating time slice subsets
time_slices <- chrono.subsets(data = BeckLee_mat99, tree = BeckLee_tree, 
                               method = "continuous", model = "proximity", 
                               time = c(120, 80, 40, 0),
                               FADLAD = BeckLee_ages)

## Bootstrapping the time slice subsets
boot.matrix(time_slices, bootstraps = 100)
```

## Disparity metrics {#disparity-metrics}

There are many ways of measuring disparity!
In brief, disparity is a summary metric that will represent an aspect of an ordinated space (e.g. a MDS, PCA, PCO, PCoA).
For example, one can look at ellipsoid hyper-volume of the ordinated space (Donohue _et al._ 2013), the sum and the product of the ranges and variances (Wills _et al._ 1994) or the median position of the elements relative to their centroid (Wills _et al._ 1994). 
Of course, there are many more examples of metrics one can use for describing some aspect of the ordinated space, with some performing better than other ones at particular descriptive tasks, and some being more generalist.
Check out this pre-print on selecting the best metric for your specific question on [biorXiv](https://www.biorxiv.org/content/10.1101/801571v1).
<!-- biorXiv preprint -->
You can also use the  [`moms` shiny app](https://tguillerme.shinyapps.io/moms/) to test which metric captures which aspect of traitspace occupancy regarding your specific space and your specific question.

Regardless, and because of this great diversity of metrics, the package `dispRity` does not have one way to measure disparity but rather proposes to facilitate users in defining their own disparity metric that will best suit their particular analysis.
In fact, the core function of the package, `dispRity`, allows the user to define any metric with the `metric` argument.
However the `metric` argument has to follow certain rules:

1.  It must be composed from one to three `function` objects;
2.  The function(s) must take as a first argument a `matrix` or a `vector`;
3.  The function(s) must be of one of the three dimension-levels described below;
4.  At least one of the functions must be of dimension-level 1 or 2 (see below).

### The function dimension-levels

The metric function dimension-levels determine the "dimensionality of decomposition" of the input matrix.
In other words, each dimension-level designates the dimensions of the output, i.e. either three (a `matrix`); two (a `vector`); or one (a single `numeric` value) dimension.

![Illustration of the different dimension-levels of functions with an input `matrix`](dispRity_fun.png)

#### Dimension-level 1 functions

A dimension-level 1 function will decompose a `matrix` or a `vector` into a single value:

```{r}
## Creating a dummy matrix
dummy_matrix <- matrix(rnorm(12), 4, 3)

## Example of dimension-level 1 functions
mean(dummy_matrix)
median(dummy_matrix)
```

Any summary metric such as mean or median are good examples of dimension-level 1 functions as they reduce the matrix to a single dimension (i.e. one value).

#### Dimension-level 2 functions

A dimension-level 2 function will decompose a `matrix` into a `vector`.

```{r}
## Defining the function as the product of rows
prod.rows <- function(matrix) apply(matrix, 1, prod)

## A dimension-level 2 metric
prod.rows(dummy_matrix)
```

Several dimension-level 2 functions are implemented in `dispRity` (see `?dispRity.metric`) such as the `variances` or `ranges` functions that calculate the variance or the range of each dimension of the ordinated matrix respectively.

#### Dimension-level 3 functions

Finally a dimension-level 3 function will transform the matrix into another matrix.
Note that the dimension of the output matrix doesn't need to match the the input matrix:

```{r}
## A dimension-level 3 metric
var(dummy_matrix)

## A dimension-level 3 metric with a forced matrix output
as.matrix(dist(dummy_matrix))
```

### `make.metric`

Of course, functions can be more complex and involve multiple operations such as the `centroids` function (see `?dispRity.metric`) that calculates the Euclidean distance between each element and the centroid of the ordinated space.
The `make.metric` function implemented in `dispRity` is designed to help test and find the dimension-level of the functions.
This function tests:

1.  If your function can deal with a `matrix` or a `vector` as an input;
2.  Your function's dimension-level according to its output (dimension-level 1, 2 or 3, see above);
3.  Whether the function can be implemented in the `dispRity` function (the function is fed into a `lapply` loop).

For example, let's see if the functions described above are the right dimension-levels:

```{r}
## Which dimension-level is the mean function? And can it be used in dispRity?
make.metric(mean)

## Which dimension-level is the prod.rows function? And can it be used in dispRity?
make.metric(prod.rows)

## Which dimension-level is the var function? And can it be used in dispRity?
make.metric(var)
```

A non verbose version of the function is also available.
This can be done using the option `silent = TRUE` and will simply output the dimension-level of the metric.

```{r}
## Testing whether mean is dimension-level 1
if(make.metric(mean, silent = TRUE) != "level1") {
    message("The metric is not dimension-level 1.")
}
## Testing whether var is dimension-level 1
if(make.metric(var, silent = TRUE) != "level1") {
    message("The metric is not dimension-level 1.")
}
```

### Metrics in the `dispRity` function

Using this metric structure, we can easily use any disparity metric in the `dispRity` function as follows:

```{r}
## Measuring disparity as the standard deviation of all the values of the
## ordinated matrix (dimension-level 1 function).
summary(dispRity(BeckLee_mat50, metric = sd))

## Measuring disparity as the standard deviation of the variance of each axis of
## the ordinated matrix (dimension-level 1 and 2 functions).
summary(dispRity(BeckLee_mat50, metric = c(sd, variances)))

## Measuring disparity as the standard deviation of the variance of each axis of
## the variance covariance matrix (dimension-level 1, 2 and 3 functions).
summary(dispRity(BeckLee_mat50, metric = c(sd, variances, var)), round = 10)
```

Note that the order of each function in the metric argument does not matter, the `dispRity` function will automatically detect the function dimension-levels (using `make.metric`) and apply them to the data in decreasing order (dimension-level 3 > 2 > 1).

```{r}
## Disparity as the standard deviation of the variance of each axis of the
## variance covariance matrix:
disparity1 <- summary(dispRity(BeckLee_mat50, metric = c(sd, variances, var)),
                      round = 10)

## Same as above but using a different function order for the metric argument
disparity2 <- summary(dispRity(BeckLee_mat50, metric = c(variances, sd, var)),
                      round = 10)

## Both ways output the same disparity values:
disparity1 == disparity2
```

In these examples, we considered disparity to be a single value.
For example, in the previous example, we defined disparity as the standard deviation of the variances of each column of the variance/covariance matrix (`metric = c(variances, sd, var)`).
It is, however, possible to calculate [disparity as a distribution](#disparity-as-a-distribution).

### Metrics implemented in `dispRity`

<!-- TODO: Sort by group of metrics -->

Several disparity metrics are implemented in the `dispRity` package.
The detailed list can be found in `?dispRity.metric` along with some description of each metric.

Level | Name | Description | Source | 
------|------|---------------------------------------------------|--------|
2 | `ancestral.dist` | The distance between an element and its ancestor | `dispRity` |
2 | `angles` | The angle of main variation of each dimensions | `dispRity` |
2 | `centroids`<sup>1</sup> | The distance between each element and the centroid of the ordinated space | `dispRity` |
1 | `convhull.surface` | The surface of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$area` |
1 | `convhull.volume` | The volume of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$vol` |
2 | `deviations` | The minimal distance between each element and a hyperplane | `dispRity` |
1 | `diagonal` | The longest distance in the ordinated space (like the diagonal in two dimensions) | `dispRity` |
2 | `displacements` | The ratio between the distance from a reference and the distance from the centroid | `dispRity` |
1 | `ellipse.volume`<sup>1</sup> | The volume of the ellipsoid of the space | Donohue _et al._ (2013) |
1 | `func.div` | The functional divergence (the ratio of deviation from the centroid) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FDiv` but without abundance)|
1 | `func.eve` | The functional evenness (the minimal spanning tree distances evenness) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FEve` but without abundance)|
1 | `mode.val` | The modal value | `dispRity` |
1 | `n.ball.volume` | The hyper-spherical (*n*-ball) volume | `dispRity` |
2 | `neighbours` | The distance to specific neighbours (e.g. the nearest neighbours - by default) | `dispRity` |
2 | `pairwise.dist` | The pairwise distances between elements | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::vegist` |
2 | `quantiles` | The *n*th quantile range per axis | `dispRity` |
2 | `radius` | The radius of each dimensions | `dispRity` |
2 | `ranges` | The range of each dimension | `dispRity` |
2 | `span.tree.length` | The minimal spanning tree length | [`ape`](https://cran.r-project.org/web/packages/vegan/index.html) (but see also [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::spantree`) |
2 | `variances` | The variance of each dimension | `dispRity` |

1: Note that by default, the centroid is the centroid of the elements.
It can, however, be fixed to a different value by using the `centroid` argument `centroids(space, centroid = rep(0, ncol(space)))`, for example the origin of the ordinated space.

2: This function uses an estimation of the eigenvalue that only works for MDS or PCoA ordinations (*not* PCA).

You can find more informations on the vast variety of metrics that you can use in your analysis in this [preprint](https://www.biorxiv.org/content/10.1101/801571v1).
<!-- biorXiv preprint -->

### Equations and implementations
Some of the functions described below are implemented in the `dispRity` package and do not require any other packages to calculate ([see implementation here](https://github.com/TGuillerme/dispRity/blob/master/R/dispRity.metric.R)).

\begin{equation}
    ancestral.dist = \sqrt{\sum_{i=1}^{n}{({d}_{n}-Ancestor_{n})^2}}
\end{equation}

\begin{equation}
    centroids = \sqrt{\sum_{i=1}^{n}{({d}_{n}-Centroid_{d})^2}}
\end{equation}

\begin{equation}
    diagonal = \sqrt{\sum_{i=1}^{d}|max(d_i) - min(k_i)|}
\end{equation}

\begin{equation}
    deviations = \frac{|Ax + By + ... + Nm + Intercept|}{\sqrt{A^2 + B^2 + ... + N^2}}
\end{equation}

\begin{equation}
    displacements = \frac{\sqrt{\sum_{i=1}^{n}{({d}_{n}-Reference_{d})^2}}}{\sqrt{\sum_{i=1}^{n}{({d}_{n}-Centroid_{k})^2}}}
\end{equation}

\begin{equation}
    ellipse.volume = \frac{\pi^{d/2}}{\Gamma(\frac{d}{2}+1)}\displaystyle\prod_{i=1}^{d} (\lambda_{i}^{0.5})
\end{equation}

\begin{equation}
    n.ball.volume = \frac{\pi^{d/2}}{\Gamma(\frac{d}{2}+1)}\displaystyle\prod_{i=1}^{d} R
\end{equation}

\begin{equation}
    radius = |\frac{\sum_{i=1}^{n}d_i}{n} - f(\mathbf{v}d)|
\end{equation}

\begin{equation}
    ranges = |max(d_i) - min(d_i)|
\end{equation}

\begin{equation}
    variances = \sigma^{2}{d_i}
\end{equation}

\begin{equation}
    span.tree.length = \mathrm{branch\ length}
\end{equation}

Where *d* is the number of dimensions,
*n* the number of elements,
$\Gamma$ is the Gamma distribution,
$\lambda_i$ is the eigenvalue of each dimensions,
$\sigma^{2}$ is their variance and
$Centroid_{k}$ is their mean,
$Ancestor_{n}$ is the coordinates of the ancestor of element $n$,
$f(\mathbf{v}k)$ is function to select one value from the vector $\mathbf{v}$ of the dimension $k$ (e.g. it's maximum, minimum, mean, etc.),
*R* is the radius of the sphere or the product of the radii of each dimensions ($\displaystyle\prod_{i=1}^{k}R_{i}$ - for a hyper-ellipsoid),
and $Reference_{k}$ is an arbitrary point's coordinates (usually 0).

### Using the different disparity metrics
Here is a brief demonstration of the main metrics implemented in `dispRity`.
First, we will create a dummy/simulated ordinated space using the `space.maker` utility function (more about that [here](#space.maker):

```{r}
## Creating a 10*5 normal space
set.seed(1)
dummy_space <- space.maker(10, 5, rnorm)
```

We will use this simulated space to demonstrate the different metrics.

#### Volumes and surface metrics
The functions `ellipse.volume`, `convhull.surface`, `convhull.volume` and `n.ball.volume`  all measure the surface or the volume of the ordinated space occupied:

Because there is only one subset (i.e. one matrix) in the dispRity object, the operations below are the equivalent of `metric(dummy_space)` (with rounding).

```{r}
## Calculating the ellipsoid volume
summary(dispRity(dummy_space, metric = ellipse.volume))
```
> WARNING: in such dummy space, this gives the estimation of the ellipsoid volume, not the real ellipsoid volume! See the cautionary note in `?ellipse.volume`.

```{r}
## Calculating the convex hull surface
summary(dispRity(dummy_space, metric = convhull.surface))

## Calculating the convex hull volume
summary(dispRity(dummy_space, metric = convhull.volume))

## Calculating the convex hull volume
summary(dispRity(dummy_space, metric = n.ball.volume))
```

The convex hull based functions are a call to the `geometry::convhulln` function with the `"FA"` option (computes total area and volume).
Also note that they are really sensitive to the size of the dataset.

> Cautionary note: measuring volumes in a high number of dimensions can be strongly affected by the [curse of dimensionality](https://en.wikipedia.org/wiki/Curse_of_dimensionality) that often results in near 0 disparity values. I strongly recommend reading [this really intuitive explanation](https://beta.observablehq.com/@tophtucker/theres-plenty-of-room-in-the-corners) from [Toph Tucker](https://github.com/tophtucker).

#### Ranges, variances, quantiles, radius, pairwise distance, neighbours, modal value and diagonal

The functions `ranges`, `variances` `radius`, `pairwise.dist`, `mode.val` and `diagonal` all measure properties of the ordinated space based on its dimensional properties (they are also less affected by the "curse of dimensionality"):

`ranges`, `variances` `quantiles` and `radius` work on the same principle and measure the range/variance/radius of each dimension:

```{r}
## Calculating the ranges of each dimension in the ordinated space
ranges(dummy_space)

## Calculating disparity as the distribution of these ranges
summary(dispRity(dummy_space, metric = ranges))

## Calculating disparity as the sum and the product of these ranges
summary(dispRity(dummy_space, metric = c(sum, ranges)))
summary(dispRity(dummy_space, metric = c(prod, ranges)))

## Calculating the variances of each dimension in the ordinated space
variances(dummy_space)

## Calculating disparity as the distribution of these variances
summary(dispRity(dummy_space, metric = variances))

## Calculating disparity as the sum and the product of these variances
summary(dispRity(dummy_space, metric = c(sum, variances)))
summary(dispRity(dummy_space, metric = c(prod, variances)))

## Calculating the quantiles of each dimension in the ordinated space
quantiles(dummy_space)

## Calculating disparity as the distribution of these variances
summary(dispRity(dummy_space, metric = quantiles))

## By default, the quantile calculated is the 95% (i.e. 95% of the data on each axis)
## this can be changed using the option quantile:
summary(dispRity(dummy_space, metric = quantiles, quantile = 50))

## Calculating the radius of each dimension in the ordinated space
radius(dummy_space)

## By default the radius is the maximum distance from the centre of
## the dimension. It can however be changed to any function:
radius(dummy_space, type = min)
radius(dummy_space, type = mean)

## Calculating disparity as the mean average radius
summary(dispRity(dummy_space, metric = c(mean, radius), type = mean))
```

The pairwise distances and the neighbours distances uses the function `vegan::vegdist` and can take the normal `vegdist` options:

```{r}
## The average pairwise euclidean distance
summary(dispRity(dummy_space, metric = c(mean, pairwise.dist)))

## The distribution of the Manhattan distances
summary(dispRity(dummy_space, metric = pairwise.dist, method = "manhattan"))

## The average nearest neighbour distances
summary(dispRity(dummy_space, metric = neighbours))

## The average furthest neighbour manhattan distances
summary(dispRity(dummy_space, metric = neighbours, which = max, method = "manhattan"))
```

Note that this function is a direct call to `vegan::vegdist(matrix, method = method, diag = FALSE, upper = FALSE, ...)`.

The `diagonal` function measures the multidimensional diagonal of the whole space (i.e. in our case the longest Euclidean distance in our five dimensional space).
The `mode.val` function measures the modal value of the matrix:

```{r}
## Calculating the ordinated space's diagonal
summary(dispRity(dummy_space, metric = diagonal))

## Calculating the modal value of the matrix
summary(dispRity(dummy_space, metric = mode.val))
```

> This metric is only a Euclidean diagonal (mathematically valid) if the dimensions within the space are all orthogonal!

#### Centroids, displacements, and ancestral distance metrics {#centroids}

The `centroids` metric allows users to measure the position of the different elements compared to a fixed point in the ordinated space.
By default, this function measures the distance between each element and their centroid (centre point):

```{r}
## The distribution of the distances between each element and their centroid
summary(dispRity(dummy_space, metric = centroids))

## Disparity as the median value of these distances
summary(dispRity(dummy_space, metric = c(median, centroids)))
```

It is however possible to fix the coordinates of the centroid to a specific point in the ordinated space, as long as it has the correct number of dimensions:

```{r}
## The distance between each element and the origin of the ordinated space
summary(dispRity(dummy_space, metric = centroids, centroid = 0))

## Disparity as the distance between each element and a specific point in space
summary(dispRity(dummy_space, metric = centroids, centroid = c(0,1,2,3,4)))
```

If you have subsets in your `dispRity` object, you can also use the `matrix.dispRity` (see [utilities](#utilities)) and `colMeans` to get the centre of a specific subgroup.
For example

```{r, warning = FALSE}
## Create a custom subsets object
dummy_groups <- custom.subsets(dummy_space, group = list("group1" = 1:5, "group2" = 6:10))
summary(dispRity(dummy_groups, metric = centroids,
                 centroid = colMeans(matrix.dispRity(dummy_groups, "group1"))))
```

The `displacements` distance is the ratio between the `centroids` distance and the `centroids` distance with `centroid = 0`.
Note that it is possible to measure a ratio from another point than `0` using the `reference` argument.
It gives indication of the relative displacement of elements in the multidimensional space: a score  >1 signifies a displacement *away* from the reference. A score of >1 signifies a displacement *towards* the reference.

```{r}
## The relative displacement of the group in space to the centre
summary(dispRity(dummy_space, metric = displacements))

## The relative displacement of the group to an arbitrary point
summary(dispRity(dummy_space, metric = displacements, reference = c(0,1,2,3,4)))
```

The `ancestral.dist` metric works on a similar principle as the `centroids` function but changes the centroid to be the coordinates of each element's ancestor.
Therefore this functions needs a tree and node coordinates as additional arguments:

```{r}
## A generating a random tree with node labels
tree <- rtree(5) ; tree$node.label <- paste0("n", 1:4)
## Adding the tip and node names to the matrix
dummy_space2 <- dummy_space[-1,]
rownames(dummy_space2) <- c(tree$tip.label, tree$node.label)

## Calculating all the ancestral nodes
all_anc_centroids <- nodes.coordinates(dummy_space2, tree, full = TRUE)

## Calculating the distances from the ancestral nodes
ancestral_dist <- dispRity(dummy_space2, metric = ancestral.dist,
                           nodes.coords = all_anc_centroids)

## The ancestral distances distributions
summary(ancestral_dist)

## Calculating disparity as the sum of the distances from all the ancestral nodes
summary(dispRity(ancestral_dist, metric = sum))
```

#### Minimal spanning tree length

The `span.tree.length` uses the `vegan::spantree` function to heuristically calculate the minimum spanning tree (the shortest multidimensional tree connecting each elements) and calculates its length as the sum of every branch lengths.

```{r}
## The length of the minimal spanning tree
summary(dispRity(dummy_space, metric = c(sum, span.tree.length)))
```

Note that because the solution is heuristic, this metric can take a long time to compute for big matrices.

#### Functional divergence and evenness

The `func.div` and `func.eve` functions are based on the `FD::dpFD` package.
They are the equivalent to `FD::dpFD(matrix)$FDiv` and `FD::dpFD(matrix)$FEve` but a bit faster (since they don't deal with abundance data).
They are pretty straightforward to use:

```{r}
## The ratio of deviation from the centroid 
summary(dispRity(dummy_space, metric = func.div))

## The minimal spanning tree distances evenness
summary(dispRity(dummy_space, metric = func.eve))

## The minimal spanning tree manhanttan distances evenness
summary(dispRity(dummy_space, metric = func.eve, method = "manhattan"))
```

#### Orientation: angles and deviations

The `angles` performs a least square regression (via the `lm` function) and returns slope of the main axis of variation for each dimension. This slope can be converted into different units, `"slope"`, `"degree"` (the default) and `"radian"`. This can be changed through the `unit` argument. 
By default, the angle is measured from the slope 0 (the horizontal line in a 2D plot) but this can be changed through the `base` argument (using the defined `unit`):

```{r}
## The distribution of each angles in degrees for each main axis in the matrix
summary(dispRity(dummy_space, metric = angles))

## The distribution of slopes deviating from the 1:1 slope:
summary(dispRity(dummy_space, metric = angles, unit = "slope", base = 1))
```

The `deviations` function is based on a similar algorithm as above but measures the deviation from the main axis (or hyperplane) of variation.
In other words, it finds the least square line (for a 2D dataset), plane (for a 3D dataset) or hyperplane (for a >3D dataset) and measures the shortest distances between every points and the line/plane/hyperplane.
By default, the hyperplane is fitted using the least square algorithm from `stats::glm`:

```{r}
## The distribution of the deviation of each point from the least square hyperplane
summary(dispRity(dummy_space), metric = deviations))
```

It is also possible to specify the hyperplane equation through the `hyperplane` equation. The equation must contain the intercept first and then all the slopes and is interpreted as $intercept + Ax + By + ... + Nd = 0$. For example, a 2 line defined as beta + intercept (e.g. $y = 2x + 1$) should be defined as `hyperplane = c(1, 2, 1)` ($2x - y + 1 = 0$).

```{r}
## The distribution of the deviation of each point from a slope (with only the two first dimensions)
summary(dispRity(dummy_space[, c(1:2)]), metric = deviations, hyperplane = c(1, 2, -1)))
```

Since both the functions `angles` and `deviations` effectively run a `lm` or `glm` to estimate slopes or hyperplanes, it is possible to use the option `significant = TRUE` to only consider slopes or intercepts that have a slope significantly different than zero using an `aov` with a significant threshold of $p = 0.05$.
Note that depending on your dataset, using and `aov` could be completely inappropriate!
In doubt, it's probably better to enter your `base` (for `angles`) or your `hyperplane` (for `deviations`) manually so you're sure you know what the function is measuring.

## Summarising dispRity data (plots)

Because of its architecture, printing `dispRity` objects only summarises their content but does not print the disparity value measured or associated analysis (more about this [here](#manipulating-dispRity-objects)).
To actually see what is in a dispRity object, one can either use the `summary` function for visualising the data in a table or `plot` to have a graphical representation of the results.

### Summarising `dispRity` data
This function is an S3 function (`summary.dispRity`) allowing users to summarise the content of `dispRity` objects that contain disparity calculations.

```{r}
## Example data from previous sections
crown_stem <- custom.subsets(BeckLee_mat50,
                                group = crown.stem(BeckLee_tree, inc.nodes = FALSE))
## Bootstrapping and rarefying these groups
boot_crown_stem <- boot.matrix(crown_stem, bootstraps = 100, rarefaction = TRUE)
## Calculate disparity
disparity_crown_stem <- dispRity(boot_crown_stem, metric = c(sum, variances))

## Creating time slice subsets
time_slices <- chrono.subsets(data = BeckLee_mat99, tree = BeckLee_tree, 
    method = "continuous", model = "proximity", time = c(120, 80, 40, 0),
    FADLAD = BeckLee_ages)
## Bootstrapping the time slice subsets
boot_time_slices <- boot.matrix(time_slices, bootstraps = 100)
## Calculate disparity
disparity_time_slices <- dispRity(boot_time_slices, metric = c(sum, variances))

## Creating time bin subsets
time_bins <- chrono.subsets(data = BeckLee_mat99, tree = BeckLee_tree, 
    method = "discrete", time = c(120, 80, 40, 0), FADLAD = BeckLee_ages,
    inc.nodes = TRUE)
## Bootstrapping the time bin subsets
boot_time_bins <- boot.matrix(time_bins, bootstraps = 100)
## Calculate disparity
disparity_time_bins <- dispRity(boot_time_bins, metric = c(sum, variances))
```

These objects are easy to summarise as follows:

```{r}
## Default summary
summary(disparity_time_slices)
```

Information about the number of elements in each subset and the observed (i.e. non-bootstrapped) disparity are also calculated.
This is specifically handy when rarefying the data for example:

```{r}
head(summary(disparity_crown_stem))
```

The summary functions can also take various options such as:

 * `quantiles` values for the confidence interval levels (by default, the 50 and 95 quantiles are calculated)
 * `cent.tend` for the central tendency to use for summarising the results (default is `median`)
 * digits` option corresponding to the number of decimal places to print (default is `2`)
 * `recall` option for printing the call of the `dispRity` object as well (default is `FALSE`)

These options can easily be changed from the defaults as follows:

```{r}
## Same as above but using the 88th quantile and the standard deviation as the summary 
summary(disparity_time_slices, quantiles = 88, cent.tend = sd)

## Printing the details of the object and digits the values to the 5th decimal place
summary(disparity_time_slices, recall = TRUE, digits = 5)
```

Note that the summary table is a `data.frame`, hence it is as easy to modify as any dataframe using `dplyr`.
You can also export it in `csv` format using `write.csv` or `write_csv` or even directly export into `LaTeX` format using the following;

```{r, eval=FALSE}
## Loading the xtable package
require(xtable)
## Converting the table in LaTeX
xtable(summary(disparity_time_slices))
```

### Plotting `dispRity` data
An alternative (and more fun!) way to display the calculated disparity is to plot the results using the S3 method `plot.dispRity`.
This function takes the same options as `summary.dispRity` along with various graphical options described in the function help files (see `?plot.dispRity`).

The plots can be of five different types:

 * `preview` for a 2d preview of the trait-space. 
 * `continuous` for displaying continuous disparity curves
 * `box`, `lines`, and `polygons` to display discrete disparity results in respectively a boxplot, confidence interval lines, and confidence interval polygons.

> This argument can be left empty. In this case, the algorithm will automatically detect the type of subsets from the `dispRity` object and plot accordingly.

It is also possible to display the number of elements in each subset (as a horizontal dotted line) using the option `elements = TRUE`.
Additionally, when the data is rarefied, one can indicate which level of rarefaction to display (i.e. only display the results for a certain number of elements) by using the `rarefaction` argument.

```{r, fig.width=8, fig.height=8}
## Graphical parameters
op <- par(mfrow = c(2, 2), bty = "n")

## Plotting continuous disparity results
plot(disparity_time_slices, type = "continuous")

## Plotting discrete disparity results
plot(disparity_crown_stem, type = "box")

## As above but using lines for the rarefaction level of 20 elements only
plot(disparity_crown_stem, type = "line", rarefaction = 20)

## As above but using polygons while also displaying the number of elements
plot(disparity_crown_stem, type = "polygon", elements = TRUE)

## Resetting graphical parameters
par(op)
```

Since `plot.dispRity` uses the arguments from the generic `plot` method, it is of course possible to change pretty much everything using the regular plot arguments:

```{r, fig.width=8, fig.height=8}
## Graphical options
op <- par(bty = "n")

## Plotting the results with some classic options from plot
plot(disparity_time_slices, col = c("blue", "orange", "green"),
    ylab = c("Some measurement"), xlab = "Some other measurement",
    main = "Many options...", ylim = c(5, 0), xlim = c(4, 0))

## Adding a legend
legend("topleft", legend = c("Central tendency",
                             "Confidence interval 1",
                            "Confidence interval 2"),
      col = c("blue", "orange", "green"), pch = 19)

## Resetting graphical parameters
par(op)
```

In addition to the classic `plot` arguments, the function can also take arguments that are specific to `plot.dispRity` like adding the number of elements or rarefaction level (as described above), and also changing the values of the quantiles to plot as well as the central tendency.

```{r, fig.width=8, fig.height=8}
## Graphical options
op <- par(bty = "n")

## Plotting the results with some plot.dispRity arguments
plot(disparity_time_slices, quantiles = c(seq(from = 10, to = 100, by = 10)),
    cent.tend = sd, type = "c", elements = TRUE, col = c("black", rainbow(10)),
    ylab = c("Disparity", "Diversity"), chrono.subsets = FALSE,
    xlab = "Time (in in units from past to present)", observed = TRUE,
    main = "Many more options...")

## Resetting graphical parameters
par(op)
```

> Note that the argument `observed = TRUE` allows to plot the disparity values calculated from the non-bootstrapped data as crosses on the plot.

For comparing results, it is also possible to add a plot to the existent plot by using `add = TRUE`:

```{r, fig.width=8, fig.height=8}
## Graphical options
op <- par(bty = "n")

## Plotting the continuous disparity with a fixed y axis
plot(disparity_time_slices, ylim = c(1, 4))
## Adding the discrete data
plot(disparity_time_bins, type = "line", ylim = c(1, 4), xlab = "", ylab = "",
    add = TRUE)

## Resetting graphical parameters
par(op)
```

Finally, if your data has been fully rarefied, it is also possible to easily look at rarefaction curves by using the `rarefaction = TRUE` argument:

```{r, fig.width=8, fig.height=8}
## Graphical options
op <- par(bty = "n")

## Plotting the rarefaction curves
plot(disparity_crown_stem, rarefaction = TRUE)

## Resetting graphical parameters
par(op)
```

### `type = preview`

Note that all the options above are plotting disparity objects for which a disparity metric *has been calculated*.
This makes totally sense for `dispRity` objects but sometimes it might be interesting to look at what the trait-space looks like before measuring the disparity.
This can be done by plotting `dispRity` objects with no calculated disparity!

For example, we might be interested in looking at how the distribution of elements change as a function of the distributions of different sub-settings.
For example custom subsets *vs.* time subsets:

```{r, fig.width=8, fig.height=4}
## Making the different subsets
cust_subsets <- custom.subsets(BeckLee_mat99, crown.stem(BeckLee_tree, inc.nodes = TRUE))
time_subsets <- chrono.subsets(BeckLee_mat99, tree = BeckLee_tree, method = "discrete", time = 5)

## Note that no disparity has been calculated here:
is.null(cust_subsets$disparity)
is.null(time_subsets$disparity)

## But we can still plot both spaces by using the default plot functions
par(mfrow = c(1,2))
## Default plotting
plot(cust_subsets)
## Plotting with more arguments
plot(time_subsets, dimensions = c(4,5), main = "Some \"low\" dimensions")
```

> DISCLAIMER: This functionality can be handy for exploring the data (e.g. to visually check whether the subset attribution worked) but it might be misleading on how the data is *actually* distributed in the multidimensional space!
Groups that don't overlap on two set dimensions can totally overlap in all other dimensions!

For `dispRity` objects that do contain disparity data, the default option is to plot your disparity data.
However you can always force the `preview` option using the following:

```{r}
par(mfrow = c(2,1))
## Default plotting
plot(disparity_time_slices, main = "Disparity through time")
## Plotting with more arguments
plot(disparity_time_slices, type = "preview", main = "Two first dimensions of the trait space")
```

## Testing disparity hypotheses

The `dispRity` package allows users to apply statistical tests to the calculated disparity to test various hypotheses.
The function `test.dispRity` works in a similar way to the `dispRity` function: it takes a `dispRity` object, a `test` and a `comparisons` argument.

The `comparisons` argument indicates the way the test should be applied to the data:

 * `pairwise` (default): to compare each subset in a pairwise manner
 * `referential`: to compare each subset to the first subset
 * `sequential`: to compare each subset to the following subset
 * `all`: to compare all the subsets together (like in analysis of variance)

It is also possible to input a list of pairs of `numeric` values or `characters` matching the subset names to create personalised tests.
Some other tests implemented in `dispRity` such as the `dispRity::null.test` have a specific way they are applied to the data and therefore ignore the `comparisons` argument. 
<!-- Add sequential test one day! -->

The `test` argument can be any statistical or non-statistical test to apply to the disparity object.
It can be a common statistical test function (e.g. `stats::t.test`), a function implemented in `dispRity` (e.g. see `?null.test`) or any function defined by the user.

This function also allows users to correct for Type I error inflation (false positives) when using multiple comparisons via the `correction` argument. 
This argument can be empty (no correction applied) or can contain one of the corrections from the `stats::p.adjust` function (see `?p.adjust`).

Note that the `test.dispRity` algorithm deals with some classical test outputs (`h.test`, `lm` and `numeric` vector) and summarises the test output.
It is, however, possible to get the full detailed output by using the options `details = TRUE`.

Here we are using the variables generated in the [section above](#summarising-dispRity-data-plots):

```{r}
## T-test to test for a difference in disparity between crown and stem mammals
test.dispRity(disparity_crown_stem, test = t.test)

## Performing the same test but with the detailed t.test output
test.dispRity(disparity_crown_stem, test = t.test, details = TRUE)

## Wilcoxon test applied to time sliced disparity with sequential comparisons,
## with Bonferroni correction
test.dispRity(disparity_time_slices, test = wilcox.test,
              comparisons = "sequential", correction = "bonferroni")

## Measuring the overlap between distributions in the time bins (using the
## implemented Bhattacharyya Coefficient function - see ?bhatt.coeff)
test.dispRity(disparity_time_bins, test = bhatt.coeff)
```

Because of the modular design of the package, tests can always be made by the user (the same way disparity metrics can be user made).
The only condition is that the test can be applied to at least two distributions.
In practice, the `test.dispRity` function will pass the calculated disparity data (distributions) to the provided function in either pairs of distributions (if the `comparisons` argument is set to `pairwise`, `referential` or `sequential`) or a table containing all the distributions (`comparisons = all`; this should be in the same format as data passed to `lm`-type functions for example).

### NPMANOVA in `dispRity`

One often useful test to apply to multidimensional data is the permutational multivariate analysis of variance based on distance matrices `vegan::adonis`.
This can be done on `dispRity` objects using the `adonis.dispRity` wrapper function.
Basically, this function takes the exact same arguments as `adonis` and a `dispRity` object for data and performs a PERMANOVA based on the distance matrix of the multidimensional space (unless the multidimensional space was already defined as a distance matrix).
The `adonis.dispRity` function uses the information from the `dispRity` object to generate default formulas:

  * If the object contains customised subsets, it applies the default formula `matrix ~ group` testing the effect of `group` as a predictor on `matrix` (called from the `dispRity` object as `data$matrix` see [`dispRitu` object details](#The-dispRity-object-content))
  * If the object contains time subsets, it applies the default formula `matrix ~ time` testing the effect of `time` as a predictor (were the different levels of `time` are the different time slices/bins)

```{r}
set.seed(1)
## Generating a random character matrix
character_matrix <- sim.morpho(rtree(20), 50, rates = c(rnorm, 1, 0))

## Calculating the distance matrix
distance_matrix <- as.matrix(dist(character_matrix))

## Creating two groups
random_groups <- list("group1" = 1:10, "group2" = 11:20)

## Generating a dispRity object
random_disparity <- custom.subsets(distance_matrix, random_groups)

## Running a default NPMANOVA
adonis.dispRity(random_disparity)
```

Of course, it is possible to pass customised formulas if the disparity object contains more more groups.
In that case the predictors must correspond to the names of the groups explained data must be set as `matrix`:

```{r}
## Creating two groups with two states each
groups <- as.data.frame(matrix(data = c(rep(1,10), rep(2,10), rep(c(1,2), 10)),
         nrow = 20, ncol = 2, dimnames = list(paste0("t", 1:20), c("g1", "g2"))))

## Creating the dispRity object
multi_groups <- custom.subsets(distance_matrix, groups)

## Running the NPMANOVA
adonis.dispRity(multi_groups, matrix ~ g1 + g2)
```

Finally, it is possible to use objects generated by `chrono.subsets`.
In this case, `adonis.dispRity` will applied the `matrix ~ time` formula by default:

```{r}
## Creating time series
time_subsets <- chrono.subsets(BeckLee_mat50, BeckLee_tree, 
     method = "discrete", inc.nodes = FALSE, time = c(100, 85, 65, 0),
     FADLAD = BeckLee_ages)

## Running the NPMANOVA with time as a predictor
adonis.dispRity(time_subsets)
```

Note that the function warns you that the input data was transformed into a distance matrix.
This is reflected in the Call part of the output (`formula = dist(matrix) ~ time`).

To use each time subset as a separate predictor, you can use the `matrix ~ chrono.subsets` formula; this is equivalent to `matrix ~ first_time_subset + second_time_subset + ...`:

```{r}
## Running the NPMANOVA with each time bin as a predictor
adonis.dispRity(time_subsets, matrix ~ chrono.subsets)
```

### `geiger::dtt` model fitting in `dispRity`

The `dtt` function from the `geiger` package is also often used to compare a trait's disparity observed in living taxa to the disparity of a simulated trait based on a given phylogeny.
The `dispRity` package proposes a wrapper function for `geiger::dtt`, `dtt.dispRity` that allows the use of any disparity metric.
Unfortunately, this implementation is slower that `geiger::dtt` (so if you're using the metrics implemented in `geiger` prefer the original version) and, as the original function, is limited to ultrametric trees (only living taxa!)...

```{r}
require(geiger)
geiger_data <- get(data(geospiza))

## Calculate the disparity of the dataset using the sum of variance
dispRity_dtt <- dtt.dispRity(data = geiger_data$dat, metric = c(sum, variances),
                             tree = geiger_data$phy, nsim = 100)

## Plotting the results
plot(dispRity_dtt, fig.width=8, fig.height=8)
```

Note that, like in the original `dtt` function, it is possible to change the evolutionary model (see `?geiger::sim.char` documentation).

### null morphospace testing with `null.test`

This test is equivalent to the test performed in @diaz2016global.
It compares the disparity measured in the observed space to the disparity measured in a set of simulated spaces.
These simulated spaces can be built with based on the hypothesis assumptions: for example, we can test whether our space is normal.

```{r}
set.seed(123)
## A "normal" multidimensional space with 50 dimensions and 10 elements
normal_space <- matrix(rnorm(1000), ncol = 50)

## Calculating the disparity as the average pairwise distances
obs_disparity <- dispRity(normal_space, metric = c(mean, pairwise.dist))

## Testing against 100 randomly generated normal spaces
(results <- null.test(obs_disparity, replicates = 100, null.distrib = rnorm))
```

Here the results show that disparity measured in our observed space is not significantly different than the one measured in a normal space.
We can then propose that our observed space is normal!

These results have an attributed `dispRity` and `randtest` class and can be plotted as `randtest` objects using the `dispRity` S3 `plot` method:

```{r, fig.width=8, fig.height=8}
## Plotting the results
plot(results, main = "Is this space normal?")
```

For more details on generating spaces see the [`space.maker`](#Simulating-multidimensional-spaces) function tutorial.

## Fitting modes of evolution to disparity data

The code used for these models is based on those developed by Gene Hunt [@hunt2006fitting; @hunt2012measuring; @hunt2015simple].
So we acknowledge and thank Gene Hunt for developing these models and writing the original R code that served as inspiration for these models.

### Simple modes of disparity change through time

#### `model.test`

Changes in disparity-through-time can follow a range of models, such as random walks, stasis, constrained evolution, trends, or an early burst model of evolution.
We will start with by fitting the simplest modes of evolution to our data.
For example we may have a null expectation of time-invariant change in disparity in which values fluctuate with a variance around the mean - this would be best describe by a Stasis model:

```{r}
## Loading premade disparity data
data(BeckLee_disparity)
disp_time <- model.test(data = BeckLee_disparity, model = "Stasis")
```

We can see the standard output from ```model.test```.
The first output message tells us it has tested for equal variances in each sample.
The model uses Bartlett's test of equal variances to assess if variances are equal, so if p > 0.05 then variance is treated as the same for all samples, but if (p < 0.05) then each bin variance is unique.
Here we have p < 0.05, so variance is not pooled between samples.

By default ```model.test``` will use Bartlett's test to assess for homogeneity of variances, and then use this to decide to pool variances or not.
This is ignored if the argument ```pool.variance``` in ```model.test``` is changed from the default ```NULL``` to ```TRUE``` or ```FALSE```.
For example, to ignore Bartlett's test and pool variances manually we would do the following:

```{r}
disp_time_pooled <- model.test(data = BeckLee_disparity, model = "Stasis", pool.variance = TRUE)
```

However, unless you have good reason to choose otherwise it is recommended to use the default of ```pool.variance = NULL```:

```{r}
disp_time <- model.test(data = BeckLee_disparity, model = "Stasis", pool.variance = NULL)
disp_time
```

The remaining output gives us the log-likelihood of the Stasis model of -59.501 (you may notice this change when we pooled variances above).
The output also gives us the small sample Akaike Information Criterion (AICc), the delta AICc (the distance from the best fitting model), and the AICc weights (~the relative support of this model compared to all models, scaled to one).

These are all metrics of relative fit, so when we test a single model they are not useful.
By using the function summary in ```dispRity``` we can see the maximum likelihood estimates of the model parameters:

```{r}
summary(disp_time)
```

So we again see the AICc, delta AICc, AICc weight, and the log-likelihood we saw previously.
We now also see the number of parameters from the model (2: theta and omega), and their estimates so the variance (omega = 0.01) and the mean (theta.1 = 3.4).

The ```model.test``` function is designed to test relative model fit, so we need to test more than one model to make relative comparisons.
So let's compare to the fit of the Stasis model to another model with two parameters: the Brownian motion.
Brownian motion assumes a constant mean that is equal to the ancestral estimate of the sequence, and the variance around this mean increases linearly with time.
The easier way to compare these models is to simply add `"BM"` to the ```models``` vector argument:

```{r}
disp_time <- model.test(data = BeckLee_disparity, model = c("Stasis", "BM"))
disp_time
```

Et voilà! Here we can see by the log-likelihood, AICc, delta AICc, and AICc weight Brownian motion has a much better relative fit to these data than the Stasis model.
Brownian motion has a relative AICc fit 366 units better than Stasis, and virtually has a AICc weight of 1.

We can also all the information about the relative fit of models alongside the maximum likelihood estimates of model parameters using the summary function

```{r}
summary(disp_time)
```

Not that because the parameters per models differ, the summary includes `NA` for inapplicable parameters per models (e.g. the theta and omega parameters from the Stasis models are inapplicable for a Brownian motion model).

We can plot the relative fit of our models using the ```plot``` function

```{r plot1, fig.align="center", fig.cap = "relative fit (AICc weight) of Stasis and Brownian models of disparity through time", echo = TRUE, fig.height = 6, fig.width = 6}
plot(disp_time)
```

Here we see and overwhelming support for the Brownian motion model.

Alternatively, we could test all available models single modes: Stasis, Brownian motion, Ornstein-Uhlenbeck (evolution constrained to an optima), Trend (increasing or decreasing mean through time), and Early Burst (exponentially decreasing rate through time)

```{r}
disp_time <- model.test(data = BeckLee_disparity, model = c("Stasis", "BM", "OU", "Trend", "EB"))
summary(disp_time)
```

These models indicate support for a Trend model, and we can plot the relative support of all model AICc weights.

```{r plot2, fig.align="center", fig.cap = "relative fit (AICc weight) of various modes of evolution", echo = TRUE, fig.height = 6, fig.width = 6}
plot(disp_time)
```

> Note that although AIC values are indicator of model best fit, it is also important to look at the parameters themselves.
For example OU can be really well supported but with an alpha parameter really close to 0, making it effectively a BM model [@Cooper2016].

Is this a trend of increasing or decreasing disparity through time? One way to find out is to look at the summary function for the Trend model:

```{r}
summary(disp_time)["Trend",]
```

This show a positive trend (0.01) of increasing disparity through time.

### Plot and run simulation tests in a single step

#### `model.test.wrapper`

Patterns of evolution can be fit using ```model.test```, but the ```model.test.wrapper``` fits the same models as `model.test` as well as running predictive tests and plots.

The predictive tests use the maximum likelihood estimates of model parameters to simulate a number of datasets (default = 1000), and analyse whether this is significantly different to the empirical input data using the Rank Envelope test [@murrell2018global].
Finally we can plot the empirical data, simulated data, and the Rank Envelope test p values.
This can all be done using the function ```model.test.wrapper```, and we will set the argument ```show.p = TRUE``` so _p_ values from the Rank Envelope test are printed on the plot:

```{r plot3, fig.align="center", fig.cap = "Empirical disparity through time (pink), simulate data based on estimated model parameters (grey), delta AICc, and range of p values from the Rank Envelope test for Trend, OU, BM, EB, and Stasis models", echo = TRUE, fig.height = 18, fig.width = 12}
disp_time <- model.test.wrapper(data = BeckLee_disparity, model = c("Stasis", "BM", "OU", "Trend", "EB"),
                                show.p = TRUE)
disp_time
```

From this plot we can see the empirical estimates of disparity through time (pink) compared to the predictive data based upon the simulations using the estimated parameters from each model.
There is no significant differences between the empirical data and simulated data, except for the Early Burst model.

Trend is the best-fitting model but the plot suggests the OU model also follows a trend-like pattern.
This is because the optima for the OU model (5.7) is different to the ancestral state 2.835 and outside the observed value.
This is potentially unrealistic, and one way to alleviate this issue is to set the optima of the OU model to equal the ancestral estimate - this is the normal practice for OU models in comparative phylogenetics.
To set the optima to the ancestral value we change the argument ```fixed.optima = TRUE```:

```{r plot4, fig.align="center", fig.cap = "Empirical disparity through time (pink), simulate data based on estimated model parameters (grey), delta AICc, and range of p values from the Rank Envelope test for Trend, OU, BM, EB, and Stasis models with the optima of the OU model set to equal the ancestral value", echo = TRUE, fig.height = 18, fig.width = 12}
disp_time <- model.test.wrapper(data = BeckLee_disparity, model = c("Stasis", "BM", "OU", "Trend", "EB"),
                                show.p = TRUE, fixed.optima = TRUE)
disp_time
```

The relative fit of the OU model is decreased by constraining the fit of the optima to equal the ancestral state value.
In fact as the OU attraction parameter (alpha) is zero, the model is equal to a Brownian motion model but is penalised by having an extra parameter. 
Note that indeed, the plots of the BM model and the OU model look nearly identical.

### Multiple modes of evolution (time shifts)

As well as fitting a single model to a sequence of disparity values we can also allow for the mode of evolution to shift at a single or multiple points in time.
The timing of a shift in mode can be based on an a prior expectation, such as a mass extinction event, or the model can test multiple points to allow to find time shift point with the highest likelihood.

Models can be fit using `model.test` but it can be more convenient to use `model.test.wrapper`.
Here we will compare the relative fit of Brownian motion, Trend, Ornstein-Uhlenbeck and a multi-mode Ornstein Uhlenbck model in which the optima changes at 66 million years ago, the Cretaceous-Palaeogene boundary.

For example, we could be testing the hypothesis that the extinction of non-avian dinosaurs allowed mammals to go from scurrying in the undergrowth (low optima/low disparity) to dominating all habitats (high optima/high disparity).
We will constrain the optima of OU model in the first time begin (i.e, pre-66 Mya) to equal the ancestral value:

```{r plot5, fig.align="center", fig.cap = "Empirical disparity through time (pink), simulate data based on estimated model parameters (grey), delta AICc, and range of p values from the Rank Envelope test for BM, Trend, OU, and multi OU models with a shift in optima allowed at 66 Ma", echo = TRUE, fig.height = 12, fig.width = 12}
disp_time <- model.test.wrapper(data = BeckLee_disparity, model = c("BM", "Trend", "OU", "multi.OU"),
                                time.split = 66, pool.variance = NULL, show.p = TRUE,
                                fixed.optima = TRUE)
disp_time
```

The multi-OU model shows an increase an optima at the Cretaceous-Palaeogene boundary, indicating a shift in disparity.
However, this model does not fit as well as a model in which there is an increasing trend through time.
We can also fit a model in which the we specify a heterogeneous model but we do not give a `time.split`.
In this instance the model will test all splits that have at least 10 time slices on either side of the split.
That's 102 potential time shifts in this example dataset so be warned, the following code will estimate 105 models!

```{r, eval = FALSE}
## An example of a time split model in which all potential splits are tested
## WARNING: this will take between 20 minutes and half and hour to run!
disp_time <- model.test.wrapper(data = BeckLee_disparity, model = c("BM", "Trend", "OU", "multi.OU"),
                                show.p = TRUE, fixed.optima = TRUE)
```

As well as specifying a multi-OU model we can run any combination of models.
For example we could fit a model at the Cretaceous-Palaeogene boundary that goes from an OU to a BM model, a Trend to an OU model, a Stasis to a Trend model or any combination you want to use.
The only model that can't be used in combination is a multi-OU model.

These can be introduced by changing the input for the models into a list, and supplying a vector with the two models.
This is easier to see with an example:

```{r plot6, fig.align="center", fig.cap = "Empirical disparity through time (pink), simulate data based on estimated model parameters (grey), delta AICc, and range of p values from the Rank Envelope test for a variety of models with a shift in optima allowed at 66 Ma", echo = TRUE, fig.height = 18, fig.width = 12}
## The models to test
my_models <- list(c("BM", "OU"),
                  c("Stasis", "OU"),
                  c("BM", "Stasis"),
                  c("OU", "Trend"),
                  c("Stasis", "BM"))

## Testing the models
disp_time <- model.test.wrapper(data = BeckLee_disparity, model = my_models, time.split = 66,
                                show.p = TRUE, fixed.optima = TRUE)
disp_time
```

### `model.test.sim`

Note that all the models above where run using the `model.test.wrapper` function that is a... wrapping function!
In practice, this function runs two main functions from the `dispRity` package and then plots the results:

 * `model.test` and
 * `model.test.sim`

The `model.test.sim` allows to simulate disparity evolution given a `dispRity` object input (as in `model.test.wrapper`) or given a model and its specification.
For example, it is possible to simulate a simple Brownian motion model (or any of the other models or models combination described above):

```{r}
## A simple BM model
model_simulation <- model.test.sim(sim = 1000, model = "BM", time.span = 50, variance = 0.1,
                                   sample.size = 100, parameters = list(ancestral.state = 0))
model_simulation
```

This will simulate 1000 Brownian motions for 50 units of time with 100 sampled elements, a variance of 0.1 and an ancestral state of 0.
We can also pass multiple models in the same way we did it for `model.test`
This model can then be summarised and plotted as most `dispRity` objects:

```{r plot7, fig.align="center", fig.cap = "A simulated Brownian motion", echo = TRUE, fig.height = 6, fig.width = 6}
## Displaying the 5 first rows of the summary
head(summary(model_simulation))

## Plotting the simulations
plot(model_simulation)
```

Note that these functions can take all the arguments that can be passed to `plot`, `summary`, `plot.dispRity` and `summary.dispRity`.

#### Simulating tested models

Maybe more interestingly though, it is possible to pass the output of `model.test` directly to `model.test.sim` to simulate the models that fits the data the best and calculate the Rank Envelope test _p_ value.
Let's see that using the simple example from the start:

```{r}
## Fitting multiple models on the data set
disp_time <- model.test(data = BeckLee_disparity, model = c("Stasis", "BM", "OU", "Trend", "EB"))
summary(disp_time)
```

As seen before, the Trend model fitted this dataset the best.
To simulate what 1000 Trend models would look like using the same parameters as the ones estimated with `model.test` (here the ancestral state being 2.839, the sigma squared beeing 0.002 and the trend of 0.01), we can simply pass this model to `model.test.sim`:

```{r}
## Simulating 1000 Trend model with the observed parameters
sim_trend <- model.test.sim(sim = 1000, model = disp_time)
sim_trend
```

By default, the model simulated is the one with the lowest AICc (`model.rank = 1`) but it is possible to choose any ranked model, for example, the OU (second one):

```{r}
## Simulating 1000 OU model with the observed parameters
sim_OU <- model.test.sim(sim = 1000, model = disp_time, model.rank = 2)
sim_OU
```

And as the example above, the simulated data can be plotted or summarised:

```{r}
head(summary(sim_trend))
head(summary(sim_OU))
```

```{r plot8, fig.align="center", fig.cap = "The best fitted model (Trend) and the observed disparity through time", echo = TRUE, fig.height = 6, fig.width = 6}
## The trend model with some graphical options
plot(sim_trend, xlab = "Time (Mya)", ylab = "sum of variances",
    col = c("#F65205", "#F38336", "#F7B27E"))

## Adding the observed disparity through time
plot(BeckLee_disparity, add = TRUE, col = c("#3E9CBA", "#98D4CF90", "#BFE4E390"))
```

## Disparity as a distribution

Disparity is often regarded as a summary value of the position of the all elements in the ordinated space.
For example, the sum of variances, the product of ranges or the median distance between the elements and their centroid will summarise disparity as a single value.
This value can be pseudo-replicated (bootstrapped) to obtain a distribution of the summary metric with estimated error.
However, another way to perform disparity analysis is to use the *whole distribution* rather than just a summary metric (e.g. the variances or the ranges).

This is possible in the `dispRity` package by calculating disparity as a dimension-level 2 metric only!
Let's have a look using our [previous example](#summarising-dispRity-data-plots) of bootstrapped time slices but by measuring the distances between each taxon and their centroid as disparity.

```{r}
## Measuring disparity as a whole distribution
disparity_centroids <- dispRity(boot_time_slices, metric = centroids)
```

The resulting disparity object is of dimension-level 2, so it can easily be transformed into a dimension-level 1 object by, for example, measuring the median distance of all these distributions:

```{r}
## Measuring median disparity in each time slice
disparity_centroids_median <- dispRity(disparity_centroids, metric = median)
```

And we can now compare the differences between these methods:

```{r}
## Summarising both disparity measurements:
## The distributions:
summary(disparity_centroids)
## The summary of the distributions (as median)
summary(disparity_centroids_median)
```

We can see that the summary message for the distribution is slightly different than before.
Here `summary` also displays the observed central tendency (i.e. the central tendency of the measured distributions).
Note that, as expected, this central tendency is the same in both metrics!

Another, maybe more intuitive way, to compare both approaches for measuring disparity is to plot the distributions:

```{r, fig.width=8, fig.height=8}
## Graphical parameters
op <- par(bty = "n", mfrow = c(1, 2))

## Plotting both disparity measurements
plot(disparity_centroids, ylab = "Distribution of all the distances")
plot(disparity_centroids_median,
     ylab = "Distribution of the medians of all the distances")

par(op)
```

We can then test for differences in the resulting distributions using `test.dispRity` and the `bhatt.coeff` test as described above.

```{r}
## Probability of overlap in the distribution of medians
test.dispRity(disparity_centroids_median, test = bhatt.coeff)
```

In this case, we are looking at the probability of overlap of the distribution of median distances from centroids among each pair of time slices.
In other words, we are measuring whether the medians from each bootstrap pseudo-replicate for each time slice overlap.
But of course, we might be interested in the actual distribution of the distances from the centroid rather than simply their central tendencies.
This can be problematic depending on the research question asked since we are effectively comparing non-independent medians distributions (because of the pseudo-replication).

One solution, therefore, is to look at the full distribution:

```{r}
## Probability of overlap for the full distributions
test.dispRity(disparity_centroids, test = bhatt.coeff)
```

These results show the actual overlap among all the measured distances from centroids concatenated across all the bootstraps.
For example, when comparing the slices 120 and 80, we are effectively comparing the 5 $\times$ 100 distances (the distances of the five elements in slice 120 bootstrapped 100 times) to the 19 $\times$ 100 distances from slice 80.
However, this can also be problematic for some specific tests since the _n_ $\times$ 100 distances are also pseudo-replicates and thus are still not independent.

A second solution is to compare the distributions to each other _for each replicate_:

```{r}
## Boostrapped probability of overlap for the full distributions
test.dispRity(disparity_centroids, test = bhatt.coeff, concatenate = FALSE)
```

These results show the median overlap among pairs of distributions in the first column (`bhatt.coeff`) and then the distribution of these overlaps among each pair of bootstraps.
In other words, when two distributions are compared, they are now compared for each bootstrap pseudo-replicate, thus effectively creating a distribution of probabilities of overlap.
For example, when comparing the slices 120 and 80, we have a mean probability of overlap of 0.28 and a probability between 0.18 and 0.43 in 50\% of the pseudo-replicates.
Note that the quantiles and central tendencies can be modified via the `conc.quantiles` option.

## Disparity from other matrices

In the example so far, disparity was measured from an ordinated multidimensional space (i.e. a PCO of the distances between taxa based on discrete morphological characters).
This is a common approach in palaeobiology, morphometrics or ecology but ordinated matrices are not mandatory for the `dispRity` package!
It is totally possible to perform the same analysis detailed above using other types of matrices as long as your elements are rows in your matrix.

For example, we can use the data set `eurodist`, an `R` inbuilt dataset that contains the distances (in km) between European cities.
We can check for example, if Northern European cities are closer to each other than Southern ones:

```{r}
## Making the eurodist data set into a matrix (rather than "dist" object)
eurodist <- as.matrix(eurodist)
eurodist[1:5, 1:5]

## The two groups of cities
Northern <- c("Brussels", "Calais", "Cherbourg", "Cologne", "Copenhagen",
              "Hamburg", "Hook of Holland", "Paris", "Stockholm")
Southern <- c("Athens", "Barcelona", "Geneva", "Gibraltar", "Lisbon", "Lyons",
              "Madrid", "Marseilles", "Milan", "Munich", "Rome", "Vienna")

## Creating the subset dispRity object
eurodist_subsets <- custom.subsets(eurodist, group = list("Northern" = Northern,
                                                        "Southern" = Southern))

## Bootstrapping and rarefying to 9 elements (the number of Northern cities)
eurodist_bs <- boot.matrix(eurodist_subsets, rarefaction = 9)

## Measuring disparity as the median distance from group's centroid
euro_disp <- dispRity(eurodist_bs, metric = c(median, centroids))

## Testing the differences using a simple wilcox.test
euro_diff <- test.dispRity(euro_disp, test = wilcox.test)
euro_diff_rar <- test.dispRity(euro_disp, test = wilcox.test, rarefaction = 9)
```

We can compare this approach to an ordination one:

```{r}
## Ordinating the eurodist matrix
euro_ord <- cmdscale(eurodist, k = nrow(eurodist) - 2)

## Calculating disparity on the bootstrapped and rarefied subset data
euro_ord_disp <- dispRity(boot.matrix(custom.subsets(euro_ord, group =
        list("Northern" = Northern, "Southern" = Southern)), rarefaction = 9),
        metric = c(median, centroids))

## Testing the differences using a simple wilcox.test
euro_ord_diff <- test.dispRity(euro_ord_disp, test = wilcox.test)
euro_ord_diff_rar <- test.dispRity(euro_ord_disp, test = wilcox.test, rarefaction = 9)
```

And visualise the differences:

```{r, fig.width=12, fig.height=12}
## Plotting the differences
par(mfrow = c(2,2), bty = "n")
## Plotting the normal disparity
plot(euro_disp, main = "Distance differences")
## Adding the p-value
text(1.5, 4000, paste0("p=",round(euro_diff[[2]][[1]], digit = 5)))
## Plotting the rarefied disparity
plot(euro_disp, rarefaction = 9, main = "Distance differences (rarefied)")
## Adding the p-value
text(1.5, 4000, paste0("p=",round(euro_diff_rar[[2]][[1]], digit = 5)))

## Plotting the ordinated disparity
plot(euro_ord_disp, main = "Ordinated differences")
## Adding the p-value
text(1.5, 1400, paste0("p=",round(euro_ord_diff[[2]][[1]], digit = 5) ))
## Plotting the rarefied disparity
plot(euro_ord_disp, rarefaction = 9, main = "Ordinated differences (rarefied)")
## Adding the p-value
text(1.5, 1400, paste0("p=",round(euro_ord_diff_rar[[2]][[1]], digit = 5) ))

```

As expected, the results are pretty similar in pattern but different in terms of scale.
The median centroids distance is expressed in km in the "Distance differences" plots and in Euclidean units of variation in the "Ordinated differences" plots.

## Disparity from multiple matrices (and multiple trees!) {#multi.input}

Since the version `1.4` of this package, it is possible to use multiple trees and multiple matrices in `dispRity` objects.
To use multiple matrices, this is rather easy: just supply a list of matrices to any of the `dispRity` functions and, as long as they have the same size and the same rownames they will be handled as a distribution of matrices.

```{r}
set.seed(1)
## Creating 3 matrices with 4 dimensions and 10 elements each (called t1, t2, t3, etc...)
matrix_list <- replicate(3, matrix(rnorm(40), 10, 4, dimnames = list(paste0("t", 1:10))),
                         simplify = FALSE)
class(matrix_list) # This is a list of matrices

## Measuring some disparity metric on one of the matrices
summary(dispRity(matrix_list[[1]], metric = c(sum, variances)))

## Measuring the same disparity metric on the three matrices
summary(dispRity(matrix_list, metric = c(sum, variances)))
```

As you can see, when measuring the sum of variances on multiple matrices, we now have a distribution of sum of variances rather than a single observed value.

Similarly as running disparity analysis using multiple matrices, you can run the `chrono.subsets` function using multiple trees.
This can be useful if you want to use a tree posterior distribution rather than a single consensus tree.
These trees can be passed to `chrono.subsets` as a `"multiPhylo"` object (with the same node and tip labels in each tree).
First let's define a function to generate multiple trees with the same labels and root ages:

```{r}
set.seed(1)
## Matches the trees and the matrices
## A bunch of trees
make.tree <- function(n, fun = rtree) {
    ## Make the tree
    tree <- fun(n)
    tree <- chronos(tree, quiet = TRUE,
                    calibration = makeChronosCalib(tree, age.min = 10, age.max = 10))
    class(tree) <- "phylo"
    ## Add the node labels
    tree$node.label <- paste0("n", 1:Nnode(tree))
    ## Add the root time
    tree$root.time <- max(tree.age(tree)$ages)
    return(tree)
}
trees <- replicate(3, make.tree(10), simplify = FALSE)
class(trees) <- "multiPhylo"
trees
```

We can now simulate some ancestral states for the matrices in the example above to have multiple matrices associated with the multiple trees.

```{r}
## A function for running the ancestral states estimations
do.ace <- function(tree, matrix) {
    ## Run one ace
    fun.ace <- function(character, tree) {
        results <- ace(character, phy = tree)$ace
        names(results) <- paste0("n", 1:Nnode(tree))
        return(results)
    }
    ## Run all ace
    return(rbind(matrix, apply(matrix, 2, fun.ace, tree = tree)))
}

## All matrices
matrices <- mapply(do.ace, trees, matrix_list, SIMPLIFY = FALSE)
```

Let's first see an example of time-slicing with one matrix and multiple trees.
This assumes that your tip values (observed) and node values (estimated) are fixed with no error on them.
It also assumes that the nodes in the matrix always corresponds to the node in the trees (in other words, the tree topologies are fixed):

```{r}
## Making three "proximity" time slices across one tree
one_tree <- chrono.subsets(matrices[[1]], trees[[1]],
                           method = "continuous", model = "proximity", time = 3)
## Making three "proximity" time slices across the three trees
three_tree <- chrono.subsets(matrices[[1]], trees,
                             method = "continuous", model = "proximity", time = 3)
## Measuring disparity as the sum of variances and summarising it
summary(dispRity(one_tree, metric = c(sum, variances)))
summary(dispRity(three_tree, metric = c(sum, variances)))
```

This results show the effect of considering a tree distribution: in the first case (`one_tree`) the time slice at 3.95 Mya has a sum of variances of 2.9 but this values goes down to 0.256 in the second case (`three_tree`) which is due to the differences in branch lengths distributions:

```{r, print = FALSE}
par(mfrow = c(3,1))
slices <- c(7.9, 3.95, 0)
fun.plot <- function(tree) {
  plot(tree)
  nodelabels(tree$node.label, cex = 0.8)
  axisPhylo()
  abline(v = tree$root.time - slices)
}
silent <- lapply(trees, fun.plot)
```

Note that in this example, the nodes are actually even different in each tree! The node `n4` for example, is not direct descendent of `t4` and `t6` in all trees!
To fix that, it is possible to input a list of trees and a list of matrices that correspond to each tree in `chrono.subsets` by using the `bind.data = TRUE` option.
In this case, the matrices need to all have the same row names and the trees all need the same labels as before:

```{r}
## Making three "proximity" time slices across three trees and three bound matrices
bound_data <- chrono.subsets(matrices, trees,
                             method = "continuous", model = "proximity",
                             time = 3, bind.data = TRUE)
## Making three "proximity" time slices across three trees and three matrices
unbound_data <- chrono.subsets(matrices, trees,
                               method = "continuous", model = "proximity",
                               time = 3, bind.data = FALSE)

## Measuring disparity as the sum of variances and summarising it
summary(dispRity(bound_data, metric = c(sum, variances)))
summary(dispRity(unbound_data, metric = c(sum, variances)))
```

Note here that the results are again rather different: with the bound data, the slices are done across the three trees and each of their corresponding matrix (resulting in three observation) which is more accurate than the previous results from `three_trees` above.
With the unbound data, the slices are done across the three trees and applied to the three matrices (resulting in 9 observations).
As we've seen before, this is incorrect in this case since the trees don't have the same topology (so the nodes selected by a slice through the second tree are not equivalent to the nodes in the first matrix) but it can be useful if the topology is fixed to integrate both uncertainty in branch length (slicing through different trees) and uncertainty from, say, ancestral states estimations (applying the slices on different matrices).